### 2018-10-04 manuscript.Rmd
initial set-up

---
title: Improving the communication and transparency of stock assessment using interactive
  visualization tools
author: "Paul M Regular, Gregory J Robertson, etc.?"
output: word_document
---

### 2018-10-04 manuscript.Rmd
Set-up document template and references

@@
 author: "Paul M Regular, Gregory J Robertson, etc.?"
-output: word_document
+date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
+output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
+bibliography: references.bib
 ---
+
+# Introduction
+
+A pimary role of stock assessment is to provide fisheries managers with the information needed to adequatly manage a fishery.
+Stock assessment involves the use of various sources of data and statistical methods to determine the status of one or more fish stocks and to make quantative predictions of the consequences of different management choices [@hilborn1992].
+
+# References

### 2018-10-05 manuscript.Rmd
Draft rough first paragraph of paper

@@
 
-A pimary role of stock assessment is to provide fisheries managers with the information needed to adequatly manage a fishery.
-Stock assessment involves the use of various sources of data and statistical methods to determine the status of one or more fish stocks and to make quantative predictions of the consequences of different management choices [@hilborn1992].
+A primary role of stock assessment is to provide fisheries managers with the information needed to adequately manage a fishery.
+Stock assessment involves the use of various sources of data and statistical methods to determine the status of one or more fish stocks and to make quantitative predictions of the consequences of different management choices [@hilborn1992].
+A wide array data may be collected for an assessment and, formally, an assessment often boils down to algorithms that convert these data to advice for policy decisions.
+In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
+
+Data from monitoring programs...
 

### 2018-10-11 manuscript.Rmd
Draft second paragraph focused on data and start thrid on models

@@
 Stock assessment involves the use of various sources of data and statistical methods to determine the status of one or more fish stocks and to make quantitative predictions of the consequences of different management choices [@hilborn1992].
-A wide array data may be collected for an assessment and, formally, an assessment often boils down to algorithms that convert these data to advice for policy decisions.
+A wide array data may be collected for an assessment and, formally, an assessment often reduces to algorithms that convert these data to advice for fisheries managers.
 In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
 
-Data from monitoring programs...
+The datasets used in stock assessments are constantly growing.
+This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
+As such, stock assessment biologists often have to manage large volumes of data from a variaty of sources.
+For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments.
+These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
+For some data-rich stocks, mark and recapture studies are also caried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
+All of the abovementioned datasets are multidimensional and as the volume and viriaty of these data increases, it becomes more difficult to be aware of the details of each dataset and to synthesize the results.
+
+Synthesizing data from multiple sources presents a key challenge to stock assessment.
+Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model.
+This approach, however, is less than ideal because information may be lost and uncertinty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
+
+... Advances in software have greatly facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. ...
+
+"We are in an age of data‐intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]
 

### 2018-10-12 manuscript.Rmd
Draft paragraph on stock assessment models

@@
 
-The datasets used in stock assessments are constantly growing.
+The data sets used in stock assessments are constantly growing.
 This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
-As such, stock assessment biologists often have to manage large volumes of data from a variaty of sources.
+As such, stock assessment biologists often have to manage large volumes of data from a variety of sources.
 For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments.
 These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
-For some data-rich stocks, mark and recapture studies are also caried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
-All of the abovementioned datasets are multidimensional and as the volume and viriaty of these data increases, it becomes more difficult to be aware of the details of each dataset and to synthesize the results.
+For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
+All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and to synthesize the results.
 
@@
 Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model.
-This approach, however, is less than ideal because information may be lost and uncertinty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
+This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
+Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
+Specifically, the application of statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
+However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017].
+A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017].
+Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000].
+These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.
 
-... Advances in software have greatly facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013]. ...
-
-"We are in an age of data‐intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]
+"We are in an age of data-intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]
 

### 2018-10-25 manuscript.Rmd
Drafting paragraph on interactive visualization tools

@@
 # Introduction
+
+In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
+Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots are divorced from the underlying data.
+This limits our ability to explore the underlying data in more detail by, for instance, zooming into areas of interest.
+New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
+In particular, browser-based data visualizations have been playing an increasing prominent role in communicating information on a wide range of topics.
+Interactive graphics are frequently utilized by media outlets, such as the New York Times, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
+
+JavaScript libraries, such as D^3^ [@bostock2011], for visualizing data on modern web-browsers now match or exceed the abilities of most platform-specific software and the development of high-level libraries ...
+plotly ...
+python ...
+R
 
@@
 
-"We are in an age of data-intensive science and big data, and ecologists must develop the capabilities to deal with their data." [@hampton2013]
+In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models.
+The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits.
+Data presented in tables is incredibly valuable, but as human beings, seeing patterns can be challenging.
+Static figures, in contrast, help reveal patterns but they are divorced from the underlying data and, as such, limit detailed explorations.
+New tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
+
+Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization
+
+https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/
+
+Good example from Keith: https://www.nature.com/articles/s41467-018-03297-7
+
+https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103
+
+https://www.nature.com/articles/d41586-018-01322-9
 

### 2018-10-30 manuscript.Rmd
Draft new first paragraph on interactive graphics

@@
 In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
-Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots are divorced from the underlying data.
-This limits our ability to explore the underlying data in more detail by, for instance, zooming into areas of interest.
+Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data.
 New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
@@
 Interactive graphics are frequently utilized by media outlets, such as the New York Times, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
+This trend has been fueled by the development of JavaScript libraries (e.g.
+D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers.
+Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python].
+By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of Big Data by facilitating the exploration and communication of complex data and models.
 
-JavaScript libraries, such as D^3^ [@bostock2011], for visualizing data on modern web-browsers now match or exceed the abilities of most platform-specific software and the development of high-level libraries ...
-plotly ...
-python ...
-R
+The field of stock assessment is no stranger to complex data and models...
 

### 2018-11-05 manuscript.Rmd
Draft a new second paragraph

@@
 
-The field of stock assessment is no stranger to complex data and models...
+The field of fisheries research has a long history of collecting and synthesizing a plethora of data using state-of-the-art computational technologies [@megrey2008].
+Fisheries science has evolved alongside the rise of the computing era, and as tools improve so does the management advice provided by fisheries scientists.
+This management advice generally culminates from a stock assessment process whereby various sources of data and statistical methods are used to model the history, current status and future status of one or more fish stocks [@hilborn1992].
+Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated stock assessment models [@maunder2013].
+This means that analyses that were once carried out independently on different data sets can now be integrated into the one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015].
+From a computational perspective, analyses of a variety of large data sets has never been easier.
+However, from a human perspective, contemporary stock assessment biologists are faced with the challenge managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
 
-A primary role of stock assessment is to provide fisheries managers with the information needed to adequately manage a fishery.
-Stock assessment involves the use of various sources of data and statistical methods to determine the status of one or more fish stocks and to make quantitative predictions of the consequences of different management choices [@hilborn1992].
-A wide array data may be collected for an assessment and, formally, an assessment often reduces to algorithms that convert these data to advice for fisheries managers.
+Note: still not liking the flow between the first and second paragraph. Still seems like I need to cut to the chase sooner...how can interactive visuals help?
+
+A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
 In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
+Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.
+
+...need to get to the point here quicker...what is the problem and how can interactive visualization tools help?
 
@@
 
-Good example from Keith: https://www.nature.com/articles/s41467-018-03297-7
+https://www.nature.com/articles/s41467-018-03297-7
 
@@
 
+"The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)
+
 # References

### 2018-11-08 manuscript.Rmd
Rough out two new sections by re-structuring some previous text

@@
 In particular, browser-based data visualizations have been playing an increasing prominent role in communicating information on a wide range of topics.
-Interactive graphics are frequently utilized by media outlets, such as the New York Times, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
+Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
 This trend has been fueled by the development of JavaScript libraries (e.g.
@@
 This management advice generally culminates from a stock assessment process whereby various sources of data and statistical methods are used to model the history, current status and future status of one or more fish stocks [@hilborn1992].
-Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated stock assessment models [@maunder2013].
+Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated assessment models [@maunder2013].
 This means that analyses that were once carried out independently on different data sets can now be integrated into the one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015].
 From a computational perspective, analyses of a variety of large data sets has never been easier.
-However, from a human perspective, contemporary stock assessment biologists are faced with the challenge managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
+However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
+Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs.
+First, we focus on data sets that are commonly used in stock assessment and provide one example where an interactive maps were used to simplify the detailed exploration of data from a long-term tagging program.
+Second, we focus on the modeling aspect of stock assessment and, again, we use an example to demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
 
-Note: still not liking the flow between the first and second paragraph. Still seems like I need to cut to the chase sooner...how can interactive visuals help?
-
-A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
-In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
-Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.
-
-...need to get to the point here quicker...what is the problem and how can interactive visualization tools help?
+## Data
 
@@
 All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and to synthesize the results.
+
+## Models
 
@@
 These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.
+
+# Other draft text
+
+A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
+In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
+Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.
 

### 2018-11-09 manuscript.Rmd
Rough draft of data section intro; set-up for tagging dashboard example

@@
 Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated assessment models [@maunder2013].
-This means that analyses that were once carried out independently on different data sets can now be integrated into the one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015].
+This means that analyses that were once carried out independently on different data sets can now be integrated into one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015].
 From a computational perspective, analyses of a variety of large data sets has never been easier.
@@
 Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs.
-First, we focus on data sets that are commonly used in stock assessment and provide one example where an interactive maps were used to simplify the detailed exploration of data from a long-term tagging program.
+First, we focus on data that are commonly used in stock assessment and provide one example where interactive maps were used to simplify the detailed exploration of data from a long-term tagging program.
 Second, we focus on the modeling aspect of stock assessment and, again, we use an example to demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
+This structure corresponds to the two general steps in the stock assessment process: 1) knowing the data, and 2) data modeling.
 
-## Data
+## Knowing the data
 
@@
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
-All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and to synthesize the results.
+All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
+This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow the construction of visual representations of the data that demonstrates large-scale patterns while also allowing the detailed exploration of the underlying data.
+For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data.
+This is exemplified by a recently developed interactive mapping tool developed for the exploration of a long-term tagging study.
 
-## Models
+- *Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?*
+- *TODO: the intro paragraph to this section could use some references if it is necessary*
+
+## Data modeling
 
@@
 These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.
+
+- *TODO: tweak this intro paragraph and work up sugue to NCAM and the dashboard*
 

### 2018-11-15 manuscript.Rmd
Rough draft of modeling section intro; set-up sugue to NCAM flexdashboard

@@
 New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
-In particular, browser-based data visualizations have been playing an increasing prominent role in communicating information on a wide range of topics.
+In particular, browser-based data visualizations have been playing an increasingly prominent role in communicating information on a wide range of topics.
 Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
@@
 All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
-This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow the construction of visual representations of the data that demonstrates large-scale patterns while also allowing the detailed exploration of the underlying data.
+This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind the plot.
 For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data.
-This is exemplified by a recently developed interactive mapping tool developed for the exploration of a long-term tagging study.
+This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.
 
@@
 Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
-Specifically, the application of statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
+Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
 However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017].
@@
 These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.
+This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice.
+Here we argue that the process of providing fisheries advice can be further streamlined by utilizing dynamic and interactive visualization tools to explore, diagnose and communicate stock assessment models.
+We hope to demonstrate this using a recently developed interactive tool for exploring an integrated assessment model for Northern cod.
 
-- *TODO: tweak this intro paragraph and work up sugue to NCAM and the dashboard*
+# Notes
+
+- Note that these tools facilitate the peer-review process and also improves the transparency and reproducibility of assessment models
+- Dynamic document can be self contained
 

### 2018-11-20 manuscript.Rmd
Write-up brief background on NCAM

@@
 
+The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history.
+As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016].
+The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
+Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
+Once the model is fit to the data, the next step is to produce visual representations of the data and model.
+An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod (**TODO: add reference**).
+The concept of using a dashboard was burrowed from the buisness community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
+Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM.
+Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document.
+We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots in the "Tagging" tab.
+
+TODO:
+- Copy the data and functions required to run the flexdashboard file here
+- Consider taking the flexdashboard outside of the vis_fit function
+- Fix up the projection bits of the code in the Rmd file
+- Take screenshots of the dashboard for static figures for the paper
+- Produce the html file for this paper and include refrence to supplment
+
 # Notes
@@
 - Dynamic document can be self contained
+- Northern cod stock is on an annual assessment schedule, therefore it makes sense to automate visuals
 

### 2018-11-21 analysis/paper/manuscript.Rmd
Re-structure files and add data and scripts required to make the NCAM dashboard

--- 2018-11-20 manuscript.Rmd
@@
 TODO:
-- Copy the data and functions required to run the flexdashboard file here
-- Consider taking the flexdashboard outside of the vis_fit function
-- Fix up the projection bits of the code in the Rmd file
-- Take screenshots of the dashboard for static figures for the paper
-- Produce the html file for this paper and include refrence to supplment
+- Take screenshots of the dashboard for static figures for the paper (use webshot?)
+- Remember to add reference to the supplement that will ultimatly include the NCAM_dashboard.html
 

### 2018-11-21 inst/rmd/vis_fit.Rmd
Re-structure files and add data and scripts required to make the NCAM dashboard

---
title: "NCAM Explorer"
output: flexdashboard::flex_dashboard
always_allow_html: yes
---

<style>
/* list */
li {
line-height: 30px;
}
</style>

```{r setup, echo=FALSE}
library(plotly)
library(Matrix)
library(knitr)
library(data.table)
library(webshot)
knitr::opts_knit$set(root.dir = tmp$output_dir)
knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
```

```{r data, include=FALSE}

cols <- viridis::viridis(4)[-1]
cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
fit <- tmp$fit # temp object from vis_fit function
if (!is.null(tmp$comp)) { comp <- tmp$comp }
if (!is.null(tmp$last)) { last <- tmp$last }
est <- as.list(fit$sd.rep, "Est")
sd <- as.list(fit$sd.rep, "Std")
tabs <- tidy_model(fit)
if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
min_year <- min(fit$tmb.data$years)
max_year <- max(fit$tmb.data$years)
ages <- fit$tmb.data$ages
years <- fit$tmb.data$years
years_plus1 <- fit$tmb.data$years_plus1

## helper functions for nice Rmarkdown text
text_range <- function(x) paste(range(x), collapse = "-")
text_c <- function(x) {
  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
}
text_units <- function(x) {
  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
  if (x >= 1000) return(paste0(round(x / 1000), "k"))
  if (x < 1000) return(as.character(round(x)))
}

```

Background
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### NCAM

<h1>NCAM: A state-space stock assessment model for Northern cod</h1>

```{r}

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)

xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
        fillcolor = 'rgba(49, 104, 142, 0.3)',
        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
  add_trace() %>%
  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
  plotly::config(displayModeBar = FALSE)

```

### General

- An *integrated* *state-space* population dynamics model
- *Integrated*
    - Uses as much information as possible
- *State-space*
    - Includes process and observation equations + error
    - Population processes are informed by observations from multiple sources
    - Process errors are necessary for realistic stochastic projections

### Data inputs

- Reported landings (`r text_range(years)`)
- Catch-at-age information (`r text_range(years)`)
- Reported landings by month (`r text_range(years)`)
- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
- Tagging data (`r text_range(years)`)

### Biological inputs

- RV survey data were used fit cohort-based models of:
   - Proportion mature at age
   - Beginning of year and mid-year stock weights at age
- Baseline natural mortality (M) values

```{r}

plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
  add_surface() %>%
  colorbar(title = "Mshift") %>%
  layout(scene = list(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Age"),
    zaxis = list(title = "Mshift")
  ))

```

### Partial catches

- User sets lower and upper bounds on total catch weight
- Bounds are used via a censored likelihood

```{r}

plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
  add_lines(name = "Lower", color = I(cols[1])) %>%
  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
  add_lines(y = rep(1, length(years)), color = I("black"),
            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
  layout(yaxis = list(title = "Catch multiplier"),
         xaxis = list(title = "Year"))

```

### Change in Q

- A shift in Q was assumed for the offshore RV survey
- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
- Q for the inshore sentinel survey was assumed to follow a random walk
    - Done to account for potential changes in migration patterns

### Tagging data

- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
- Applies adjustments for tagging mortality, tag loss and reporting rates
- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
- The Baranov catch equation was applied to age-based tag survivors
- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock

### References

Cadigan, N. 2015. A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates. Canadian Journal of Fisheries and Aquatic Sciences 73 (2). NRC Research Press: 296-308.

Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.

Catch
================================================================================

Row
--------------------------------------------------------------------------------

### Landings

```{r}

plot_landings(fit)

```

> Y-axes in log scale

Row {.tabset}
--------------------------------------------------------------------------------

### Observed and predicted catch at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
                    pred = fit$rep$pred_prop_catch),
               ages = ages, years = years)
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = cols[2], pred_col = cols[1]) %>%
  layout(yaxis = list(rangemode = "tozero"))

```

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$CA_resid,
                    fit = fit$rep$pred_crl),
               ages = ages[-length(ages)], years = years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

```

> based on continuation ratio logits

### Catch at age surface (proportion)

```{r}
plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
```

<!-- ### Catch at age surface (numbers) -->

<!-- ```{r} -->
<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
<!-- ``` -->

<!-- > Note: model fits to composition - this plot is included for illustration -->

RV Survey
================================================================================

Row
--------------------------------------------------------------------------------

### RV survey index

```{r}

obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
obs[obs == 0] <- NA
pred[pred == 0] <- NA
d <- data.frame(year = years, obs = obs, pred = pred)

# ## RV biomass
# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
# d <- data.frame(year = years, obs = obs, pred = pred)
# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)

plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
              showlegend = TRUE, type = "log")

```

> Y-axes in log scale

### Standardized log residuals

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    zero = fit$tmb.data$ind_zero,
                    wt = fit$tmb.data$est_wt),
               ages = ages, years = years)
d <- d[d$wt == 1, ]
bubble_plot(d)

```

Row {.tabset}
--------------------------------------------------------------------------------

### RV survey index at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
                    pred = exp(fit$rep$Elog_RV_index),
                    wt = fit$tmb.data$est_wt),
               ages = ages, years = years)
d$obs[d$wt == 0] <- NA
d$pred[d$wt == 0] <- NA
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Mean numbers per tow",
               obs_col = cols[2], pred_col = cols[1], type = "log")

```

> Y-axes in log scale

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$RV_resid,
                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
                    fit = fit$rep$Elog_RV_index),
               ages = ages, years = years)
d <- d[d$keep, ]
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

```

SN Survey
================================================================================

Row
--------------------------------------------------------------------------------

### Sentinel gillnet survey index

```{r}

obs <- colSums(fit$tmb.data$SN_GN_index)
pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)

plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
              showlegend = TRUE, type = "log")

```

> Y-axes in log scale

### Standardized log residuals

```{r}

d <- tidy_mat(fit$rep$log_SN_GN_resid,
              ages = fit$tmb.data$SN_ages,
              years = fit$tmb.data$SN_years, value = "res")
bubble_plot(d)

```

Row {.tabset}
--------------------------------------------------------------------------------

### Sentinel gillnet survey index at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
                    pred = exp(fit$rep$Elog_SN_GN_index)),
               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
facet_obs_pred(d, by = "age",
               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
               nrows = 4, ylab = "Sentinel survey index",
               obs_col = cols[2], pred_col = cols[1], type = "log")

```

> Y-axes in log scale

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
                    fit = fit$rep$Elog_SN_GN_index),
               ages = fit$tmb.data$SN_ages,
               years = fit$tmb.data$SN_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

```

SS Survey
================================================================================

Row
--------------------------------------------------------------------------------

### Smith sound acoustic biomass estimates

```{r}

d <- data.frame(year = fit$tmb.data$SS_bms_years,
                obs = fit$tmb.data$SSbms,
                pred = fit$rep$ESSbms)
plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
              showlegend = TRUE)

```

### Standardized log residuals

```{r}

d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
              ages = fit$tmb.data$SS_ages,
              years = fit$tmb.data$SS_age_years, value = "res")
bubble_plot(d)

```

Row {.tabset}
--------------------------------------------------------------------------------

### Smith Sound trawl numbers at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
                    pred = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)
facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
               obs_col = cols[2], pred_col = cols[1])

```

### Residuals by year, cohort, age and expected values

```{r}

d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
                    fit = fit$rep$ESS_agecomps),
               ages = fit$tmb.data$SS_ages,
               years = fit$tmb.data$SS_age_years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = cols[1])

```

Tagging {data-orientation=rows}
================================================================================

```{r}

## Prep tag data and fits
tag_dat <- data.frame(
  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
             rep("post 1996", length(fit$tmb.data$tag_exp)))
)
tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
                      "age:", tag_dat$age, "</br>",
                      "delay:", tag_dat$delay, "</br>")

Udev <- data.frame(
  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
)

Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
names(Fay)[names(Fay) == "year"] <- "rel_year"

tag_dat <- merge(tag_dat, Udev, by = "exp")
tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)

## Share data across multiple plots
d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)

```

Row {data-height=700 .tabset}
--------------------------------------------------------------------------------

### Observed and predicted tag recaptures by experimemnt

```{r}

y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
y_tickvals <- sqrt(y_ticktext)

rel_plot <- base %>%
  filter(rel_year == cap_year) %>%
  group_by(exp) %>%
  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N released",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year"))

y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
y_tickvals <- sqrt(y_ticktext)

cap_plot <- base %>%
  group_by(exp, rel_year, cap_year) %>%
  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
            period = unique(period)) %>%
  group_by(exp) %>%
  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
            alpha = 0.7, showlegend = FALSE, symbol = NULL,
            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y2",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N recaptured",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year"))

subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
        nrows = 2, heights = c(0.3, 0.7))

```

> Y-axis is in square root scale

<!-- ### Animation -->

<!-- ```{r} -->

<!-- sum_d <- tag_dat %>% -->
<!--   group_by(exp, cap_year) %>% -->
<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->

<!-- sum_d %>%  -->
<!--   group_by(exp) %>%  -->
<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
<!--   layout(yaxis = list(title = "N recaptured", -->
<!--                       tickvals = y_tickvals, -->
<!--                       ticktext = y_ticktext, -->
<!--                       range = range(y_tickvals)), -->
<!--          xaxis = list(title = "Year")) -->

<!-- ``` -->

Row
--------------------------------------------------------------------------------

### Tag reporting rate

```{r}

cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
rr <- data.frame(year = fit$tmb.data$SRR_years,
                 est = fit$tmb.data$log_SRR_est,
                 sd = sqrt(diag(cov_mat)))
rr$lwr <- rr$est - 1.96 * rr$sd
rr$upr <- rr$est + 1.96 * rr$sd
rr <- trans_est(rr)
rr$model <- "Random Walk"

ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
                      est = est$log_SRR, sd = sd$log_SRR)
ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
ncam_rr <- trans_est(ncam_rr)
ncam_rr$model <- "NCAM"

rr <- rbind(rr, ncam_rr)

plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)

```

### Initial F deviations for tagging experiments

```{r}

base %>%
  distinct(rel_year, Udev, exp, period) %>%
  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
              showlegend = FALSE) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "Initial F deviation"),
         xaxis = list(title = "Year"))

```

### Observed vs. predicted recaptures

```{r}

x_ticktext <- c(5, pretty(tag_dat$pred, 10))
y_ticktext <- c(5, pretty(tag_dat$obs, 10))
x_tickvals <- sqrt(x_ticktext)
y_tickvals <- sqrt(y_ticktext)

base %>%
  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
                       line = list(width = 1)),
         xaxis = list(title = "Predicted number of recaptures",
                      tickvals = x_tickvals,
                      ticktext = x_ticktext),
         yaxis = list(title = "Observed number of recaptures",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext))

```

> Axes are in square root scale

### Standardized residuals vs. predicted recaptures

```{r}

y_ticktext <- c(5, pretty(tag_dat$res, 10))
y_tickvals <- sqrt(x_ticktext)

base %>%
  add_markers( x = ~sqrt(pred), y = ~res,
               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
  layout(xaxis = list(title = "Predicted number of recaptures",
                      tickvals = x_tickvals,
                      ticktext = x_ticktext),
         yaxis = list(title = "Standardized residuals"))

```

> X-axis is in square root scale

Catchability
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Fishery selectivity

```{r}

S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
plot_surface(S, ages = ages, years = years, name = "S")

```

Row {.tabset}
--------------------------------------------------------------------------------

### RV survey catchability

```{r}

q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
plot_surface(q, ages = ages, years = years, name = "q")

```

### Proportion of the stock in Smith Sound

```{r}
p <- matrix(0, nrow = length(ages), ncol = length(years),
            dimnames = list(ages, years))
ss_p <- fit$rep$SSD
ss_years <- fit$tmb.data$SS_years
ss_ages <- fit$tmb.data$SS_ages
rownames(ss_p) <- ss_ages
colnames(ss_p) <- ss_years
p[as.character(ss_ages), as.character(ss_years)] <- ss_p
plot_surface(p, ages = ages, years = years, name = "p")

```

### Sentinel gillnet survey catchability

```{r}

log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
log_qa <- fit$rep$log_SN_GN_Qa
q <- exp(log_qy + log_qa)
plot_surface(q, ages = fit$tmb.data$SN_ages,
             years = fit$tmb.data$SN_years,
             name = "q")

```

Trends
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Stock status

```{r}

## Note: biomass is in tonnes and numbers are in thousands
bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
             data.frame(metric = "Ages 2+", tabs$log_biomass))
abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
             data.frame(metric = "Ages 2+", tabs$log_abundance))

bms_plot <- trans_est(bms) %>%
  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
             annotate = TRUE)

abn_plot <- trans_est(abn, scale = 0.001) %>%
  plot_trend(ylab = "Abundance", group = "metric", col = cols,
             annotate = TRUE)

blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
           color = I("black"), textposition = "right", hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

### Abundance at age

```{r}
plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
```

### Biomass at age

```{r}
plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
```

### SSB at age

```{r}
plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
```

Row {.tabset}
--------------------------------------------------------------------------------

### Mortality Rates

```{r}

f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
m <- data.frame(metric = "Ages 5+", tabs$log_aveM)

z_plot <- trans_est(z) %>%
  plot_trend(ylab = "Average Z", group = "metric",
             annotate = TRUE, col = cols)

m_plot <- trans_est(m) %>%
  plot_trend(ylab = "Average M", group = "metric",
             annotate = TRUE, col = cols)

f_plot <- trans_est(f) %>%
  plot_trend(ylab = "Average F", group = "metric",
             annotate = TRUE, col = cols)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

### Z at age

```{r}
plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
```

### M at age

```{r}
plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
```

### F at age

```{r}
plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
```

### Z, M and F

```{r}
plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
                 ages = ages, years = years)
```

> Z = purples, M = blues, F = reds

### Process error

```{r}

pe <- fit$rep$pe
plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
              color = I("black"), opacity = 0.8)

```

### Process error (bubble plot)

```{r}
pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
               ages = 2:8, value = "res")
bubble_plot(pe)
```

Retro {`r if (is.null(fit$retro)) ".hidden"`}
================================================================================

Row
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(fit$retro)}

bms_plot <- trans_est(retro_tabs$log_ssb) %>%
  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])

abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])

blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Row
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(fit$retro)}

z_plot <- trans_est(retro_tabs$log_aveZ) %>%
  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])

m_plot <- trans_est(retro_tabs$log_aveM) %>%
  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])

f_plot <- trans_est(retro_tabs$log_aveF) %>%
  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Comps {`r if (is.null(tmp$comp)) ".hidden"`}
================================================================================

Row
--------------------------------------------------------------------------------

### Stock status

```{r, eval = !is.null(tmp$comp)}

abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
  plot_comps(ylab = "Recruitment (age 2)", col = cols)

bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)

blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)

subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Row
--------------------------------------------------------------------------------

### Mortality Rates

```{r, eval = !is.null(tmp$comp)}

z_plot <- trans_est(comp_tabs$log_aveZ) %>%
  plot_comps(ylab = "Average Z (ages 5+)", col = cols)

m_plot <- trans_est(comp_tabs$log_aveM) %>%
  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)

f_plot <- trans_est(comp_tabs$log_aveF) %>%
  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)

subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)

```

Productivity
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Recruitment

```{r}
trans_est(tabs$log_recruitment, scale = 0.001) %>%
  plot_trend(ylab = "Recruitment", col = cols[1])
```

### SSB

```{r}
trans_est(tabs$log_ssb) %>%
  plot_trend(ylab = "SSB (t)", col = cols[1])
```

### RPS

```{r}
trans_est(tabs$log_rps) %>%
  filter(year >= min_year + min(ages)) %>% 
  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
  layout(xaxis = list(range = c(min_year, max_year + 1)))
```

### SSB x Recruitment

```{r}

ssb <- trans_est(tabs$log_ssb)
rec <- trans_est(tabs$log_recruitment, scale = 0.001)
rps <- trans_est(tabs$log_rps)
names(ssb) <- paste0("ssb_", names(ssb))
names(rec) <- paste0("rec_", names(rec))
names(rps) <- paste0("rps_", names(rps))
ssb$year <- ssb$ssb_year + min(ages)
rec$year <- rec$rec_year
rps$year <- rps$rps_year
sr <- merge(ssb, rec, by = "year")
sr <- merge(sr, rps, by = "year")
sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)

plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
                 col = cols[1])

```

### SSB x RPS

```{r}
p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
                 col = cols[1])
p
```

### log(SSB) x log(RPS)

```{r}
p %>% layout(yaxis = list(type = "log"),
             xaxis = list(type = "log"))
```

```{r}

message("Bits of the projection code is currently not working. Specifically the plots that require the frame argument in plotly.")

```

<!-- Assumptions  {data-navmenu="Projections" .storyboard} -->
<!-- ================================================================================ -->

<!-- ### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers -->

<!-- ```{r} -->

<!-- Cp <- fit$tmb.data$Cp -->
<!-- Cp_string <- text_c(Cp) -->
<!-- Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1)) -->
<!-- Lp_string <- paste(text_c(Lp), "t") -->

<!-- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>% -->
<!--   filter(year >= max(years) - 4) %>% -->
<!--   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>% -->
<!--   add_lines(alpha = 0.5) %>% -->
<!--   filter(year == max(years)) %>% -->
<!--   add_lines(showlegend = FALSE, line = list(width = 3)) %>% -->
<!--   layout(xaxis = list(title = "Age")) -->

<!-- ``` -->

<!-- *** -->
<!-- - The following catch multipliers were applied: `r #Cp_string` -->
<!-- - This translates to `r #Lp_string` of catch in the projection years -->

<!-- ### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994 -->

<!-- ```{r} -->

<!-- if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic") -->

<!-- aveM <- tabs$log_aveM -->
<!-- aveM <- exp(mean(aveM$est[!aveM$year %in% 1991:1994])) -->

<!-- trans_est(tabs$log_aveM) %>% -->
<!--   plot_trend(ylab = "Average M", col = cols[1]) %>% -->
<!--   add_segments(x = 1995, xend = max(years) + 3, y = aveM, yend = aveM, -->
<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
<!--   add_segments(x = 1990, xend = 1995, y = aveM, yend = aveM, -->
<!--                name = "long-term mean", color = I("black"), -->
<!--                line = list(width = 1, dash = "dot")) %>% -->
<!--   add_segments(x = min(years), xend = 1990, y = aveM, yend = aveM, -->
<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
<!--   add_text(x = max(years) + 0.2, y = aveM, text = "long-term mean", textposition = "bottom right", -->
<!--            color = I("black")) %>% -->
<!--   layout(xaxis = list(range = c(min(years), max(years) + 3))) -->

<!-- ``` -->

<!-- ### Assume a relationship between recruits per spawner and number of spawners to project recruits in the future -->

<!-- ```{r} -->

<!-- if (fit$tmb.data$Rp_mode != 3) stop("TODO: make the recruitment mode documentation dynamic") -->

<!-- rps_tab <- tabs$log_rps -->
<!-- names(rps_tab) <- c("year", "rps", "rps_sd", "rps_lwr", "rps_upr") -->
<!-- rps_tab <- rps_tab[rps_tab$year >= 1985, ] -->
<!-- rps_tab$year <- rps_tab$year - 2 # scale back two year to allign with spawners that produced the recruits -->
<!-- ssa_tab <- tabs$log_ssa -->
<!-- names(ssa_tab) <- c("year", "ssa", "ssa_sd", "ssa_lwr", "ssa_upr") -->
<!-- d <- merge(ssa_tab, rps_tab, by = "year") -->
<!-- rec_tab <- tabs$log_recruitment -->
<!-- names(rec_tab) <- c("year", "rec", "rec_sd", "rec_lwr", "rec_upr") -->
<!-- rec_tab <- rec_tab[rec_tab$year >= 1985, ] -->
<!-- rec_tab$year <- rec_tab$year - 2 # scale back two years -->
<!-- d <- merge(d, rec_tab, by = "year") -->
<!-- d$excluded <- d$year %in% 1991:1994 -->

<!-- sub_d <- d[!d$year %in% 1991:1994, ] -->
<!-- lin_rel <- lm(rps ~ ssa, data = sub_d) -->
<!-- # coefficients(lin_rel) -->

<!-- preds <- data.frame(year = NA, rps_sd = NA, ssa_sd = NA, ssa = seq(min(d$ssa), max(d$ssa), length.out = 100)) -->
<!-- preds$fits <- predict(lin_rel, newdata = preds) -->
<!-- preds$excluded <- FALSE -->

<!-- plot_ly(data = d, x = ~ssa, y = ~rps, text = ~year, -->
<!--         type = "scatter", mode = "markers", name = "est", -->
<!--         color = ~excluded, colors = c(cols[2], "grey"), -->
<!--         error_y = list(array = ~ qnorm(0.975) * rps_sd, -->
<!--                        width = 0, thickness = 1), -->
<!--         error_x = list(array = ~ qnorm(0.975) * ssa_sd, -->
<!--                        width = 0, thickness = 1)) %>% -->
<!--   add_text(text = ~year, textfont = list(color = toRGB("grey50")), textposition = "top right", -->
<!--            error_y = list(visible = FALSE), -->
<!--            error_x = list(visible = FALSE)) %>% -->
<!--   layout(title = "Spawners vs Recruits / Spawner", -->
<!--          xaxis = list(title = "log(Spawners)"), -->
<!--          yaxis = list(title = "log(Recruits / Spawner)", zerolinecolor = toRGB("lightgrey")), -->
<!--          showlegend = FALSE) %>% -->
<!--   add_lines(data = preds, x = ~ssa, y = ~fits, name = "fit", color = I(cols[1])) -->

<!-- ``` -->

<!-- *** -->
<!-- - Grey points indicate years excluded from the regressionn -->

<!-- ### Assume averages weights and proportions mature from the last three years -->

<!-- ```{r} -->

<!-- mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>% -->
<!--   filter(year >= max(years) - 2) %>% -->
<!--   plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>% -->
<!--   add_lines(alpha = 0.5) %>% -->
<!--   group_by(age) %>% -->
<!--   summarise(mat = mean(mat), year = "3y mean") %>% -->
<!--   add_lines(line = list(width = 3)) %>% -->
<!--   layout(xaxis = list(title = "Age"), -->
<!--          yaxis = list(title = "Proportion mature")) -->

<!-- wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>% -->
<!--   filter(year >= max(years) - 2) %>% -->
<!--   plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols), -->
<!--           showlegend = FALSE) %>% -->
<!--   add_lines(alpha = 0.5) %>% -->
<!--   group_by(age) %>% -->
<!--   summarise(weight = mean(weight), year = "3y mean") %>% -->
<!--   add_lines(line = list(width = 3)) %>% -->
<!--   layout(xaxis = list(title = "Age"), -->
<!--          yaxis = list(title = "Weight (kg)")) -->

<!-- subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE) -->

<!-- ``` -->

<!-- Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
<!-- ================================================================================ -->

<!-- ### Catch multiplier projections for SSB / Blim from the last RAP -->

<!-- ```{r, eval = !is.null(tmp$last)} -->

<!-- d <- trans_est(last_tabs$Cproj_log_ssb_lrp) -->
<!-- d$scenario <- factor(d$scenario) -->
<!-- d <- d[d$year <= max(years) + 1, ] -->

<!-- past_d <- trans_est(last_tabs$log_ssb_lrp) -->
<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
<!-- past_d <- do.call(rbind, past_d) -->
<!-- d <- rbind(past_d, d) -->
<!-- xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1) -->
<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->

<!-- current_d <- trans_est(tabs$log_ssb_lrp) -->
<!-- current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s)) -->
<!-- current_d <- do.call(rbind, current_d) -->
<!-- current_d$model <- "current RAP" -->
<!-- d$model <- "last RAP" -->
<!-- d <- rbind(d, current_d) -->

<!-- plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model", -->
<!--           vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE) -->

<!-- ``` -->

<!-- Results {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
<!-- ================================================================================ -->

<!-- Row {.tabset} -->
<!-- -------------------------------------------------------------------------------- -->

<!-- ### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~ -->

<!-- ```{r} -->

<!-- d <- trans_est(tabs$Cproj_log_ssb_ratio) -->
<!-- d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100 -->
<!-- d$scenario <- factor(d$scenario) -->
<!-- xlim <- c(max(years) + 0.9, max(years) + 4.1) -->
<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->

<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%", -->
<!-- #           vline = max(years) + 1, col = cols[1]) %>% -->
<!-- #   layout(yaxis = list(tickvals = seq(-200, 200, by = 25))) -->
<!-- message("plot_proj not working properly at the moment") -->

<!-- ``` -->

<!-- ### SSB / Blim -->

<!-- ```{r} -->

<!-- d <- trans_est(tabs$Cproj_log_ssb_lrp) -->
<!-- d$scenario <- factor(d$scenario) -->

<!-- past_d <- trans_est(tabs$log_ssb_lrp) -->
<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
<!-- past_d <- do.call(rbind, past_d) -->
<!-- d <- rbind(past_d, d) -->
<!-- xlim <- c(max(years) - 2.1, max(years) + 4.1) -->
<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->

<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", -->
<!-- #           vline = max(years) + 1, hline = 1, col = cols[1]) -->
<!-- message("plot_proj not working properly at the moment") -->

<!-- ``` -->

<!-- Row {.tabset} -->
<!-- -------------------------------------------------------------------------------- -->

<!-- ### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~) -->

<!-- ```{r} -->

<!-- growth <- tabs$Cproj_log_ssb_ratio -->
<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->

<!-- d <- growth[growth$year != max(years) + 1, ] -->
<!-- d$scenario <- factor(d$scenario) -->
<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
<!--   add_heatmap() %>% -->
<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
<!--   layout(yaxis = list(title = "Catch multiplier"), -->
<!--          xaxis = list(title = "Year")) -->

<!-- ``` -->

<!-- ### Prob(SSB~yp~ &ge; B~lim~) -->

<!-- ```{r} -->

<!-- growth <- tabs$Cproj_log_ssb_lrp -->
<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->

<!-- d <- growth[growth$year != max(years) + 1, ] -->
<!-- d$scenario <- factor(d$scenario) -->
<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
<!--   add_heatmap() %>% -->
<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
<!--   layout(yaxis = list(title = "Catch multiplier"), -->
<!--          xaxis = list(title = "Year")) -->

<!-- ``` -->

Inputs {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Landings

```{r}
kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
```

### Catch at age

```{r}
kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
```

<!-- ### Monthly landings -->

<!-- ```{r} -->
<!-- d <- inputs$catch.month -->
<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
<!-- ``` -->

### RV index

```{r}
d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
```

### SN index

```{r}
d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
mat <- d[, as.character(3:10)]
rownames(mat) <- d$year
kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
```

### SS index

```{r}
d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
names(d) <- c("Year", "Month", "Biomass", "SE^2^")
kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
```

### SS age composition

```{r}
kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
```

### Weight at age

```{r}
d <- inputs$pwts
d <- d[d$year <= max(inputs$index$year), ]
d <- data.table::dcast(d, year ~ age, value.var = "weight")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
```

### Maturity at age

```{r}
d <- inputs$mats
d <- d[d$year <= max(inputs$index$year), ]
d <- data.table::dcast(d, year ~ age, value.var = "mat")
mat <- d[, as.character(2:14)]
rownames(mat) <- d$year
kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div. 2J+3KL from DFO RV autumn bottom trawl surveys. Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
```

### Tag summary

```{r}
d <- inputs$mtag1 %>% 
  group_by(rel_year, exp) %>% 
  summarise(NAFO = unique(div),
            Nrel = sum(unique(released)), 
            Ncap = sum(recaptured))
names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
```

Settings {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Catch bounds

```{r}
d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
```

### Baseline M

```{r}
m <- fit$tmb.data$M_matrix
rownames(m) <- ages
colnames(m) <- years
kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
```

Output {data-navmenu="Tables"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Parameter estimates

```{r}

par_tab <- tidy_pars(fit)
rownames(par_tab) <- NULL
options(knitr.kable.NA = '')
kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")

```

> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial

### Parameter estimates (retro)

```{r}

par_tabs <- lapply(names(fit$retro), function(nm) {
  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
  d$id <- seq(nrow(d))
  d
})
par_tabs <- do.call(rbind, par_tabs)
final_par_tab <- data.frame(tyear = max(years), par_tab)
final_par_tab$id <- seq(nrow(final_par_tab))
par_tabs <- rbind(par_tabs, final_par_tab)
par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
par_tabs$Quantity <- as.character(par_tabs$Quantity)
par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
par_tabs$Symbol <- as.character(par_tabs$Symbol)
par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
rownames(par_tabs) <- NULL
names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"

kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")

```

> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial

### Stock size

```{r}

tab_name <- c("log_abundance" = "Abundance (millions)", 
              "log_recruitment" = "Recruits (age 2; millions)", 
              "log_biomass" = "Biomass (kt)", 
              "log_ssa" = "Spawners (millions)", 
              "log_ssb" = "SSB (kt)", 
              "log_ssb_lrp" = "SSB/Blim n(%)")

nms <- names(tab_name)
status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))

for (n in nms[-1]) {
  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
  status_tab <- merge(status_tab, tab, by = "Year")
}

knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")

```

### Rates

```{r}

tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
              "log_aveM" = "Average M (ages 5+)", 
              "log_aveF" = "Average F (ages 5+)")

nms <- names(tab_name)
rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))

for (n in nms[-1]) {
  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
  rate_tab <- merge(rate_tab, tab, by = "Year")
}

knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")

```

### Abundance-at-age

```{r}
N_matrix <- fit$rep$N_matrix
dimnames(N_matrix) <- list(age = ages, year = years_plus1)
knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
```

### Mature abundance-at-age

```{r}
ssa_matrix <- fit$rep$ssa_matrix
dimnames(ssa_matrix) <- list(age = ages, year = years_plus1)
knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
```

### Biomass-at-age

```{r}
B_matrix <- fit$rep$biomass_matrix
dimnames(B_matrix) <- list(age = ages, year = years_plus1)
knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
```

### Mature biomass-at-age

```{r}
ssb_matrix <- fit$rep$ssb_matrix
dimnames(ssb_matrix) <- list(age = ages, year = years_plus1)
knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
```

### Z-at-age

```{r}
Z_matrix <- fit$rep$Z_matrix
dimnames(Z_matrix) <- list(age = ages, year = years)
knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
```

### M-at-age

```{r}
M_matrix <- fit$rep$Mpe_matrix
dimnames(M_matrix) <- list(age = ages, year = years)
M_matrix <- M_matrix[as.character(2:8), ]
rownames(M_matrix) <- c(as.character(2:7), "8+")
knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
```

### F-at-age

```{r}
F_matrix <- fit$rep$F_matrix
dimnames(F_matrix) <- list(age = ages, year = years)
knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
```

### 2018-11-21 R/plot.R

@@ -1,0 +1,645 @@
+
+#' Helper function for plotly labels
+#'
+#' @param  label   Label for plot
+#' @param  offset  Number of pixels to offset from axis
+#'
+#' @return A list to supply to \code{annotations} argument in \code{plotly::layout}
+#'
+#' @export
+#'
+
+top_label <- function(label, offset = 0) {
+  list(
+    text = label,
+    font = list(size = 14),
+    xref = "paper",
+    yref = "paper",
+    yanchor = "bottom",
+    xanchor = "center",
+    align = "center",
+    yshift = -offset,
+    x = 0.5,
+    y = 1,
+    showarrow = FALSE
+  )
+}
+
+#' @export
+#' @rdname top_label
+bottom_label <- function(label, offset = 20) {
+  list(
+    text = label,
+    font = list(size = 14),
+    xref = "paper",
+    yref = "paper",
+    yanchor = "top",
+    yshift = -offset,
+    xanchor = "center",
+    align = "center",
+    x = 0.5,
+    y = 0,
+    showarrow = FALSE
+  )
+}
+
+#' @export
+#' @rdname top_label
+left_label <- function(label, offset = 20) {
+  list(
+    text = label,
+    font = list(size = 14),
+    xref = "paper",
+    yref = "paper",
+    yanchor = "center",
+    xanchor = "right",
+    align = "center",
+    xshift = -offset,
+    x = 0,
+    y = 0.5,
+    textangle = -90,
+    showarrow = FALSE
+  )
+}
+
+#' @export
+#' @rdname top_label
+right_label <- function(label, offset = 0) {
+  list(
+    text = label,
+    font = list(size = 14),
+    xref = "paper",
+    yref = "paper",
+    yanchor = "center",
+    xanchor = "left",
+    align = "center",
+    xshift = -offset,
+    x = 1,
+    y = 0.5,
+    textangle = -90,
+    showarrow = FALSE
+  )
+}
+
+
+#' Plot trends
+#'
+#' @param data         Output produced by \code{\link{tidy_model}}
+#' @param value        Name of value to plot
+#' @param group        Name of column to group by
+#' @param ylab         Label for the y axis
+#' @param title        Plot title
+#' @param ci           How should confidence interevals be represented? ("ribbon", "line", "none")
+#' @param slider       Add range slider?
+#' @param annotate     Add annotations
+#' @param showlegend   Show legend
+#' @param lwd          Line width
[…]

### 2018-11-21 R/tidy.R

@@ -1,0 +1,280 @@
+
+#' Makes one data.frame from a list of many
+#'
+#' @param l         A list containing data.frames.
+#' @param add_id    If the list is named and add_id is TRUE, an id column will be added.
+#' @export
+#'
+
+rbind_list <- function(l, add_id = TRUE) {
+  all <- do.call("rbind", l)
+  if (add_id) { all$id <- rep(names(l), sapply(l, nrow)) }
+  rownames(all) <- NULL
+  all
+}
+
+
+#' Tidy value output from sdreport
+#'
+#' @param model  model object produced by \code{\link{fit_model}}
+#'
+#' @return Returns a list of data.frames with estimates (est), standard deviations (sd),
+#' lower (lwr) and upper (upr) 95% confidence intervals of reported objects
+#'
+#' @export
+#'
+
+tidy_sdreport <- function(model) {
+
+  ## Extract key objects
+  sd.rep <- model$sd.rep
+  tmb.data <- model$tmb.data
+  values <- sd.rep$value
+  sds <- sd.rep$sd
+  lwr <- values - qnorm(0.975) * sds
+  upr <- values + qnorm(0.975) * sds
+  value_names <- sort(unique(names(values)))
+
+  ## Split sdreport by name
+  split_values <- split(unname(values), names(values))
+  split_sds <- split(unname(sds), names(values))
+  split_lwr <- split(unname(lwr), names(values))
+  split_upr <- split(unname(upr), names(values))
+
+  ## Convert to data.frame and add key identifiers
+  l <- vector("list", length(split_values))
+  names(l) <- names(split_values)
+  for (nm in value_names) {
+    vals <- data.frame(est = split_values[[nm]], sd = split_sds[[nm]], lwr = split_lwr[[nm]], upr = split_upr[[nm]])
+    if (grepl("^Cproj|^Fproj|^Hproj", nm)) {
+      if (grepl("^Cproj", nm)) scenario <- tmb.data$Cp
+      if (grepl("^Fproj", nm)) scenario <- tmb.data$Fp
+      if (grepl("^Hproj", nm)) scenario <- tmb.data$HCR
+      vals <- cbind(expand.grid(scenario = scenario, year = tmb.data$proj_years), vals)
+    } else {
+      if (length(split_values[[nm]]) == tmb.data$Y) {
+        vals <- cbind(data.frame(year = tmb.data$years), vals)
+      }
[…]

### 2018-11-21 R/vis_fit.R

@@ -1,0 +1,47 @@
+#' Make a flexdashboard for visualizing the model fit
+#'
+#' Assumes the working directory is the project directory
+#'
+#' @param fit            Object produced by \code{\link{fit_model}}
+#' @param comp           Named list of model objects to compare
+#' @param last           Model object from the last assessment
+#' @param output_file    Name of file to export using \code{\link{rmarkdown::render}}.
+#'                       If NULL, flexdashboard will be rendered using \code{\link{rmarkdown::run}}
+#' @param output_format  R markdown output format (see \code{\link{rmarkdown::run}})
+#'
+#' @export
+#'
+
+vis_fit <- function(fit = NULL, comp = NULL, last = NULL, output_file = NULL, output_format = NULL) {
+
+  pkg <- c("rmarkdown", "shiny", "flexdashboard", "plotly", "viridis")
+  for (p in pkg) {
+    if (!requireNamespace(p, quietly = TRUE)) {
+      stop(paste(p, "is needed for vis_fit to work. Please install it."), call. = FALSE)
+    }
+  }
+
+  ## simplify file and dir specification
+  if (!is.null(output_file)) {
+    output_dir <- normalizePath(dirname(output_file))
+    output_file <- basename(output_file)
+  }
+
+  ## make a tmp object in the global environment for use by rmarkdown
+  ## (rmarkdown::run likes to operate on objects in the global environment)
+  tmp <- as.list(environment())
+  assign("tmp", tmp, globalenv())
+
+  ## keep most objects created in the rmd file local (i.e. not in the global)
+  local({
+    if (is.null(output_file)) {
+      rmarkdown::run(file = "inst/rmd/vis_fit.Rmd")
+    } else {
+      rmarkdown::render(input = "inst/rmd/vis_fit.Rmd",
+                        output_file = output_file,
+                        output_dir = output_dir,
+                        output_format = output_format)
+    }
+  })
+
+}

### 2018-11-21 analysis/dashboards/make_NCAM_dashboard.R

@@ -1,0 +1,23 @@
+
+## load NCAM model inputs
+## (object is named inputs)
+load("data/inputs.RData")
+
+## load NCAM model output
+## (object is named mshift)
+load("data/mshift.RData")
+
+## source plotting functions
+## (these functions are used in the Rmarkdown file that makes the dashboard)
+source("R/plot.R")
+
+## source some data management functions
+## (these functions are used to "tidy" the model output)
+source("R/tidy.R")
+
+## source function for making the dashboard
+## (this function uses rmarkdown::run or rmarkdown::render)
+source("R/vis_fit.R")
+
+## Run the vis_fit function to make the dashboard
+vis_fit(mshift, output_file = "analysis/dashboards/NCAM_dashboard.html")

### 2018-12-11 analysis/paper/manuscript.Rmd
Tweak ms and start producing webshots of the NCAM dashboard

@@
 Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python].
-By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of Big Data by facilitating the exploration and communication of complex data and models.
+By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of big data by facilitating the exploration and communication of complex data and models.
 
@@
 Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000].
-These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs and model outputs.
+These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs, model structures and outputs.
 This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice.
@@
 Once the model is fit to the data, the next step is to produce visual representations of the data and model.
-An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod (**TODO: add reference**).
-The concept of using a dashboard was burrowed from the buisness community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
+An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod [@dwyer2019].
+The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
 Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM.
@@
 - Take screenshots of the dashboard for static figures for the paper (use webshot?)
-- Remember to add reference to the supplement that will ultimatly include the NCAM_dashboard.html
+- Remember to add reference to the supplement that will ultimately include the NCAM_dashboard.html
 

### 2018-12-11 analysis/paper/figures.R

@@ -1,0 +1,18 @@
+
+library(webshot)
+
+## helper function for exporting dashboard tabs using webshot
+dashboard_webshot <- function(in_file = NULL, out_file = NULL, 
+                              vwidth = 1200, vheight = 800, zoom = 10, delay = 10,
+                              ...) {
+    
+    webshot(url = file.path("file://", getwd(), in_file),
+            file = out_file,
+            delay = delay, vwidth = vwidth, vheight = vheight, zoom = zoom)
+}
+
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#tagging",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_tagging_tab.png")
+
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#rv-survey",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png")

### 2019-01-09 analysis/paper/manuscript.Rmd
Make a panel figure for NCAM dashboard screenshots

@@
 
-TODO:
-- Take screenshots of the dashboard for static figures for the paper (use webshot?)
-- Remember to add reference to the supplement that will ultimately include the NCAM_dashboard.html
+The NCAM dashboard contains a series of tabs, the first of which provides a series of tarse point-form background on the model (Figure \@ref(fig:background)).
+
+TODO: Aggregate screenshots into one a, b, c, etc. image.
+
+```{r background, echo = FALSE, fig.cap = "- Screenshots of NCAM dashboard."}
+knitr::include_graphics("figures/NCAM_dashboard_background_tab.png")
+```
+
+...useful format for meetings because a self-contained html can be produced for participants to independently scrutinize. This html has been included as a supplement (Supplement 2)...
 

### 2019-01-09 analysis/paper/figures.R

@@ -1,5 +1,6 @@
 
 library(webshot)
+library(magick)
 
 ## helper function for exporting dashboard tabs using webshot
 dashboard_webshot <- function(in_file = NULL, out_file = NULL, 
@@ -11,8 +12,39 @@ dashboard_webshot <- function(in_file = NULL, out_file = NULL,
             delay = delay, vwidth = vwidth, vheight = vheight, zoom = zoom)
 }
 
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#tagging",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_tagging_tab.png")
+## take screenshots of NCAM dashboard
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#background",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_background_tab.png")
 
 dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#rv-survey",
                   out_file = "analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png")
+
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#tagging",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_tagging_tab.png")
+
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#output",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_tables_tab.png")
+
+## combine screenshots into one image
+img_a <- image_read("analysis/paper/figures/NCAM_dashboard_background_tab.png") %>% 
+    image_border("white", "300x50") %>% 
+    image_annotate("a)", color = "black", size = 300)
+img_b <- image_read("analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png") %>% 
+    image_border("white", "300x50") %>% 
+    image_annotate("b)", color = "black", size = 300)
+img_c <- image_read("analysis/paper/figures/NCAM_dashboard_tagging_tab.png") %>% 
+    image_border("white", "300x50") %>% 
+    image_annotate("c)", color = "black", size = 300)
+img_d <- image_read("analysis/paper/figures/NCAM_dashboard_tables_tab.png") %>% 
+    image_border("white", "300x50") %>% 
+    image_annotate("d)", color = "black", size = 300)
+
+side_a <- image_append(image_scale(c(img_a, img_b), "x3000"))
[…]

### 2019-01-14 analysis/paper/manuscript.Rmd
Draft section on NCAM dashboard

--- 2018-12-11 analysis/paper/manuscript.Rmd
@@
 ---
-
-# Introduction
 
@@
 
-## Knowing the data
+# Knowing the data
 
@@
 
-- *Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?*
-- *TODO: the intro paragraph to this section could use some references if it is necessary*
+> Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?
 
-## Data modeling
+> TODO: the intro paragraph to this section could use some references if it is necessary
+
+# Data modeling
 
@@
 Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document.
-We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots in the "Tagging" tab.
+We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots.
 
-TODO:
-- Take screenshots of the dashboard for static figures for the paper (use webshot?)
-- Remember to add reference to the supplement that will ultimately include the NCAM_dashboard.html
+The NCAM dashboard, which is included in Supplement 2 as a self-contained html file, contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
+Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data.
+For instance, the "RV survey" tab includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)).
+The dashboard also includes tabs focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity").
+Finally, some results from a retrospective analyses are included under the "Retro" tab and key inputs and outputs are included as tables under the "Tables" tab.
+The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail.
+Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.
+This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst.
+Such access improves the transparancy of the stock assessment model which, in turn, leads to richer discussion and interogation of the biological and statistical rigor of the model.
 
-# Notes
+> TODO: should I add detail of questions not previously raised? If so, I can't quite remember.
 
-- Note that these tools facilitate the peer-review process and also improves the transparency and reproducibility of assessment models
-- Dynamic document can be self contained
-- Northern cod stock is on an annual assessment schedule, therefore it makes sense to automate visuals
+```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
+Age-disaggregated observed (dots) and perdicted (lines) values from the DFO RV survey are shown in the right panel.
+The tab on the right panel including residual plots are hidden here, but asscessable in the dashboard.'}
+knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
+```
 
-# Other draft text
+```{r trends, echo = FALSE, fig.cap = '- Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
+knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
+```
 
-A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
+# Open stock-assessment
+
+> TODO: discuss this concept in the context of open science and open stock assessment.
+
+> **Other draft text**
+
+> A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
 In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
@@
 
-In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models.
+> In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models.
 The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits.
@@
 
-Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization
+> **Links to some handy references**
 
-https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/
+> Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization
 
-https://www.nature.com/articles/s41467-018-03297-7
+> https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/
 
-https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103
+> https://www.nature.com/articles/s41467-018-03297-7
 
-https://www.nature.com/articles/d41586-018-01322-9
+> https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103
 
-"The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)
+> https://www.nature.com/articles/d41586-018-01322-9
+
+> "The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)
 

### 2019-01-14 analysis/paper/figures.R

@@ -4,7 +4,7 @@ library(magick)
 
 ## helper function for exporting dashboard tabs using webshot
 dashboard_webshot <- function(in_file = NULL, out_file = NULL, 
-                              vwidth = 1200, vheight = 800, zoom = 10, delay = 10,
+                              vwidth = 1200, vheight = 800, zoom = 3, delay = 10,
                               ...) {
     
     webshot(url = file.path("file://", getwd(), in_file),
@@ -16,11 +16,17 @@ dashboard_webshot <- function(in_file = NULL, out_file = NULL,
 dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#background",
                   out_file = "analysis/paper/figures/NCAM_dashboard_background_tab.png")
 
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#catch",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_catch_tab.png")
+
 dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#rv-survey",
                   out_file = "analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png")
 
 dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#tagging",
                   out_file = "analysis/paper/figures/NCAM_dashboard_tagging_tab.png")
+
+dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#trends",
+                  out_file = "analysis/paper/figures/NCAM_dashboard_trends_tab.png")
 
 dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#output",
                   out_file = "analysis/paper/figures/NCAM_dashboard_tables_tab.png")

### 2019-02-15 analysis/paper/manuscript.Rmd
Add sections from Greg and Keith, and address some of Keith's suggestions

@@
   visualization tools
-author: "Paul M Regular, Gregory J Robertson, etc.?"
+author: "Paul M Regular, Gregory J Robertson, Keith P. Lewis, etc?"
 date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
@@
 
-In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
-Scientific data are often presented as tabular summaries and static plots, however, seeing patterns in tables can be difficult and static plots preclude detailed explorations of the underlying data.
-New open source tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
-In particular, browser-based data visualizations have been playing an increasingly prominent role in communicating information on a wide range of topics.
-Interactive graphics are frequently utilized by media outlets, such as the FiveThirtyEight, and scientists from a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [@yeatman2018; @tushar2017; @wick2015].
-This trend has been fueled by the development of JavaScript libraries (e.g.
-D^3^ [@bostock2011], plotly [@plotly], leaflet [@leaflet], highcharts [@highcharts]) that enable the production of dynamic and interactive data visualizations in modern web-browsers.
-Interfaces to JavaScript libraries have also been developed to allow interactive plots to be coded using commonly used languages such as R [@R] or Python [@python].
-By laying the technical infrastructure for creating interactive graphics, these new tools are ushering in a new era of big data by facilitating the exploration and communication of complex data and models.
+Quantitative stock assessment plays a central role in modern fisheries management.
+Over time and as new methods are developed, biologists have more and more data available to inform stock assessments.
+These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
+Concurrently, analysis methods have improved, and techniques are now available that can integrate many data sources into one stock assessment model; state-space models being one popular class of such models.
+For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on the status of a stock, to presenting the mass of information produced by increasingly complex statistical models to end-users in a meaningful way.
+Traditional methods of presenting information at stakeholder meetings and to fisheries managers, like written reports full of tables, or with presentations with endless figures, are simply no longer sufficient to convey the richness of the information.
+More importantly, these static presentation formats tend to stifle meaningful discussion of model outputs, and important questions and assumptions of the models that should be discussed tend to get lost under the sheer volume of output.
+A solution to this problem of communicating complex output and results to encourage engaging discussion can be through using the effective use of interactive visualization tools.
+These are common tools we use every day on an endless array of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
 
-The field of fisheries research has a long history of collecting and synthesizing a plethora of data using state-of-the-art computational technologies [@megrey2008].
-Fisheries science has evolved alongside the rise of the computing era, and as tools improve so does the management advice provided by fisheries scientists.
-This management advice generally culminates from a stock assessment process whereby various sources of data and statistical methods are used to model the history, current status and future status of one or more fish stocks [@hilborn1992].
-Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated assessment models [@maunder2013].
-This means that analyses that were once carried out independently on different data sets can now be integrated into one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015].
-From a computational perspective, analyses of a variety of large data sets has never been easier.
-However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
 Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs.
@@
 Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
-However, integrated analyses are not a panacea because model misspecifications and data conflicts are an inevitable consequence of simplifying reality to a small series of equations [@maunder2017].
-A potential solution to this quandary to use a superensemble model, whereby multiple models with different structures are run and their predictions are supplied as covariates to an additional superensemble model [@anderson2017].
-Ensemble approaches reduce the risk of picking the "wrong" model and also expands the range of hypotheses explored [@dietterich2000].
-These advances greatly improve our ability to assess the status and trends of fish populations, however, modern stock assessment biologists are now faced with the overwhelming task of understanding an ever expanding array of data inputs, model structures and outputs.
-This challenge is compounded by the ever increasing demand for rapid and sound fisheries advice.
-Here we argue that the process of providing fisheries advice can be further streamlined by utilizing dynamic and interactive visualization tools to explore, diagnose and communicate stock assessment models.
-We hope to demonstrate this using a recently developed interactive tool for exploring an integrated assessment model for Northern cod.
+From a computational perspective, analyses of a variety of large data sets has never been easier.
+However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
+By using a recently developed interactive tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that such tools can streamline the process of understanding and communicating results from complex stock assessment models.
 
@@
 Once the model is fit to the data, the next step is to produce visual representations of the data and model.
-An interactive and self-contained "dashboard" was developed for this purpose for the last assessment of Northern cod [@dwyer2019].
+A standard approach would be to produce static presentations and documents with a series of figures and tables.
+With large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analysist and all the stakeholders involved in the stock assessment process.
+First, it is not feasible for the analysist to include and describe all figures and tables produced into a single document.
+Second, it is difficult for stakeholders to efficiently digest the information that has been compressed into a series of static slides or pages.
+Interactive documents provide a potential solution to this problem as they allow much more information to be contained and assessible on a single screen.
+
+In an effort to make the results from the Northern cod assessment model more accessible and transparant, an interactive and self-contained "dashboard" was developed for the last assessment of Northern cod [@dwyer2019].
 The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
@@
 This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst.
-Such access improves the transparancy of the stock assessment model which, in turn, leads to richer discussion and interogation of the biological and statistical rigor of the model.
+Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and interogation of the biological and statistical rigor of the model.
 
@@
 ```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
-Age-disaggregated observed (dots) and perdicted (lines) values from the DFO RV survey are shown in the right panel.
-The tab on the right panel including residual plots are hidden here, but asscessable in the dashboard.'}
+Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.
+The tab on the right panel including residual plots are hidden here, but assessable in the dashboard.'}
 knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
@@
 
-> TODO: discuss this concept in the context of open science and open stock assessment.
+decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics.
+The majority of this sequence of events, sometimes called “the data pipeline”, have not traditionally been part of the peer-review process which sees only the end products.
+However, decisions made at these steps increasingly influence the outcome of the study.
+Gelman coined the term “the forked garden path” to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path.
+Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible, even if it is one’s own analysis.
+A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science.
+Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work (Lewis et al.
+2018).
+
+We submit that the dashboard tool is also a major step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
+Consider the following process.
+Through the Canadian Science Advisory Secretariat (CSAS), DFO Science provides peer-reviewed science advice to fisheries managers at regular meetings (Regional Assessment Process).
+These meetings are attended by a variety of stake holders including harvesters, academics, and NGOs.
+Documents summarizing these processes (research documents, stock assessment reports, and proceedings summaries) are available to the public through the CSAS website.
+Ergo the CSAS process strives to uphold the ideal of transparency.
+However, as it currently stands, only the end products of an assessment are open and transparent; as with virtually all scientific analyses, the data pipeline of stock assessment is hidden from review.
+
+With the exception of data management, the dashboard removes much this problem – all analyses and diagnostics can be readily reviewed.
+In terms of data management, this point is largely ameliorated because the dashboard tool is constructed in Program R using R Markdown and the code for creating the dashboard and all associated analyses are available on Github.
+Data are also available through information requests.
+So the entire data pipeline is transparent as well as reproducible, i.e., the requirements of open science are fulfilled.
+
+The dashboard also promotes accessibility to the stock assessment.
+Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted (Banks 20XX).
+In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model).
+However, the dashboard removes many of these difficulties.
+The html file contains all of the relevant analyses.
+The number of analyses can be expanded upon request and can be distributed prior to meetings and reviewers can assess the results and diagnostics at their leisure.
+
+We submit that the plotly/flex dashboard tool is an important step forward in transparent and reproducible science.
+Further, this tool makes stock assessments more accessible to stakeholders.
+Beyond stock assessments, it this tool could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.
 

### 2019-03-05 analysis/paper/manuscript.Rmd
More tweaks given Keith's suggestions. Next: revise Keith's open stock-assessment section

@@
 However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
-By using a recently developed interactive tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that such tools can streamline the process of understanding and communicating results from complex stock assessment models.
+By using a recently developed interactive tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that interactive visualization tools can streamline the process of understanding and communicating results from complex stock assessment models.
 
@@
 A standard approach would be to produce static presentations and documents with a series of figures and tables.
-With large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analysist and all the stakeholders involved in the stock assessment process.
-First, it is not feasible for the analysist to include and describe all figures and tables produced into a single document.
+With large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and all the stakeholders involved in the stock assessment process.
+First, it is not feasible for the analyst to include and describe all figures and tables produced into a single document.
 Second, it is difficult for stakeholders to efficiently digest the information that has been compressed into a series of static slides or pages.
-Interactive documents provide a potential solution to this problem as they allow much more information to be contained and assessible on a single screen.
+Interactive documents provide a potential solution to this problem as they allow much more information to be contained and accessible on a single screen.
 
-In an effort to make the results from the Northern cod assessment model more accessible and transparant, an interactive and self-contained "dashboard" was developed for the last assessment of Northern cod [@dwyer2019].
+In the pursuit for an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the last assessment of Northern cod [@dwyer2019].
 The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
-Many tools have been developed to facilitate the development of data dashboards and we leveraged R-based tools to construct a tool for exploring the input and output of NCAM.
+Many tools have been created to facilitate the development of data dashboards and we leveraged R-based packages to construct a tool for exploring the input and output of NCAM.
 Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document.
 We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots.
+Via R Markdown [@markdown], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
 
-The NCAM dashboard, which is included in Supplement 2 as a self-contained html file, contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
+The NCAM dashboard, which is included as a self-contained html file in Supplement 2, contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
 Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data.
@@
 The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail.
+For instance, the ability to zoom allows users to investigate specific patterns and tooltabs display the values behind specific points.
 Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.
 This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst.
-Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and interogation of the biological and statistical rigor of the model.
+Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.
 

### 2019-03-06 analysis/paper/manuscript.Rmd
Modify Keith's section and produce the second draft of the manuscript

@@
 
-decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics.
+The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics.
 The majority of this sequence of events, sometimes called “the data pipeline”, have not traditionally been part of the peer-review process which sees only the end products.
@@
 A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science.
-Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work (Lewis et al.
-2018).
+Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@lewis2018].
 
@@
 Through the Canadian Science Advisory Secretariat (CSAS), DFO Science provides peer-reviewed science advice to fisheries managers at regular meetings (Regional Assessment Process).
-These meetings are attended by a variety of stake holders including harvesters, academics, and NGOs.
+These meetings are attended by a variety of stakeholders including harvesters, academics, and NGOs.
 Documents summarizing these processes (research documents, stock assessment reports, and proceedings summaries) are available to the public through the CSAS website.
-Ergo the CSAS process strives to uphold the ideal of transparency.
+Ergo the CSAS process strives to uphold the idea of transparency.
 However, as it currently stands, only the end products of an assessment are open and transparent; as with virtually all scientific analyses, the data pipeline of stock assessment is hidden from review.
 
-With the exception of data management, the dashboard removes much this problem – all analyses and diagnostics can be readily reviewed.
+The dashboard removes much of this problem by providing access to all model inputs, outputs and diagnostics in an open and easy to use interactive document.
 In terms of data management, this point is largely ameliorated because the dashboard tool is constructed in Program R using R Markdown and the code for creating the dashboard and all associated analyses are available on Github.
@@
 The dashboard also promotes accessibility to the stock assessment.
-Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted (Banks 20XX).
+Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted [@banks2011].
 In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model).
@@
 
-We submit that the plotly/flex dashboard tool is an important step forward in transparent and reproducible science.
+We submit that the dashboard tool is an important step forward in transparent and reproducible science.
 Further, this tool makes stock assessments more accessible to stakeholders.
-Beyond stock assessments, it this tool could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.
-
-> **Other draft text**
-
-> A wide array data may be collected for an assessment and, formally, the process often reduces to algorithms that convert these data to advice for fisheries managers.
-In some cases, particularly for commercially valuable species, this means that hundreds or thousands of historical data points from the monitoring program of a stock gets reduced into a single policy value, such as a recommended catch quota [@maunder2009].
-Sound management advice therefore hinges on the ability of stock assessment biologists to manage, interpret and communicate large volumes of data from a variety of sources.
-
-> In this data-intensive era of stock assessment, one of the difficulties becomes visualizing the inputs and outputs to integrated stock assessment models.
-The standard approach has been to create a series of tables and static plots to help assess the inputs and model fits.
-Data presented in tables is incredibly valuable, but as human beings, seeing patterns can be challenging.
-Static figures, in contrast, help reveal patterns but they are divorced from the underlying data and, as such, limit detailed explorations.
-New tools are bridging this gap by making it easier to build interactive figures [@perkel2018].
-
-> **Links to some handy references**
-
-> Interactive visuals...visual representation of the data and model that help us interpret complex data at a glance https://campus.sagepub.com/blog/3-benefits-of-interactive-visualization
-
-> https://centricdigital.com/blog/digital-trends/the-value-of-interactive-data-visualization-and-why-it-matters-to-business/
-
-> https://www.nature.com/articles/s41467-018-03297-7
-
-> https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/120103
-
-> https://www.nature.com/articles/d41586-018-01322-9
-
-> "The ultimate objective of Fisheries Science is to inform management" (http://www.fao.org/tempref/docrep/fao/008/a0212e/a0212E02.pdf)
+Beyond stock assessments, interactive documents could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.
 

### 2019-04-09 analysis/paper/manuscript.Rmd
Add Greg's section to the paper. Next: edit open stock-assessment section

@@
 ---
-title: Improving the communication and transparency of stock assessment using interactive
+title: Improving the communication and accessibility of stock assessment using interactive
   visualization tools
-author: "Paul M Regular, Gregory J Robertson, Keith P. Lewis, etc?"
+author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, etc?"
 date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
@@
 These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
-Concurrently, analysis methods have improved, and techniques are now available that can integrate many data sources into one stock assessment model; state-space models being one popular class of such models.
-For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on the status of a stock, to presenting the mass of information produced by increasingly complex statistical models to end-users in a meaningful way.
-Traditional methods of presenting information at stakeholder meetings and to fisheries managers, like written reports full of tables, or with presentations with endless figures, are simply no longer sufficient to convey the richness of the information.
-More importantly, these static presentation formats tend to stifle meaningful discussion of model outputs, and important questions and assumptions of the models that should be discussed tend to get lost under the sheer volume of output.
+Concurrently, analysis methods have evolved with techniques now available that integrate many data sources into one stock assessment model.
+For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on stock status, to presenting the mass of information produced by increasingly complex statistical models to end-users in a meaningful way.
+Traditional methods of presenting information at stakeholder meetings and to fisheries managers are simply no longer sufficient to convey the richness of the information.
+More importantly, static presentation formats tend to stifle meaningful discussion of model results, and important questions and assumptions of the models that should be discussed tend to get lost under the sheer volume of output.
 A solution to this problem of communicating complex output and results to encourage engaging discussion can be through using the effective use of interactive visualization tools.
-These are common tools we use every day on an endless array of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
+These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
 
@@
 
-The data sets used in stock assessments are constantly growing.
-This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
+from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
 As such, stock assessment biologists often have to manage large volumes of data from a variety of sources.
-For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments.
-These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
+Time series of reported landings and catch-at-age are often analyzed in conjunction with data from “fishery-independent” surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
@@
 
-> Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?
+Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day.
+The tagging and recovery data is captured in standardized database, with fields typical of most tagging programs.
+This data base has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
+The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.
+To begin to explore and understand this large data set, we built a simple shiny [@chang2018a] application using the mapping package leaflet [@cheng2018], all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio].
+As the application developed, it was incorporated into a shinydashboard [@chang2018b], as a means to quickly and dynamically subset the data (e.g.
+ranges of release and capture years, specific geographic locations).
+Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging.
+The markercluster (https://github.com/Leaflet/Leaflet.markercluster) function available in leaflet was particularly useful as a means to dynamically scale the level of pooling of spatial points (Figure \@ref(fig:map)).
+This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, and explore options on how to spatially pool the data for subsequent demographic analysis.
+Further tabs were added to provide basic summaries of the selected data (Figure \@ref(fig:summary)).
 
-> TODO: the intro paragraph to this section could use some references if it is necessary
+```{r map, echo = FALSE, fig.cap = '- Screen shot of cod tag mapping tool using shinydashboard and leaflet.
+The markerCluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
+Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released and date captured), which is particular useful when error checking.'}
+knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
+```
+
+```{r summary, echo = FALSE, fig.cap = '- Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
+In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
+knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")
+```
 
@@
 With large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and all the stakeholders involved in the stock assessment process.
-First, it is not feasible for the analyst to include and describe all figures and tables produced into a single document.
+First, it is not feasible for the analyst to include and describe all figures and tables produced in a single document.
 Second, it is difficult for stakeholders to efficiently digest the information that has been compressed into a series of static slides or pages.
@@
 
-In the pursuit for an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the last assessment of Northern cod [@dwyer2019].
+In the pursuit for an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019].
 The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
-Many tools have been created to facilitate the development of data dashboards and we leveraged R-based packages to construct a tool for exploring the input and output of NCAM.
-Specifically we utilized the flexdashboard [@allaire2017] package to group interactive plotly-based [@plotly] visuals into a dynamic document.
+We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document.
 We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots.
-Via R Markdown [@markdown], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
+Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
 
@@
 The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail.
-For instance, the ability to zoom allows users to investigate specific patterns and tooltabs display the values behind specific points.
+For instance, the ability to zoom allows users to investigate specific patterns and tooltip display the values behind specific points.
 Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.

### 2019-04-10 analysis/paper/manuscript.Rmd
Revise the open stock assessment section to clue up the next draft

@@
 
-from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
+The data sets used in stock assessments are constantly growing.
+This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
 As such, stock assessment biologists often have to manage large volumes of data from a variety of sources.
@@
 The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail.
-For instance, the ability to zoom allows users to investigate specific patterns and tooltip display the values behind specific points.
+For instance, the ability to zoom allows users to investigate specific patterns and small boxes, called tooltips, display more information about the points users hover over.
 Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.
@@
 Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.
-
-> TODO: should I add detail of questions not previously raised? If so, I can't quite remember.
 
@@
 
-# Open stock-assessment
+# Towards open stock-assessment
 
@@
 However, decisions made at these steps increasingly influence the outcome of the study.
-Gelman coined the term “the forked garden path” to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path.
+Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path.
 Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible, even if it is one’s own analysis.
@@
 
-We submit that the dashboard tool is also a major step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
+We submit that the dashboard tool is also a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
 Consider the following process.
@@
 These meetings are attended by a variety of stakeholders including harvesters, academics, and NGOs.
-Documents summarizing these processes (research documents, stock assessment reports, and proceedings summaries) are available to the public through the CSAS website.
-Ergo the CSAS process strives to uphold the idea of transparency.
-However, as it currently stands, only the end products of an assessment are open and transparent; as with virtually all scientific analyses, the data pipeline of stock assessment is hidden from review.
-
-The dashboard removes much of this problem by providing access to all model inputs, outputs and diagnostics in an open and easy to use interactive document.
-In terms of data management, this point is largely ameliorated because the dashboard tool is constructed in Program R using R Markdown and the code for creating the dashboard and all associated analyses are available on Github.
-Data are also available through information requests.
-So the entire data pipeline is transparent as well as reproducible, i.e., the requirements of open science are fulfilled.
-
-The dashboard also promotes accessibility to the stock assessment.
-Even if one is able to download code from a content management system (e.g., GitHub or Bitbucket), submit a data request, and manipulate the data so that it conforms to the code, it is unlikely that an individual has the time to fully recreate a particular analysis in its entirety when most reviewers are already overcommitted [@banks2011].
+All participants attending a CSAS meeting are encouraged to question, comment and constructively challenge the science presented and the goal is to provide the best possible science advice to the Minister, managers, stakeholders and the public.
+While the entire data-pipeline can be revealed to all by sharing code and data, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
 In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model).
 However, the dashboard removes many of these difficulties.
-The html file contains all of the relevant analyses.
-The number of analyses can be expanded upon request and can be distributed prior to meetings and reviewers can assess the results and diagnostics at their leisure.
+The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document.
+The number of analyses can be expanded upon request and can be distributed prior to meetings and participants can assess the results and diagnostics at their leisure.
+Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
 
-We submit that the dashboard tool is an important step forward in transparent and reproducible science.
+We must acknowledge there are costs to adopting this approach of using the interactive tools.
+The first and obvious one being that staff are required to have both the time and training to effectively to develop these tools.
+In our two examples, the developers were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
+Most programming issues were readily addressed through simple internet searchers, but some learning time was certainly needed (a few weeks).
+As more people in the fisheries community develop and use these tools, we expect tool develop speed and quality will only increase.
+It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
+Scientific exploration is rarely a linear process, and the ability to jump around a dashboard is a more in tune with the natural way scientist’s pursue ideas.
+In contrast, a slide presentation is effectively a linear and a single way of developing a scientific narrative; a method hardly conducive to exploring complex data problems.
+
+In summary, we submit that the dashboard tool is an important step forward in transparent and reproducible science.
 Further, this tool makes stock assessments more accessible to stakeholders.

### 2019-05-06 analysis/paper/manuscript.Rmd
Another round of edits

@@
 bibliography: references.bib
+csl: apa.csl
 ---
@@
 Quantitative stock assessment plays a central role in modern fisheries management.
-Over time and as new methods are developed, biologists have more and more data available to inform stock assessments.
+Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments.
 These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
-Concurrently, analysis methods have evolved with techniques now available that integrate many data sources into one stock assessment model.
-For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on stock status, to presenting the mass of information produced by increasingly complex statistical models to end-users in a meaningful way.
-Traditional methods of presenting information at stakeholder meetings and to fisheries managers are simply no longer sufficient to convey the richness of the information.
-More importantly, static presentation formats tend to stifle meaningful discussion of model results, and important questions and assumptions of the models that should be discussed tend to get lost under the sheer volume of output.
+Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model.
+For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on stock status, to presenting the large quantity of information produced by increasingly complex statistical models to end-users in a meaningful way.
+Traditional methods of presenting information at stakeholder meetings and to fisheries managers are simply insufficient to convey the richness of the information available.
+More importantly, static presentation formats tend to stifle meaningful discussion of model results, and important questions and assumptions that should be discussed tend to get lost under the sheer volume of output.
 A solution to this problem of communicating complex output and results to encourage engaging discussion can be through using the effective use of interactive visualization tools.
@@
 
-Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs.
-First, we focus on data that are commonly used in stock assessment and provide one example where interactive maps were used to simplify the detailed exploration of data from a long-term tagging program.
-Second, we focus on the modeling aspect of stock assessment and, again, we use an example to demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
-This structure corresponds to the two general steps in the stock assessment process: 1) knowing the data, and 2) data modeling.
+Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs.
+First, we focus on data that are commonly presented in stock assessments and use interactive maps to simplify the detailed exploration of data from a long-term tagging program.
+Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
+This structure corresponds to two important  steps in the stock assessment process: 1) knowing the data, and 2) data modeling.
 
@@
 
-The data sets used in stock assessments are constantly growing.
-This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs.
-As such, stock assessment biologists often have to manage large volumes of data from a variety of sources.
+The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both.
+As such, stock assessment biologists need to manage large volumes of data from a variety of sources.
 Time series of reported landings and catch-at-age are often analyzed in conjunction with data from “fishery-independent” surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
@@
 All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
-This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind the plot.
-For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data.
+This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot.
+For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
 This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.
@@
 Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day.
-The tagging and recovery data is captured in standardized database, with fields typical of most tagging programs.
+The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs.
 This data base has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
@@
 The markercluster (https://github.com/Leaflet/Leaflet.markercluster) function available in leaflet was particularly useful as a means to dynamically scale the level of pooling of spatial points (Figure \@ref(fig:map)).
-This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, and explore options on how to spatially pool the data for subsequent demographic analysis.
+This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, as well as explore options on how to spatially pool the data for subsequent demographic analysis.
 Further tabs were added to provide basic summaries of the selected data (Figure \@ref(fig:summary)).
@@
 ```{r map, echo = FALSE, fig.cap = '- Screen shot of cod tag mapping tool using shinydashboard and leaflet.
-The markerCluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
-Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released and date captured), which is particular useful when error checking.'}
+The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
+Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date captured), which is particularly useful when error checking.'}
 knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
@@
 As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016].
-The "base case" model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
-Using a series of observation equations, this TMB based model reduces thousands of historical data points from these monitoring programs into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
-Once the model is fit to the data, the next step is to produce visual representations of the data and model.
-A standard approach would be to produce static presentations and documents with a series of figures and tables.
-With large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and all the stakeholders involved in the stock assessment process.
-First, it is not feasible for the analyst to include and describe all figures and tables produced in a single document.
+The model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
+Using a series of observation equations, this TMB based model reduces thousands of historical data points into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
+Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
+The usual approach would be to produce static presentations and documents with a series of figures and tables, but with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and all the stakeholders involved.
+First, it is no longer feasible for the analyst to include and describe every figure and table produced in a single document.
 Second, it is difficult for stakeholders to efficiently digest the information that has been compressed into a series of static slides or pages.
@@
 
-In the pursuit for an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019].
-The concept of using a dashboard was burrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
+In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained “dashboard” was developed for the 2018 assessment of Northern cod [@dwyer2019].
+The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
 We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document.
@@
 ```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
-Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.
-The tab on the right panel including residual plots are hidden here, but assessable in the dashboard.'}
+Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
 knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
@@
 
-The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, as well as the formal analyses and associated diagnostics.
-The majority of this sequence of events, sometimes called “the data pipeline”, have not traditionally been part of the peer-review process which sees only the end products.
-However, decisions made at these steps increasingly influence the outcome of the study.
-Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions one makes along the garden path.
-Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible, even if it is one’s own analysis.
-A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science.
-Proponents of open science and reproducible research posit a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@lewis2018].
+The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, in addition to the formal analyses and associated diagnostics [@lewis2018 and references therin].
+The majority of this sequence of events, sometimes called “the data pipeline” [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis.
+However, decisions made along the data pipeline increasingly influence the outcome of the study.
+Gelman and Loken [-@gelman2014] coined the term “garden of forking paths” to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis.
+Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible.
+A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015].
+Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@lewis2018].
 
-We submit that the dashboard tool is also a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
-Consider the following process.
-Through the Canadian Science Advisory Secretariat (CSAS), DFO Science provides peer-reviewed science advice to fisheries managers at regular meetings (Regional Assessment Process).
-These meetings are attended by a variety of stakeholders including harvesters, academics, and NGOs.
-All participants attending a CSAS meeting are encouraged to question, comment and constructively challenge the science presented and the goal is to provide the best possible science advice to the Minister, managers, stakeholders and the public.
+We  believe that the dashboard tool is a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
 While the entire data-pipeline can be revealed to all by sharing code and data, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
-In the case of NCAM, considerable statistical experience and expertise is required simply to run the model (or perhaps to even tweak the model).
+In the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
 However, the dashboard removes many of these difficulties.
-The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document.
-The number of analyses can be expanded upon request and can be distributed prior to meetings and participants can assess the results and diagnostics at their leisure.
+The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document, and can be distributed prior to meetings and participants can assess the results and diagnostics at their leisure.
 Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
 
-We must acknowledge there are costs to adopting this approach of using the interactive tools.
+We must acknowledge there are costs to adopting the use of interactive tools.
 The first and obvious one being that staff are required to have both the time and training to effectively to develop these tools.
@@
 It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
-Scientific exploration is rarely a linear process, and the ability to jump around a dashboard is a more in tune with the natural way scientist’s pursue ideas.
+Scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
 In contrast, a slide presentation is effectively a linear and a single way of developing a scientific narrative; a method hardly conducive to exploring complex data problems.
 
-In summary, we submit that the dashboard tool is an important step forward in transparent and reproducible science.
-Further, this tool makes stock assessments more accessible to stakeholders.
-Beyond stock assessments, interactive documents could be used for a more fulsome review of manuscripts (i.e., reviewers can now assess claims like “we examined the residuals and found that the assumptions of the model were met”) in many fields.
-
 # References

### 2019-05-07 analysis/paper/manuscript.Rmd
Edits following meeting with Greg and Keith on the paper

@@
 Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model.
-For well-monitored stocks, the challenge has shifted from having sufficient data and information to provide sound advice on stock status, to presenting the large quantity of information produced by increasingly complex statistical models to end-users in a meaningful way.
-Traditional methods of presenting information at stakeholder meetings and to fisheries managers are simply insufficient to convey the richness of the information available.
-More importantly, static presentation formats tend to stifle meaningful discussion of model results, and important questions and assumptions that should be discussed tend to get lost under the sheer volume of output.
-A solution to this problem of communicating complex output and results to encourage engaging discussion can be through using the effective use of interactive visualization tools.
+For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way.
+Traditional formats (e.g.
+slides) of presenting information at stakeholder meetings, and to fisheries managers, are simply insufficient to convey the richness of the information available.
+More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output.
+A solution to this problem is the effective use of interactive visualization tools.
 These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
@@
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
-All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
+All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
 This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot.
-For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
+For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
 This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.
@@
 To begin to explore and understand this large data set, we built a simple shiny [@chang2018a] application using the mapping package leaflet [@cheng2018], all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio].
-As the application developed, it was incorporated into a shinydashboard [@chang2018b], as a means to quickly and dynamically subset the data (e.g.
+Shiny provides a means for the user to easily interface with the application, including options to quickly and dynamically subset the data (e.g.
 ranges of release and capture years, specific geographic locations).
+As the application developed, it was incorporated into shinydashboard [@chang2018b], to take advantage of the number of layout options available in shinydashboard.
 Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging.
@@
 From a computational perspective, analyses of a variety of large data sets has never been easier.
-However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers.
-By using a recently developed interactive tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that interactive visualization tools can streamline the process of understanding and communicating results from complex stock assessment models.
+However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of understanding and integrating data from multiple sources into a single model and communicating the methods and results to stakeholders and fisheries managers.
+Using a recently developed tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that interactive visualization tools can help meet these challenges.
 
@@
 Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
-The usual approach would be to produce static presentations and documents with a series of figures and tables, but with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and all the stakeholders involved.
+The usual approach would be to produce static presentations and documents with a series of figures and tables, but with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved.
 First, it is no longer feasible for the analyst to include and describe every figure and table produced in a single document.
@@
 Finally, some results from a retrospective analyses are included under the "Retro" tab and key inputs and outputs are included as tables under the "Tables" tab.
-The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the dashboard permits easy and efficient access to significantly more detail.
-For instance, the ability to zoom allows users to investigate specific patterns and small boxes, called tooltips, display more information about the points users hover over.
+The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the plots permit easy and efficient access to significantly more detail.
 Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.
@@
 
-The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, in addition to the formal analyses and associated diagnostics [@lewis2018 and references therin].
+The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018].
 The majority of this sequence of events, sometimes called “the data pipeline” [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis.
@@
 A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015].
-Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@lewis2018].
+Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009].
 
@@
 
+# Conclusions
+
 We must acknowledge there are costs to adopting the use of interactive tools.
@@
 Scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
-In contrast, a slide presentation is effectively a linear and a single way of developing a scientific narrative; a method hardly conducive to exploring complex data problems.
 

### 2019-05-07 analysis/paper/manuscript.Rmd
Draft an abstract

@@
 csl: apa.csl
+abstract: |
+  **Abstract**: In the age of data-intensive science and big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
+The field of stock assessment is no exception as the volume and variety of the data used in the field has grown alongside the computational methods used to integrate them.
+While this growth in data and model complexity has mediated some of the difficulties in assessing the status of many stocks, the process of communicating the results to stakeholders and fisheries managers in a meaningful way has become more daunting.
+The traditional approach of presenting information across a series of static slides often fail to convey the richness of information available and, as such, important patterns and details are easily overlooked.
+Here we contend that this problem can be mediated through the effective use of new open source tools for building interactive visualizations.
+These tools facilitate a deeper and more collective understanding of both the data and models used to inform stock assessments as they allow a broader audience to conduct detailed explorations of the results.
+*...need a punchy last line...*
 ---

### 2019-05-09 analysis/paper/manuscript.Rmd
Minor tweaks following edits from Greg and Keith

--- 2019-05-07 analysis/paper/manuscript.Rmd
@@
 csl: apa.csl
+abstract: |
+  **Abstract**: In the age of big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
+The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
+While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to stakeholders and fisheries managers in a meaningful way has become more daunting.
+The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked.
+Here we contend that this problem can be mediated through the effective use of new open source tools for building interactive visualizations.
+These tools allow a broader audience to conduct detailed explorations of the results, leading to a deeper and collective understanding of both the data and models used to inform stock assessments.
+As a consequence, the peer review process is more open and accessible and the resulting science advice is improved and widely supported.
 ---
@@
 This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
-Such issues have largely been curtailed in contemporary stock assessments thanks to advances in software that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
+Such issues have largely been curtailed in contemporary stock assessments thanks to advances in statistical computing that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
 Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
@@
 
-The NCAM dashboard, which is included as a self-contained html file in Supplement 2, contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
+The NCAM dashboard (Supplement 2) contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
 Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data.
@@
 We must acknowledge there are costs to adopting the use of interactive tools.
-The first and obvious one being that staff are required to have both the time and training to effectively to develop these tools.
-In our two examples, the developers were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
-Most programming issues were readily addressed through simple internet searchers, but some learning time was certainly needed (a few weeks).
-As more people in the fisheries community develop and use these tools, we expect tool develop speed and quality will only increase.
+The first and obvious one being that staff are required to have both the time and training to effectively develop these tools.
+In our two examples, the developers (first and second authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
+While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g.
+JavaScript, CSS, etc.).
+Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches.
+We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents.
 It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.

### 2019-05-13 analysis/paper/manuscript.Rmd
Add keywords and acknowledgements

@@
 ---
-title: Improving the communication and accessibility of stock assessment using interactive
-  visualization tools
-author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, etc?"
-date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
+title: |
+  Improving the communication and accessibility of stock assessment using interactive visualization tools
+author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis"
+date: |
+  Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
 output: 
@@
 abstract: |
+  **Running title**: Interactive stock assessment
+
+  **Conflict of Interest Statement**: The authors have no conflicts of interest to declare.
+
   **Abstract**: In the age of big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
@@
 As a consequence, the peer review process is more open and accessible and the resulting science advice is improved and widely supported.
+
+  **Keywords**: dashboard, data exploration, dynamic documents, integrated analysis, model validation, open stock assessment
+
 ---
@@
 
+# Acknowledgements
+
+We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
+
 # References

### 2019-05-13 analysis/paper/cover_letter.Rmd
Add cover letter

---
output: word_document
---

`r format(Sys.time(), '%Y-%m-%d')`

Professor Paul J B Hart  
Fish and Fisheries  
Department of Biology  
University of Leicester  
University Road  
Leicester  
LE1 7RH, UK

Dr. Hart,

We thank you for taking the time to examine our initial idea for our paper "Improving the communication and accessibility of stock assessment using interactive visualization tools". Based on your feedback, we have assembled a paper we feel fits the theme of Ghoti papers, and points towards ways of improving the stock assessment process. We re-iterate that this paper is not technical in nature, but mostly to encourage readers to pursue investigating these interactive visualization tools, and to convince readers that these tools are very much accessible to practicing biologists.

I look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067

### 2019-09-06 analysis/paper/manuscript.Rmd
Modifications for initial submission to Fish and Fisheries
Integrate Bob's paragraph and figures on his RStrap dashboard

@@
   Improving the communication and accessibility of stock assessment using interactive visualization tools
-author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis"
+author: "Paul M. Regular, Gregory J. Robertson, Robert Rogers, Keith P. Lewis"
 date: |
@@
 
+  **Data Availability Statement**: Data sharing is not applicable to this article as no new data were created or analyzed in this study.
+
   **Abstract**: In the age of big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
@@
 ---
+
+<!-- READ AND REFERENCE: -->
+<!-- - https://www.sciencedirect.com/science/article/pii/S1364815210000551 -->
+
+<!-- AIM FOR SHORT COMMUNICATION IN Environmental Modeling & Software -->
 
@@
 As such, stock assessment biologists need to manage large volumes of data from a variety of sources.
-Time series of reported landings and catch-at-age are often analyzed in conjunction with data from “fishery-independent” surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
+Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
@@
 For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
-This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.
+This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine "fishery independent" survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
+
+<!-- should the concept of a dashboard be introduced here? -->
+
+Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018].
+Currently, these data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981].
+Both the inputs and outputs from the analysis are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
+In order for people to quickly and reliability explore these data without iteratively modifying R scripts, an app was built using a combination of shiny [@chang2018a], crosstalk [@cheng2016], flexdashboard [@allaire2017], plotly [@sievert2018], and R Markdown [@allaire2018] all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio].
+The app started as a way to visualize estimates of biomass and abundance trends of specific species using the flexdashboard package, however, shiny was latter applied as it allows stratified analyses to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session.
+Both crosstalk and plotly were incorporated to allow the user to interact with the data.
+The app contains four primary tabs ("Survey Indices", "Age & Length Distributions", "Recruitment", and "Help") containing stock level estimates of biomass (Figure \@ref(fig:biomass)) and abundance, mean numbers and weights per standardized survey tow, stock level biomass broken out by division, age and length frequencies (Figure \@ref(fig:freq)), and recruitment indices.
+The number of potential inputs to RStrap are rather large and most do not need to be modified.
+Some inputs however, do require modification.
+The input interface on the RStrap Explorer app is divided into two parts: basic and advanced inputs.
+Basic inputs are those like species, season, and survey year(s).
+Advanced inputs allow the user to select unconverted data (default) or converted data (which allows for a single complete time-series to be used as opposed to time-series broken by changes in survey gear) and select whether they would like analyses conducted by length and/or age, among others.
+This app was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.
+
+```{r biomass, echo = FALSE, fig.cap = '- Screen shot of RStrap Explorer tool, using flexdashboard, plotly, and shiny.
+On the left side of the figure, one can see the "Basic Inputs" sidebar, which is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest.
+The "Biomass Index" sub-tab is illustrated here.
+This sub-tab shows the unconverted time-series of biomass data with 95% confidence intervals (Engle and Campelen) for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889).
+All trend lines can be toggled off using plotly’s dynamic user interface.
+The user also has the ability to hover over the points to compare estimates and export the figure as a png.'}
+knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
+```
+
+```{r freq, echo = FALSE, fig.cap = '- "Age and Length Frequencies" tab of the RStrap Explorer tool.
+This tab illustrates the length frequency distributions of both cod (*Gadus morhua*; species code 438) and American plaice (*Hippoglossoides platessoides*; species code 889) binned into 2 cm length bins.
+The binning is easily modified using the "Length Group" field in the "Advanced Options" drop down menu (shown on left side).'}
+knitr::include_graphics("figures/RStrap_dashboard_freq.png")
+```
 
@@
 
-```{r map, echo = FALSE, fig.cap = '- Screen shot of cod tag mapping tool using shinydashboard and leaflet.
+```{r map, echo = FALSE, fig.cap = 'Screen shot of cod tag mapping tool using shinydashboard and leaflet.
 The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
@@
 
-```{r summary, echo = FALSE, fig.cap = '- Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
+```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
 In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
@@
 
-In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained “dashboard” was developed for the 2018 assessment of Northern cod [@dwyer2019].
+In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019].
 The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
@@
 
-```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
+```{r rv, echo = FALSE, fig.cap = 'Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
 Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
@@
 
-```{r trends, echo = FALSE, fig.cap = '- Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
+```{r trends, echo = FALSE, fig.cap = 'Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
 knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
@@
 The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018].
-The majority of this sequence of events, sometimes called “the data pipeline” [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis.
+The majority of this sequence of events, sometimes called "the data pipeline" [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis.
 However, decisions made along the data pipeline increasingly influence the outcome of the study.
-Gelman and Loken [-@gelman2014] coined the term “garden of forking paths” to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis.
+Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis.
 Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible.

### 2019-09-10 inst/rmd/vis_fit.Rmd
Update NCAM dashboard and include Comps and Projection tabs this time

@@
 library(data.table)
+library(dplyr)
 library(webshot)
@@
 
-cols <- viridis::viridis(4)[-1]
+cols <-  viridis::viridis(4)[c(2, 3, 1, 4)]
 cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
@@
 
-Row
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
 
+### Numbers at age
+
+```{r}
+
+d <- tidy_mats(list(obs = t(inputs$catch.num[as.character(years), as.character(ages)]),
+                    pred = fit$rep$pred_catch),
+               ages = ages, years = years)
+d$cohort <- factor(d$year - d$age)
+d$age <- factor(d$age)
+
+facet_obs_pred(d, by = "age", type = "log",
+               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
+               nrows = 6, ylab = "Proportion",
+               obs_col = cols[2], pred_col = cols[1]) %>%
+  layout(yaxis = list(rangemode = "tozero"))
+
+```
+
+> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
+
+### Predictions by age
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "log", group = "age",
+              group_cols = viridis::viridis(nlevels(d$age), 
+                                            direction = 1, end = 0.8),
+              ylab = "Numbers")
+```
+
+> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
+
+### Predictions by cohort
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "log", group = "cohort",
+              group_cols = viridis::viridis(nlevels(d$cohort), 
+                                            direction = 1, end = 0.8),
+              ylab = "Numbers")
+```
+
+> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
+
+### Numbers at age surface
+
+```{r}
+plot_surface(mat = t(inputs$catch.num[as.character(years), as.character(ages)]), 
+             ages = ages, years = years, name = "N")
+```
+
+> Note: model fits to composition - this plot is included for illustration
+
 Row {.tabset}
@@
 
-### Observed and predicted catch at age
+### Proportions at age
 
@@
                ages = ages, years = years)
+d$cohort <- factor(d$year - d$age)
+d$age <- factor(d$age)
+
 facet_obs_pred(d, by = "age",
@@
 ```
+
+### Fits by age
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "linear", group = "age",
+              group_cols = viridis::viridis(nlevels(d$age), 
+                                            direction = 1, end = 0.8),
+              ylab = "Proportion", stack_group = TRUE)
+```
+
+> Note: values are stacked
+
+### Fits by cohort
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "linear", group = "cohort",
+              group_cols = viridis::viridis(nlevels(d$cohort), 
+                                            direction = 1, end = 0.8),
+              ylab = "Proportion", stack_group = TRUE)
+```
+
+> Note: values are stacked
 
@@
 
-### Catch at age surface (proportion)
+### Proportions at age surface
 
@@
 
-<!-- ### Catch at age surface (numbers) -->
-
-<!-- ```{r} -->
-<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
-<!-- ``` -->
-
-<!-- > Note: model fits to composition - this plot is included for illustration -->
-
-RV Survey
+RV Survey {data-navmenu="Surveys"}
 ================================================================================
@@
 d$pred[d$wt == 0] <- NA
+d$cohort <- factor(d$year - d$age)
+d$age <- factor(d$age)
+
 facet_obs_pred(d, by = "age",
@@
 
+```
+
+> Y-axes in log scale
+
+### Fits by age
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "log", group = "age",
+              group_cols = viridis::viridis(nlevels(d$age), 
+                                            direction = 1, end = 0.8),
+              ylab = "Mean numbers per tow")
+```
+
+> Y-axes in log scale
+
+### Fits by cohort
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE,
+              type = "log", group = "cohort",
+              group_cols = viridis::viridis(nlevels(d$cohort), 
+                                            direction = -1, end = 0.8),
+              ylab = "Mean numbers per tow")
 ```
@@
 
-SN Survey
+SN Survey {data-navmenu="Surveys"}
 ================================================================================
@@
                ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
+d$cohort <- factor(d$year - d$age)
+d$age <- factor(d$age)
+
 facet_obs_pred(d, by = "age",
@@
 
+```
+
+> Y-axes in log scale
+
+### Fits by age
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "log", group = "age",
+              group_cols = viridis::viridis(nlevels(d$age), 
+                                            direction = 1, end = 0.8),
+              ylab = "Sentinel survey index")
+```
+
+> Y-axes in log scale
+
+### Fits by cohort
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE,
+              type = "log", group = "cohort",
+              group_cols = viridis::viridis(nlevels(d$cohort), 
+                                            direction = -1, end = 0.8),
+              ylab = "Sentinel survey index")
 ```
@@
 
-SS Survey
+SS Survey {data-navmenu="Surveys"}
 ================================================================================
@@
 d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)
+base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% 
+  group_by(exp) %>% 
+  highlight(off = "plotly_relayout")
 
@@
   summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
+  add_markers(x = ~jitter(rel_year, factor = 0.5), 
+              y = ~sqrt(tot_rel), color = I(cols[1]),
               text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
@@
 
-<!-- ### Animation -->
-
-<!-- ```{r} -->
-
-<!-- sum_d <- tag_dat %>% -->
-<!--   group_by(exp, cap_year) %>% -->
-<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->
-
-<!-- sum_d %>%  -->
-<!--   group_by(exp) %>%  -->
-<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
-<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
-<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
-<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
-<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
-<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
-<!--   layout(yaxis = list(title = "N recaptured", -->
-<!--                       tickvals = y_tickvals, -->
-<!--                       ticktext = y_ticktext, -->
-<!--                       range = range(y_tickvals)), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
+### Animation
+
+```{r}
+
+sum_d <- tag_dat %>%
+  group_by(exp, cap_year) %>%
+  summarise(sum_pred = sum(pred), sum_obs = sum(obs))
+
+sum_d %>%
+  group_by(exp) %>%
+  plot_ly(x = ~cap_year, frame = ~exp, hoverinfo = "x+y+text",
+          showlegend = FALSE) %>%
+  add_markers(y = ~sqrt(sum_obs), color = I(cols[1]),
+              text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
+  add_lines(y = ~sqrt(sum_pred), color = I(cols[2]),
+            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
+  animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>%
+  layout(yaxis = list(title = "N recaptured",
+                      tickvals = y_tickvals,
+                      ticktext = y_ticktext,
+                      range = range(y_tickvals)),
+         xaxis = list(title = "Year"))
+
+```
 
@@
 base %>%
-  distinct(rel_year, Udev, exp, period) %>%
+  filter(rel_year == cap_year) %>%
   add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
@@
 
+<!-- ### Exploitable biomass at age -->
+
+<!-- ```{r} -->
+<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
+<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
+<!-- E <- B * S -->
+<!-- plot_surface(E, ages = ages, years = years, name = "Exploitable biomass") -->
+<!-- ``` -->
+
 Row {.tabset}
@@
 pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
-             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
+plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], 
+             ages = fit$tmb.data$pe_ages, name = "pe", 
+             col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
   add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
@@
 pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = 2:8, value = "res")
+               ages = fit$tmb.data$pe_ages, value = "res")
 bubble_plot(pe)
@@
 
-Row
+### Mohn's rho
+
+```{r, eval = !is.null(fit$retro)}
+
+bms_plot <- trans_est(retro_tabs$log_ssb) %>%
+  plot_mohns(col = cols[1], ylab = "SSB")
+
+abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
+  plot_mohns(ylab = "Recruitment (age 2)", col = cols[1])
+
+blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
+  plot_mohns(ylab = "SSB/Blim", col = cols[1])
+
+subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Relative difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
+
+### Difference
+
+```{r, eval = !is.null(fit$retro)}
+
+bms_plot <- retro_tabs$log_ssb %>%
+  plot_retro_deltas(in_col = cols[1], ylab = "log(SSB)")
+
+abn_plot <- retro_tabs$log_recruitment %>%
+  plot_retro_deltas(ylab = "log(Recruitment)", in_col = cols[1])
+
+blim_plot <- retro_tabs$log_ssb_lrp %>%
+  plot_retro_deltas(ylab = "log(SSB/Blim)", in_col = cols[1])
+
+subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
+
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
 ```
+
+### Mohn's rho
+
+```{r, eval = !is.null(fit$retro)}
+
+z_plot <- trans_est(retro_tabs$log_aveZ) %>%
+  plot_mohns(ylab = "Average Z (ages 5+)", col = cols[1])
+
+m_plot <- trans_est(retro_tabs$log_aveM) %>%
+  plot_mohns(ylab = "Average M (ages 5+)", col = cols[1])
+
+f_plot <- trans_est(retro_tabs$log_aveF) %>%
+  plot_mohns(ylab = "Average F (ages 5+)", col = cols[1])
+
+subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Relative difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
+
+### Difference
+
+```{r, eval = !is.null(fit$retro)}
+
+z_plot <- retro_tabs$log_aveZ %>%
+  plot_retro_deltas(ylab = "log(Average Z)", in_col = cols[1])
+
+m_plot <- retro_tabs$log_aveM %>%
+  plot_retro_deltas(ylab = "log(Average M)", in_col = cols[1])
+
+f_plot <- retro_tabs$log_aveF %>%
+  plot_retro_deltas(ylab = "log(Average F)", in_col = cols[1])
+
+subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
 
@@
                  text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
-                 col = cols[1])
+                 col = ~year, clab = "Year")
 
@@
                  text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
-                 col = cols[1])
+                 col = ~year, clab = "Year")
 p
@@
 
-```{r}
-
-message("Bits of the projection code is currently not working. Specifically the plots that require the frame argument in plotly.")
-
-```
-
-<!-- Assumptions  {data-navmenu="Projections" .storyboard} -->
-<!-- ================================================================================ -->
-
-<!-- ### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers -->
-
-<!-- ```{r} -->
-
-<!-- Cp <- fit$tmb.data$Cp -->
-<!-- Cp_string <- text_c(Cp) -->
-<!-- Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1)) -->
-<!-- Lp_string <- paste(text_c(Lp), "t") -->
-
-<!-- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>% -->
-<!--   filter(year >= max(years) - 4) %>% -->
-<!--   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   filter(year == max(years)) %>% -->
-<!--   add_lines(showlegend = FALSE, line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age")) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - The following catch multipliers were applied: `r #Cp_string` -->
-<!-- - This translates to `r #Lp_string` of catch in the projection years -->
-
-<!-- ### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994 -->
-
-<!-- ```{r} -->
-
-<!-- if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic") -->
-
-<!-- aveM <- tabs$log_aveM -->
-<!-- aveM <- exp(mean(aveM$est[!aveM$year %in% 1991:1994])) -->
-
-<!-- trans_est(tabs$log_aveM) %>% -->
-<!--   plot_trend(ylab = "Average M", col = cols[1]) %>% -->
-<!--   add_segments(x = 1995, xend = max(years) + 3, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_segments(x = 1990, xend = 1995, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), -->
-<!--                line = list(width = 1, dash = "dot")) %>% -->
-<!--   add_segments(x = min(years), xend = 1990, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_text(x = max(years) + 0.2, y = aveM, text = "long-term mean", textposition = "bottom right", -->
-<!--            color = I("black")) %>% -->
-<!--   layout(xaxis = list(range = c(min(years), max(years) + 3))) -->
-
-<!-- ``` -->
-
-<!-- ### Assume a relationship between recruits per spawner and number of spawners to project recruits in the future -->
-
-<!-- ```{r} -->
-
-<!-- if (fit$tmb.data$Rp_mode != 3) stop("TODO: make the recruitment mode documentation dynamic") -->
-
-<!-- rps_tab <- tabs$log_rps -->
-<!-- names(rps_tab) <- c("year", "rps", "rps_sd", "rps_lwr", "rps_upr") -->
-<!-- rps_tab <- rps_tab[rps_tab$year >= 1985, ] -->
-<!-- rps_tab$year <- rps_tab$year - 2 # scale back two year to allign with spawners that produced the recruits -->
-<!-- ssa_tab <- tabs$log_ssa -->
-<!-- names(ssa_tab) <- c("year", "ssa", "ssa_sd", "ssa_lwr", "ssa_upr") -->
-<!-- d <- merge(ssa_tab, rps_tab, by = "year") -->
-<!-- rec_tab <- tabs$log_recruitment -->
-<!-- names(rec_tab) <- c("year", "rec", "rec_sd", "rec_lwr", "rec_upr") -->
-<!-- rec_tab <- rec_tab[rec_tab$year >= 1985, ] -->
-<!-- rec_tab$year <- rec_tab$year - 2 # scale back two years -->
-<!-- d <- merge(d, rec_tab, by = "year") -->
-<!-- d$excluded <- d$year %in% 1991:1994 -->
-
-<!-- sub_d <- d[!d$year %in% 1991:1994, ] -->
-<!-- lin_rel <- lm(rps ~ ssa, data = sub_d) -->
-<!-- # coefficients(lin_rel) -->
-
-<!-- preds <- data.frame(year = NA, rps_sd = NA, ssa_sd = NA, ssa = seq(min(d$ssa), max(d$ssa), length.out = 100)) -->
-<!-- preds$fits <- predict(lin_rel, newdata = preds) -->
-<!-- preds$excluded <- FALSE -->
-
-<!-- plot_ly(data = d, x = ~ssa, y = ~rps, text = ~year, -->
-<!--         type = "scatter", mode = "markers", name = "est", -->
-<!--         color = ~excluded, colors = c(cols[2], "grey"), -->
-<!--         error_y = list(array = ~ qnorm(0.975) * rps_sd, -->
-<!--                        width = 0, thickness = 1), -->
-<!--         error_x = list(array = ~ qnorm(0.975) * ssa_sd, -->
-<!--                        width = 0, thickness = 1)) %>% -->
-<!--   add_text(text = ~year, textfont = list(color = toRGB("grey50")), textposition = "top right", -->
-<!--            error_y = list(visible = FALSE), -->
-<!--            error_x = list(visible = FALSE)) %>% -->
-<!--   layout(title = "Spawners vs Recruits / Spawner", -->
-<!--          xaxis = list(title = "log(Spawners)"), -->
-<!--          yaxis = list(title = "log(Recruits / Spawner)", zerolinecolor = toRGB("lightgrey")), -->
-<!--          showlegend = FALSE) %>% -->
-<!--   add_lines(data = preds, x = ~ssa, y = ~fits, name = "fit", color = I(cols[1])) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - Grey points indicate years excluded from the regressionn -->
-
-<!-- ### Assume averages weights and proportions mature from the last three years -->
-
-<!-- ```{r} -->
-
-<!-- mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(mat = mean(mat), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Proportion mature")) -->
-
-<!-- wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols), -->
-<!--           showlegend = FALSE) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(weight = mean(weight), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Weight (kg)")) -->
-
-<!-- subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- ### Catch multiplier projections for SSB / Blim from the last RAP -->
-
-<!-- ```{r, eval = !is.null(tmp$last)} -->
-
-<!-- d <- trans_est(last_tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- d <- d[d$year <= max(years) + 1, ] -->
-
-<!-- past_d <- trans_est(last_tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- current_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s)) -->
-<!-- current_d <- do.call(rbind, current_d) -->
-<!-- current_d$model <- "current RAP" -->
-<!-- d$model <- "last RAP" -->
-<!-- d <- rbind(d, current_d) -->
-
-<!-- plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model", -->
-<!--           vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Results {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~ -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_ratio) -->
-<!-- d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100 -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- xlim <- c(max(years) + 0.9, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%", -->
-<!-- #           vline = max(years) + 1, col = cols[1]) %>% -->
-<!-- #   layout(yaxis = list(tickvals = seq(-200, 200, by = 25))) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- ### SSB / Blim -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-
-<!-- past_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(years) - 2.1, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", -->
-<!-- #           vline = max(years) + 1, hline = 1, col = cols[1]) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_ratio -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-<!-- ### Prob(SSB~yp~ &ge; B~lim~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_lrp -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
+Assumptions  {data-navmenu="Projections" .storyboard}
+================================================================================
+
+### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers
+
+```{r}
+
+Cp <- fit$tmb.data$Cp
+Cp_string <- text_c(Cp)
+Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1))
+Lp_string <- paste(text_c(Lp), "t")
+
+tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>%
+  filter(year >= max(years) - 4) %>%
+  plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>%
+  add_lines(alpha = 0.5) %>%
+  filter(year == max(years)) %>%
+  add_lines(showlegend = FALSE, line = list(width = 3)) %>%
+  layout(xaxis = list(title = "Age"))
+
+```
+
+***
+- The following catch multipliers were applied: `r Cp_string`
+- This translates to `r Lp_string` of catch in the projection years
+
+### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994
+
+```{r}
+
+if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic")
+
+Mpe <- fit$rep$Mpe_matrix
+Mproj <- replicate(fit$tmb.data$Yp, fit$tmb.data$pnm)
+bpe <- fit$rep$bpe[, (fit$tmb.data$Y + 1):(fit$tmb.data$Y + fit$tmb.data$Yp)]
+Mproj <- Mproj * exp(bpe)
+Mproj <- cbind(Mpe, Mproj)
+proj_years <- fit$tmb.data$proj_years
+
+plot_surface(mat = Mproj, ages = ages, years = c(years, proj_years), name = "M")
+
+```
+
+### Assume mean recruitment from the last three years
+
+```{r}
+
+last_three <- tail(fit$tmb.data$years, 3)
+mean_rec <- exp(mean(tabs$log_recruitment$est[tabs$log_recruitment$year %in% last_three]))
+mean_rec <- data.frame(year = c(last_three, last_three[3] + fit$tmb.data$Yp), 
+                       mean_rec = mean_rec * 1000, grp = "mean")
+trans_est(tabs$log_recruitment, scale = 0.001) %>%
+  plot_trend(ylab = "Recruitment", col = cols[1]) %>% 
+  add_lines(data = mean_rec, x = ~year, y = ~mean_rec, name = "mean",
+            color = I("black"), line = list(width = 1)) %>% 
+  add_text(x = max(years) + 1.2, y = mean_rec$mean_rec[1], 
+           text = "mean recruitment", 
+           textposition = "bottom right",
+           color = I("black")) %>%
+  layout(xaxis = list(range = c(min(years), max(years) + fit$tmb.data$Yp)))
+
+```
+
+### Assume averages weights and proportions mature from the last three years
+
+```{r}
+
+mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>%
+  filter(year >= max(years) - 2) %>%
+  plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>%
+  add_lines(alpha = 0.5) %>%
+  group_by(age) %>%
+  summarise(mat = mean(mat), year = "3y mean") %>%
+  add_lines(line = list(width = 3)) %>%
+  layout(xaxis = list(title = "Age"),
+         yaxis = list(title = "Proportion mature"))
+
+wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>%
+  filter(year >= max(years) - 2) %>%
+  plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols),
+          showlegend = FALSE) %>%
+  add_lines(alpha = 0.5) %>%
+  group_by(age) %>%
+  summarise(weight = mean(weight), year = "3y mean") %>%
+  add_lines(line = list(width = 3)) %>%
+  layout(xaxis = list(title = "Age"),
+         yaxis = list(title = "Weight (kg)"))
+
+subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE)
+
+```
+
+Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`}
+================================================================================
+
+### Catch multiplier projections for SSB / Blim from the last RAP
+
+```{r, eval = !is.null(tmp$last)}
+
+d <- trans_est(last_tabs$Cproj_log_ssb_lrp)
+d$scenario <- factor(d$scenario)
+d$type <- "Projected"
+
+past_d <- trans_est(last_tabs$log_ssb_lrp)
+past_d <- past_d[past_d$year != max(past_d$year), ] # drop the last year here (it's in the projection output)
+past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
+past_d <- do.call(rbind, past_d)
+past_d$type <- "Estimated"
+d <- rbind(past_d, d)
+xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
+ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
+
+current_d <- trans_est(tabs$log_ssb_lrp)
+current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s))
+current_d <- do.call(rbind, current_d)
+current_d$model <- "current RAP"
+current_d$type <- "Estimated"
+d$model <- "last RAP"
+d <- rbind(d, current_d)
+
+plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model",
+          vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE,
+          darken_proj = TRUE)
+
+```
+
+Retro projections {data-navmenu="Projections" `r if (is.null(fit$retro)) ".hidden"`}
+================================================================================
+
+### Catch multiplier projections for SSB / Blim from retrospective fits
+
+```{r, eval = !is.null(fit$retro)}
+
+d <- trans_est(retro_tabs$Cproj_log_ssb_lrp)
+d$terminal_year <- as.numeric(as.character(d$terminal_year))
+d$scenario <- factor(d$scenario)
+d$type <- "Projected"
+
+past_d <- trans_est(retro_tabs$log_ssb_lrp)
+past_d$terminal_year <- as.numeric(as.character(past_d$terminal_year))
+past_d <- past_d[past_d$year <= past_d$terminal_year, ] # drop the last year (it's in the projection output)
+past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
+past_d <- do.call(rbind, past_d)
+past_d$type <- "Estimated"
+d <- rbind(past_d, d)
+d <- d[order(d$scenario, d$terminal_year, d$year), ]
+
+# xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
+# ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
+
+plot_proj(data = d, ylab = "SSB/Blim", group = "terminal_year",
+          hline = 1, col = rev(viridis::viridis(length(unique(d$terminal_year)))), 
+          showlegend = TRUE, markers = FALSE, darken_proj = TRUE)
+
+```
+
+Results {data-navmenu="Projections"}
+================================================================================
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~
+
+```{r}
+
+d <- trans_est(tabs$Cproj_log_ssb_ratio)
+d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100
+d$scenario <- factor(d$scenario)
+xlim <- c(max(years) + 0.9, max(years) + 4.1)
+ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
+
+plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%",
+          vline = max(years) + 1, col = cols[1]) %>%
+  layout(yaxis = list(tickvals = seq(-200, 200, by = 25)))
+
+```
+
+### SSB / Blim
+
+```{r}
+
+d <- trans_est(tabs$Cproj_log_ssb_lrp)
+d <- d[d$year != max(years_plus1), ] # drop the first year (it's in the time series)
+d$type <- "Projected"
+d$scenario <- factor(d$scenario)
+
+## Append full time series to projections
+past_d <- trans_est(tabs$log_ssb_lrp)
+past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
+past_d <- do.call(rbind, past_d)
+past_d$type <- ifelse(past_d$year == max(years_plus1), "Projected", "Estimated")
+d <- rbind(past_d, d)
+xlim <- c(max(years) - 2.1, max(years) + 4.1)
+ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
+
+plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim",
+          vline = max(years) + 1, hline = 1, col = cols[1], darken_proj = TRUE)
+
+```
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~)
+
+```{r}
+
+growth <- tabs$Cproj_log_ssb_ratio
+growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)
+
+d <- growth[growth$year != max(years) + 1, ]
+d$scenario <- factor(d$scenario)
+plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
+  add_heatmap() %>%
+  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
+                  showarrow = FALSE, font = list(color = "white")) %>%
+  colorbar(title = "P", limits = c(0, 1)) %>%
+  layout(yaxis = list(title = "Catch multiplier"),
+         xaxis = list(title = "Year", 
+                      tickvals = min(d$year):max(d$year)))
+
+```
+
+### Prob(SSB~yp~ &ge; B~lim~)
+
+```{r}
+
+growth <- tabs$Cproj_log_ssb_lrp
+growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)
+
+d <- growth[growth$year != max(years) + 1, ]
+d$scenario <- factor(d$scenario)
+plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
+  add_heatmap() %>%
+  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
+                  showarrow = FALSE, font = list(color = "white")) %>%
+  colorbar(title = "P", limits = c(0, 1)) %>%
+  layout(yaxis = list(title = "Catch multiplier"),
+         xaxis = list(title = "Year",
+                      tickvals = min(d$year):max(d$year)))
+
+```
 
@@
 ```{r}
-N_matrix <- fit$rep$N_matrix
-dimnames(N_matrix) <- list(age = ages, year = years_plus1)
+N_matrix <- rbind(fit$rep$N_matrix, fit$rep$abundance)
+dimnames(N_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
 knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
@@
 ```{r}
-ssa_matrix <- fit$rep$ssa_matrix
-dimnames(ssa_matrix) <- list(age = ages, year = years_plus1)
+ssa_matrix <- rbind(fit$rep$ssa_matrix, fit$rep$ssa)
+dimnames(ssa_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
 knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
@@
 ```{r}
-B_matrix <- fit$rep$biomass_matrix
-dimnames(B_matrix) <- list(age = ages, year = years_plus1)
+B_matrix <- rbind(fit$rep$biomass_matrix, fit$rep$biomass)
+dimnames(B_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
 knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
@@
 ```{r}
-ssb_matrix <- fit$rep$ssb_matrix
-dimnames(ssb_matrix) <- list(age = ages, year = years_plus1)
+ssb_matrix <- rbind(fit$rep$ssb_matrix, fit$rep$ssb)
+dimnames(ssb_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
 knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
+```
+
+<!-- ### Exploitable biomass -->
+
+<!-- ```{r} -->
+<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
+<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
+<!-- E <- B * S -->
+<!-- ebms_matrix <- rbind(E, colSums(E)) -->
+<!-- dimnames(ebms_matrix) <- list(age = c(ages, "Total"), year = years) -->
+<!-- knitr::kable(t(ebms_matrix) / 1000, -->
+<!--              caption = "Exploitable biomass estimates (kt).", digits = 3) -->
+<!-- ``` -->
+
+### Predicted catch
+
+```{r}
+pred_catch_wt_matrix <- rbind(fit$rep$pred_catch_wt_matrix, fit$rep$pred_catch_wt)
+dimnames(pred_catch_wt_matrix) <- list(age = c(ages, "Total"), year = years)
+knitr::kable(t(pred_catch_wt_matrix) / 1000, 
+             caption = "NCAM predicted catch at age (kt).", digits = 3)
 ```
@@
 ```{r}
+pe_ages <- fit$tmb.data$pe_ages
 M_matrix <- fit$rep$Mpe_matrix
 dimnames(M_matrix) <- list(age = ages, year = years)
-M_matrix <- M_matrix[as.character(2:8), ]
-rownames(M_matrix) <- c(as.character(2:7), "8+")
+M_matrix <- M_matrix[as.character(pe_ages), ]
+rownames(M_matrix) <- c(as.character(head(pe_ages, -1)), paste0(tail(pe_ages, 1), "+"))
 knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)

### 2019-09-10 R/plot.R

@@ -184,12 +184,98 @@ plot_retro <- function(data, col = '#1f77b4', ...) {
   data[data$terminal_year != max_year, c("lwr", "upr")] <- NA
   p <- plot_trend(data, group = "terminal_year", col = col, ...)
   data$grp <- data$terminal_year
-  marker_data <- data %>% group_by(terminal_year) %>% filter(row_number() == n())
+  marker_data <- data %>% group_by(terminal_year) %>% filter(dplyr::row_number() == n())
   p %>% add_markers(data = marker_data, x = ~year, y = ~est)
 
 }
 
 
+#' Plot relative differences from retrospective fits
+#'
+#' @description Plot of the relative differences of the estimates from the
+#' model with all years included against estimates from a series of retrospective
+#' "peals" or "folds" of the data.
+#'
+#' @param data       Data produced by \code{\link{tidy_retro}}
+#' @param ylab       Label for the y axis
+#' @param col        Plot color
+#'
+#' @export
+#'
+
+plot_mohns <- function(data = NULL, ylab = "Relative difference", col = '#1f77b4') {
+
+  d <- data
+  d$terminal_year <- as.numeric(as.character(d$terminal_year))
+  if (max(d$year) > max(d$terminal_year)) {
+    d$terminal_year <- d$terminal_year + 1 # plus one if the estimates are steped forward by one year
+  }
+  td <- d[d$terminal_year == max(d$terminal_year), c("year", "est")]   # terminal data
+  rd <- d[d$terminal_year != max(d$terminal_year) &
+            d$year == d$terminal_year,
+          c("year", "est", "terminal_year")]                           # retrospective data
+  cd <- merge(rd, td, by = "year", suffixes = c("_r", "_t"))           # combined data (r = retro, t = terminal)
+  cd$pdiff <- (cd$est_r - cd$est_t) / cd$est_t
+
+  mohns <- mean(cd$pdiff)
+
+  plot_ly(data = cd, x = ~year, y = ~pdiff,
+          hoverinfo = "x+y", showlegend = FALSE,
+          color = I(col)) %>%
+    add_markers() %>%
+    add_segments(x = ~year, xend = ~year, y = 0, yend = ~pdiff) %>%
+    layout(xaxis = list(title = "Year"),
+           yaxis = list(title = ylab),
+           annotations = top_label(paste("Mohn's rho =", round(mohns, 3))))
+
+}
+
+#' Plot differences from retrospective fits
+#'
[…]

### 2019-09-10 R/tidy.R

@@ -176,16 +176,17 @@ tidy_mats <- function(mats, ages = NULL, years = NULL) {
     tidy_mat(mats[[m]], ages = ages, years = years, value = m)
   })
 
-  d <- merge(d_list[[1]], d_list[[2]], by = c("age", "year"))
-  if (length(d_list) > 2) {
-    d <- lapply(d_list[3:length(d_list)], function(x, y) merge(x, y, by = c("age", "year")), d)
-    d <- d[[1]]
+  for (i in seq_along(d_list)[-1]) {
+    if (i == 2) {
+      d <- merge(d_list[[1]], d_list[[2]], by = c("age", "year"))
+    } else {
+      d <- merge(d, d_list[[i]], by = c("age", "year"))
+    }
   }
 
   d
 
 }
-
 
 
 
@@ -220,8 +221,8 @@ tidy_pars <- function(model) {
                "log_std_D" = "1 - exp - D variance parameter - $\\sigma_{D}$",
                "logit_ar_D_age" = "1 - plogis - Age correlation in D - $\\varphi_{D,age}$",
                "logit_ar_D_year" = "1 - plogis - Year correlation in D - $\\varphi_{D,year}$",
-               "log_R1" = "1000000 - exp - Mean log-recruitment (pre 1992; million) - $r_1$",
-               "log_R2" = "1000000 - exp - Mean log-recruitment (post 1992; million) - $r_2$",
+               "log_R1" = "1000000 - exp - Mean recruitment (pre 1992; billion) - $r_1$",
+               "log_R2" = "1000000 - exp - Mean recruitment (post 1992; billion) - $r_2$",
                "log_sd_log_R" = "1 - exp - Variance of log-recruitment - $\\sigma_{r}$",
                "log_std_old_Udev" = "1 - exp - Variance of tagging F deviation from stock F (pre 1997) - $\\sigma_{f_{x,1}}$",
                "log_std_Udev" = "1 - exp - Variance of tagging F deviation from stock F (post 1997) - $\\sigma_{f_{x,2}}$",

### 2019-09-10 analysis/dashboards/make_NCAM_dashboard.R

@@ -3,9 +3,10 @@
 ## (object is named inputs)
 load("data/inputs.RData")
 
-## load NCAM model output
-## (object is named mshift)
-load("data/mshift.RData")
+## load NCAM model output from the last two years
+## (object names match file names)
+load("data/ncam_2018.RData")
+load("data/ncam_2019.RData")
 
 ## source plotting functions
 ## (these functions are used in the Rmarkdown file that makes the dashboard)
@@ -20,4 +21,12 @@ source("R/tidy.R")
 source("R/vis_fit.R")
 
 ## Run the vis_fit function to make the dashboard
-vis_fit(mshift, output_file = "analysis/dashboards/NCAM_dashboard.html")
+## (comp - named list of model output to compare)
+## (last - model output from the last assessment)
+vis_fit(ncam_2019, 
+        last = ncam_2018,                     
+        comp = list("NCAM 2018" = ncam_2018, 
+                    "NCAM 2019" = ncam_2019),
+        output_file = "analysis/dashboards/NCAM_dashboard.html")
+
+

### 2019-09-11 analysis/paper/manuscript.Rmd
First pass at adressing comments from Reviewer 1

@@
 
-  **Abstract**: In the age of big data, scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
+  **Abstract**: Scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
 The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
@@
 Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs.
-First, we focus on data that are commonly presented in stock assessments and use interactive maps to simplify the detailed exploration of data from a long-term tagging program.
+First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program.
 Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
@@
 For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
-This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine "fishery independent" survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
+This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
 
 <!-- should the concept of a dashboard be introduced here? -->
+
+## Survey data
 
@@
 ```
+
+## Tagging data
 
@@
 
-The NCAM dashboard (Supplement 2) contains a series of tabs, the first of which provides terse point-form background on the model (tab named "Background").
-Subsequent tabs provide a series of diagnostic plots for assessing model fits to the catch ("Catch"), research vessel autumn trawl survey ("RV survey"), Sentinel fishery survey ("SN survey"), inshore acoustic survey ("SS survey"), and tagging ("Tagging") data.
-For instance, the "RV survey" tab includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)).
-The dashboard also includes tabs focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity").
-Finally, some results from a retrospective analyses are included under the "Retro" tab and key inputs and outputs are included as tables under the "Tables" tab.
-The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, the interactive nature of the plots permit easy and efficient access to significantly more detail.
-Because the dashboard is contained in an html file, it acts as an interactive assessment document that can be shared for peer-review.
-This allows colleagues and stakeholders to independently scrutinize details of the data and model that are not easily accessed by users other than the analyst.
+The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background").
+Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data.
+For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)).
+The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity").
+Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
+The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides: 1) interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
+both broader patterns and finer details in the data are accessible via zooming and tooltips); and, 2) the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides.
+Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and model to model comparisons.
+The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file.
+This allows others to independently scrutinize details of the data and model that are typically only accessible to the analyst.
 Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.
+For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model.
+These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussion helped the assessment biologists, managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables [e.g Prob(SSB~2020~ &ge; B~lim~)].
 
@@
 
-We  believe that the dashboard tool is a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
-While the entire data-pipeline can be revealed to all by sharing code and data, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
+The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (e.g.
+_________).
+We  believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
+While these tools may not reveal the entire data-pipeline, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
 In the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
 However, the dashboard removes many of these difficulties.
-The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document, and can be distributed prior to meetings and participants can assess the results and diagnostics at their leisure.
+The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document that can be distributed prior to meetings and, as such, participants can assess the results and diagnostics at their leisure.
 Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
@@
 The first and obvious one being that staff are required to have both the time and training to effectively develop these tools.
-In our two examples, the developers (first and second authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
+In our two examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
 While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g.

### 2019-09-13 analysis/paper/manuscript.Rmd
Add more references to the intro as suggested by Reviewer 4

@@
 
-Quantitative stock assessment plays a central role in modern fisheries management.
+Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992].
 Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments.
 These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
-Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model.
+Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2009; @maunder2013].
 For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way.
@@
 More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output.
-A solution to this problem is the effective use of interactive visualization tools.
+A solution to this problem is the effective use of interactive visualization tools [@keena2016].
 These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
+As a result, an increasing number of scientists across a wide range of disciplines, including fisheries research, are starting to apply interactive visualization tools to explore and communicate their results [e.g.
+@jones2016; @yeatman2018; @letcher2018].
 

### 2019-09-13 analysis/paper/manuscript.Rmd
Address comments from reviewer 3 (namely add an appendix listing some examples) and reviewer 2

@@
 These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
-Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2009; @maunder2013].
+Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013].
 For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way.
@@
 These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
-As a result, an increasing number of scientists across a wide range of disciplines, including fisheries research, are starting to apply interactive visualization tools to explore and communicate their results [e.g.
+As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g.
 @jones2016; @yeatman2018; @letcher2018].
+This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).
 
@@
 A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015].
-Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009].
+Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009; @lowndes2017].
 
-The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (e.g.
-_________).
+The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (see Appendix A).
 We  believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
@@
 We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
+We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and two anonymous reviewers.
+
+# Appendix A
+
+Partial list of applications developed to explore and communicate marine and fisheries science:
+
+- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
+- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
+- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
+- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
+- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
+- VISA tool (e.g.
+https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"
 

### 2019-09-24 analysis/paper/manuscript.Rmd
Address Keith's comments on the ms down to the Survey data section

@@
 
-<!-- READ AND REFERENCE: -->
-<!-- - https://www.sciencedirect.com/science/article/pii/S1364815210000551 -->
-
-<!-- AIM FOR SHORT COMMUNICATION IN Environmental Modeling & Software -->
+*The greatest value of a picture is when it forces us to notice what we never expected to see.*  
+- Tukey [-@tukey1977]
 
@@
 Traditional formats (e.g.
-slides) of presenting information at stakeholder meetings, and to fisheries managers, are simply insufficient to convey the richness of the information available.
-More importantly, these static and linear formats tend to stifle meaningful discussion as important details and patterns tend to get lost under the sheer volume of output.
+slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
 A solution to this problem is the effective use of interactive visualization tools [@keena2016].
@@
 Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
-This structure corresponds to two important  steps in the stock assessment process: 1) knowing the data, and 2) data modeling.
+This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
 
-# Knowing the data
+# Data exploration
 
@@
 All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
-This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot.
+It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988].
 For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
-This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
+These abilities are illustrated  by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
 

### 2019-09-30 analysis/paper/manuscript.Rmd
Address Keith's suggestions down to the Tagging section

@@
 A solution to this problem is the effective use of interactive visualization tools [@keena2016].
-These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community.
+These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018].
 As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g.
@@
 Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs.
-First, we focus on data that are commonly presented in stock assessments and use interactive plots to simplify the detailed exploration of data from long-term monitoring programs program.
-Second, we focus on the modeling aspect of stock assessment and demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.
+We especially highlight the use of interactive visuals inside dynamic documents called "dasbhoards".
+The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
+We first demonstrate the application of these tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
+Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dasbhoards can be used to explore, diagnose and communicate results from an integrated assessment model.
 This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
+In each case, the dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).
+
+```{r pkg, results = "asis", echo = FALSE}
+
+## Package name and description
+pkg <- c(shiny = 'Makes it incredibly easy to build interactive web applications with R.
+Automatic "reactive" binding between inputs and outputs and extensive prebuilt widgets make it possible to build beautiful, responsive, and powerful applications with minimal effort [@chang2018a]',
+         shinydashboard = "Create dashboards with 'Shiny'. This package provides a theme on top of 'Shiny', making it easy to create attractive dashboards [@chang2018b].",
+         rmarkdown = "Convert R Markdown documents into a variety of formats [@allaire2018].",
+         flexdashboard = "Format for converting an R Markdown document to a grid oriented dashboard.
+The dashboard flexibly adapts the size of it's components to the containing web page [@allaire2017].",
+         plotly = "Easily translate 'ggplot2' graphs to an interactive web-based version and/or create custom web-based visualizations directly from R [@sievert2018].",
+         leaflet = "Create and customize interactive maps using the 'Leaflet' JavaScript library and the 'htmlwidgets' package.
+These maps can be used directly from the R console, from 'RStudio', in Shiny applications and R [@cheng2018].",
+         crosstalk = 'Provides building blocks for allowing HTML widgets to communicate with each other, with Shiny or without (i.e. static .html files) [@cheng2016].')
+
+tab <- data.frame(names(pkg), unname(pkg))
+names(tab) <- NULL
+# names(tab) <- c("**Package**", "**Description**")
+tab[, 1] <- paste0("**`", tab[, 1], "`**")
+
+knitr::kable(tab, caption = "Names and descriptions of the key packages used to build the interactive applications presented in this paper.
+Descriptions are directly from the `DESCRIPTION` file of each package.")
+
+```
 
@@
 
-<!-- should the concept of a dashboard be introduced here? -->
-
 ## Survey data
@@
 Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018].
-Currently, these data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981].
-Both the inputs and outputs from the analysis are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
-In order for people to quickly and reliability explore these data without iteratively modifying R scripts, an app was built using a combination of shiny [@chang2018a], crosstalk [@cheng2016], flexdashboard [@allaire2017], plotly [@sievert2018], and R Markdown [@allaire2018] all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio].
-The app started as a way to visualize estimates of biomass and abundance trends of specific species using the flexdashboard package, however, shiny was latter applied as it allows stratified analyses to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session.
+These data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981].
+Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
+In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)).
+RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results.
+The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session.
 Both crosstalk and plotly were incorporated to allow the user to interact with the data.
-The app contains four primary tabs ("Survey Indices", "Age & Length Distributions", "Recruitment", and "Help") containing stock level estimates of biomass (Figure \@ref(fig:biomass)) and abundance, mean numbers and weights per standardized survey tow, stock level biomass broken out by division, age and length frequencies (Figure \@ref(fig:freq)), and recruitment indices.
+
+RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis.
 The number of potential inputs to RStrap are rather large and most do not need to be modified.
 Some inputs however, do require modification.
-The input interface on the RStrap Explorer app is divided into two parts: basic and advanced inputs.
+The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs.
 Basic inputs are those like species, season, and survey year(s).
-Advanced inputs allow the user to select unconverted data (default) or converted data (which allows for a single complete time-series to be used as opposed to time-series broken by changes in survey gear) and select whether they would like analyses conducted by length and/or age, among others.
-This app was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.
+Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on.
+This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.
 
-```{r biomass, echo = FALSE, fig.cap = '- Screen shot of RStrap Explorer tool, using flexdashboard, plotly, and shiny.
-On the left side of the figure, one can see the "Basic Inputs" sidebar, which is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest.
+```{r biomass, echo = FALSE, fig.cap = '- Screen shot of the "Survey Indices" page of the RStrap Explorer application.
+The "Basic Inputs" sidebar is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest.
 The "Biomass Index" sub-tab is illustrated here.
-This sub-tab shows the unconverted time-series of biomass data with 95% confidence intervals (Engle and Campelen) for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889).
+This sub-tab shows a time-series of biomass data with 95% confidence intervals, with a break in the lines when the survey gear changed from an Engle to Campelen trawl, for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889).
 All trend lines can be toggled off using plotly’s dynamic user interface.
-The user also has the ability to hover over the points to compare estimates and export the figure as a png.'}
+The user also has the ability to hover over the points to compare estimates and export the figure as a png file.'}
 knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
@@
 In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019].
-The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
 We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document.
@@
 - The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
+- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo) from the Fisheries Center, University of British Columbia
+- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/) from Fisheries and Oceans Canada, Maritimes Region
 - Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana

### 2019-10-02 analysis/paper/manuscript.Rmd
Address the remainder of Keith's edits/comments

@@
 The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.
-To begin to explore and understand this large data set, we built a simple shiny [@chang2018a] application using the mapping package leaflet [@cheng2018], all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio].
-Shiny provides a means for the user to easily interface with the application, including options to quickly and dynamically subset the data (e.g.
+To begin to explore and understand this large data set, we built a simple application using shiny and leaflet (Table \@ref(tab:pkg)) that allows a user to quickly and dynamically subset the data (e.g.
 ranges of release and capture years, specific geographic locations).
-As the application developed, it was incorporated into shinydashboard [@chang2018b], to take advantage of the number of layout options available in shinydashboard.
+As the application developed, it was incorporated into shinydashboard, to take advantage of the number of layout options available in shinydashboard (Table \@ref(tab:pkg)).
 Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging.
@@
 However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of understanding and integrating data from multiple sources into a single model and communicating the methods and results to stakeholders and fisheries managers.
-Using a recently developed tool for exploring the integrated assessment model for Northern cod as an example, we hope to demonstrate that interactive visualization tools can help meet these challenges.
+This challenge was palpable for data-rich case of Northern cod.
 
-The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada, perhaps by virtue of its history.
+The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada.
 As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016].
@@
 Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
-The usual approach would be to produce static presentations and documents with a series of figures and tables, but with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved.
+The traditional approach would involve producing static presentations and documents with a series of figures and tables.
+However, with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved for at least two reasons.
 First, it is no longer feasible for the analyst to include and describe every figure and table produced in a single document.
@@
 
-In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019].
-We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document.
-We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots.
-Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
+In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019].
+We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document.
+We also used the crosstalk package to link the data displayed across multiple plots.
+Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
 
@@
 both broader patterns and finer details in the data are accessible via zooming and tooltips); and, 2) the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides.
-Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and model to model comparisons.
+Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.
 The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file.
-This allows others to independently scrutinize details of the data and model that are typically only accessible to the analyst.
+This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst.
 Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.
 For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model.
-These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussion helped the assessment biologists, managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables [e.g Prob(SSB~2020~ &ge; B~lim~)].
+These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussion helped the assessment biologists, managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
 
@@
 We believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
-While these tools may not reveal the entire data-pipeline, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
-In the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
-However, the dashboard removes many of these difficulties.
-The html file contains all model inputs, outputs and diagnostics in an open and easy to use interactive document that can be distributed prior to meetings and, as such, participants can assess the results and diagnostics at their leisure.
+Although these tools do not reveal the entire data pipeline, in conjunction with posting code and data on-line, they represent a significant advance over more traditional methods.
+Similarly, these tools do not fully alleviate pragmatic issues associated with reproducible research.
+For example, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
+Further, in the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
+However, the dashboard approach removes many of these obstacles.
+The html file contains all model inputs, outputs and diagnostics in an open, and easy to use, interactive document that can be distributed prior to meetings.
+As such, participants can assess the results and diagnostics at their leisure.
 Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
@@
 It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
-Scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
+Finally, scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
 

### 2019-10-04 analysis/dashboards/getting_started.Rmd
Draft an intro on a getting started dashboard

---
title: "Dashboards for stock assessments: Getting started"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
```

Background
=====================================

Column {.tabset}
-------------------------------------

### Concept

- Browser-based dashboards and interactive visualizations are becoming increasingly common and accessible  
- With a relatively shallow learning-curve, an [**R**](https://www.r-project.org/) user can use the [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/) packages to generate interactive dashboards for exploring data and models
- The concept is demonstrated here using a simple data-set and linear regression; this can be extended to more complex cases

### flexdashboard

- Uses [**rmarkdown**](https://rmarkdown.rstudio.com/) to render a group of related figures, tables and text into a dashboard  
- Layout is flexible and the components automatically resize to fill the browser and adapt to mobile displays
- Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com)
- Optionally use [**shiny**](http://shiny.rstudio.com/) or [**crosstalk**](https://rstudio.github.io/crosstalk/) to bolster interactivity

> Source: https://rmarkdown.rstudio.com/flexdashboard/

### plotly

- A graphing package that works like other R plots expect it produces interactive visualizations
- The package allows the user to create interactive web graphics from [**ggplot2**](https://ggplot2.tidyverse.org/) graphs
- Also provides a more 'direct' link to the core [plotly.js](https://plot.ly/javascript/) JavaScript library using syntax inspired by the grammar of graphics

> Source: https://plotly-r.com

### crosstalk

- Enables cross-widget interactions by linking brushing and/or filtering across multiple views
- i.e. Interactions with one plot can affect change in another plot

> Source: https://rstudio.github.io/crosstalk/

### 2019-10-15 analysis/paper/manuscript.Rmd
Address latest comments/edits from Greg and Bob

@@
 bibliography: references.bib
-csl: apa.csl
+csl: ices.csl
 abstract: |
@@
 
-  **Data Availability Statement**: Data sharing is not applicable to this article as no new data were created or analyzed in this study.
+  **Data Availability Statement**: Data sharing is not applicable to this article as no data were generated and no analysis to draw scientific inference was conducted in this study.
+
+  **Corresponding author**: Paul M.
+Regular, Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St.
+John's, Newfoundland and Labrador, A1C 5X1, Canada, E-mail: Paul.Regular@dfo-mpo.gc.ca, Phone: 1-709-772-2067
 
@@
 The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
-While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to stakeholders and fisheries managers in a meaningful way has become more daunting.
+While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to colleagues, stakeholders and fisheries managers in a meaningful way has become more daunting.
 The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked.
@@
 Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs.
-We especially highlight the use of interactive visuals inside dynamic documents called "dasbhoards".
+We especially highlight the use of interactive visuals inside dynamic documents called "dashboards".
 The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
 We first demonstrate the application of these tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
-Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dasbhoards can be used to explore, diagnose and communicate results from an integrated assessment model.
+Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dashboards can be used to explore, diagnose and communicate results from an integrated assessment model.
 This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
@@
          flexdashboard = "Format for converting an R Markdown document to a grid oriented dashboard.
-The dashboard flexibly adapts the size of it's components to the containing web page [@allaire2017].",
+The dashboard flexibly adapts the size of its components to the containing web page [@allaire2017].",
          plotly = "Easily translate 'ggplot2' graphs to an interactive web-based version and/or create custom web-based visualizations directly from R [@sievert2018].",
@@
 RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results.
-The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session.
+The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session.
 Both crosstalk and plotly were incorporated to allow the user to interact with the data.
@@
 RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis.
-The number of potential inputs to RStrap are rather large and most do not need to be modified.
-Some inputs however, do require modification.
 The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs.
@@
 Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on.
-This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.
+This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough the data checking and analysis required for formal stock assessment begins.
 
@@
 However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of understanding and integrating data from multiple sources into a single model and communicating the methods and results to stakeholders and fisheries managers.
-This challenge was palpable for data-rich case of Northern cod.
+This challenge was palpable for the data-rich case of Northern cod.
 
@@
 The first and obvious one being that staff are required to have both the time and training to effectively develop these tools.
-In our two examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but had their formal training in field-based population ecology.
+In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but all had their formal training in field-based population ecology.
 While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g.

### 2019-10-16 analysis/dashboards/getting_started.Rmd
Draft a 'getting_started' dashboard

@@
   flexdashboard::flex_dashboard:
+    orientation: rows
     source_code: embed
@@
 ```{r setup, include=FALSE}
-library(flexdashboard)
+library(plotly)
+library(crosstalk)
 ```
@@
 Background
-=====================================
+=======================================================================
 
-Column {.tabset}
--------------------------------------
+Row {data-height=300}
+-----------------------------------------------------------------------
 
-### Concept
+### Dashboards
 
@@
 - With a relatively shallow learning-curve, an [**R**](https://www.r-project.org/) user can use the [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/) packages to generate interactive dashboards for exploring data and models
-- The concept is demonstrated here using a simple data-set and linear regression; this can be extended to more complex cases
+- The concept is demonstrated on the [Skeleton](#skeleton) page using a simple data-set and linear regression; this can be extended to more complex cases
+
+> Note: this document is meant to be a high-level overview of constructing a dashboard using [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/).
+This is not a definitive guide to each package nor is this the only approach to constructing a dashboard using R.
+Please refer to the links provided for more details on how to use the packages highlighted here.
+
+Row {data-height=500}
+-----------------------------------------------------------------------
 
@@
 - Uses [**rmarkdown**](https://rmarkdown.rstudio.com/) to render a group of related figures, tables and text into a dashboard  
-- Layout is flexible and the components automatically resize to fill the browser and adapt to mobile displays
+- Layout is flexible and the components automatically re-size to fill the browser and adapt to mobile displays
 - Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com)
@@
 - i.e. Interactions with one plot can affect change in another plot  
+- Supports a wide range of [htmlwidgets](http://www.htmlwidgets.org/index.html), such as [**plotly**](https://plotly-r.com) and [**leaflet**](https://rstudio.github.io/leaflet/)
 
 > Source: https://rstudio.github.io/crosstalk/
+
+Skeleton {data-orientation=columns}
+=======================================================================
+
+Column {.tabset}
+-------------------------------------
+
+### Code
+
+```{r echo = FALSE, comment = ""}
+cat(htmltools::includeText("skeleton.Rmd"))
+```
+
+> This will generate a simple flexdashboard with two interactive figures (duplicated to the right)
+
+### Details
+
+This skeleton is a verbatim copy of an R markdown file (extension `.Rmd`) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`.
+Like most R markdown files, it includes three types of content:
+
+1. A YAML header surrounded by `----`
+2. R code chunks surrounded by `` ``` ``
+3. Text following `markdown` formatting
+
+The YAML header includes the metadata for the file, such as the document title and output format:
+
+````
+---
+title: "flexdashboard + plotly + crosstalk"
+output: flexdashboard::flex_dashboard
+---
+````
+
+While only a title and format was specified in the skeleton, many other options are available (e.g. author, date).
+
+The next section is a chunk of R code:
+
+````
+```{r setup, include=FALSE} `r ''`
+library(plotly)
+library(crosstalk)
+lw <- read.csv("length-weight_data.csv")
+shared_lw <- SharedData$new(lw, ~record)
+```
+````
+
+This is where the `plotly` and `crosstalk` packages are loaded along with some sample data.
+The sample data includes length and weight data (columns `length` and `weight`, respectively) along with fitted values and residuals from a length-weight regression (columns called `fit` and `res`, respectively).
+The `SharedData` function from `crosstalk` package is also used here to generate a data object that can be "shared" across independent plots.
+
+Following this chunk of R code is a level 2 markdown header that tells flexdashboard to introduce a column break and place the subsequent components into separate tabs:
+
+````
+Column {.tabset}
+-------------------------------------
+````
+
+Following this break, independent tabs are defined using three hashtags followed by an optional tab name (level 3 markdown header).
+This header can be followed by either markdown text or R code.
+Here, two tabs are generated with `plotly` plots using shared data from `crosstalk`:
+
+````
+### Length-Weight relationship
+
+```{r} `r ''`
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
+    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
+```
+
+### Residuals vs. fitted values
+
+```{r} `r ''`
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~fit, y = ~res)
+```
+
+````
+
+The native syntax of `plotly` was inspired by the Grammar of Graphics and, as such, its general structure will be familiar to those who have used `ggplot2`.
+The package has also been structured to be pipe (`%>%`) and `dplyr` friendly, making the code more intuitive and efficient.
+
+Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), these plots are rendered into independent tabs (shown to the right) and clicking a specific point in one plot will highlight the corresponding point in the other plot.
+In short, `flexdashboard` sets up the structure of the document, `plotly` produces the interactive figures and `crosstalk` connects the plots held in independent tabs.
+
+Of course, this is only a rudimentary overview of what is possible with `flexdashboard`, `plotly` and `crosstalk`.
+A wide range of layout options are possible using `flexdashboard`, `plotly` can produce more than just scatter and line plots and `crosstalk` can connect various widgets.
+See the links provided on the [Background](#background) page for more details on each package.
+The hope here is that this skeleton serves as a starting point from which to build more elaborate dashboards tailored to specific needs.
+
+Column {.tabset}
+-------------------------------------
+
+```{r}
+lw <- read.csv("length-weight_data.csv")
+shared_lw <- SharedData$new(lw, ~record)
+```
+
+### Length-Weight relationship
+
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
+    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
+```
+
+### Residuals vs. fitted values
+
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~fit, y = ~res)
+```

### 2019-10-16 analysis/dashboards/skeleton.Rmd
Draft a 'getting_started' dashboard

--- 2019-10-04 analysis/dashboards/getting_started.Rmd
@@
 ---
-title: "Dashboards for stock assessments: Getting started"
-output: 
-  flexdashboard::flex_dashboard:
-    source_code: embed
+title: "flexdashboard + plotly + crosstalk"
+output: flexdashboard::flex_dashboard
 ---
@@
 ```{r setup, include=FALSE}
-library(flexdashboard)
+library(plotly)
+library(crosstalk)
+lw <- read.csv("length-weight_data.csv")
+shared_lw <- SharedData$new(lw, ~record)
 ```
-
-Background
-=====================================
 
@@
 
-### Concept
+### Length-Weight relationship
 
-- Browser-based dashboards and interactive visualizations are becoming increasingly common and accessible  
-- With a relatively shallow learning-curve, an [**R**](https://www.r-project.org/) user can use the [**flexdashboard**](https://rmarkdown.rstudio.com/flexdashboard/), [**plotly**](https://plotly-r.com) and [**crosstalk**](https://rstudio.github.io/crosstalk/) packages to generate interactive dashboards for exploring data and models
-- The concept is demonstrated here using a simple data-set and linear regression; this can be extended to more complex cases
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
+    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
+```
 
-### flexdashboard
+### Residuals vs. fitted values
 
-- Uses [**rmarkdown**](https://rmarkdown.rstudio.com/) to render a group of related figures, tables and text into a dashboard  
-- Layout is flexible and the components automatically resize to fill the browser and adapt to mobile displays
-- Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com)
-- Optionally use [**shiny**](http://shiny.rstudio.com/) or [**crosstalk**](https://rstudio.github.io/crosstalk/) to bolster interactivity
-
-> Source: https://rmarkdown.rstudio.com/flexdashboard/
-
-### plotly
-
-- A graphing package that works like other R plots expect it produces interactive visualizations
-- The package allows the user to create interactive web graphics from [**ggplot2**](https://ggplot2.tidyverse.org/) graphs
-- Also provides a more 'direct' link to the core [plotly.js](https://plot.ly/javascript/) JavaScript library using syntax inspired by the grammar of graphics
-
-> Source: https://plotly-r.com
-
-### crosstalk
-
-- Enables cross-widget interactions by linking brushing and/or filtering across multiple views
-- i.e. Interactions with one plot can affect change in another plot
-
-> Source: https://rstudio.github.io/crosstalk/
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~fit, y = ~res)
+```

### 2019-10-17 analysis/dashboards/getting_started.Rmd
Minor tweak of the 'getting_started' dashboard

@@
 
-The native syntax of `plotly` was inspired by the Grammar of Graphics and, as such, its general structure will be familiar to those who have used `ggplot2`.
+The native syntax of `plotly` was inspired by the grammar of graphics and, as such, its general structure will be familiar to those who have used `ggplot2`.
 The package has also been structured to be pipe (`%>%`) and `dplyr` friendly, making the code more intuitive and efficient.
+Though the script ends with the residual vs.
+fitted plot, the dashboard can easily be extended to include other diagnostic plots, such as a histogram of the residuals.
 
-Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), these plots are rendered into independent tabs (shown to the right) and clicking a specific point in one plot will highlight the corresponding point in the other plot.
+Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), a stand-alone html document will be produced with the plots rendered into independent tabs (shown to the right).
+Clicking a specific point in one plot will highlight the corresponding point in the other plot.
 In short, `flexdashboard` sets up the structure of the document, `plotly` produces the interactive figures and `crosstalk` connects the plots held in independent tabs.

### 2019-10-17 analysis/dashboards/getting_started.Rmd
Draft an rmarkdown basics page in the getting_started dashboard

@@
 ---
+
+<style>
+h5 {display: inline;}
+</style>
 
@@
 
-- A graphing package that works like other R plots expect it produces interactive visualizations
+- A graphing package that works like other R plots except it produces interactive visualizations
 - The package allows the user to create interactive web graphics from [**ggplot2**](https://ggplot2.tidyverse.org/) graphs
@@
 1. A YAML header surrounded by `----`
-2. R code chunks surrounded by `` ``` ``
-3. Text following `markdown` formatting
+2. Text following `markdown` formatting
+3. R code chunks surrounded by `` ``` ``
 
@@
 ```
+
+rmarkdown {data-navmenu="Basics"}
+=======================================================================
+
+Row {data-height=250}
+-------------------------------------
+
+### R markdown
+
+The package `rmarkdown` enables the writing of `markdown` text and `R` code in the same document (extension `.Rmd`). Most R markdown files includes three types of content:
+
+1. A YAML header surrounded by `----`
+2. Text following `markdown` formatting
+3. R code chunks surrounded by `` ``` ``
+
+Row {data-height=600}
+-------------------------------------
+
+### YAML
+
+The first step in setting up an R markdown file is to define a YAML header. In the [skeleton](#skeleton), it looks like this:
+
+````
+---
+title: "flexdashboard + plotly + crosstalk"
+output: flexdashboard::flex_dashboard
+---
+````
+
+This is where the output type is defined along with other items such as title, author, date, etc.
+A `flex_dashboard` is requested here.
+There are are growing number of output types supported by R markdown (e.g.
+html_document, pdf_document, word_document).
+
+### Markdown
+
+Next, write plain text using markdown syntax to describe how to format the text in the final document.
+
+- `Plain text` simply translates to Plain text in the default font of the output document.
+- Italics and bold are specified using asterisk and/or underscores:
+    - `*italic*` or `_italic_` $\rightarrow$ *italic*
+    - `**bold**` or `__bold__` $\rightarrow$ **bold**
+    - `**_bold-italic_** ` $\rightarrow$ **_bold-italic_** 
+- Equations can be specified using LaTeX syntax:
+    - `$W = \alpha L ^{\beta}$` $\rightarrow$ $W = \alpha L ^{\beta}$
+- Ordered and unordered lists are generated using numbers and dashes or asterisk, respectively:
+    - `1. ordered item 1` $\rightarrow$ 1. &nbsp; ordered item 1
+    - `- unordered item` $\rightarrow$ • &nbsp; unordered item
+- Headers are define using hashtags:
+   - `##### Header 5` $\rightarrow$ <h5> Header 5 </h5>
+   - Note that markdown headers with three hashtags or less have specific formatting functions in `flexdashboard`
+
+### R
+
+Finally, mix in `R` code by surrounding chunks of code using one backtick for inline code (e.g.`` `r knitr::inline_expr("1+1")` `` will print 2 in the output) or three backticks to run several lines of code and/or display a table or plot.
+The below chunk will print the top three rows of the `cars` data-set:
+
+````
+```{r} `r ''`
+head_cars <- head(cars, 3)
+head_cars
+```
+````
+Resulting in the following output:
+
+```{r}
+head_cars <- head(cars, 3)
+head_cars
+```

### 2019-10-18 analysis/dashboards/getting_started.Rmd
Draft a flexdashboard basics page to the getting_started dashboard

@@
 - Layout is flexible and the components automatically re-size to fill the browser and adapt to mobile displays
-- Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com)
+- Supports a wide range of components, including base plot, ggplots, gauges, tables and [htmlwidgets](http://www.htmlwidgets.org/index.html) such as [**plotly**](https://plotly-r.com), [**leaflet**](https://rstudio.github.io/leaflet/) and [**DT**](https://rstudio.github.io/DT/)
 - Optionally use [**shiny**](http://shiny.rstudio.com/) or [**crosstalk**](https://rstudio.github.io/crosstalk/) to bolster interactivity
@@
 - i.e. Interactions with one plot can affect change in another plot  
-- Supports a wide range of [htmlwidgets](http://www.htmlwidgets.org/index.html), such as [**plotly**](https://plotly-r.com) and [**leaflet**](https://rstudio.github.io/leaflet/)
+- Supports a wide range of [htmlwidgets](http://www.htmlwidgets.org/index.html), such as [**plotly**](https://plotly-r.com), [**leaflet**](https://rstudio.github.io/leaflet/) and [**DT**](https://rstudio.github.io/DT/)
 
@@
 Finally, mix in `R` code by surrounding chunks of code using one backtick for inline code (e.g.`` `r knitr::inline_expr("1+1")` `` will print 2 in the output) or three backticks to run several lines of code and/or display a table or plot.
-The below chunk will print the top three rows of the `cars` data-set:
+The chunk below will print the top three rows of the `cars` data-set:
 
@@
 ```
+
+flexdashboard {data-navmenu="Basics"}
+=======================================================================
+
+Row {data-height=200}
+-------------------------------------
+
+### flexdashboard
+
+The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard.
+Using a combination of `markdown` syntax and `R` code (i.e.
+R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on.
+The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard.
+Some layout and component options are highlited below.
+
+Row {data-height=600}
+-------------------------------------
+
+### Layout
+
+Individual charts are defined using markdown's level 3 header (`### Chart title`) and they are, by default, stacked vertically within columns defined using markdown's level 2 header (`---------`).
+The code below will create a `flexdashboard` with two stacked columns and three charts:
+
+````
+---
+title: "Layout example 1"
+output: flexdashboard::flex_dashboard
+---
+
+Column
+-------------------------------------
+
+### Chart 1
+
+Column
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+````
+
+The output from this example has been duplicated [here](#layout-example-1).
+
+Alternatively, charts can be organized by row by modifying the YAML header:
+
+````
+---
+title: "Layout example 2"
+output: 
+  flexdashboard::flex_dashboard:
+    orientation: rows
+---
+
+Row
+-------------------------------------
+
+### Chart 1
+
+Row
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+````
+The output from this example has been duplicated [here](#layout-example-2).
+
+### Components
+
+The components of a dashboard can hold a wide range of outputs.
+The code below generates a mixture of tabular and graphical output.
+The table is generated following `markdown` syntax with inline `R` code and the interactive plot is generated using `plotly`.
+
+````
+---
+title: "Components example"
+output: flexdashboard::flex_dashboard
+---
+
+Column {data-width=300}
+-------------------------------------
+
+### Table
+
+Summary statistics of the `volcano` data-set
+
+| Statistic     | Elevation                   |
+|:------------- |:--------------------------- |
+| Min           | `r knitr::inline_expr('min(volcano)')`            | 
+| Median        | `r knitr::inline_expr('median(volcano)')`         |
+| Mean          | `r knitr::inline_expr('round(mean(volcano))')`    |
+| Max           | `r knitr::inline_expr('max(volcano)')`            |
+
+Column {data-width=700}
+-------------------------------------
+
+### Plot
+
+```{r} `r ''`
+plotly::plot_ly(z = volcano, type = "surface")
+```
+
+````
+
+The output from this example has been duplicated [here](#components-example). Note the use of the `data-width` attribute to make the table chart relatively narrow.
+
+Layout example 1 {data-orientation=columns .hidden}
+=======================================================================
+
+Column
+-------------------------------------
+
+### Chart 1
+
+Column
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+Layout example 2 {data-orientation=rows .hidden}
+=======================================================================
+
+Row
+-------------------------------------
+
+### Chart 1
+
+Row
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+Components example {data-orientation=columns .hidden}
+=======================================================================
+
+Column {data-width=300}
+-------------------------------------
+
+### Table
+
+Summary statistics of the `volcano` data-set
+
+| Statistic     | Elevation                   |
+|:------------- |:--------------------------- |
+| Min           | `r min(volcano)`            |
+| Median        | `r median(volcano)`         |
+| Mean          | `r round(mean(volcano))`    |
+| Max           | `r max(volcano)`            |
+
+Column {data-width=700}
+-------------------------------------
+
+### Plot
+
+```{r}
+plotly::plot_ly(z = volcano, type = "surface")
+```

### 2019-10-18 analysis/dashboards/getting_started.Rmd
Draft a plotly basics page to the getting_started dashboard

@@
 > Source: https://rstudio.github.io/crosstalk/
+
+rmarkdown {data-navmenu="Basics"}
+=======================================================================
+
+Row {data-height=250}
+-------------------------------------
+
+### R markdown
+
+The package `rmarkdown` enables the writing of `markdown` text and `R` code in the same document (extension `.Rmd`). Most R markdown files includes three types of content:
+
+1. A YAML header surrounded by `----`
+2. Text following `markdown` formatting
+3. R code chunks surrounded by `` ``` ``
+
+Row {data-height=600}
+-------------------------------------
+
+### YAML
+
+The first step in setting up an R markdown file is to define a YAML header. In the [skeleton](#skeleton), it looks like this:
+
+````
+---
+title: "flexdashboard + plotly + crosstalk"
+output: flexdashboard::flex_dashboard
+---
+````
+
+This is where the output type is defined along with other items such as title, author, date, etc.
+A `flex_dashboard` is requested here.
+There are are growing number of output types supported by R markdown (e.g.
+html_document, pdf_document, word_document).
+
+### Markdown
+
+Next, write plain text using markdown syntax to describe how to format the text in the final document.
+
+- `Plain text` simply translates to Plain text in the default font of the output document.
+- Italics and bold are specified using asterisk and/or underscores:
+    - `*italic*` or `_italic_` $\rightarrow$ *italic*
+    - `**bold**` or `__bold__` $\rightarrow$ **bold**
+    - `**_bold-italic_** ` $\rightarrow$ **_bold-italic_** 
+- Equations can be specified using LaTeX syntax:
+    - `$W = \alpha L ^{\beta}$` $\rightarrow$ $W = \alpha L ^{\beta}$
+- Ordered and unordered lists are generated using numbers and dashes or asterisk, respectively:
+    - `1. ordered item 1` $\rightarrow$ 1. &nbsp; ordered item 1
+    - `- unordered item` $\rightarrow$ • &nbsp; unordered item
+- Headers are define using hashtags:
+   - `##### Header 5` $\rightarrow$ <h5> Header 5 </h5>
+   - Note that markdown headers with three hashtags or less have specific formatting functions in `flexdashboard`
+
+### R
+
+Finally, mix in `R` code by surrounding chunks of code using one backtick for inline code (e.g.`` `r knitr::inline_expr("1+1")` `` will print 2 in the output) or three backticks to run several lines of code and/or display a table or plot.
+The chunk below will print the top three rows of the `cars` data-set:
+
+````
+```{r} `r ''`
+head_cars <- head(cars, 3)
+head_cars
+```
+````
+Resulting in the following output:
+
+```{r}
+head_cars <- head(cars, 3)
+head_cars
+```
+
+flexdashboard {data-navmenu="Basics"}
+=======================================================================
+
+Row {data-height=200}
+-------------------------------------
+
+### flexdashboard
+
+The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard.
+Using a combination of `markdown` syntax and `R` code (i.e.
+R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on.
+The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard.
+Some layout and component options are highlighted below.
+
+Row {data-height=600}
+-------------------------------------
+
+### Layout
+
+Individual charts are defined using markdown's level 3 header (`### Chart title`) and they are, by default, stacked vertically within columns defined using markdown's level 2 header (`---------`).
+The code below will create a `flexdashboard` with two stacked columns and three charts:
+
+````
+---
+title: "Layout example 1"
+output: flexdashboard::flex_dashboard
+---
+
+Column
+-------------------------------------
+
+### Chart 1
+
+Column
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+````
+
+The output from this example has been duplicated [here](#layout-example-1).
+
+Alternatively, charts can be organized by row by modifying the YAML header:
+
+````
+---
+title: "Layout example 2"
+output: 
+  flexdashboard::flex_dashboard:
+    orientation: rows
+---
+
+Row
+-------------------------------------
+
+### Chart 1
+
+Row
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+````
+The output from this example has been duplicated [here](#layout-example-2).
+
+### Components
+
+The components of a dashboard can hold a wide range of outputs.
+The code below generates a mixture of tabular and graphical output.
+The table is generated following `markdown` syntax with inline `R` code and the interactive plot is generated using `plotly`.
+
+````
+---
+title: "Components example"
+output: flexdashboard::flex_dashboard
+---
+
+Column {data-width=300}
+-------------------------------------
+
+### Table
+
+Summary statistics of the `volcano` data-set
+
+| Statistic     | Elevation                   |
+|:------------- |:--------------------------- |
+| Min           | `r knitr::inline_expr('min(volcano)')`            | 
+| Median        | `r knitr::inline_expr('median(volcano)')`         |
+| Mean          | `r knitr::inline_expr('round(mean(volcano))')`    |
+| Max           | `r knitr::inline_expr('max(volcano)')`            |
+
+Column {data-width=700}
+-------------------------------------
+
+### Plot
+
+```{r} `r ''`
+plotly::plot_ly(z = volcano, type = "surface")
+```
+
+````
+
+The output from this example has been duplicated [here](#components-example). Note the use of the `data-width` attribute to make the table chart relatively narrow.
+
+plotly {data-navmenu="Basics" data-orientation=columns}
+=======================================================================
+
+Column {data-width=250}
+-------------------------------------
+
+### plotly
+
+There are two ways to produce interactive graphics using `plotly`:
+
+1. Using the function `ggplotly` to convert plots from `ggplot` to `plotly` objects
+2. Using the native syntax of the `plotly` package
+
+Both options follow the layered grammar of graphics, which is a generic tool for concisely describing the components of a graphic, such as the:
+
+- Data and mappings (aesthetics)
+- Geometric objects (points, lines, polygons)
+- Scales (color, size, shape, axes)
+
+Column {data-width=400 .tabset}
+-------------------------------------
+
+### ggplotly
+
+Using the `ggplotly` function, a user can construct a plot using `ggplot2` and supply the plot object to `ggplotly` to convert it to an interactive `plotly` graphic:
+
+```{r, echo = TRUE, eval = FALSE}
+p <- ggplot(data = iris, 
+            mapping = aes(x = Sepal.Length, 
+                          y = Petal.Length, 
+                          color = Species)) +
+  geom_point() +
+  scale_color_manual(values = viridis::viridis(3))
+ggplotly(p)
+```
+
+#### Data and mapping
+
+Here, the `ggplot` function is used to set-up the base layer of the plot as well as the aesthetics.
+That is, `iris` data are supplied to the function and specific columns in the `iris` data-set are mapped to `x`, `y` and `color` aesthetics.
+
+#### Geometric objects
+
+Next points are added using the `geom_point` function. Many other geometries can be used to visually represent the data, such as lines (`geom_line`), bars (`geom_bar`), text (`geom_text`).
+
+#### Scales
+
+Finally, colors were modified using `scale_color_manual`.
+While the colors are manually defined here, `ggplot2` includes built in color scales such as `scale_color_grey`.
+Scales used for the `x` and `y` aesthetics can also be modified using functions such as `scale_x_log10`.
+
+### Output
+
+```{r}
+p <- ggplot(data = iris, 
+            mapping = aes(x = Sepal.Length, 
+                          y = Petal.Length, 
+                          color = Species)) +
+  geom_point() +
+  scale_color_manual(values = viridis::viridis(3))
+ggplotly(p)
+```
+
+Column {data-width=400 .tabset}
+-------------------------------------
+
+### plot_ly
+
+Using functions from the `plotly` package, `plotly` plots can be directly created using syntax similar to `ggplot2`:
+
+```{r, echo = TRUE, eval = FALSE}
+plot_ly(data = iris, 
+        x = ~Sepal.Length, 
+        y = ~Petal.Length, 
+        color = ~Species,
+        colors = viridis::viridis(3)) %>% 
+  add_markers()
+```
+
+#### Data and mapping
+
+The `plot_ly` function is analogus to the `ggplot` function from `ggplot2` and, like `ggplot`, this function sets-up the base layer of the plot.
+Again, the `iris` data is supplied and specific columns are mapped to `x`, `y` and `color` aesthetics.
+
+#### Geometric objects
+
+Under `plotly`, geometries are added using functions with the `add_` prefix rather that the `geom_` prefix of used in `ggplot2`.
+Points are added here using `add_markers` and, like `ggplot2`, several other geometries can be specified, such as lines (`add_lines`), bars (`add_bars`), text (`add_text`).
+
+#### Scales
+
+Unlike `ggplot2`, scales are not modified using specific functions with a `scale_` prefix, rather, scales are specified in the `plot_ly` function call itself or via the `layout` function.
+Here, colors are specified using the `colors` argument in the `plot_ly` function.
+
+### Output
+
+```{r}
+plot_ly(data = iris, 
+        x = ~Sepal.Length, 
+        y = ~Petal.Length, 
+        color = ~Species,
+        colors = viridis::viridis(3)) %>% 
+  add_markers()
+```
 
@@
 
-rmarkdown {data-navmenu="Basics"}
-=======================================================================
-
-Row {data-height=250}
--------------------------------------
-
-### R markdown
-
-The package `rmarkdown` enables the writing of `markdown` text and `R` code in the same document (extension `.Rmd`). Most R markdown files includes three types of content:
-
-1. A YAML header surrounded by `----`
-2. Text following `markdown` formatting
-3. R code chunks surrounded by `` ``` ``
-
-Row {data-height=600}
--------------------------------------
-
-### YAML
-
-The first step in setting up an R markdown file is to define a YAML header. In the [skeleton](#skeleton), it looks like this:
-
-````
----
-title: "flexdashboard + plotly + crosstalk"
-output: flexdashboard::flex_dashboard
----
-````
-
-This is where the output type is defined along with other items such as title, author, date, etc.
-A `flex_dashboard` is requested here.
-There are are growing number of output types supported by R markdown (e.g.
-html_document, pdf_document, word_document).
-
-### Markdown
-
-Next, write plain text using markdown syntax to describe how to format the text in the final document.
-
-- `Plain text` simply translates to Plain text in the default font of the output document.
-- Italics and bold are specified using asterisk and/or underscores:
-    - `*italic*` or `_italic_` $\rightarrow$ *italic*
-    - `**bold**` or `__bold__` $\rightarrow$ **bold**
-    - `**_bold-italic_** ` $\rightarrow$ **_bold-italic_** 
-- Equations can be specified using LaTeX syntax:
-    - `$W = \alpha L ^{\beta}$` $\rightarrow$ $W = \alpha L ^{\beta}$
-- Ordered and unordered lists are generated using numbers and dashes or asterisk, respectively:
-    - `1. ordered item 1` $\rightarrow$ 1. &nbsp; ordered item 1
-    - `- unordered item` $\rightarrow$ • &nbsp; unordered item
-- Headers are define using hashtags:
-   - `##### Header 5` $\rightarrow$ <h5> Header 5 </h5>
-   - Note that markdown headers with three hashtags or less have specific formatting functions in `flexdashboard`
-
-### R
-
-Finally, mix in `R` code by surrounding chunks of code using one backtick for inline code (e.g.`` `r knitr::inline_expr("1+1")` `` will print 2 in the output) or three backticks to run several lines of code and/or display a table or plot.
-The chunk below will print the top three rows of the `cars` data-set:
-
-````
-```{r} `r ''`
-head_cars <- head(cars, 3)
-head_cars
-```
-````
-Resulting in the following output:
-
-```{r}
-head_cars <- head(cars, 3)
-head_cars
-```
-
-flexdashboard {data-navmenu="Basics"}
-=======================================================================
-
-Row {data-height=200}
--------------------------------------
-
-### flexdashboard
-
-The `flexdashboard` package can be used to render groups of related text, figures and tables into a dashboard.
-Using a combination of `markdown` syntax and `R` code (i.e.
-R markdown), this package facilitates a wide range of layout options and each component can include output from packages such as `plotly`, `leaflet`, `ggplot2`, and so on.
-The package also integrates nicely with `shiny` and `crosstalk`, providing options for increasing the interactivity of the dashboard.
-Some layout and component options are highlited below.
-
-Row {data-height=600}
--------------------------------------
-
-### Layout
-
-Individual charts are defined using markdown's level 3 header (`### Chart title`) and they are, by default, stacked vertically within columns defined using markdown's level 2 header (`---------`).
-The code below will create a `flexdashboard` with two stacked columns and three charts:
-
-````
----
-title: "Layout example 1"
-output: flexdashboard::flex_dashboard
----
-
-Column
--------------------------------------
-
-### Chart 1
-
-Column
--------------------------------------
-
-### Chart 2
-
-### Chart 3
-
-````
-
-The output from this example has been duplicated [here](#layout-example-1).
-
-Alternatively, charts can be organized by row by modifying the YAML header:
-
-````
----
-title: "Layout example 2"
-output: 
-  flexdashboard::flex_dashboard:
-    orientation: rows
----
-
-Row
--------------------------------------
-
-### Chart 1
-
-Row
--------------------------------------
-
-### Chart 2
-
-### Chart 3
-
-````
-The output from this example has been duplicated [here](#layout-example-2).
-
-### Components
-
-The components of a dashboard can hold a wide range of outputs.
-The code below generates a mixture of tabular and graphical output.
-The table is generated following `markdown` syntax with inline `R` code and the interactive plot is generated using `plotly`.
-
-````
----
-title: "Components example"
-output: flexdashboard::flex_dashboard
----
-
-Column {data-width=300}
--------------------------------------
-
-### Table
-
-Summary statistics of the `volcano` data-set
-
-| Statistic     | Elevation                   |
-|:------------- |:--------------------------- |
-| Min           | `r knitr::inline_expr('min(volcano)')`            | 
-| Median        | `r knitr::inline_expr('median(volcano)')`         |
-| Mean          | `r knitr::inline_expr('round(mean(volcano))')`    |
-| Max           | `r knitr::inline_expr('max(volcano)')`            |
-
-Column {data-width=700}
--------------------------------------
-
-### Plot
-
-```{r} `r ''`
-plotly::plot_ly(z = volcano, type = "surface")
-```
-
-````
-
-The output from this example has been duplicated [here](#components-example). Note the use of the `data-width` attribute to make the table chart relatively narrow.
-
 Layout example 1 {data-orientation=columns .hidden}

### 2019-10-18 analysis/dashboards/getting_started.Rmd
Draft a crosstalk basics page to the getting_started dashboard

@@
 library(plotly)
+library(leaflet)
 library(crosstalk)
@@
 There are are growing number of output types supported by R markdown (e.g.
-html_document, pdf_document, word_document).
+`html_document`, `pdf_document`, `word_document`).
 
@@
 ```{r, echo = TRUE, eval = FALSE}
+library(ggplot2)
 p <- ggplot(data = iris, 
@@
   scale_color_manual(values = viridis::viridis(3))
-ggplotly(p)
+plotly::ggplotly(p)
 ```
@@
 ```{r}
+library(ggplot2)
 p <- ggplot(data = iris, 
@@
 ```{r, echo = TRUE, eval = FALSE}
+library(plotly)
 plot_ly(data = iris, 
@@
 
-The `plot_ly` function is analogus to the `ggplot` function from `ggplot2` and, like `ggplot`, this function sets-up the base layer of the plot.
+The `plot_ly` function is analogous to the `ggplot` function from `ggplot2` and, like `ggplot`, this function sets-up the base layer of the plot.
 Again, the `iris` data is supplied and specific columns are mapped to `x`, `y` and `color` aesthetics.
@@
 ```{r}
+library(plotly)
 plot_ly(data = iris, 
@@
   add_markers()
+```
+
+crosstalk {data-navmenu="Basics" data-orientation=columns}
+=======================================================================
+
+Column
+-------------------------------------
+
+### crosstalk
+
+The `crosstalk` package enables HTML widgets, such as `plotly` and `leaflet`, to communicate with each other without `shiny`.
+As such, self-contained html files with similar behaviour to `shiny` applications can be built using `crosstalk`.
+
+A first step to using `crosstalk` is to set-up a shared data-set using the `SharedData` function:
+
+```{r, echo = TRUE}
+library(crosstalk)
+sub_quakes <- subset(quakes, stations %in% c(10, 20, 50, 100))
+shared_quakes <- SharedData$new(sub_quakes)
+```
+
+The `quakes` data is used for this example.
+As shown to the right using `leaflet` and `plotly`, HTML widgets using this shared data will be linked and selections in one plot will affect the other plot.
+
+The `crosstalk` package also includes a series of functions for filtering the shared data (`filter_checkbox`, `filter_slider` and `filter_select`).
+When used in a `flexdashboard`, and other interactive contexts, the following lines of code will produce the check-box, slider and drop-down filters shown below.
+
+```{r, echo = TRUE, eval = FALSE}
+filter_checkbox("stations", "Stations", shared_quakes, 
+                ~stations, inline = TRUE)
+filter_slider("depth", "Depth", shared_quakes, ~depth)
+filter_select("mag", "Magnitude", shared_quakes, 
+              ~cut(mag, breaks = seq(1, 10, 0.5), right = FALSE))
+```
+
+```{r}
+filter_checkbox("stations", "Stations", shared_quakes, 
+                ~stations, inline = TRUE)
+filter_slider("depth", "Depth", shared_quakes, ~depth)
+filter_select("mag", "Magnitude", shared_quakes, 
+              ~cut(mag, breaks = seq(1, 10, 0.5), right = FALSE))
+```
+
+Column
+--------------------------------------
+
+### leaflet
+
+```{r}
+leaflet(shared_quakes) %>%
+  addTiles() %>%
+  addCircleMarkers(stroke = FALSE,
+                   fillOpacity = 0.5,
+                   color = ~viridis::viridis(4)[factor(stations)])
+```
+
+### plotly
+
+```{r}
+plot_ly(data = shared_quakes) %>% 
+  add_markers(x = ~depth, 
+              y = ~mag, 
+              color = ~factor(stations),
+              colors = viridis::viridis(4),
+              marker = list(size = 15))
 ```
@@
 
-This skeleton is a verbatim copy of an R markdown file (extension `.Rmd`) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`.
+This skeleton is a verbatim copy of an R markdown file (`` `r xfun::embed_file("skeleton.Rmd", text = "skeleton.Rmd")` ``) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`.
 Like most R markdown files, it includes three types of content:
@@
 
-This is where the `plotly` and `crosstalk` packages are loaded along with some sample data.
+This is where the `plotly` and `crosstalk` packages are loaded along with some sample data (`` `r xfun::embed_file("length-weight_data.csv", text = "length-weight_data.csv")` ``).
 The sample data includes length and weight data (columns `length` and `weight`, respectively) along with fitted values and residuals from a length-weight regression (columns called `fit` and `res`, respectively).

### 2019-10-21 analysis/dashboards/getting_started.Rmd
Shift the structure of the getting_started dashboard and add a quick start tab to the skeleton page

@@
     orientation: rows
-    source_code: embed
 ---
@@
 
-rmarkdown {data-navmenu="Basics"}
+TEMP {.hidden}
+=======================================================================
+
+### leaflet
+
+```{r, eval = TRUE, include=FALSE}
+leaflet::leaflet() # subsequent leaflet would not render withouth an eariler call to leaflet for some unknown reason?
+```
+
+Skeleton {data-orientation=columns}
+=======================================================================
+
+Column {.tabset}
+-------------------------------------
+
+### Code
+
+```{r echo = FALSE, comment = ""}
+cat(htmltools::includeText("skeleton.Rmd"))
+```
+
+> This will generate a simple flexdashboard with two interactive figures (duplicated to the right)
+
+### Quick start
+
+**1\.** `r xfun::embed_files(c("skeleton.Rmd", "length-weight_data.csv"), text = "Download code and data")`
+
+**2\.** Open `skeleton.Rmd` in Rstudio
+
+**3\.** Install packages: `install.packages(c("flexdashboard", "plotly", "crosstalk"))`
+
+**4\.** Knit file by typing `Ctrl+Shift+K` or use the `r knitr::include_graphics("knit_button.png", dpi = 175)` button
+
+### Details
+
+This skeleton is a verbatim copy of an R markdown file (`` `r xfun::embed_file("skeleton.Rmd", text = "skeleton.Rmd")` ``) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`.
+Like most R markdown files, it includes three types of content:
+
+1. A YAML header surrounded by `----`
+2. Text following `markdown` formatting
+3. R code chunks surrounded by `` ``` ``
+
+The YAML header includes the metadata for the file, such as the document title and output format:
+
+````
+---
+title: "flexdashboard + plotly + crosstalk"
+output: flexdashboard::flex_dashboard
+---
+````
+
+While only a title and format was specified in the skeleton, many other options are available (e.g. author, date).
+
+The next section is a chunk of R code:
+
+````
+```{r setup, include=FALSE} `r ''`
+library(plotly)
+library(crosstalk)
+lw <- read.csv("length-weight_data.csv")
+shared_lw <- SharedData$new(lw, ~record)
+```
+````
+
+This is where the `plotly` and `crosstalk` packages are loaded along with some sample data (`` `r xfun::embed_file("length-weight_data.csv", text = "length-weight_data.csv")` ``).
+The sample data includes length and weight data (columns `length` and `weight`, respectively) along with fitted values and residuals from a length-weight regression (columns called `fit` and `res`, respectively).
+The `SharedData` function from `crosstalk` package is also used here to generate a data object that can be "shared" across independent plots.
+
+Following this chunk of R code is a level 2 markdown header that tells flexdashboard to introduce a column break and place the subsequent components into separate tabs:
+
+````
+Column {.tabset}
+-------------------------------------
+````
+
+Following this break, independent tabs are defined using three hashtags followed by an optional tab name (level 3 markdown header).
+This header can be followed by either markdown text or R code.
+Here, two tabs are generated with `plotly` plots using shared data from `crosstalk`:
+
+````
+### Length-Weight relationship
+
+```{r} `r ''`
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
+    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
+```
+
+### Residuals vs. fitted values
+
+```{r} `r ''`
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~fit, y = ~res)
+```
+
+````
+
+The native syntax of `plotly` was inspired by the grammar of graphics and, as such, its general structure will be familiar to those who have used `ggplot2`.
+The package has also been structured to be pipe (`%>%`) and `dplyr` friendly, making the code more intuitive and efficient.
+Though the script ends with the residual vs.
+fitted plot, the dashboard can easily be extended to include other diagnostic plots, such as a histogram of the residuals.
+
+Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), a stand-alone html document will be produced with the plots rendered into independent tabs (shown to the right).
+Clicking a specific point in one plot will highlight the corresponding point in the other plot.
+In short, `flexdashboard` sets up the structure of the document, `plotly` produces the interactive figures and `crosstalk` connects the plots held in independent tabs.
+
+Of course, this is only a rudimentary overview of what is possible with `flexdashboard`, `plotly` and `crosstalk`.
+A wide range of layout options are possible using `flexdashboard`, `plotly` can produce more than just scatter and line plots and `crosstalk` can connect various widgets.
+See the links provided on the [Background](#background) page for more details on each package.
+The hope here is that this skeleton serves as a starting point from which to build more elaborate dashboards tailored to specific needs.
+
+Column {.tabset}
+-------------------------------------
+
+```{r}
+lw <- read.csv("length-weight_data.csv")
+shared_lw <- SharedData$new(lw, ~record)
+```
+
+### Length-Weight relationship
+
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
+    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
+```
+
+### Residuals vs. fitted values
+
+```{r}
+plot_ly(data = shared_lw) %>% 
+    add_markers(x = ~fit, y = ~res)
+```
+
+rmarkdown {data-navmenu="Extra"}
 =======================================================================
@@
 
-flexdashboard {data-navmenu="Basics"}
+flexdashboard {data-navmenu="Extra"}
 =======================================================================
@@
 
-plotly {data-navmenu="Basics" data-orientation=columns}
+plotly {data-navmenu="Extra" data-orientation=columns}
 =======================================================================
@@
 
-crosstalk {data-navmenu="Basics" data-orientation=columns}
+crosstalk {data-navmenu="Extra" data-orientation=columns}
 =======================================================================
@@
 
-Skeleton {data-orientation=columns}
-=======================================================================
-
-Column {.tabset}
--------------------------------------
-
-### Code
-
-```{r echo = FALSE, comment = ""}
-cat(htmltools::includeText("skeleton.Rmd"))
-```
-
-> This will generate a simple flexdashboard with two interactive figures (duplicated to the right)
-
-### Details
-
-This skeleton is a verbatim copy of an R markdown file (`` `r xfun::embed_file("skeleton.Rmd", text = "skeleton.Rmd")` ``) set-up to produce a simple `flexdashboard` with interactive `plotly` visuals connected by `crosstalk`.
-Like most R markdown files, it includes three types of content:
-
-1. A YAML header surrounded by `----`
-2. Text following `markdown` formatting
-3. R code chunks surrounded by `` ``` ``
-
-The YAML header includes the metadata for the file, such as the document title and output format:
-
-````
----
-title: "flexdashboard + plotly + crosstalk"
-output: flexdashboard::flex_dashboard
----
-````
-
-While only a title and format was specified in the skeleton, many other options are available (e.g. author, date).
-
-The next section is a chunk of R code:
-
-````
-```{r setup, include=FALSE} `r ''`
-library(plotly)
-library(crosstalk)
-lw <- read.csv("length-weight_data.csv")
-shared_lw <- SharedData$new(lw, ~record)
-```
-````
-
-This is where the `plotly` and `crosstalk` packages are loaded along with some sample data (`` `r xfun::embed_file("length-weight_data.csv", text = "length-weight_data.csv")` ``).
-The sample data includes length and weight data (columns `length` and `weight`, respectively) along with fitted values and residuals from a length-weight regression (columns called `fit` and `res`, respectively).
-The `SharedData` function from `crosstalk` package is also used here to generate a data object that can be "shared" across independent plots.
-
-Following this chunk of R code is a level 2 markdown header that tells flexdashboard to introduce a column break and place the subsequent components into separate tabs:
-
-````
-Column {.tabset}
--------------------------------------
-````
-
-Following this break, independent tabs are defined using three hashtags followed by an optional tab name (level 3 markdown header).
-This header can be followed by either markdown text or R code.
-Here, two tabs are generated with `plotly` plots using shared data from `crosstalk`:
-
-````
-### Length-Weight relationship
-
-```{r} `r ''`
-plot_ly(data = shared_lw) %>% 
-    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
-    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
-```
-
-### Residuals vs. fitted values
-
-```{r} `r ''`
-plot_ly(data = shared_lw) %>% 
-    add_markers(x = ~fit, y = ~res)
-```
-
-````
-
-The native syntax of `plotly` was inspired by the grammar of graphics and, as such, its general structure will be familiar to those who have used `ggplot2`.
-The package has also been structured to be pipe (`%>%`) and `dplyr` friendly, making the code more intuitive and efficient.
-Though the script ends with the residual vs.
-fitted plot, the dashboard can easily be extended to include other diagnostic plots, such as a histogram of the residuals.
-
-Once this script is "Knit" (`Ctrl+Shift+K` in Rstudio), a stand-alone html document will be produced with the plots rendered into independent tabs (shown to the right).
-Clicking a specific point in one plot will highlight the corresponding point in the other plot.
-In short, `flexdashboard` sets up the structure of the document, `plotly` produces the interactive figures and `crosstalk` connects the plots held in independent tabs.
-
-Of course, this is only a rudimentary overview of what is possible with `flexdashboard`, `plotly` and `crosstalk`.
-A wide range of layout options are possible using `flexdashboard`, `plotly` can produce more than just scatter and line plots and `crosstalk` can connect various widgets.
-See the links provided on the [Background](#background) page for more details on each package.
-The hope here is that this skeleton serves as a starting point from which to build more elaborate dashboards tailored to specific needs.
-
-Column {.tabset}
--------------------------------------
-
-```{r}
-lw <- read.csv("length-weight_data.csv")
-shared_lw <- SharedData$new(lw, ~record)
-```
-
-### Length-Weight relationship
-
-```{r}
-plot_ly(data = shared_lw) %>% 
-    add_markers(x = ~length, y = ~weight, name = "Observed") %>% 
-    add_lines(x = ~length, y = ~exp(fit), name = "Predicted")
-```
-
-### Residuals vs. fitted values
-
-```{r}
-plot_ly(data = shared_lw) %>% 
-    add_markers(x = ~fit, y = ~res)
-```
-
 Layout example 1 {data-orientation=columns .hidden}

### 2019-10-23 analysis/paper/cover_letter.Rmd
Draft a new cover letter

@@
 
+Howard I. Browman, Ph.D.  
+Editor-in-Chief  
+ICES Journal of Marine Science  
+Institute of Marine Research  
+Marine Ecosystem Acoustics Group  
+Austevoll Research Station  
+5392 Storebø  
+Norway
+
 `r format(Sys.time(), '%Y-%m-%d')`
 
-Professor Paul J B Hart  
-Fish and Fisheries  
-Department of Biology  
-University of Leicester  
-University Road  
-Leicester  
-LE1 7RH, UK
+Dr. Browman,
 
-Dr. Hart,
+I am writing to submit our manuscript entitled, **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** for consideration as a **_Quo Vadimus_** article in **ICES Journal of Marine Science**.
+In our manuscript, we discuss the merits of applying modern interactive visualization tools to stock assessment processes and we hope to convince readers that these tools, which we use every day on a range of web sites, are very accessible to practicing biologists.
+We argue that the application of these tools alleviates some of the difficulty in presenting the mass of information produced by complex monitoring programs and statistical models to end-users in a meaningful, interactive and engaging way.
+To exemplify this point, we describe the structure of three interactive dashboards developed to aid the exploration and communication of results from 1) a fishery-independent trawl survey, 2) a mark-recapture tagging program, and 3) a stock assessment model.
+In each case we discuss the efficacy of applying such tools in a stock assessment context and contrast this approach with more traditional modes of communicating the results from complex monitoring programs and stock assessment models.
 
-We thank you for taking the time to examine our initial idea for our paper "Improving the communication and accessibility of stock assessment using interactive visualization tools".
-Based on your feedback, we have assembled a paper we feel fits the theme of Ghoti papers, and points towards ways of improving the stock assessment process.
-We re-iterate that this paper is not technical in nature, but mostly to encourage readers to pursue investigating these interactive visualization tools, and to convince readers that these tools are very much accessible to practicing biologists.
+We believe that this is a fitting and timely article for the Journal's **_Quo Vadimus_** type as there are a growing number of scientists from a wide range of disciplines applying these tools to explore and communicate their results.
+We can personally attest that the application of interactive tools has helped us in our daily work and we hope that our perspective will appeal to readers of **ICES Journal of Marine Science** and encourage the application of interactive tools.
 
-I look forward to your decision.
+We look forward to your decision.
 

### 2019-10-31 analysis/presentation/interactive_tools.Rmd
Start drafting a seminar presentation on this paper

---
title: "Interactive Tools for Science Advice"
author: "Paul M. Regular, Gregory J. Robertson, Robert Rogers, Keith P. Lewis"
date: "Ocean and Ecosystem Science Seminar Series | BIO | 2019-11-06"
output:
  ioslides_presentation:
    incremental: yes
    css: style.css
    logo: dfo_logo.png
    widescreen: yes
bibliography: references.bib
csl: csas.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.height = 5, 
                      fig.width = 10)
knitr::opts_knit$set(root.dir = "../..")

```

```{r data}
## xaringan::inf_mr()
library(NCAM)
library(plotly)
load("data/NCAM_2018.RData")
fit <- ncam_2018
tabs <- tidy_model(fit)
ages <- fit$tmb.data$ages
years <- fit$tmb.data$years
years_plus1 <- fit$tmb.data$years_plus1
```

```{r bg-fig, eval = FALSE}

d <- trans_est(tabs$Cproj_log_ssb_lrp)
d <- d[d$year != max(years_plus1), ] # drop the first year (it's in the time series)
d$type <- "Projected"
d$scenario <- factor(d$scenario)

## Append full time series to projections
past_d <- trans_est(tabs$log_ssb_lrp)
past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
past_d <- do.call(rbind, past_d)
past_d$type <- ifelse(past_d$year == max(years_plus1), "Projected", "Estimated")
d <- rbind(past_d, d)
xlim <- c(max(years) - 2.1, max(years) + 4.1)
ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])

d <- d[d$scenario == 1, ]

xax <- yax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)
xax$range <- c(1995, 2025)
yax$range <- c(0, 2)

p <- plot_ly(x = ~year, hoverinfo = "none")

interval <- seq(0.9, 0, length.out = 7)
val <- 1 - ((1 - interval) / 2)
col <- RColorBrewer::brewer.pal(length(interval), "Reds") # viridis::viridis(length(interval))
alpha <- seq(0.1, 1, lenth.out = length(interval))

for (i in seq_along(val)) {
  d$upr <- exp(log(d$est) + qnorm(val[i]) * d$sd)
  d$lwr <- exp(log(d$est) - qnorm(val[i]) * d$sd)
  p <- p %>% 
    add_ribbons(ymin = ~upr, ymax = ~lwr, data = d,
                line = list(width = 0),
                fillcolor = toRGB(col[i]))
}

p <- p %>% 
  hide_guides() %>% 
  layout(xaxis = xax, yaxis = yax, 
         plot_bgcolor = "transparent", 
         paper_bgcolor = "transparent",
         margin = list(t = 0, l = 0, r = 0, b = 0, pad = 0)) %>% 
  plotly::config(displayModeBar = FALSE)

orca(p, file = "analysis/presentation/figures/bg-fig.png", 
     height = 700, width = 1100, scale = 6)

```

## Background {data-background="figures/bg-fig.png"}

*The greatest value of a picture is when it forces us to notice*   
*what we never expected to see.*  
&nbsp;&nbsp;&nbsp;- Tukey (1977)

- Discovery is one of the most exciting parts of science
- Visulizations are one of the tools we use to reveal patterns
- Models allow us to simplify and describe these patterns
- At various stages, we need to communicate our findings

### 2019-11-01 analysis/presentation/interactive_tools.Rmd
Draft seminar slides up to the objectives

@@
 ## xaringan::inf_mr()
+
 library(NCAM)
 library(plotly)
+library(ggplot2)
+library(emojifont)
+
+htmltools::tagList(rmarkdown::html_dependency_font_awesome())
+
 load("data/NCAM_2018.RData")
@@
 - Discovery is one of the most exciting parts of science
-- Visulizations are one of the tools we use to reveal patterns
+- Visualizations are one of the tools we use to reveal patterns
 - Models allow us to simplify and describe these patterns
 - At various stages, we need to communicate our findings
+
+## Communicating information
+
+Simplified workflow:
+
+<hr>
+
+<div style="float: left; width: 20%; text-align: center;">
+Data  
+<font size="20"><i class="fa fa-database"></i></font>
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+Models  
+<font size="20"><i class="fa fa-chart-area"></i></font>
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+Discuss  
+<font size="20"><i class="fa fa-comments"></i></font>
+</div>
+
+<hr>
+
+<div>
+- Connections may not come easy
+    - Growing volume and variety of data
+    - Increasingly complex models
+    - Diverse backgrounds
+
+</div>
+
+## Stock assessments <i class="fa fa-fish"></i>
+
+<div style="float: left; width: 65%;">
+
+- Communicating stock assessment results isn't getting easier  
+    - Growing volume and variate of data  
+    - Increasingly complex models  
+    - Diverse backgrounds  
+- Traditional formats sometimes 
+    - Overwhelm participants with endless tables and figures
+    - Fail to convey the richness of information available
+
+</div>
+
+<div style="float: right; width: 35%;">
+
+```{r, out.width = "100%"}
+knitr::include_graphics("figures/doc_screenshots.PNG")
+```
+
+</div>
+
+## Interactive tools
+
+<div style="float: left; width: 55%;">
+
+- Simplify communication and improve accessibility
+- Common tools that are used daily on a range of websites
+- Their use is no longer restricted to website developers
+    - Integrated into software commonly used by the research community
+    - Surprisingly easy to generate interactive dashboards
+    - Growing number of users
+
+</div>
+
+<div style="float: right; width: 40%; padding-left: 5%">
+
+```{r, out.width = "100%"}
+knitr::include_graphics("figures/canada_poll_tracker.PNG")
+```
+
+</div>
+
+## Objective
+
+- Here we aim to demonstrate that interactive tools simplify our workflow, from data and model exploration to communication

### 2019-11-03 analysis/presentation/interactive_tools.Rmd
Draft seminar presentation up to tagging data section

@@
 ---
+
+<style>
+@import url('https://fonts.googleapis.com/css?family=Montserrat');
+</style>
 
@@
 ```{r, out.width = "100%"}
-knitr::include_graphics("figures/canada_poll_tracker.PNG")
+# knitr::include_graphics("figures/canada_poll_tracker.PNG")
+knitr::include_url("https://newsinteractives.cbc.ca/elections/poll-tracker/canada/", height = "300px")
 ```
@@
 
-- Here we aim to demonstrate that interactive tools simplify our workflow, from data and model exploration to communication
+- Using examples we aim to demonstrate that interactive tools can simplify our workflow, from data and model exploration to communication
+<br>
+- Examples:
+    1. **RStrap Explorer:** A tool for examinging fishery-independent survey data
+    2. **Tag Explorer:** A mapping tool developed for the exploration of a long-term tagging study
+    3. **NCAM Explorer:** A tool for exploring results from the Northern Cod Assessment Model
+
+## RStrap Explorer
+
+<div style="float: left; width: 50%;">
+
+- Fisheries-independent surveys have become a mainstay in stock assessment
+- Stratified-random surveys have been conducted by DFO for > 40 years
+- Typically analyzed using stratified analyses (RStrap in NL region)
+
+</div>
+
+<div style="float: right; width: 45%; padding-left: 5%">
+
+<br>
+<img src="figures/teleost_twitter_crop.jpg" width="100%"/>
+<font size="1"> 
+<br>
+&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
+</font>
+
+</div>
+
+## RStrap Explorer {.smaller}
+
+> - Running a series of stratified analysis can be cumbersome
+
+```{r, eval = FALSE, echo = TRUE}
+
+library(Rstrap)
+
+## Load survey data
+load("converted_set_details_2017-04-27.Rdata")
+load("converted_length-frequency_data_2017-04-27.Rdata")
+load("age-growth_data_2017-04-27.Rdata")
+
+## 3NO Witch Flounder
+witch <- strat.fun(setdet = setdet, data.series = "Campelen", 
+                   species = 890, survey.year = 1995:2012, 
+                   season = "spring", NAFOdiv = c("3N", "3O"))
+
+## 3K Roughhead Grenadier (length disaggregated)
+grenadier <- strat.fun(setdet = setdet, lf = lf, data.series = "Campelen", 
+                       program = "strat2 & strat1", species = 474, survey.year = 2012, 
+                       season = "fall", NAFOdiv = "3K", strat = 622:654, sex = "unsexed", 
+                       length.group = 0.5, group.by = "length")
+
+## 2J3KL Greenland Halibut (length and age disaggregated)
+turbot <- strat.fun(setdet = setdet, lf = lf, ag = ag, data.series = "Campelen", 
+                    program = "strat2 & strat1", species = 892, survey.year = 1998:2012, 
+                    season = "fall", NAFOdiv = c("2J", "3K", "3L"), 
+                    sex = c("male","female","unsexed"), length.group = 1, 
+                    group.by = "length & age")
+
+## 3Ps Atlantic Cod (length and age disaggregated)
+cod <- strat.fun(setdet = setdet, lf = lf, ag = ag, data.series = "Campelen", 
+                 program = "strat2 & strat1", species = 438, survey.year = 1998:2012, 
+                 season = "spring", NAFOdiv = "3P", sex = c("male","female","unsexed"), 
+                 length.group = 1, group.by = "length & age")
+
+```
+
+- This doesn't include code for plotting these results
+
+## RStrap Explorer
+
+- A dashboard was created using the **`shiny`**, **`flexdashboard`**, **`plotly`**, and **`crosstalk`** packages to dynamically run and explore RStrap results
+- Four primary pages:
+    1. **Survey Indices** contains stock level estimates of biomass and abundance
+    2. **Age & Length Distributions** contains length and age frequency plots 
+    3. **Recruitment** displays recruitment indices
+    4. **Help** provides additional context to the survey data and analysis
+
+## RStrap Explorer {data-background="figures/RStrap_dashboard_biomass.png" data-background-position="center bottom"}
+
+## RStrap Explorer {data-background="figures/RStrap_dashboard_freq.png" data-background-position="center bottom"}
+
+## Tag Explorer
+
+<div style="float: left; width: 55%;">
+
+- For some data-rich stocks, mark and recapture studies are carried out to estimate movement, growth rate, natural mortality, etc.
+- There is a long-standing tagging program for Northern cod
+    - \>60 years, >60,000 records, 1,000 - 10,000 tags deployed annually
+
+</div>
+
+<div style="float: right; width: 40%; padding-left: 5%">
+
+<img src="figures/cod_tagging.jpg" width="100%"/>
+<font size="1"> 
+Photo: John Brattey 
+</font>
+
+</div>
+
+## Tag Explorer
+
+> - Spatial explorations of these data have been limited since the seminal work by Taggart et al. 1995
+
+<div style="float: left; width: 25.6%; padding: 1%">
+<div class="drop-shadow" style="background-color: rgb(236, 192, 107);">
+<img src="figures/taggart_1.png" width="100%"/>
+</div>
+</div>
+
+<div style="float: left; width: 26.4%; padding: 1%">
+
+<div class="drop-shadow">
+<img src="figures/taggart_2.png" width="100%"/>
+</div>
+
+</div>
+
+<div style="float: left; width: 42%; padding: 1%">
+
+<div class="drop-shadow">
+<img src="figures/taggart_3.png" width="100%"/>
+</div>
+
+</div>
+
+- **Nearly 500 pages!**
+
+## Tag Explorer
+
+- A dashboard was created using the **`shiny`**, **`shinydashboard`**, and **`leaflet`** packages to dynamically explore this large dataset
+- Contains
+    - One primary page, **Main map**, where release and recovery locations are mapped
+    - A series of summary pages where these data are summarised
+
+## Tag Explorer {data-background="figures/tagging_dashboard_map_tab.png" data-background-position="center bottom"}
+
+## Tag Explorer {data-background="figures/tagging_dashboard_summary_tab.png" data-background-position="center bottom"}

### 2019-11-04 analysis/presentation/interactive_tools.Rmd
Draft seminar up to NCAM dashboard section

@@
 
-orca(p, file = "analysis/presentation/figures/bg-fig.png", 
+orca(p, file = "analysis/presentation/graphics/bg-fig.png", 
      height = 700, width = 1100, scale = 6)
@@
 
-## Background {data-background="figures/bg-fig.png"}
+## Background {data-background="graphics/bg-fig.png"}
 
@@
 ```{r, out.width = "100%"}
-knitr::include_graphics("figures/doc_screenshots.PNG")
+knitr::include_graphics("graphics/doc_screenshots.PNG")
 ```
@@
 ```{r, out.width = "100%"}
-# knitr::include_graphics("figures/canada_poll_tracker.PNG")
+# knitr::include_graphics("graphics/canada_poll_tracker.PNG")
 knitr::include_url("https://newsinteractives.cbc.ca/elections/poll-tracker/canada/", height = "300px")
@@
 - Examples:
-    1. **RStrap Explorer:** A tool for examinging fishery-independent survey data
+    1. **RStrap Explorer:** A tool for examining fishery-independent survey data
     2. **Tag Explorer:** A mapping tool developed for the exploration of a long-term tagging study
@@
 <br>
-<img src="figures/teleost_twitter_crop.jpg" width="100%"/>
+<img src="graphics/teleost_twitter_crop.jpg" width="100%"/>
 <font size="1"> 
@@
 
-## RStrap Explorer {data-background="figures/RStrap_dashboard_biomass.png" data-background-position="center bottom"}
+## RStrap Explorer {data-background="graphics/RStrap_dashboard_biomass.png" data-background-position="center bottom"}
 
-## RStrap Explorer {data-background="figures/RStrap_dashboard_freq.png" data-background-position="center bottom"}
+## RStrap Explorer {data-background="graphics/RStrap_dashboard_freq.png" data-background-position="center bottom"}
 
@@
 
-<img src="figures/cod_tagging.jpg" width="100%"/>
+<img src="graphics/cod_tagging.jpg" width="100%"/>
 <font size="1"> 
@@
 <div class="drop-shadow" style="background-color: rgb(236, 192, 107);">
-<img src="figures/taggart_1.png" width="100%"/>
+<img src="graphics/taggart_1.png" width="100%"/>
 </div>
@@
 <div class="drop-shadow">
-<img src="figures/taggart_2.png" width="100%"/>
+<img src="graphics/taggart_2.png" width="100%"/>
 </div>
@@
 <div class="drop-shadow">
-<img src="figures/taggart_3.png" width="100%"/>
+<img src="graphics/taggart_3.png" width="100%"/>
 </div>
@@
 
-- A dashboard was created using the **`shiny`**, **`shinydashboard`**, and **`leaflet`** packages to dynamically explore this large dataset
+- A dashboard was created using the **`shiny`**, **`shinydashboard`**, and **`leaflet`** packages to dynamically explore this large data-set
 - Contains
     - One primary page, **Main map**, where release and recovery locations are mapped
-    - A series of summary pages where these data are summarised
+    - A series of summary pages where these data are summarized
 
-## Tag Explorer {data-background="figures/tagging_dashboard_map_tab.png" data-background-position="center bottom"}
+## Tag Explorer {data-background="graphics/tagging_dashboard_map_tab.png" data-background-position="center bottom"}
 
-## Tag Explorer {data-background="figures/tagging_dashboard_summary_tab.png" data-background-position="center bottom"}
+## Tag Explorer {data-background="graphics/tagging_dashboard_summary_tab.png" data-background-position="center bottom"}
+
+## NCAM Explorer
+
+<div style="float: left; width: 50%; padding-right: 5%;">
+
+- Integrated stock assessment models are becoming more common
+- Advances in computational power and methods $\rightarrow$ possibilities
+- Requires a solid understanding of the inputs, methods and outputs
+    - Understanding is a prerequisite for communication
+
+</div>
+
+<div class="drop-shadow" style="float: right; width: 40%; padding: 2.5%; font-size: 8px;">
+
+**Process equations**  
+<br>
+Stochastic cohort model with a plus group to model the unobserved states:
+$$\log(N_{a,y}) = \left\{\begin{matrix}
+ \log(N_{a-1,y-1}) - Z_{a-1,y-1} + \delta_{a,y}, & a < A \\
+ \log\{N_{a-1,y-1}\exp(-Z_{a-1,y-1}) + N_{a,y-1}\exp(-Z_{a,y-1})\} + \delta_{a,y}, & a = A.
+\end{matrix}\right. $$
+
+The ages are 1-10+ and years are 1975-2015. $Z_{a,y} = F_{a,y} + M_{a,y}$, where $M_{a,y} = 0.2$ is the base case assumption.
+
+Recruitments, $N_{1,1}, ... N_{1,Y}$, are treated as uncorrelated lognormal random variables
+$$\log(N_{1,y}) \overset{iid}\sim N(r, \sigma_{r}^2).$$
+
+Catches are modeled using the Baranov catch equation,
+$$C_{a,y} = N_{a,y}\{1 - \exp(-Z_{a,y})\}F_{a,y}/Z_{a,y}.$$
+
+Fishing moralities are modeled as a stochastic process, with
+$$Cov\{\log(F_{a,y}),\log(F_{a-j,y-k})\} = \frac{\sigma_{F}^2 \varphi_{F,a}^j \varphi_{F,y}^k}{(1-\varphi_{F,a}^2)(1-\varphi_{F,y}^2)}.$$
+
+**Observation equations**  
+<br>
+The model predicted catch for survey $s$ is
+$$\log(I_{s,a,y}) = \log(q_{s,a}) + \log(N_{a,y}) - t_{s,y}Z_{a,y} + \varepsilon_{s,a,y}, ~~ \varepsilon_{s,a,y} \overset{iid}\sim N(0, \sigma_{s,G(a)}^2).$$
+
+Survey variance was split out and self-weighted by age groups 1-3, 4-7, and 8-10+
+
+Total catch and age compositions were treated separately. Total catch was modeled as lognormal, 
+$$\log(C_{obs,y}) = \log(C_{y}) + \varepsilon_{C,y}, ~~ \varepsilon_{C,y} \overset{iid}\sim N(0, \sigma_{C}^2)$$
+Age compositions were modeled as multiplicative logistic normal with a censored component for zero's
+
+</div>
+
+## NCAM Explorer <img src="graphics/cod.png" height="100px" style="margin-bottom: -35px; margin-left: 10px;">
+
+- **Challenge:** the data-rich case of Northern cod
+    - Assessed using an age-structured, state-space assessment model, called NCAM, which integrates:
+        - Research vessel autumn trawl surveys (1983-present)
+        - Sentinel fishery surveys (1995-present)
+        - Inshore acoustic surveys (1995-2009)
+        - Fishery catch-at-age compositions and partial fishery landings (1983-present)
+        - Tagging data (1983-present)
+
+## NCAM Explorer
+
+> - **Communication:** standard approach involves compiling and presenting static documentation
+
+<hr>
+
+<div style="float: left; width: 20%; text-align: center;">
+Code  
+<font size="20"><i class="fa fa-file-code"></i></font>
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+Plots & Tables  
+<font size="20"><i class="fa fa-file-image"></i></font>
+<font size="20"><i class="fa fa-file-excel"></i></font>
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 20%; text-align: center;">
+Documentation  
+<font size="20"><i class="fa fa-file-word"></i></font>
+<font size="20"><i class="fa fa-file-powerpoint"></i></font>
+</div>
+
+<hr>
+
+- Involves a lot of copy-pasting
+- Important details may not be visible
+- Linear format may be the best way to explore ideas
+
+## NCAM Explorer
+
+> - A dashboard was created using the **`flexdashboard`**, **`plotly`**, and **`crosstalk`** packages to explore and communicate results from NCAM
+
+<hr>
+
+<div style="float: left; width: 33.3333%; text-align: center;">
+Code  
+<font size="20"><i class="fa fa-file-code"></i></font>
+</div>
+
+<div style="float: left; width: 33.3333%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 33.3333%; text-align: center;">
+HTML  
+<font size="20"><i class="far fa-file-code"></i></font>
+</div>
+
+<hr>
+
+- Contains a series of pages with plots and tables typically presented at assessment meetings
+
+## NCAM Explorer
+
+<iframe src="graphics/NCAM_dashboard.html" style="height: 700px; width: 1100px; margin-top: -150px; margin-left: -60px; margin-right: -60px; position: absolute; z-index: 1;"></iframe>

### 2019-11-04 analysis/presentation/graphics/make_NCAM_dashboard.R

@@ -1,0 +1,29 @@
+
+## load NCAM model inputs
+## (object is named inputs)
+load("data/inputs.RData")
+
+## load NCAM model output from the last two years
+## (object names match file names)
+load("data/ncam_2018.RData")
+load("data/ncam_2019.RData")
+
+## source plotting functions
+## (these functions are used in the Rmarkdown file that makes the dashboard)
+source("R/plot.R")
+
+## source some data management functions
+## (these functions are used to "tidy" the model output)
+source("R/tidy.R")
+
+## source function for making the dashboard
+## (this function uses rmarkdown::run or rmarkdown::render)
+source("R/vis_fit.R")
+
+## Run the vis_fit function to make the dashboard
+## (last - model output from the last assessment)
+vis_fit(ncam_2019, 
+        last = ncam_2018,                     
+        output_file = "analysis/presentation/graphics/NCAM_dashboard.html")
+
+

### 2019-11-04 analysis/presentation/interactive_tools.Rmd
Complete draft of seminar

@@
 <iframe src="graphics/NCAM_dashboard.html" style="height: 700px; width: 1100px; margin-top: -150px; margin-left: -60px; margin-right: -60px; position: absolute; z-index: 1;"></iframe>
+
+## Interactive vs. static
+
+- Static formats are useful for explaining ideas
+- Interactive formats are useful for exploring ideas
+    - Easy and efficient access to the details via drop-down filters and data-rich illustrations
+    - Navigation bar circumvents scrolling through pages or digging through files
+    - **Bonus:** automation circumvents copy-pasting
+
+## Towards open stock-assessment
+
+<hr>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Data  
+<font size="20"><i class="fa fa-database"></i></font>
+</div>
+
+<div style="float: left; width: 5%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Code  
+<font size="20"><i class="fa fa-code"></i></font>
+</div>
+
+<div style="float: left; width: 5%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Models  
+<font size="20"><i class="fa fa-chart-area"></i></font>
+</div>
+
+<div style="float: left; width: 5%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Code  
+<font size="20"><i class="fa fa-code"></i></font>
+</div>
+
+<div style="float: left; width: 5%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Document  
+<font size="20"><i class="fa fa-file"></i></font>
+</div>
+
+<div style="float: left; width: 5%; text-align: center;">
+<br>
+$\rightarrow$
+</div>
+
+<div style="float: left; width: 12.5%; text-align: center;">
+Discuss  
+<font size="20"><i class="fa fa-comments"></i></font>
+</div>
+
+<hr>
+
+- Although these tools do not reveal the entire "data pipeline", transparency is improved
+- Accessible to a broad audience $\rightarrow$ interdisciplinary input 
+- Greater access $\rightarrow$ richer peer-review
+- Richer peer-review $\rightarrow$ improved and widely supported science advice
+
+## Conclusion
+
+- There is a cost to developing interactive tools
+- However, the learning curve is relatively shallow
+   - The tools highlighted here were developed by ecologists with R experience
+- Benefits > cost
+    - Efficient delivery of products
+    - Richer explorations and discussions
+    - Collective understanding
+
+## Acknowledgements
+
+We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
+
+# Thank you!

### 2019-11-06 analysis/dashboards/getting_started.Rmd
Minor tweaks to getting started dashboard following workshop

@@
 lw <- read.csv("length-weight_data.csv")
-shared_lw <- SharedData$new(lw, ~record)
+shared_lw <- SharedData$new(lw)
 ```
@@
 The output from this example has been duplicated [here](#layout-example-2).
+
+Also note that multiple pages can be generated using markdown's level 1 header:
+
+````
+---
+title: "Layout example 1"
+output: flexdashboard::flex_dashboard
+---
+
+Page 1
+====================================
+
+Column
+-------------------------------------
+
+### Chart 1
+
+Page 2
+====================================
+
+Column
+-------------------------------------
+
+### Chart 2
+
+### Chart 3
+
+````
 

### 2019-11-30 analysis/paper/cover_letter.Rmd
Prep material for submission to CJFAS

@@
 
-Howard I. Browman, Ph.D.  
-Editor-in-Chief  
-ICES Journal of Marine Science  
-Institute of Marine Research  
-Marine Ecosystem Acoustics Group  
-Austevoll Research Station  
-5392 Storebø  
-Norway
+Editorial office  
+Canadian Journal of Fisheries and Aquatic Sciences  
+Canadian Science Publishing (NRC Research Press)  
+65 Auriga Drive, Suite 203   
+Ottawa, ON K2E 7W6  
+Canada
 
@@
 
-Dr. Browman,
+Dear Editors,
 
-I am writing to submit our manuscript entitled, **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** for consideration as a **_Quo Vadimus_** article in **ICES Journal of Marine Science**.
+I am writing to submit our manuscript entitled, **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** for consideration as a **_Perspectives_** article in **Canadian Journal of Fisheries and Aquatic Sciences**.
 In our manuscript, we discuss the merits of applying modern interactive visualization tools to stock assessment processes and we hope to convince readers that these tools, which we use every day on a range of web sites, are very accessible to practicing biologists.
@@
 
-We believe that this is a fitting and timely article for the Journal's **_Quo Vadimus_** type as there are a growing number of scientists from a wide range of disciplines applying these tools to explore and communicate their results.
-We can personally attest that the application of interactive tools has helped us in our daily work and we hope that our perspective will appeal to readers of **ICES Journal of Marine Science** and encourage the application of interactive tools.
+We believe that this is a fitting and timely article for the Journal's **_Perspectives_** type as there are a growing number of scientists from a wide range of disciplines applying these tools to explore and communicate their results.
+We can personally attest that the application of interactive tools has helped us in our daily work and we hope that our perspective will appeal to readers of **Canadian Journal of Fisheries and Aquatic Sciences** and encourage the application of interactive tools.
 

### 2019-11-30 analysis/paper/manuscript.Rmd
Prep material for submission to CJFAS

@@
 bibliography: references.bib
-csl: ices.csl
+csl: cjfas.csl
 abstract: |
@@
 The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
-While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating the results to colleagues, stakeholders and fisheries managers in a meaningful way has become more daunting.
+While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating results from multiple monitoring programs and integrated assessment models to colleagues, stakeholders and fisheries managers in a meaningful way has become more daunting.
 The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked.
@@
 These tools allow a broader audience to conduct detailed explorations of the results, leading to a deeper and collective understanding of both the data and models used to inform stock assessments.
-As a consequence, the peer review process is more open and accessible and the resulting science advice is improved and widely supported.
+As a consequence, the results of the peer review process are more open and accessible.
 
@@
 As such, stock assessment biologists need to manage large volumes of data from a variety of sources.
-Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
+Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in relative abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
@@
 The model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
-Using a series of observation equations, this TMB based model reduces thousands of historical data points into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
+Using a series of observation equations, this TMB based model simplifies thousands of historical data points into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
 Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
@@
 Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
-The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents, however, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides: 1) interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
-both broader patterns and finer details in the data are accessible via zooming and tooltips); and, 2) the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides.
+The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents.
+However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.
+First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
+both broader patterns and finer details in the data are accessible via zooming and tooltips).
+Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g.
+slide shows can easily be made using R markdown).
 Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.
@@
 For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model.
-These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussion helped the assessment biologists, managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
+These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
+Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieve concerns that something is hidden.
 
@@
 Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches.
+There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools.
 We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents.
 It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
-Finally, scientific exploration is rarely a linear process, and the ability to navigate around a dashboard fits the natural way people pursue ideas.
 
@@
 We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
-We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and two anonymous reviewers.
+We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and four anonymous reviewers.
 

### 2019-11-30 analysis/paper/CJFAS_submission/make_NCAM_dashboard.R

@@ -1,0 +1,32 @@
+
+## load NCAM model inputs
+## (object is named inputs)
+load("data/inputs.RData")
+
+## load NCAM model output from the last two years
+## (object names match file names)
+load("data/ncam_2018.RData")
+load("data/ncam_2019.RData")
+
+## source plotting functions
+## (these functions are used in the Rmarkdown file that makes the dashboard)
+source("R/plot.R")
+
+## source some data management functions
+## (these functions are used to "tidy" the model output)
+source("R/tidy.R")
+
+## source function for making the dashboard
+## (this function uses rmarkdown::run or rmarkdown::render)
+source("R/vis_fit.R")
+
+## Run the vis_fit function to make the dashboard
+## (comp - named list of model output to compare)
+## (last - model output from the last assessment)
+vis_fit(ncam_2019, 
+        last = ncam_2018,                     
+        comp = list("NCAM 2018" = ncam_2018, 
+                    "NCAM 2019" = ncam_2019),
+        output_file = "analysis/dashboards/NCAM_dashboard.html")
+
+

### 2019-12-03 analysis/paper/manuscript.Rmd
Minor tweaks for submission to CJFAS

@@
 abstract: |
-  **Running title**: Interactive stock assessment
-
-  **Conflict of Interest Statement**: The authors have no conflicts of interest to declare.
-
-  **Data Availability Statement**: Data sharing is not applicable to this article as no data were generated and no analysis to draw scientific inference was conducted in this study.
 
@@
 The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
-While this growth in data and model complexity has improved the assessments of many stocks, the process of communicating results from multiple monitoring programs and integrated assessment models to colleagues, stakeholders and fisheries managers in a meaningful way has become more daunting.
+While this growth in data and model complexity has improved many stock assessments, the process of communicating key results to colleagues and stakeholders in a meaningful way has become more daunting.
 The traditional approach of presenting information across a series of static slides often fails to convey the richness of information available and, as such, important patterns and details are easily overlooked.
@@
 These tools allow a broader audience to conduct detailed explorations of the results, leading to a deeper and collective understanding of both the data and models used to inform stock assessments.
-As a consequence, the results of the peer review process are more open and accessible.
-
-  **Keywords**: dashboard, data exploration, dynamic documents, integrated analysis, model validation, open stock assessment
+As a consequence, the peer review process is more open and accessible.
 

### 2020-02-29 analysis/paper/manuscript.Rmd
Augment introduction following latest review

@@
 Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013].
-For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way.
+While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way.
 Traditional formats (e.g.
@@
 @jones2016; @yeatman2018; @letcher2018].
-This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).
 
-Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs.
-We especially highlight the use of interactive visuals inside dynamic documents called "dashboards".
+Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science.
+For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and test custom parameterizations of a Von Bertalanffy Growth model [@catalan2018; see Appendix A for more examples].
+While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts.
+These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.
+
+We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
+Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model.
+This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
+In both cases interactive visuals are nested inside dynamic documents called "dashboards".
 The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
-We first demonstrate the application of these tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
-Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dashboards can be used to explore, diagnose and communicate results from an integrated assessment model.
-This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
-In each case, the dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).
+These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).
 
@@
 We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
-We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and four anonymous reviewers.
+We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar, Daniel Goethel and seven anonymous reviewers.
 
@@
 
-- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
-- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
-- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
-- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
-- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo) from the Fisheries Center, University of British Columbia
-- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/) from Fisheries and Oceans Canada, Maritimes Region
-- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
+- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/); a benthic-focused tool to assist with marine license application and compliance monitoring.
+- Species Dashboard (https://shiny.marine.ie/speciesdash/); a tool for exploring biological fisheries data collected in Irish waters
+- The Stock Book (https://shiny.marine.ie/stockbook/); a tool for accessing up-to-date information on the status of fish stocks in Irish waters.
+- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
+- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.
+- EcoCast Explorer (https://coastwatch.pfeg.noaa.gov/ecocast/explorer.html); an application for displaying results from a dynamic ocean management tool focused on bycatch avoidance [see @hazen2018 for more detail].
+- B-VBGM!
+Bayesian Von Bertalanffy Growth Model tool (https://fishecology.shinyapps.io/B-VBGM/); a supplement Catalán et al.
+[-@catalan2018] that facilitates the exploration of the model described in the paper.
+- Beyond Temperature tool (https://heatherwelch.shinyapps.io/beyond_temperature/); a supplement to McHenry et al.
+[-@mchenry2019] that allows the exploration of habitat suitability hindcasts and projections by species and model type.
+- BRSApp (https://catalinagomez.shinyapps.io/BRSApp/); a supplement to Gomez et al. [-@gomez2016] that allows the exploration of behavioural response severity scores of marine mammals to noise.
 - VISA tool (e.g.
-https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"
+https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice (developed to meet an "EU request on dissemination of ICES advice beyond pdf files").
 

### 2020-03-02 analysis/paper/manuscript.Rmd
Add a few extra lines to the conclusions and tweak the intro

@@
 Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science.
-For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and test custom parameterizations of a Von Bertalanffy Growth model [@catalan2018; see Appendix A for more examples].
+For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples).
 While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts.
@@
 These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).
+See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.
 
@@
 
-We must acknowledge there are costs to adopting the use of interactive tools.
+Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information.
+However, we must acknowledge there are costs to adopting the use of interactive tools.
 The first and obvious one being that staff are required to have both the time and training to effectively develop these tools.
-In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but all had their formal training in field-based population ecology.
+Ideally, professional programmers and graphic designers would be hired to construct dashboards and other interactive info-graphics [@mcinerny2014], but fisheries management organizations rarely have the resources to fund such initiatives.
+Consequently, the process of communicating fisheries science tends to be the responsibility of fisheries scientists who lack formal training in these fields.
+In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology.
 While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g.
@@
 - Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
+- Toy Tuna MSE application (https://puntapps.shinyapps.io/tunamse/); a toy MSE example built to allow users to enter various management procedures and assess their performance.
 - Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.

### 2020-03-02 analysis/paper/manuscript.Rmd
Clarify the intent/audience of the NCAM dashboard

@@
 This sub-tab shows a time-series of biomass data with 95% confidence intervals, with a break in the lines when the survey gear changed from an Engle to Campelen trawl, for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889).
-All trend lines can be toggled off using plotly’s dynamic user interface.
-The user also has the ability to hover over the points to compare estimates and export the figure as a png file.'}
+All trend lines can be toggled off using plotly’s dynamic user interface.'}
 knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
@@
 
-In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019].
+In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019].
 We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document.
@@
 Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
+The intent of this tool was not to replace detailed documentation on the model, however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.
 

### 2020-03-03 analysis/paper/manuscript.Rmd
Add a paragraph to the open stock assessment section

@@
 
+Pragmatic issues aside, it is not hard to picture a future whereby the full data to document pipeline is open and accessible.
+Open resources are actively being developed and used to improve accessibility to all the data, methods and results that feed into stock assessments [e.g.
+the Transparent Assessment Framework, https://taf.ices.dk/app/about; @magnusson2020].
+Likewise, tools are being developed to facilitate the creation of reproducible advisory documents [e.g.
+@ices2018; @anderson2020; @regular2020].
+These stand-alone advisory documents can be interactive, as is the case for the VISA tool [@ices2018] or, as discussed here, they can be supplemented by interactive tools.
+The technology behind these initiatives is rapidly developing and, at the same time, society is becoming increasingly reliant on screens.
+We may therefore reach a point where interactive documents become commonplace through increasingly open stock assessment processes.
+
 # Conclusions
@@
 - VISA tool (e.g.
-https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice (developed to meet an "EU request on dissemination of ICES advice beyond pdf files").
+https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice [developed to meet an "EU request on dissemination of ICES advice beyond pdf files"; @ices2018].
 

### 2020-03-03 analysis/paper/manuscript.Rmd
Change placement of new paragraph in the 'Towards open stock assessment' section

@@
 
-# Towards open stock-assessment
+# Towards open stock assessment
 
@@
 
-The tools presented here are additions to the growing number of interactive stock assessment tools being developed and applied across the globe (see Appendix A).
-We believe that such tools represent an important a step forward in terms of open science and reproducible research in addition to being a major step forward in how stock assessments are presented and critiqued.
-Although these tools do not reveal the entire data pipeline, in conjunction with posting code and data on-line, they represent a significant advance over more traditional methods.
-Similarly, these tools do not fully alleviate pragmatic issues associated with reproducible research.
+A culture of openness is actively evolving in the field of stock assessment and, at this point, it is not hard to picture a future where it is commonplace for the the full data to document pipeline to be open and accessible.
+Open resources are actively being developed and used to improve accessibility to all the data, methods and results that feed into stock assessments [e.g.
+the Transparent Assessment Framework, https://taf.ices.dk/app/about; @magnusson2020].
+Likewise, tools are being developed to facilitate the creation of reproducible advisory documents [e.g.
+@ices2018; @anderson2020; @regular2020], and these documents can be interactive, as is the case for the VISA tool [@ices2018], or they can be supplemented by interactive tools.
+There is plenty of opportunity for synergies between these initiatives, as most are being developed using R, and all details and decisions can be traceable and accessible using an online repository such as GitHub.
+We therefore foresee a future where stock assessments become increasingly open and reproducible.
+
+Providing open access to code and data, however, does not necessarily make a stock assessment more accessible.
+We believe that interactive tools, such as those presented here and in Appendix A, represent an important a step forward in terms of open science as they make the peer review process more efficient.
+Although these tools do not reveal the entire data pipeline, a wide range of peers can easily access and review details that may be inaccessible even if the pipeline is open.
 For example, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
@@
 Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
-
-Pragmatic issues aside, it is not hard to picture a future whereby the full data to document pipeline is open and accessible.
-Open resources are actively being developed and used to improve accessibility to all the data, methods and results that feed into stock assessments [e.g.
-the Transparent Assessment Framework, https://taf.ices.dk/app/about; @magnusson2020].
-Likewise, tools are being developed to facilitate the creation of reproducible advisory documents [e.g.
-@ices2018; @anderson2020; @regular2020].
-These stand-alone advisory documents can be interactive, as is the case for the VISA tool [@ices2018] or, as discussed here, they can be supplemented by interactive tools.
-The technology behind these initiatives is rapidly developing and, at the same time, society is becoming increasingly reliant on screens.
-We may therefore reach a point where interactive documents become commonplace through increasingly open stock assessment processes.
 

### 2020-03-03 analysis/paper/manuscript.Rmd
Add link to GitHub repo to paper

@@
 While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way.
-Traditional formats (e.g.
-slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
+Visualisation play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g.
+slides) are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
 A solution to this problem is the effective use of interactive visualization tools [@keena2016].
@@
 Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.
+
 The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file.
@@
 ```
+
+# Generic vs. bespoke
+
+The tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations.
 
@@
 In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology.
-While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g.
+While some learning time was needed, the learning curve was relatively minor because there was no need to learn a new programming language (e.g.
 JavaScript, CSS, etc.).
-Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches.
+We can confirm that an R user can construct a basic dashboard within a day, using Supplement 1 as a guide, and a more custom dashboard in a matter of weeks.
+Moreover, a growing user base means that most programming issues can be readily addressed through simple internet searches.
 There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools.
+Once over the initial learning curve, the time investment may not be much greater than one would invest into a standard presentation.
 We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents.
@@
 We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar, Daniel Goethel and seven anonymous reviewers.
+
+All code used to generate this paper, Supplement 1 ("Getting Started" dashboard) and Supplement 2 ("NCAM explorer" dashboard) are available at https://github.com/PaulRegular/interactive-stock-assessment.
 

### 2020-03-04 analysis/paper/manuscript.Rmd
Add new section to the discussion on 'bespoke to general tools'

@@
 While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way.
-Visualisation play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g.
+Visualizations play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g.
 slides) are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
@@
 
-# Generic vs. bespoke
+# From bespoke to general tools
 
-The tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations.
+The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations.
+Consequently, there exists the possibility of generalizing these tools to simplify their application to other regions.
+Region specific data management and modelling approaches, however, will surely complicate matters as tools built for one region may be difficult to apply in another region.
+Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications to be applied to another model being developed.
+Nevertheless, considerable chunks of code could be recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard.
+This effort was one step forward in improving the generality of the tool and another step, which is actively being worked on by a colleague in IFREMER, is to modify the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014].
+Though more time is required to make the dashboard code more general as the developer must ensure the code is robust to a wider range of scenarios, substantial value could be added by making it more widely applicable.
+Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g.
+StoX software; @johnsen2019].
+These are, of course, early days in the development and application of interactive tools to stock assessment.
+Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards.
+Regardless of the approach, it is important that the path taken is transparent and reproducible.
 
@@
 
-All code used to generate this paper, Supplement 1 ("Getting Started" dashboard) and Supplement 2 ("NCAM explorer" dashboard) are available at https://github.com/PaulRegular/interactive-stock-assessment.
+All code used to generate this paper, Supplement 1 (getting started dashboard) and Supplement 2 (NCAM explorer) are available at https://github.com/PaulRegular/interactive-stock-assessment.
 

### 2020-03-05 analysis/paper/manuscript.Rmd
Draft responses up to minor comments on NCAM explorer

@@
 For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples).
-While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts.
+While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore data and models in ways not amenable to static formats and, in the process, these heuristic improve accessibility and augment communication efforts.
 These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.
@@
 Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
-The intent of this tool was not to replace detailed documentation on the model, however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.
+The intent of this tool was not to replace detailed documentation on the model [i.e.
+@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.
 
@@
 Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.
+For instance, explorations of the residual plots in the dashboard allowed the authors to quickly identify conflicts between two surveys used in the model.
+Potential solutions to this problem were then explored by modifying the configuration of the model and these changes were quickly evaluated using a dashboard that is produced in seconds.
 
@@
 Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards.
-Regardless of the approach, it is important that the path taken is transparent and reproducible.
+Regardless of the approach, one should strive to make the path taken transparent and reproducible.
 

### 2020-03-06 analysis/paper/manuscript.Rmd
First cut at responses to reviewer comments

@@
 Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018].
-These data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981].
+These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981].
 Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
@@
 The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs.
-This data base has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
+This database has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
 The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.

### 2020-03-06 analysis/paper/response_letter.Rmd
Draft response letter and manuscript sent to co-authors

---
output:
  word_document: default
  md_document: default
---

Daniel Goethel, PhD  
Southeast Fisheries Science Center  
National Marine Fisheries Service  
National Oceanic and Atmospheric Administration  
75 Virginia Beach Drive  
Miami, FL 33133  
USA

`r format(Sys.time(), '%Y-%m-%d')`

Dear Dr. Goethel,

Thank you for considering another revision of manuscript cjfas-2019-0424,  **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** by Paul M. Regular, Gregory J. Robertson, Robert Rogers and Keith P. Lewis. We are very grateful for the constructive and detailed reviews provided and we have made every effort to do justice to the reviewer's suggestions. In the process, we believe our manuscript has been greatly improved. Most notably we have expanded our introduction and discussion to include a more through exploration of the literature, emphasizing advances in fisheries science. We have also clarified the intent and audience for the tools highlighted in our manuscript and these changes should curtail impressions that these tools are supposed to be stand-alone documents. These changes should also alleviate impressions that interactive tools serve as a complete replacement for static documents when, in actual fact, they are most often used to complement static documentation. Please see below for more details on the changes we made in response to the reviews.

We submit this revised manuscript for your consideration and look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067

----------------------------------------------------------------------------------------------------

Associate Editor:

Dr. Regular,

Thank you for your submission to the Canadian Journal of Fisheries and Aquatic Sciences. The submitted manuscript has now undergone extensive and thorough review by three independent reviewers. Each of the reviewers have indicated that the manuscript addresses a timely topic worthy of publication in CJFAS. However, a number of concerns were raised regarding overlooked issues related to communication and visualization. After reviewing the article myself, I agree with the reviewers that this perspectives article addresses an important subject that would be of interest to the broad readership of CJFAS, but it is slightly limited in its review of the existing literature on data visualization in fisheries applications and does not tackle some critical problems with utilizing interactive visualization tools. I have summarized the reviewers’ comments along with my own below. Given the concerns raised, I have suggested that the article would be suitable for future publication after major revision and further referee assignment.

Although this is a perspective piece and I appreciated the succinctness of the manuscript, I think that a more thorough exploration of the literature on data visualization and stakeholder communication is warranted. As reviewer 1 points out, there have been a number of papers on this subject in recent years including Miller et al. (2019), which was part of a larger MSE special issue with a strong focus on stakeholder engagement (https://www.nrcresearchpress.com/toc/cjfas-mse/01/01). Other articles discussing communication in MSEs (e.g., Punt et al., 2017; https://academic.oup.com/icesjms/article/74/2/499/2907904) have also highlighted the importance of innovative and interactive visualization techniques. Similarly, the marine policy and socioeconomic literature has many examples of how data visualization can improve stakeholder interactions within participatory modeling initiatives (e.g., Levontin et al. 2017; https://www.sciencedirect.com/science/article/pii/S0308597X1630793X?via%3Dihub). I would suggest exploring the literature a bit more in depth and using these to expand the introduction and discussion.

*We thank you for alerting us to these papers and have incorporated them where appropriate.*

I think the introduction could include further discussion (with citations) of why improved data visualization techniques are warranted and how they can improve stakeholder engagement.

*We now reference Levontin et al. 2017 in our introduction and we have added a general paragraph to the that highlights the diversity of existing interactive applications in fisheries research. With these changes, we have tried maintain the succinctness and generality of the introduction while adding more discussion of the fisheries research literature.*

Similarly, the discussion could be expanded to better highlight some of the issues with developing these type of graphic tools. For instance, as noted by reviewer 1, scientists are not necessarily the best equipped to develop visualizations due to lack of expertise and time constraints. Ideally, graphic designers could be consulted to help design visuals that take into account the needs of all stakeholders, but the cost of hiring graphic designers may be prohibitive in some circumstances.

*We agree, and we have addressed this oversight in the _Conclusion_ section by noting that professional programmers and graphic designers would, ideally, be hired to construct interactive tools, however, it is rare for RFMOs to fund such initiatives. Most often, the task of communicating stock assessment results is left up to scientists without formal training in programming or graphic design. In our case, for instance, we would happily consult graphic design experts if we had funding to do so; alas, this is not the case and improvements must come from self-critical judgement, based on what we have gleaned from books on data visualization (e.g. The Visual Display of Quantitative Information by Edward Tufte, Data Visualization by Kieran Healy), along with constructive feedback from peers (e.g. Reviewer 1). In reality, all scientists have to communicate their research to the best of their ability, and the point of the paper is that interactive tools can be created by scientists without formal training in programming or graphic design. To use an analogy, consulting a chef will improve ones cooking, but one does not need to be a chef to make a meal. We have modified our conclusion to bolster this point.*

*Regarding time constraints, we have expanded our discussion on the learning curve and emphasize that an R user can start producing basic dashboards within a day using Supplement 1 as a guide. We also contend that, once over the learning curve, the time investment may not be significantly different than the investment one would make constructing a static presentation.*

Reviewer 1 notes that some of the language may need to be toned down and further supported by the literature. Interactive visualization tools are clearly useful and helpful for presenting the results of complex modeling approaches. However, further implementation is probably needed before any categorical statements can be made along the lines of transparency and accessibility of assessment models. Overall, I would also like to see more discussion of similar approaches and how they have been received by stakeholders. You provide a table in the appendix, but these were not really discussed in-depth. Using the case studies to support your points was very helpful and gave the manuscript a good flow, but I think highlighting a few other examples (spanning a variety of geographic areas and management agencies) within each sub-section (along with any difficulties encountered) would solidify the paper.

*We have gone through our manuscript and we have tempered several conclusions. Some remain as our perspective is based on extensive verbal feedback at meetings from stakeholders and following tutorials the primary author has provided on this topic. In the "Towards open stock assessment" section, we also clarify why we think these interactive tools improve accessibility.*

*We acknowledge that a formal survey would be more scientific and it would offer insights into ways of improving the dashboards, however, such a formal approach was outside the scope of this work. Likewise, a jurisdictional survey to learn about and speak to experiences outside our region would be an excellent addition to the paper, however, this too was outside the scope of this work. If we implemented such surveys we would likely be submitting this paper as a research article with stronger categorical statements.*

Similarly, there are a number of discussion points raised by the reviewers that should be incorporated into the manuscript. For instance, reviewer 3 suggests discussion on how to incorporate graphics that compare results across multiple model runs, how to streamline versioning and code updates, how to ensure compatibility among graphics code and assessment outputs as either are updated or reparametrized, and what the benefits/difficulties are of developing generic versus study-specific visualization tools.

*These are important points, several of which the first author coincidentally discussed with colleagues while this paper was in review. We have added a new section, "From bespoke to general tools", where we discuss the challenges and potential benefits of developing generic tools. We also acknowledge that the tools can be modified to visualize multiple model runs. In the following section, "Towards open stock assessment", we note that versioning can be traced using GitHub. The only item we did not cover was ensuring compatibility; this is where automated testing via TravisCI and the testthat package are sure to help, however, we did not add this to the discussion because 1) these are programming tools the authors have limited experience with, and 2) it is potentially superfluous to the main message of the paper.*

I also agree that the link between interactive and static reports or presentations is unclear. It would seem that there needs to be a mix of approaches, but this is not really discussed. Similarly, what tips would the authors give for presenting results using interactive graphics? One element that is often useful with static presentations is the building of a ‘story or progress line’ (especially in model building or continuity exercises), which might be difficult to replicate if only using interactive graphics. It does not seem like an all or nothing approach is warranted, but that is the impression that I received from the article. I would suggest carefully reworking the discussion of how interactive graphics can be used when presenting to stakeholders and how static and interactive methods can be used synergistically.

*This is a fair point and, in hindsight, we see why the paper leaves this impression. To try and rectify this, we have clarified the intent of interactive tools in several places in the paper. We now explicitly note in the introduction that these tools are not meant to replace static modes of communication, rather they are often intended to supplement standard documents. This point is reiterated in the  "Towards open stock assessment" section, where we also highlight an example from Appendix A to make the point that advisory documents can be interactive, and these tend to be linear in structure. As with the previous iteration, we contend that a non-linear format can be useful for evaluating a stock assessment model. That being said, we take your point that a story may be more suitable in some cases. This format is possible with interactive documents as well because dashboards can be nested within html slide-shows, or dashboards can include "story-boards". These are very flexible tools and in this iteration of the manuscript we hope we have clarified their flexibility without discounting static formats. In short, we want to make the point that interactive tools are a potentially powerful tool to add to a biologists’ toolkit.*

Finally, all of the reviewers have provided useful suggestions for improving the graphics, especially those utilized within the manuscript. Please carefully consider these and revise as warranted.

*We have made every effort to improve the graphics as suggested and while not compromising the readability and utility of the dashboard tool.  Please see responses to individual Reviewers below.*

Similarly, please provide access to the source code (r code) via an online database (e.g., GitHub) as requested by reviewer 1.

*We completely agree with the Reviewer.  The data and code used to produce the manuscript, supplement 1, and supplement 2 are available on GitHub.*

If the authors are able to revise the manuscript to adequately address the reviewers’ concerns, then it should be suitable for future publication.

Thank you,  
Daniel Goethel

Comments to the Author: See attached.

Reviewer: 1

Comments to the Author

Summary

This paper states a hard to disagree with proposition that fisheries science can benefit from data visualisation tools that are widely available and applied in other fields. It correctly situates data visualisation within the ambition to move stock assessment towards a model of reproducible research that is now commonly taught in beginner data science courses. The authors describe various efforts they have made in developing visualisation tools, such as interactive dashboards, and contextualise these as steps towards stock assessments meeting the standards of reproducible research.

The paper is clearly written, and is important for its promotion of timely developments within fisheries science.

*We thank the Reviewer for the kind comments.*

Major problems

The main shortcoming of the paper is that the authors don’t have sufficient relevant expertise to discuss matters related to design and lack skills necessary to implement effective visualisations. Yet this is the crux of the paper and visualisations are offered as the main results or innovations being presented. Typography, the use of colour and text, the layout, aesthetics, ergonomic organisation of information and the use of interactivity in the tools presented – all fall below professional standards, preventing these tools from being effective at communication. This is potentially damaging to the author’s cause. What the authors need is to work with a graphic designer and address these shortcomings before the paper can be published. The authors’ aim is to enable ‘a broader audience to conduct detailed exploration of the results’ so that they can arrive at a ‘deeper and collective understanding’ of stock assessments – programming and science knowledge are insufficient for this task, the authors need to work alongside experts on visual communication.

*Please see the response to these comments in the AE section above.*

Another issue, is accessibility to the code which is essential for reproducible research. Only one of the tools was available for review as a website, no source code was provided along with the paper and it seems that the code is not intended for release alongside the paper, e.g. via a link to GitHub. The publication of this paper should be made contingent on the availability of the code. The data provided along with the code can be simulated if there are privacy concerns.

*Please see the response to these comments in the AE section above.*

There is no discussion of data conflicts, or data weighting issues, that plague many integrated stock assessments.

*While we have not gone into a detailed discussion of data conflicts, as that has been covered quite nicely by Maunder and Piner (Maunder, M.N. and Piner, K.R., 2017. Dealing with data conflicts in statistical inference of population assessment models that integrate information from multiple diverse data sets. Fisheries Research, 192, pp.16-27), we have noted that NCAM explorer helped us efficiently identify conflicts between two of the surveys included in the model. Data weighting issues, however, are not mentioned in the context of this paper as NCAM is self-weighting.*

Ability to examine a variety of graphs in pairs would be an important feature enabling the user to examine contribution of individual data sources and detect data conflicts of model misspecifications. Features to examine convergence, boundary behaviour, comparisons with simpler models or external data might be good to consider as additional functionality.

*We agree with these points. These tools are a work in progress and we continue to add functionality to them. We have now added a new section ("From bespoke to general tools") where we discuss some of our experiences with making these tools more general. We acknowledge that these are early days and much work has yet to be done; still, we are hopeful that we have included enough to spur discussions among the broader community and catalyze the development of more features.*

Minor problems

The authors need to be more tentative in some of their conclusions, such as ‘peer review process is more open and accessible’ as a result. Further testing with more and varied stakeholders is necessary to demonstrate that these tools are indeed achieving the aims to make stock assessments more reproducible, easier to critique by a wider range of stakeholders, and as a result some systemic changes in the processes of performing stock assessments are occurring.

*Please see previous comments to the AE.*

The authors are missing some important references on using visualisation tools in fisheries, e.g. Miller. S.K., Anganuzzi, A., Butterworth, D.S., Davies, C.R., Donovan, G.P., Nickson, A., Rademeyer, R.A., Restrepo, V. 2018. Improving communication: the key to more effective MSE processes. Canadian Journal of Fisheries and Aquatic Sciences, https://doi.org/10.1139/cjfas-2018-0134.

*We thank the Reviewer for alerting us to this paper and have added where appropriate.*

Comments on the NCAM tool It’s great to see more interactive visualisation tools being developed for fisheries science. I greatly enjoyed playing with the tool. Please find below some comments and suggestions for how to improve it, but I would strongly recommend getting professional help from a designer. Please upload the full code to a GitHub directory and provide a link, use simulated data if necessary to protect confidentiality of any of the data that is not publicly available.

*We thank the Reviewer for the kinds words.  Please see the response to these comments in the AE section above.*

The tool should be made easily transferrable to other case studies otherwise the contribution made by the paper is diminished.

*The code for the tools are available for other researchers and stock assessment biologists to adapt to their own work flows.  Further, the tutorial that is provided should ensure that biologists can learn this approach in a short period of time.  Finally, these tools are just examples of a more general approach – they are not meant to be the only possible solution.*

*In the comments below, the Reviewer makes a number of good suggestions about the layout of the dashboard.  We will deal with these individually. *

*However, we argue that there may be a misunderstanding that we hope to rectify.  Many of the below suggestions seem to be based on the premise that the dashboard is a stand alone piece of work like a paper or book when, in reality, it is better to think of it as a PowerPoint replacement. There are DFO research papers, stock assessment reports, and primary publications that support the dashboard. Reviewers, stakeholders and participants generally have a working knowledge of this documentation and, as such, excessive texts and guides are not required. Moreover, excessive texts may be a distraction as the presenter actively walks participants through the latest model results (i.e. wordy PowerPoints are generally discouraged). Therefore, although these suggestions are greatly appreciated, not all of them are relevant and we have tried to make this clear in the text by stating that the tool was "built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model."  We also state the tool was not built to replace detailed documentation on the model. We refer to this paragraph as the "stand alone comment" below.*

Background tab:

-NCAM tab needs more information about the tool overall and a better image.

*See "stand alone" comment with regards to the more information suggestion.*

*As for a better picture, we agree that it would be nice to have, however, for the time being, we have opted to focus on changes that improve the functionality of the tool.*

-General tab similarly should have more text and an overall guide with images and links to the rest of the tool

*See "stand alone" comment above.*

Data inputs tab:

-Should have information about the data, how it is collected and processed with links to raw data and a code for processing if any was used

*See "stand alone" comment above.*

-Something about data quality, uncertainties, and potential conflicts in data should be mentioned here

*See "stand alone" comment above.*

Biological inputs tab:

-Confusing graphics perhaps because of a lack of explanatory text – this is a problem for most of the pages, two dimensional graphics should be added for clarity.

*See "stand alone" comment above.*

Partial catches:

-Need explanation as well as context, the tool overall should be able to tell a story by itself but it feels like a somewhat chaotic assembly of cut out figures from a stock assessment report.

*See "stand alone" comment above.*

Change in Q:

-Similarly you have an entire page to explain the context related to catchability, but there is insufficient information to be able to critically engage with the visualisation

*See "stand alone" comment above.*

Tagging:

-Likewise, there is a whole page to explain the issues of using tagging data in stock assessment – it seems this would be beneficial

*See "stand alone" comment above.*

References:

-Should be more references and they should be linked to the documents.

*These are the key references for the accepted assessment model.*

Catch tab:

-Everywhere in the tool there is insufficient labelling and cryptic legends, for example it is not clear what is meant by ‘bounds’ or what is represented by the shaded area -the choices of colour and line types are questionable and inconsistent throughout the tool -there are two shaded areas in the bottom graph (absolute landings) without a key

*TODO: Paul*

On the right side:

-there should be consistency in how observations and estimation are signified, going from a dashed line to solid circles adds unnecessary confusion -the graphs can have better titles such as ‘Age 2’ rather than ‘2’, for instance -other graphs, such as ‘fits by age’ are unwieldly and confusing, this, like a majority of graphs, does not represent a good use of data visualisation principles and betrays the authors’ lack of experience with graphic design. When there is a lot of data in one graph, making sure that interactivity works well is paramount, currently selecting a specific line is not so clear. Use different colour scheme.

*TODO: Paul*

Overall there are too many graphs, offered in such a way as it is difficult to prioritise information. The design can be far more ergonomic, with graphs organised more helpfully and with a sense of hierarchy and a narrative, encouraging to explore key results first, while offering extra levels of details as a choice to those who want to pursue a particular matter further. There as an inconsistent and suboptimal from a design point of view use of colour. A data visualisation professional should be able to help with making a good use of colour to encode information and to help the user navigate through results without getting overwhelmed. I would particularly recommend enabling the user to make side by side comparisons of different combination of graphs that are currently in fixed/distinct places in the tool.

*Undoubtedly, there are many ways to organize this information and we have done our best to organize it in a way that makes sense to us. That is, the pages follow the general order in which our stock assessments tend to flow: 1) assess fits and residuals, 2) assess retrospective plots and derived quantities, 3) if 1 and 2 pass peer review, assess the projections. We know the tool is far from perfect and agree that it would benefit from input from programmers and/or graphic design experts but, unfortunately, we lack funding to pursue such consultations. This is a reality that many biologists face and a key point of our paper is that this limitation should not preclude the use of these tools.*

*See also previous comments to the AE.*

Reviewer: 2

Comments to the Author

This is a clearly written and organized manuscript that presents an application of novel, open-source tools to disseminate stock assessment inputs and results. The interactive visualization tools were developed by others but modified by the authors to suit their needs; the original developers are credited appropriately. The authors make a strong case for using these tools for visualization of stock assessment results, and in that way, this paper is fitting for a perspective since it promotes a new paradigm for dissemination of stock assessments. There are no new research results, nor is there anything controversial in this piece. An important and laudable message of the paper is that these tools can help the transition to "a culture of open science and reproducible research."

*We thank the Reviewer for the kind comments.*

The manuscript repeatedly mentions the usefulness of these tools for stakeholders, both for data exploration and data modeling. In the data exploration figures (Figures 1 through 3), the interfaces would be more digestible to stakeholders if they included an annotation of what the user is viewing. For example, the Figure 1 view could include a note specifying that the error bars represent 95% confidence intervals, and species names should be used rather than, or in addition to, codes. Figure 2 could be improved by including units for the length bins, and Figure 3 should have a legend for the kernals. Each could also have a text field that gives an overview of the what the figure is showing and how a user can interact with the figure.

*TODO?: Greg & Bob*

More broadly, the NCAM Explorer (Supplement 2) would also be more functional with increased annotation. The background tab provides important context but should be expanded beyond the "terse point-form" to serve as a user guide for each tab. These improvements apply to the tools themselves rather than to the manuscript.

*We agree with these comments and have added text to the Background tab and added references for background.  See "stand alone" comment above.  We have tried to strike a balance between addressing the spirit of the Reviewers comments and unnecessarily cluttering the dashboard.*

The one area where I think the manuscript should be expanded is in the "Towards open stock-assessment" section. The main reason given for increasing the transparency of stock assessment through tools such as these is to improve peer review. However, as mentioned at several other points in the manuscript, these tools are also valuable for increasing stakeholder engagement in and understanding of stock assessment. That function warrants equal attention to peer review in this section and could present ideas for how to introduce these tools to stakeholders and how to help them sift through the large volume of data to find what is most relevant to them.

*TODO?: Keith*

Lastly, Supplement 1 (getting_started.html) isn’t referenced anywhere in the paper. It can be removed from the supplementary files.

*We have now added a reference to Supplement 1 to the introduction and conclusion.*

Reviewer: 3

Comments to the Author

This manuscript illustrates a novel approach to exploring and communicating data and model results related to fisheries stock assessments. Communication methods for complex patterns in data or stock assessment results is not a frequent topic covered in fisheries science journals but it is an area where new ideas are rapidly developing, much further progress is needed, and greater prominence in the literature is deserved. I highly recommend publishing this manuscript.

*We thank the Reviewer for the kind comments.*

The availability of the NCAM Explorer file NCAM_dashboard.html (referenced in the manuscript as Supplement 2) is vital for the understanding of the interactive nature of the tools developed. It would be useful to reference the other supplemental file getting_started.html in the text and include as an additional supplement an example of the RStrap Explorer as well in addition to the static figures in the text.

*We agree with the Reviewers comments and we have included a reference to Supplement 1 in the introduction and conclusion. RStrap Explorer is harder to include as a supplement as it is Shiny based and must be hosted on a server. At the moment, the use of this tool has been limited to local computers; we have yet to host the tool on an online server.*

I raise a few minor specific issues below, but the area where I would suggest the greatest change is adding to the introduction and conclusion further discussion of a few larger questions that this new approach raises. First, the visualization and communication needs are very broad within the stock assessment community, but the example tools presented are specific to analyses used in the region where the authors work: the RStrap analysis of survey data and the NCAM assessment model. I would have like to see some discussion about the trade-offs associated with developing generic dashboards that could be applied to other survey or assessment tools vs. developing separate dashboard applications for each database or stock assessment model. Comparing results from multiple stock assessment models which might be applied to the same stock is another challenge that might be discussed.

*Please see previous comments to the AE.*

Second, I would have liked more discussion of version control and archiving. The supplemental file NCAM_dashboard.html includes a comparison of two models "NCAM 2018" and "NCAM 2019" but in the assessment process there is typically a proliferation of numerous alternative models being considered in any given year. Exploration of these alternative models would presumably lead to a proliferation of these html dashboard files. Keeping track of which files show the results from which models is a challenge not addressed in the manuscript (and not obvious in the supplemental file). Furthermore, keeping track of the software version for the dashboard software (and R packages on which it depends) and format of the database or assessment model and maintaining compatibility between the two is a challenge not discussed here.

*This is an excellent point, however, we do discuss this issue as this approach does not offer a novel solution to file organization. The fact is that there is little difference between the file management with this approach versus a more standard approach. Through pre-dashboard days, different model runs were nested within independent folders and within these folders were figure and table folders. The only difference with the dashboard approach is that key figures and tables are held in the html file (i.e. there is no longer figure and table folders). A convention we apply to help keep track of things is to use a consistent name for the folder, the RData file containing all the inputs and outputs, and the html file. Though an interesting and important challenge, we did not cover this as file management and project conventions is a potential tangent that may add little to the paper as potential users are likely going to apply their own personal preferences.*

*Regarding software version, we now note in the conclusion that these files can be traceable and accessible via GitHub.*

Third, the link between this tool and more traditional static reports could be also addressed. The manuscript presents a compelling argument in favor of interactive dashboards as a tool for the rich volume of information needing to be explored in fisheries models, but it is not suggested that static reports will go away. It is noted that "The user also has the ability to hover over the points to compare estimates and export the figure as a png file" (line 110) but this would not be an efficient way to produce figures for a static report. Further discussion of how the same tools used for the dashboard view could used to make the development of static reports more efficient may be worth including in the discussion.

*We agree, it would not be very efficient to manually export individual plots. We have removed this note and have expanded our introductory text to clarify how synergies can be created between static and interactive documents.*

Specific issues: Line 29: "For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way." I think this is a mischaracterization of the state of the science. Perhaps Northern Cod is an exception, but I believe the challenge of having adequate information to provide sound advice remains as present as ever, even for relatively well-monitored stocks, while the second of meaningfully presenting the resulting information has grown beside it.

*Fair point. We have rephrased the sentence to read: "While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way."*

Line 81: the reference to a 1981 report for information on the "locally developed R package called RStrap" seems inadequate. Perhaps "locally developed" means only locally available and thus no citation for the R package is available (in contrast to the other packages cited in the manuscript).

*This is correct.  We now simply state that it is a local package.*

Line 121: "data base" should be one word for consistency with earlier occurrence on the same line.

*We have made the suggested change.*

Lines 262: The statement "Although these tools do not reveal the entire data pipeline" may be contradicted by the ICES Transparent Assessment Framework (https://taf.ices.dk/app/about) which archives the full process of data processing to running assessment models in a reproducible way and should be added to Appendix A.

*We thank the Reviewer for drawing our attention to the TAF. We have now re-structured our "Towards open stock assessment" section to highlight TAF and several other open science initiatives in fisheries research.*

Figure 2: I realize this figure is only for illustrative purposes, but it might be good to add a note about the strange pattern in the length compositions of American plaice in the 1980s. The spiky pattern with peaks in the same bins each year seems like an artifact of rounding error or conversion from some other type of length measurement.

*TODO?: Bob*

Again, I think this manuscript is a valuable contribution to the field, is of broad interest, and I encourage it to be published with or without the refinements proposed above.

*We thank the Reviewer for the kind comments.*

### 2020-03-14 inst/rmd/background.Rmd
Revamped NCAM explorer

Background
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### NCAM

<h1>NCAM: A state-space stock assessment model for Northern cod</h1>

```{r}

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)

xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
        fillcolor = toRGB(pred_col, alpha = 0.2),
        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
  add_trace() %>%
  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
  plotly::config(displayModeBar = FALSE)

```

### General

- An *integrated* *state-space* population dynamics model
- *Integrated*
    - Uses as much information as possible
- *State-space*
    - Includes process and observation equations + error
    - Population processes are informed by observations from multiple sources
    - Process errors are necessary for realistic stochastic projections
- See Cadigan 2015 [1] and 2016 [2] for details

### Data inputs

- Reported landings (`r text_range(years)`)
- Catch-at-age information (`r text_range(years)`)
- Reported landings by month (`r text_range(years)`)
- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
- Tagging data

### Biological inputs

::: {}

- RV survey data were used fit cohort-based models of:
   - Proportion mature at age
   - Beginning of year and mid-year stock weights at age [3]
   `r if (min(years) < 1983) "   - **Note that hindcast values were used for years that predate the RV survey (1983)**"`
- Baseline natural mortality (M) values
   `r if (min(years) < 1983) "   - **Assumed the same baseline M values for years prior to the start year of the base case NCAM model (1983)**"`

```{r}

plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 300, width = 500) %>%
  add_surface() %>%
  colorbar(title = "Mshift") %>%
  layout(scene = list(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Age"),
    zaxis = list(title = "Mshift")
  ))

```

:::

### Partial catches

::: {}

- User sets lower and upper bounds on total catch weight

```{r}

plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
  add_lines(name = "Lower", color = I(pos_col)) %>%
  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(neg_col)) %>%
  add_lines(y = rep(1, length(years)), color = I("black"),
            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
  layout(yaxis = list(title = "Catch multiplier"),
         xaxis = list(title = "Year"))

```

<br>

- Bounds are used via a censored likelihood
`r if (min(years) < 1983) "   - **Assumed the same starting bounds as those chosen for the base case NCAM model for years prior to 1983**"`
    - $l\left(C_{01}, \ldots, C_{o Y} | \boldsymbol{\theta}\right)=\sum_{y=1}^{Y} \log \left\{\Phi_{N}\left[\frac{\log \left(U_{Y} / C_{Y}\right)}{\sigma_{C}}\right]-\Phi_{N}\left[\frac{\log \left(C_{o Y} / C_{Y}\right)}{\sigma_{C}}\right]\right\}$

```{r, fig.height = 3.5, fig.width = 8}

C <- seq(400, 900, length = 500)
U <- 750
L <- 500
CV <- seq(0.005, 0.1, by = 0.005)
den <- vector("list", length(CV))
for (i in seq_along(CV)) {
  LZU <- (log(U) - log(C)) / CV[i]
  LZL <- (log(L) - log(C)) / CV[i]
  den[[i]] <- pnorm(LZU) - pnorm(LZL)
}
d <- data.frame(Catch = rep(C, length(C)), 
                Likelihood = unlist(den), 
                CV = rep(CV, each = length(C)) * 100)

plot_ly(data = d, hoverinfo = "none", showlegend = FALSE, height = 300, width = 500) %>% 
  add_lines(x = ~Catch, y = ~Likelihood, frame = ~CV, color = I(pred_col)) %>% 
  add_segments(x = L, y = 0, xend = L, yend = 1.05,
               color = I("black")) %>% 
  add_text(x = L, y = 1.09, text = "Lower bound") %>% 
  add_segments(x = U, y = 0, xend = U, yend = 1.05,
               color = I("black")) %>% 
  add_text(x = U, y = 1.09, text = "Upper bound") %>% 
  animation_opts(transition = 0) %>% 
  animation_button(visible = FALSE) %>% 
  animation_slider(currentvalue = list(suffix = "%")) %>% 
  layout(plot_bgcolor = "transparent", 
         paper_bgcolor = "transparent") %>% 
  config(displaylogo = FALSE, collaborate = FALSE, displayModeBar = FALSE)

```

<br>

- Age composition modeled using continuation-ratio logits
    - $l(\{X_{oa,y}\}|\boldsymbol{\theta})=\Sigma^{A-1}_{a=1}\Sigma^{Y}_{y=1}log\left[\varphi_N\left(\frac{X_{oa,y}-X_{a,y}}{\sigma_{Pa}}\right)\right]$

:::

### Surveys

- Survey estimates are assumed to be log-normally distributed
    - $E(I_{s,a,y})=q_{s,a}N_{a,y} exp(-t_{s,y}Z_{a,y})$
    - $l\left(\left\{I_{s, a, y}\right\} | \boldsymbol{\theta}\right)=\sum_{a} \sum_{y} \log \left(\varphi_{N}\left\{\frac{\log \left(I_{s, a y}\right)-E\left[\log \left(I_{s, a, y}\right)\right]}{\sigma_{s, I}}\right\}\right)$

- Shift q for the offshore RV survey by estimating the proportion of the stock in Smith Sound $(D_{a,y})$
    - $E\left(I_{\mathrm{RV}, a, y}\right)=\left\{\begin{array}{ll}{q_{a} N_{a, y} \exp \left(-t Z_{a, y}\right),} & {y \neq 1995, \ldots, 2009} \\ {q_{a}\left(1-D_{a, y}\right) N_{a, y} \exp \left(-t Z_{a, y}\right),} & {y=1995, \ldots, 2009}\end{array}\right.$

- Shift q for the inshore sentinel survey by estimating year effects that follow a random walk $(q_{s,y})$
    - $log(q_{s,a,y})=log(q_{s,a})+log(q_{s,y})$

### Tagging data

- Age of tagged fish ~ length-at-release | Von Bertalanffy growth model [4]
- Age-based tag returns in each experiment $(x)$ modeled using Baranov catch equation
    - $R_{x, a, y}=\lambda_{x,y}N_{x, a, y}\left\{1-\exp \left(-Z_{x, a, y}\right)\right\} \frac{F_{x, a, y}}{Z_{x, a, y}}$
        - note adjustment for reporting rate $(\lambda_{x,y})$

- Adjust for incomplete mixing in release year: experiment $~F~$ is assumed to be random deviations from whole stock $~F~$
    - $\log \left(F_{x, a, y}\right) \sim N\left\{\log \left(F_{a, y}\right), \sigma_{f x}^{2}\right\}$

- Adjust for partial exposure to the fishery $(s_{x,y})$ and other mortality sources$(t_{x,y})$, tag loss $(\gamma_y)$ and tagging mortality $(u)$
    - $Z_{x, a, y} = s_{x,y} F_{x, a, y} + t_{x,y} M_{a,y} + \gamma_y + u$

### Population process

- Cohort equation
    - $N_{a,y}=N_{a-1,y-1}exp(-Z_{a-1,y-1})$

- Baranov catch equation
    - $C_{a,y}=N_{a,y}[1-exp(-Z_{a,y})]\frac{F_{a,y}}{Z_{a,y}}$

- Process error $(\delta)$ is added to baseline $m$ values
    - $log(M_{a,y})=log(m_{a,y})+\delta_{a,y}$

- $\delta$ and $F$ modeled using 2D age-year AR1 process
    - $\operatorname{Cov}\left(\delta_{a, y}, \delta_{a-j, y-k}\right)=\frac{\sigma_{\delta}^{2} \varphi_{\delta, \mathrm{age}} \varphi_{\delta, \mathrm{year}}^{k}}{\left(1-\varphi_{\delta, \mathrm{age}}^{2}\right)\left(1-\varphi_{\delta, \mathrm{year}}^{2}\right)}, \quad \operatorname{Corr}\left(\delta_{a, y}, \delta_{a-j, y-k}\right)=\varphi_{\delta, \mathrm{age}}^{j} \varphi_{\delta, \text { year }}^{k}$
    - $\operatorname{Cov}\left[\log \left(F_{a, y}\right), \log \left(F_{a-j, y-k}\right)\right]=\frac{\sigma_{F}^{2} \varphi_{F, \text { age }}^{j} \varphi_{F \text { year }}^{k}}{\left(1-\varphi_{\text {Frage }}^{2}\right)\left(1-\varphi_{F, \text { year }}^{2}\right)}$

### Summary

- $N$ informed by reported landings and survey indices
    - **Challenges:** bounds and catchability
- $F$ and $M$ informed by catch composition, survey indices and tagging data
    - **Challenges:** catchability, reporting rate, incomplete mixing
- Different parameters confound population estimates from each data-set
    - Statistical triangulation $\rightarrow$ underlying population process
    - Same process $\therefore$ conflicting trends $\rightarrow$ poor fits

### References

1. Cadigan, N. G. A state-space stock assessment model for northern cod, including under-reported catches and variable natural mortality rates. Canadian Journal of Fisheries and Aquatic Sciences 73, 296–308 (2015).

2. Cadigan, N. G. Updates to a northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Canadian Science Advisory Secretariat Research Document 2016/022, 58 p. (2016).

3. Cadigan, N. G. Weight-at-age growth models and forecasts for northern cod (*Gadus morhua*). DFO Canadian Science Advisory Secretariat Research Document 2016/016, 19 p. (2016).

4. Cadigan, N. G. & Konrad, C. A cohort time-series von bertalanffy growth model for northern cod (*Gadus morhua*), and estimation of the age of tagged cod. DFO Canadian Science Advisory Secretariat Research Document 2016/017, 37 p. (2016).

### 2020-03-14 inst/rmd/fishery.Rmd
Revamped NCAM explorer

Landings {data-navmenu="Fishery"}
================================================================================

Row {.tabset .tabset-dropdown}
--------------------------------------------------------------------------------

### Landings

```{r}

plot_landings(fit, cols = c(pred_col, obs_col))

```

> Y-axes in log scale

### Numbers at age

```{r}

d <- tidy_mats(list(obs = t(inputs$catch.num[as.character(years), as.character(ages)]),
                    pred = fit$rep$pred_catch),
               ages = ages, years = years)
d$cohort <- factor(d$year - d$age)
d$age <- factor(d$age)

facet_obs_pred(d, by = "age", type = "log",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = obs_col, pred_col = pred_col) %>%
  layout(yaxis = list(rangemode = "tozero"))

```

> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration

### Numbers at age (frame)

```{r}
plot_obs_pred(data = d, showlegend = TRUE, type = "log",
              frame = "age", ylab = "Numbers",
              obs_col = obs_col, pred_col = pred_col)

```

> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration

Row {.tabset}
--------------------------------------------------------------------------------

### Proportions at age

```{r}

d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
                    pred = fit$rep$pred_prop_catch),
               ages = ages, years = years)
d$cohort <- factor(d$year - d$age)
d$age <- factor(d$age)

facet_obs_pred(d, by = "age",
               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
               nrows = 6, ylab = "Proportion",
               obs_col = obs_col, pred_col = pred_col) %>%
  layout(yaxis = list(rangemode = "tozero"))

```

### Proportions at age (frame)

```{r}

plot_obs_pred(data = d, showlegend = TRUE, type = "linear",
              frame = "age", ylab = "Proportion",
              obs_col = obs_col, pred_col = pred_col)

```

### Residuals

```{r}

d <- tidy_mats(list(res = fit$rep$CA_resid,
                    fit = fit$rep$pred_crl),
               ages = ages[-length(ages)], years = years)
d$cohort <- d$year - d$age
d$rec <- seq(nrow(d))
plot_resid(d, col = obs_col, line_col = pred_col)

```

> based on continuation ratio logits

Selectivity {data-navmenu="Fishery"}
================================================================================

Row {.tabset}
--------------------------------------------------------------------------------

### Proportions at age surface

```{r}
plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
```

<!-- ### Numbers at age surface -->

<!-- ```{r} -->
<!-- plot_surface(mat = t(inputs$catch.num[as.character(years), as.character(ages)]),  -->
<!--              ages = ages, years = years, name = "N") -->
<!-- ``` -->

<!-- > Note: model fits to composition - this plot is included for illustration -->

Row {.tabset}
--------------------------------------------------------------------------------

### Fishery selectivity

```{r}

S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
plot_surface(S, ages = ages, years = years, name = "S")

```

### 2020-03-14 inst/rmd/projections.Rmd
Revamped NCAM explorer

--- 2019-09-10 inst/rmd/vis_fit.Rmd
@@
----
-title: "NCAM Explorer"
-output: flexdashboard::flex_dashboard
-always_allow_html: yes
----
-
-<style>
-/* list */
-li {
-line-height: 30px;
-}
-</style>
-
-```{r setup, echo=FALSE}
-library(plotly)
-library(Matrix)
-library(knitr)
-library(data.table)
-library(dplyr)
-library(webshot)
-knitr::opts_knit$set(root.dir = tmp$output_dir)
-knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
-```
-
-```{r data, include=FALSE}
-
-cols <-  viridis::viridis(4)[c(2, 3, 1, 4)]
-cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
-fit <- tmp$fit # temp object from vis_fit function
-if (!is.null(tmp$comp)) { comp <- tmp$comp }
-if (!is.null(tmp$last)) { last <- tmp$last }
-est <- as.list(fit$sd.rep, "Est")
-sd <- as.list(fit$sd.rep, "Std")
-tabs <- tidy_model(fit)
-if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
-if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
-if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
-min_year <- min(fit$tmb.data$years)
-max_year <- max(fit$tmb.data$years)
-ages <- fit$tmb.data$ages
-years <- fit$tmb.data$years
-years_plus1 <- fit$tmb.data$years_plus1
-
-## helper functions for nice Rmarkdown text
-text_range <- function(x) paste(range(x), collapse = "-")
-text_c <- function(x) {
-  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
-}
-text_units <- function(x) {
-  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
-  if (x >= 1000) return(paste0(round(x / 1000), "k"))
-  if (x < 1000) return(as.character(round(x)))
-}
-
-```
-
-Background
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### NCAM
-
-<h1>NCAM: A state-space stock assessment model for Northern cod</h1>
-
-```{r}
-
-ax <- list(
-  title = "",
-  zeroline = FALSE,
-  showline = FALSE,
-  showticklabels = FALSE,
-  showgrid = FALSE
-)
-
-xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
-plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
-        fillcolor = 'rgba(49, 104, 142, 0.3)',
-        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
-  add_trace() %>%
-  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
-         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
-  plotly::config(displayModeBar = FALSE)
-
-```
-
-### General
-
-- An *integrated* *state-space* population dynamics model
-- *Integrated*
-    - Uses as much information as possible
-- *State-space*
-    - Includes process and observation equations + error
-    - Population processes are informed by observations from multiple sources
-    - Process errors are necessary for realistic stochastic projections
-
-### Data inputs
-
-- Reported landings (`r text_range(years)`)
-- Catch-at-age information (`r text_range(years)`)
-- Reported landings by month (`r text_range(years)`)
-- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
-- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
-- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
-- Tagging data (`r text_range(years)`)
-
-### Biological inputs
-
-- RV survey data were used fit cohort-based models of:
-   - Proportion mature at age
-   - Beginning of year and mid-year stock weights at age
-- Baseline natural mortality (M) values
-
-```{r}
-
-plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
-  add_surface() %>%
-  colorbar(title = "Mshift") %>%
-  layout(scene = list(
-    xaxis = list(title = "Year"),
-    yaxis = list(title = "Age"),
-    zaxis = list(title = "Mshift")
-  ))
-
-```
-
-### Partial catches
-
-- User sets lower and upper bounds on total catch weight
-- Bounds are used via a censored likelihood
-
-```{r}
-
-plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
-  add_lines(name = "Lower", color = I(cols[1])) %>%
-  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
-  add_lines(y = rep(1, length(years)), color = I("black"),
-            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Change in Q
-
-- A shift in Q was assumed for the offshore RV survey
-- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
-- Q for the inshore sentinel survey was assumed to follow a random walk
-    - Done to account for potential changes in migration patterns
-
-### Tagging data
-
-- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
-- Applies adjustments for tagging mortality, tag loss and reporting rates
-- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
-- The Baranov catch equation was applied to age-based tag survivors
-- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock
-
-### References
-
-Cadigan, N.
-2015.
-A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates.
-Canadian Journal of Fisheries and Aquatic Sciences 73 (2).
-NRC Research Press: 296-308.
-
-Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.
-
-Catch
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-
-plot_landings(fit)
-
-```
-
-> Y-axes in log scale
-
-### Numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = t(inputs$catch.num[as.character(years), as.character(ages)]),
-                    pred = fit$rep$pred_catch),
-               ages = ages, years = years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age", type = "log",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Predictions by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Numbers")
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Predictions by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = 1, end = 0.8),
-              ylab = "Numbers")
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Numbers at age surface
-
-```{r}
-plot_surface(mat = t(inputs$catch.num[as.character(years), as.character(ages)]), 
-             ages = ages, years = years, name = "N")
-```
-
-> Note: model fits to composition - this plot is included for illustration
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Proportions at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
-                    pred = fit$rep$pred_prop_catch),
-               ages = ages, years = years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "linear", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Proportion", stack_group = TRUE)
-```
-
-> Note: values are stacked
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "linear", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = 1, end = 0.8),
-              ylab = "Proportion", stack_group = TRUE)
-```
-
-> Note: values are stacked
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$CA_resid,
-                    fit = fit$rep$pred_crl),
-               ages = ages[-length(ages)], years = years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-> based on continuation ratio logits
-
-### Proportions at age surface
-
-```{r}
-plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
-```
-
-RV Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### RV survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
-pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
-obs[obs == 0] <- NA
-pred[pred == 0] <- NA
-d <- data.frame(year = years, obs = obs, pred = pred)
-
-# ## RV biomass
-# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# d <- data.frame(year = years, obs = obs, pred = pred)
-# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    zero = fit$tmb.data$ind_zero,
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d <- d[d$wt == 1, ]
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
-                    pred = exp(fit$rep$Elog_RV_index),
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d$obs[d$wt == 0] <- NA
-d$pred[d$wt == 0] <- NA
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Mean numbers per tow",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Mean numbers per tow")
-```
-
-> Y-axes in log scale
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE,
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = -1, end = 0.8),
-              ylab = "Mean numbers per tow")
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
-                    fit = fit$rep$Elog_RV_index),
-               ages = ages, years = years)
-d <- d[d$keep, ]
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SN Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$SN_GN_index)
-pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
-d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$log_SN_GN_resid,
-              ages = fit$tmb.data$SN_ages,
-              years = fit$tmb.data$SN_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
-                    pred = exp(fit$rep$Elog_SN_GN_index)),
-               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
-               nrows = 4, ylab = "Sentinel survey index",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Sentinel survey index")
-```
-
-> Y-axes in log scale
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE,
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = -1, end = 0.8),
-              ylab = "Sentinel survey index")
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
-                    fit = fit$rep$Elog_SN_GN_index),
-               ages = fit$tmb.data$SN_ages,
-               years = fit$tmb.data$SN_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SS Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Smith sound acoustic biomass estimates
-
-```{r}
-
-d <- data.frame(year = fit$tmb.data$SS_bms_years,
-                obs = fit$tmb.data$SSbms,
-                pred = fit$rep$ESSbms)
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
-              showlegend = TRUE)
-
-```
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
-              ages = fit$tmb.data$SS_ages,
-              years = fit$tmb.data$SS_age_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Smith Sound trawl numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
-                    pred = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
-               obs_col = cols[2], pred_col = cols[1])
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
-                    fit = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-Tagging {data-orientation=rows}
-================================================================================
-
-```{r}
-
-## Prep tag data and fits
-tag_dat <- data.frame(
-  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
-  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
-  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
-  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
-  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
-  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
-  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
-  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
-  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
-             rep("post 1996", length(fit$tmb.data$tag_exp)))
-)
-tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
-tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
-                      "age:", tag_dat$age, "</br>",
-                      "delay:", tag_dat$delay, "</br>")
-
-Udev <- data.frame(
-  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
-  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
-)
-
-Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
-names(Fay)[names(Fay) == "year"] <- "rel_year"
-
-tag_dat <- merge(tag_dat, Udev, by = "exp")
-tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
-tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
-
-## Share data across multiple plots
-d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% 
-  group_by(exp) %>% 
-  highlight(off = "plotly_relayout")
-
-```
-
-Row {data-height=700 .tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted tag recaptures by experimemnt
-
-```{r}
-
-y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
-y_tickvals <- sqrt(y_ticktext)
-
-rel_plot <- base %>%
-  filter(rel_year == cap_year) %>%
-  group_by(exp) %>%
-  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~jitter(rel_year, factor = 0.5), 
-              y = ~sqrt(tot_rel), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N released",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
-y_tickvals <- sqrt(y_ticktext)
-
-cap_plot <- base %>%
-  group_by(exp, rel_year, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
-            period = unique(period)) %>%
-  group_by(exp) %>%
-  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
-              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
-            alpha = 0.7, showlegend = FALSE, symbol = NULL,
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y2",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
-        nrows = 2, heights = c(0.3, 0.7))
-
-```
-
-> Y-axis is in square root scale
-
-### Animation
-
-```{r}
-
-sum_d <- tag_dat %>%
-  group_by(exp, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs))
-
-sum_d %>%
-  group_by(exp) %>%
-  plot_ly(x = ~cap_year, frame = ~exp, hoverinfo = "x+y+text",
-          showlegend = FALSE) %>%
-  add_markers(y = ~sqrt(sum_obs), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(y = ~sqrt(sum_pred), color = I(cols[2]),
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>%
-  layout(yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Tag reporting rate
-
-```{r}
-
-cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
-rr <- data.frame(year = fit$tmb.data$SRR_years,
-                 est = fit$tmb.data$log_SRR_est,
-                 sd = sqrt(diag(cov_mat)))
-rr$lwr <- rr$est - 1.96 * rr$sd
-rr$upr <- rr$est + 1.96 * rr$sd
-rr <- trans_est(rr)
-rr$model <- "Random Walk"
-
-ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
-                      est = est$log_SRR, sd = sd$log_SRR)
-ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
-ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
-ncam_rr <- trans_est(ncam_rr)
-ncam_rr$model <- "NCAM"
-
-rr <- rbind(rr, ncam_rr)
-
-plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)
-
-```
-
-### Initial F deviations for tagging experiments
-
-```{r}
-
-base %>%
-  filter(rel_year == cap_year) %>%
-  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
-              showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "Initial F deviation"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Observed vs. predicted recaptures
-
-```{r}
-
-x_ticktext <- c(5, pretty(tag_dat$pred, 10))
-y_ticktext <- c(5, pretty(tag_dat$obs, 10))
-x_tickvals <- sqrt(x_ticktext)
-y_tickvals <- sqrt(y_ticktext)
-
-base %>%
-  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
-              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       line = list(width = 1)),
-         xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Observed number of recaptures",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext))
-
-```
-
-> Axes are in square root scale
-
-### Standardized residuals vs. predicted recaptures
-
-```{r}
-
-y_ticktext <- c(5, pretty(tag_dat$res, 10))
-y_tickvals <- sqrt(x_ticktext)
-
-base %>%
-  add_markers( x = ~sqrt(pred), y = ~res,
-               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Standardized residuals"))
-
-```
-
-> X-axis is in square root scale
-
-Catchability
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Fishery selectivity
-
-```{r}
-
-S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
-plot_surface(S, ages = ages, years = years, name = "S")
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey catchability
-
-```{r}
-
-q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
-plot_surface(q, ages = ages, years = years, name = "q")
-
-```
-
-### Proportion of the stock in Smith Sound
-
-```{r}
-p <- matrix(0, nrow = length(ages), ncol = length(years),
-            dimnames = list(ages, years))
-ss_p <- fit$rep$SSD
-ss_years <- fit$tmb.data$SS_years
-ss_ages <- fit$tmb.data$SS_ages
-rownames(ss_p) <- ss_ages
-colnames(ss_p) <- ss_years
-p[as.character(ss_ages), as.character(ss_years)] <- ss_p
-plot_surface(p, ages = ages, years = years, name = "p")
-
-```
-
-### Sentinel gillnet survey catchability
-
-```{r}
-
-log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
-log_qa <- fit$rep$log_SN_GN_Qa
-q <- exp(log_qy + log_qa)
-plot_surface(q, ages = fit$tmb.data$SN_ages,
-             years = fit$tmb.data$SN_years,
-             name = "q")
-
-```
-
-Trends
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r}
-
-## Note: biomass is in tonnes and numbers are in thousands
-bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
-             data.frame(metric = "Ages 2+", tabs$log_biomass))
-abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
-             data.frame(metric = "Ages 2+", tabs$log_abundance))
-
-bms_plot <- trans_est(bms) %>%
-  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
-             annotate = TRUE)
-
-abn_plot <- trans_est(abn, scale = 0.001) %>%
-  plot_trend(ylab = "Abundance", group = "metric", col = cols,
-             annotate = TRUE)
-
-blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
-  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
-  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
-           color = I("black"), textposition = "right", hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Abundance at age
-
-```{r}
-plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
-```
-
-### Biomass at age
-
-```{r}
-plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
-```
-
-### SSB at age
-
-```{r}
-plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
-```
-
-<!-- ### Exploitable biomass at age -->
-
-<!-- ```{r} -->
-<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
-<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
-<!-- E <- B * S -->
-<!-- plot_surface(E, ages = ages, years = years, name = "Exploitable biomass") -->
-<!-- ``` -->
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r}
-
-f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
-z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
-m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
-
-z_plot <- trans_est(z) %>%
-  plot_trend(ylab = "Average Z", group = "metric",
-             annotate = TRUE, col = cols)
-
-m_plot <- trans_est(m) %>%
-  plot_trend(ylab = "Average M", group = "metric",
-             annotate = TRUE, col = cols)
-
-f_plot <- trans_est(f) %>%
-  plot_trend(ylab = "Average F", group = "metric",
-             annotate = TRUE, col = cols)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Z at age
-
-```{r}
-plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
-```
-
-### M at age
-
-```{r}
-plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
-```
-
-### F at age
-
-```{r}
-plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
-```
-
-### Z, M and F
-
-```{r}
-plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
-                 ages = ages, years = years)
-```
-
-> Z = purples, M = blues, F = reds
-
-### Process error
-
-```{r}
-
-pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], 
-             ages = fit$tmb.data$pe_ages, name = "pe", 
-             col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
-  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
-              color = I("black"), opacity = 0.8)
-
-```
-
-### Process error (bubble plot)
-
-```{r}
-pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = fit$tmb.data$pe_ages, value = "res")
-bubble_plot(pe)
-```
-
-Retro {`r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Mohn's rho
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_mohns(col = cols[1], ylab = "SSB")
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_mohns(ylab = "Recruitment (age 2)", col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_mohns(ylab = "SSB/Blim", col = cols[1])
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Relative difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
-
-### Difference
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- retro_tabs$log_ssb %>%
-  plot_retro_deltas(in_col = cols[1], ylab = "log(SSB)")
-
-abn_plot <- retro_tabs$log_recruitment %>%
-  plot_retro_deltas(ylab = "log(Recruitment)", in_col = cols[1])
-
-blim_plot <- retro_tabs$log_ssb_lrp %>%
-  plot_retro_deltas(ylab = "log(SSB/Blim)", in_col = cols[1])
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Mohn's rho
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_mohns(ylab = "Average Z (ages 5+)", col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_mohns(ylab = "Average M (ages 5+)", col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_mohns(ylab = "Average F (ages 5+)", col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Relative difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
-
-### Difference
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- retro_tabs$log_aveZ %>%
-  plot_retro_deltas(ylab = "log(Average Z)", in_col = cols[1])
-
-m_plot <- retro_tabs$log_aveM %>%
-  plot_retro_deltas(ylab = "log(Average M)", in_col = cols[1])
-
-f_plot <- retro_tabs$log_aveF %>%
-  plot_retro_deltas(ylab = "log(Average F)", in_col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
-
-Comps {`r if (is.null(tmp$comp)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(tmp$comp)}
-
-abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
-  plot_comps(ylab = "Recruitment (age 2)", col = cols)
-
-bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
-  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)
-
-blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
-  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(tmp$comp)}
-
-z_plot <- trans_est(comp_tabs$log_aveZ) %>%
-  plot_comps(ylab = "Average Z (ages 5+)", col = cols)
-
-m_plot <- trans_est(comp_tabs$log_aveM) %>%
-  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)
-
-f_plot <- trans_est(comp_tabs$log_aveF) %>%
-  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Productivity
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Recruitment
-
-```{r}
-trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1])
-```
-
-### SSB
-
-```{r}
-trans_est(tabs$log_ssb) %>%
-  plot_trend(ylab = "SSB (t)", col = cols[1])
-```
-
-### RPS
-
-```{r}
-trans_est(tabs$log_rps) %>%
-  filter(year >= min_year + min(ages)) %>% 
-  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
-  layout(xaxis = list(range = c(min_year, max_year + 1)))
-```
-
-### SSB x Recruitment
-
-```{r}
-
-ssb <- trans_est(tabs$log_ssb)
-rec <- trans_est(tabs$log_recruitment, scale = 0.001)
-rps <- trans_est(tabs$log_rps)
-names(ssb) <- paste0("ssb_", names(ssb))
-names(rec) <- paste0("rec_", names(rec))
-names(rps) <- paste0("rps_", names(rps))
-ssb$year <- ssb$ssb_year + min(ages)
-rec$year <- rec$rec_year
-rps$year <- rps$rps_year
-sr <- merge(ssb, rec, by = "year")
-sr <- merge(sr, rps, by = "year")
-sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)
-
-plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
-                 col = ~year, clab = "Year")
-
-```
-
-### SSB x RPS
-
-```{r}
-p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
-                 col = ~year, clab = "Year")
-p
-```
-
-### log(SSB) x log(RPS)
-
-```{r}
-p %>% layout(yaxis = list(type = "log"),
-             xaxis = list(type = "log"))
-```
-
 Assumptions  {data-navmenu="Projections" .storyboard}
@@
 tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>%
-  filter(year >= max(years) - 4) %>%
+  filter(year >= max(years) - 3) %>%
   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>%
@@
 trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1]) %>% 
+  plot_trend(ylab = "Recruitment", col = pred_col) %>% 
   add_lines(data = mean_rec, x = ~year, y = ~mean_rec, name = "mean",
@@
 plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%",
-          vline = max(years) + 1, col = cols[1]) %>%
+          vline = max(years) + 1, col = pred_col) %>%
   layout(yaxis = list(tickvals = seq(-200, 200, by = 25)))
@@
 plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim",
-          vline = max(years) + 1, hline = 1, col = cols[1], darken_proj = TRUE)
+          vline = max(years) + 1, hline = 1, col = pred_col, darken_proj = TRUE)
 
@@
 ```
-
-Inputs {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
-```
-
-### Catch at age
-
-```{r}
-kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
-```
-
-<!-- ### Monthly landings -->
-
-<!-- ```{r} -->
-<!-- d <- inputs$catch.month -->
-<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
-<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
-<!-- ``` -->
-
-### RV index
-
-```{r}
-d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
-```
-
-### SN index
-
-```{r}
-d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
-mat <- d[, as.character(3:10)]
-rownames(mat) <- d$year
-kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
-```
-
-### SS index
-
-```{r}
-d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
-names(d) <- c("Year", "Month", "Biomass", "SE^2^")
-kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
-```
-
-### SS age composition
-
-```{r}
-kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
-```
-
-### Weight at age
-
-```{r}
-d <- inputs$pwts
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "weight")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
-```
-
-### Maturity at age
-
-```{r}
-d <- inputs$mats
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "mat")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div.
-2J+3KL from DFO RV autumn bottom trawl surveys.
-Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
-```
-
-### Tag summary
-
-```{r}
-d <- inputs$mtag1 %>% 
-  group_by(rel_year, exp) %>% 
-  summarise(NAFO = unique(div),
-            Nrel = sum(unique(released)), 
-            Ncap = sum(recaptured))
-names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
-kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
-```
-
-Settings {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Catch bounds
-
-```{r}
-d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
-kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
-```
-
-### Baseline M
-
-```{r}
-m <- fit$tmb.data$M_matrix
-rownames(m) <- ages
-colnames(m) <- years
-kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
-```
-
-Output {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Parameter estimates
-
-```{r}
-
-par_tab <- tidy_pars(fit)
-rownames(par_tab) <- NULL
-options(knitr.kable.NA = '')
-kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Parameter estimates (retro)
-
-```{r}
-
-par_tabs <- lapply(names(fit$retro), function(nm) {
-  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
-  d$id <- seq(nrow(d))
-  d
-})
-par_tabs <- do.call(rbind, par_tabs)
-final_par_tab <- data.frame(tyear = max(years), par_tab)
-final_par_tab$id <- seq(nrow(final_par_tab))
-par_tabs <- rbind(par_tabs, final_par_tab)
-par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-par_tabs$Quantity <- as.character(par_tabs$Quantity)
-par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
-par_tabs$Symbol <- as.character(par_tabs$Symbol)
-par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-rownames(par_tabs) <- NULL
-names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"
-
-kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Stock size
-
-```{r}
-
-tab_name <- c("log_abundance" = "Abundance (millions)", 
-              "log_recruitment" = "Recruits (age 2; millions)", 
-              "log_biomass" = "Biomass (kt)", 
-              "log_ssa" = "Spawners (millions)", 
-              "log_ssb" = "SSB (kt)", 
-              "log_ssb_lrp" = "SSB/Blim n(%)")
-
-nms <- names(tab_name)
-status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
-names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
-  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  status_tab <- merge(status_tab, tab, by = "Year")
-}
-
-knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Rates
-
-```{r}
-
-tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
-              "log_aveM" = "Average M (ages 5+)", 
-              "log_aveF" = "Average F (ages 5+)")
-
-nms <- names(tab_name)
-rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
-names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  rate_tab <- merge(rate_tab, tab, by = "Year")
-}
-
-knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Abundance-at-age
-
-```{r}
-N_matrix <- rbind(fit$rep$N_matrix, fit$rep$abundance)
-dimnames(N_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
-knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Mature abundance-at-age
-
-```{r}
-ssa_matrix <- rbind(fit$rep$ssa_matrix, fit$rep$ssa)
-dimnames(ssa_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
-knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Biomass-at-age
-
-```{r}
-B_matrix <- rbind(fit$rep$biomass_matrix, fit$rep$biomass)
-dimnames(B_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
-knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Mature biomass-at-age
-
-```{r}
-ssb_matrix <- rbind(fit$rep$ssb_matrix, fit$rep$ssb)
-dimnames(ssb_matrix) <- list(age = c(ages, "Total"), year = years_plus1)
-knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
-```
-
-<!-- ### Exploitable biomass -->
-
-<!-- ```{r} -->
-<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
-<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
-<!-- E <- B * S -->
-<!-- ebms_matrix <- rbind(E, colSums(E)) -->
-<!-- dimnames(ebms_matrix) <- list(age = c(ages, "Total"), year = years) -->
-<!-- knitr::kable(t(ebms_matrix) / 1000, -->
-<!--              caption = "Exploitable biomass estimates (kt).", digits = 3) -->
-<!-- ``` -->
-
-### Predicted catch
-
-```{r}
-pred_catch_wt_matrix <- rbind(fit$rep$pred_catch_wt_matrix, fit$rep$pred_catch_wt)
-dimnames(pred_catch_wt_matrix) <- list(age = c(ages, "Total"), year = years)
-knitr::kable(t(pred_catch_wt_matrix) / 1000, 
-             caption = "NCAM predicted catch at age (kt).", digits = 3)
-```
-
-### Z-at-age
-
-```{r}
-Z_matrix <- fit$rep$Z_matrix
-dimnames(Z_matrix) <- list(age = ages, year = years)
-knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
-```
-
-### M-at-age
-
-```{r}
-pe_ages <- fit$tmb.data$pe_ages
-M_matrix <- fit$rep$Mpe_matrix
-dimnames(M_matrix) <- list(age = ages, year = years)
-M_matrix <- M_matrix[as.character(pe_ages), ]
-rownames(M_matrix) <- c(as.character(head(pe_ages, -1)), paste0(tail(pe_ages, 1), "+"))
-knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
-```
-
-### F-at-age
-
-```{r}
-F_matrix <- fit$rep$F_matrix
-dimnames(F_matrix) <- list(age = ages, year = years)
-knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
-```

### 2020-03-14 inst/rmd/surveys.Rmd
Revamped NCAM explorer

--- 2018-11-21 inst/rmd/vis_fit.Rmd
@@
----
-title: "NCAM Explorer"
-output: flexdashboard::flex_dashboard
-always_allow_html: yes
----
-
-<style>
-/* list */
-li {
-line-height: 30px;
-}
-</style>
-
-```{r setup, echo=FALSE}
-library(plotly)
-library(Matrix)
-library(knitr)
-library(data.table)
-library(webshot)
-knitr::opts_knit$set(root.dir = tmp$output_dir)
-knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
-```
-
-```{r data, include=FALSE}
-
-cols <- viridis::viridis(4)[-1]
-cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
-fit <- tmp$fit # temp object from vis_fit function
-if (!is.null(tmp$comp)) { comp <- tmp$comp }
-if (!is.null(tmp$last)) { last <- tmp$last }
-est <- as.list(fit$sd.rep, "Est")
-sd <- as.list(fit$sd.rep, "Std")
-tabs <- tidy_model(fit)
-if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
-if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
-if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
-min_year <- min(fit$tmb.data$years)
-max_year <- max(fit$tmb.data$years)
-ages <- fit$tmb.data$ages
-years <- fit$tmb.data$years
-years_plus1 <- fit$tmb.data$years_plus1
-
-## helper functions for nice Rmarkdown text
-text_range <- function(x) paste(range(x), collapse = "-")
-text_c <- function(x) {
-  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
-}
-text_units <- function(x) {
-  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
-  if (x >= 1000) return(paste0(round(x / 1000), "k"))
-  if (x < 1000) return(as.character(round(x)))
-}
-
-```
-
-Background
+RV Survey  {data-navmenu="Surveys"}
 ================================================================================
@@
 Row {.tabset}
---------------------------------------------------------------------------------
-
-### NCAM
-
-<h1>NCAM: A state-space stock assessment model for Northern cod</h1>
-
-```{r}
-
-ax <- list(
-  title = "",
-  zeroline = FALSE,
-  showline = FALSE,
-  showticklabels = FALSE,
-  showgrid = FALSE
-)
-
-xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
-plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
-        fillcolor = 'rgba(49, 104, 142, 0.3)',
-        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
-  add_trace() %>%
-  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
-         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
-  plotly::config(displayModeBar = FALSE)
-
-```
-
-### General
-
-- An *integrated* *state-space* population dynamics model
-- *Integrated*
-    - Uses as much information as possible
-- *State-space*
-    - Includes process and observation equations + error
-    - Population processes are informed by observations from multiple sources
-    - Process errors are necessary for realistic stochastic projections
-
-### Data inputs
-
-- Reported landings (`r text_range(years)`)
-- Catch-at-age information (`r text_range(years)`)
-- Reported landings by month (`r text_range(years)`)
-- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
-- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
-- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
-- Tagging data (`r text_range(years)`)
-
-### Biological inputs
-
-- RV survey data were used fit cohort-based models of:
-   - Proportion mature at age
-   - Beginning of year and mid-year stock weights at age
-- Baseline natural mortality (M) values
-
-```{r}
-
-plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
-  add_surface() %>%
-  colorbar(title = "Mshift") %>%
-  layout(scene = list(
-    xaxis = list(title = "Year"),
-    yaxis = list(title = "Age"),
-    zaxis = list(title = "Mshift")
-  ))
-
-```
-
-### Partial catches
-
-- User sets lower and upper bounds on total catch weight
-- Bounds are used via a censored likelihood
-
-```{r}
-
-plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
-  add_lines(name = "Lower", color = I(cols[1])) %>%
-  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
-  add_lines(y = rep(1, length(years)), color = I("black"),
-            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Change in Q
-
-- A shift in Q was assumed for the offshore RV survey
-- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
-- Q for the inshore sentinel survey was assumed to follow a random walk
-    - Done to account for potential changes in migration patterns
-
-### Tagging data
-
-- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
-- Applies adjustments for tagging mortality, tag loss and reporting rates
-- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
-- The Baranov catch equation was applied to age-based tag survivors
-- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock
-
-### References
-
-Cadigan, N.
-2015.
-A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates.
-Canadian Journal of Fisheries and Aquatic Sciences 73 (2).
-NRC Research Press: 296-308.
-
-Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.
-
-Catch
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-
-plot_landings(fit)
-
-```
-
-> Y-axes in log scale
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted catch at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
-                    pred = fit$rep$pred_prop_catch),
-               ages = ages, years = years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$CA_resid,
-                    fit = fit$rep$pred_crl),
-               ages = ages[-length(ages)], years = years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-> based on continuation ratio logits
-
-### Catch at age surface (proportion)
-
-```{r}
-plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
-```
-
-<!-- ### Catch at age surface (numbers) -->
-
-<!-- ```{r} -->
-<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
-<!-- ``` -->
-
-<!-- > Note: model fits to composition - this plot is included for illustration -->
-
-RV Survey
-================================================================================
-
-Row
 --------------------------------------------------------------------------------
@@
 
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
+plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "RV survey index",
               showlegend = TRUE, type = "log")
@@
 
-### Standardized log residuals
+### Index at age
 
@@
 
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    zero = fit$tmb.data$ind_zero,
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d <- d[d$wt == 1, ]
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
+d <- frame_d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
                                pred = exp(fit$rep$Elog_RV_index),
@@
 d$pred[d$wt == 0] <- NA
+d$cohort <- frame_d$cohort <- factor(d$year - d$age)
+d$age <- frame_d$age <- factor(d$age)
+
 facet_obs_pred(d, by = "age",
@@
                nrows = 6, ylab = "Mean numbers per tow",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
+               obs_col = obs_col, pred_col = pred_col, type = "log")
 
@@
 
-### Residuals by year, cohort, age and expected values
+### Index at age (frame)
+
+```{r}
+plot_obs_pred(data = frame_d, showlegend = TRUE, 
+              type = "log", frame = "age",
+              obs_col = obs_col, pred_col = pred_col,
+              ylab = "Mean numbers per tow")
+```
+
+> Y-axes in log scale
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Residuals
 
@@
 d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
+plot_resid(d, col = obs_col, line_col = pred_col)
 
@@
 
-SN Survey
+### Residuals (bubble)
+
+```{r}
+
+d <- tidy_mats(list(res = fit$rep$RV_resid,
+                    zero = fit$tmb.data$ind_zero,
+                    wt = fit$tmb.data$est_wt),
+               ages = ages, years = years)
+d <- d[d$wt == 1, ]
+bubble_plot(d, pos_col = pos_col, neg_col = neg_col, na_col = obs_col)
+
+```
+
+SN Survey {data-navmenu="Surveys"}
 ================================================================================
 
-Row
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
 
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
+plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "Sentinel survey index",
               showlegend = TRUE, type = "log")
@@
 
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$log_SN_GN_resid,
-              ages = fit$tmb.data$SN_ages,
-              years = fit$tmb.data$SN_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index at age
+### Index at age
 
@@
                nrows = 4, ylab = "Sentinel survey index",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
+               obs_col = obs_col, pred_col = pred_col, type = "log")
 
@@
 
-### Residuals by year, cohort, age and expected values
+### Index at age (frame)
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              type = "log", frame = "age",
+              obs_col = obs_col, pred_col = pred_col,
+              ylab = "Sentinel survey index")
+```
+
+> Y-axes in log scale
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Residuals
 
@@
 
-SS Survey
+### Residuals (bubble)
+
+```{r}
+
+d <- tidy_mat(fit$rep$log_SN_GN_resid,
+              ages = fit$tmb.data$SN_ages,
+              years = fit$tmb.data$SN_years, value = "res")
+bubble_plot(d, pos_col = pos_col, neg_col = neg_col)
+
+```
+
+SS Survey {data-navmenu="Surveys"}
 ================================================================================
 
-Row
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
                 pred = fit$rep$ESSbms)
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
+plot_obs_pred(d, obs_col = obs_col, pred_col = pred_col, ylab = "Biomass (t)",
               showlegend = TRUE)
@@
 
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
-              ages = fit$tmb.data$SS_ages,
-              years = fit$tmb.data$SS_age_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Smith Sound trawl numbers at age
+### Numbers at age
 
@@
                years = fit$tmb.data$SS_age_years)
+
 facet_obs_pred(d, by = "age",
@@
                nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
-               obs_col = cols[2], pred_col = cols[1])
+               obs_col = obs_col, pred_col = pred_col)
 
@@
 
-### Residuals by year, cohort, age and expected values
+### Numbers at age (frame)
+
+```{r}
+plot_obs_pred(data = d, showlegend = TRUE, 
+              frame = "age",
+              obs_col = obs_col, pred_col = pred_col,
+              ylab = "Numbers caught")
+```
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Residuals
 
@@
 
-Tagging {data-orientation=rows}
-================================================================================
+### Residuals (bubble)
 
@@
 
-## Prep tag data and fits
-tag_dat <- data.frame(
-  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
-  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
-  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
-  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
-  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
-  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
-  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
-  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
-  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
-             rep("post 1996", length(fit$tmb.data$tag_exp)))
-)
-tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
-tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
-                      "age:", tag_dat$age, "</br>",
-                      "delay:", tag_dat$delay, "</br>")
-
-Udev <- data.frame(
-  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
-  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
-)
-
-Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
-names(Fay)[names(Fay) == "year"] <- "rel_year"
-
-tag_dat <- merge(tag_dat, Udev, by = "exp")
-tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
-tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
-
-## Share data across multiple plots
-d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)
+d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
+              ages = fit$tmb.data$SS_ages,
+              years = fit$tmb.data$SS_age_years, value = "res")
+bubble_plot(d, pos_col = pos_col, neg_col = neg_col)
 
@@
 
-Row {data-height=700 .tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted tag recaptures by experimemnt
-
-```{r}
-
-y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
-y_tickvals <- sqrt(y_ticktext)
-
-rel_plot <- base %>%
-  filter(rel_year == cap_year) %>%
-  group_by(exp) %>%
-  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N released",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
-y_tickvals <- sqrt(y_ticktext)
-
-cap_plot <- base %>%
-  group_by(exp, rel_year, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
-            period = unique(period)) %>%
-  group_by(exp) %>%
-  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
-              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
-            alpha = 0.7, showlegend = FALSE, symbol = NULL,
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y2",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
-        nrows = 2, heights = c(0.3, 0.7))
-
-```
-
-> Y-axis is in square root scale
-
-<!-- ### Animation -->
-
-<!-- ```{r} -->
-
-<!-- sum_d <- tag_dat %>% -->
-<!--   group_by(exp, cap_year) %>% -->
-<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->
-
-<!-- sum_d %>%  -->
-<!--   group_by(exp) %>%  -->
-<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
-<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
-<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
-<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
-<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
-<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
-<!--   layout(yaxis = list(title = "N recaptured", -->
-<!--                       tickvals = y_tickvals, -->
-<!--                       ticktext = y_ticktext, -->
-<!--                       range = range(y_tickvals)), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
+Catchability {data-navmenu="Surveys"}
+================================================================================
 
 Row
---------------------------------------------------------------------------------
-
-### Tag reporting rate
-
-```{r}
-
-cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
-rr <- data.frame(year = fit$tmb.data$SRR_years,
-                 est = fit$tmb.data$log_SRR_est,
-                 sd = sqrt(diag(cov_mat)))
-rr$lwr <- rr$est - 1.96 * rr$sd
-rr$upr <- rr$est + 1.96 * rr$sd
-rr <- trans_est(rr)
-rr$model <- "Random Walk"
-
-ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
-                      est = est$log_SRR, sd = sd$log_SRR)
-ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
-ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
-ncam_rr <- trans_est(ncam_rr)
-ncam_rr$model <- "NCAM"
-
-rr <- rbind(rr, ncam_rr)
-
-plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)
-
-```
-
-### Initial F deviations for tagging experiments
-
-```{r}
-
-base %>%
-  distinct(rel_year, Udev, exp, period) %>%
-  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
-              showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "Initial F deviation"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Observed vs. predicted recaptures
-
-```{r}
-
-x_ticktext <- c(5, pretty(tag_dat$pred, 10))
-y_ticktext <- c(5, pretty(tag_dat$obs, 10))
-x_tickvals <- sqrt(x_ticktext)
-y_tickvals <- sqrt(y_ticktext)
-
-base %>%
-  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
-              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       line = list(width = 1)),
-         xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Observed number of recaptures",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext))
-
-```
-
-> Axes are in square root scale
-
-### Standardized residuals vs. predicted recaptures
-
-```{r}
-
-y_ticktext <- c(5, pretty(tag_dat$res, 10))
-y_tickvals <- sqrt(x_ticktext)
-
-base %>%
-  add_markers( x = ~sqrt(pred), y = ~res,
-               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Standardized residuals"))
-
-```
-
-> X-axis is in square root scale
-
-Catchability
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Fishery selectivity
-
-```{r}
-
-S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
-plot_surface(S, ages = ages, years = years, name = "S")
-
-```
-
-Row {.tabset}
 --------------------------------------------------------------------------------
@@
 ```
+
+Row {.tabset}
+--------------------------------------------------------------------------------
 
@@
 ```
-
-Trends
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r}
-
-## Note: biomass is in tonnes and numbers are in thousands
-bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
-             data.frame(metric = "Ages 2+", tabs$log_biomass))
-abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
-             data.frame(metric = "Ages 2+", tabs$log_abundance))
-
-bms_plot <- trans_est(bms) %>%
-  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
-             annotate = TRUE)
-
-abn_plot <- trans_est(abn, scale = 0.001) %>%
-  plot_trend(ylab = "Abundance", group = "metric", col = cols,
-             annotate = TRUE)
-
-blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
-  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
-  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
-           color = I("black"), textposition = "right", hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Abundance at age
-
-```{r}
-plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
-```
-
-### Biomass at age
-
-```{r}
-plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
-```
-
-### SSB at age
-
-```{r}
-plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r}
-
-f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
-z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
-m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
-
-z_plot <- trans_est(z) %>%
-  plot_trend(ylab = "Average Z", group = "metric",
-             annotate = TRUE, col = cols)
-
-m_plot <- trans_est(m) %>%
-  plot_trend(ylab = "Average M", group = "metric",
-             annotate = TRUE, col = cols)
-
-f_plot <- trans_est(f) %>%
-  plot_trend(ylab = "Average F", group = "metric",
-             annotate = TRUE, col = cols)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Z at age
-
-```{r}
-plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
-```
-
-### M at age
-
-```{r}
-plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
-```
-
-### F at age
-
-```{r}
-plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
-```
-
-### Z, M and F
-
-```{r}
-plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
-                 ages = ages, years = years)
-```
-
-> Z = purples, M = blues, F = reds
-
-### Process error
-
-```{r}
-
-pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
-             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
-  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
-              color = I("black"), opacity = 0.8)
-
-```
-
-### Process error (bubble plot)
-
-```{r}
-pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = 2:8, value = "res")
-bubble_plot(pe)
-```
-
-Retro {`r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Comps {`r if (is.null(tmp$comp)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(tmp$comp)}
-
-abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
-  plot_comps(ylab = "Recruitment (age 2)", col = cols)
-
-bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
-  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)
-
-blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
-  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(tmp$comp)}
-
-z_plot <- trans_est(comp_tabs$log_aveZ) %>%
-  plot_comps(ylab = "Average Z (ages 5+)", col = cols)
-
-m_plot <- trans_est(comp_tabs$log_aveM) %>%
-  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)
-
-f_plot <- trans_est(comp_tabs$log_aveF) %>%
-  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Productivity
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Recruitment
-
-```{r}
-trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1])
-```
-
-### SSB
-
-```{r}
-trans_est(tabs$log_ssb) %>%
-  plot_trend(ylab = "SSB (t)", col = cols[1])
-```
-
-### RPS
-
-```{r}
-trans_est(tabs$log_rps) %>%
-  filter(year >= min_year + min(ages)) %>% 
-  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
-  layout(xaxis = list(range = c(min_year, max_year + 1)))
-```
-
-### SSB x Recruitment
-
-```{r}
-
-ssb <- trans_est(tabs$log_ssb)
-rec <- trans_est(tabs$log_recruitment, scale = 0.001)
-rps <- trans_est(tabs$log_rps)
-names(ssb) <- paste0("ssb_", names(ssb))
-names(rec) <- paste0("rec_", names(rec))
-names(rps) <- paste0("rps_", names(rps))
-ssb$year <- ssb$ssb_year + min(ages)
-rec$year <- rec$rec_year
-rps$year <- rps$rps_year
-sr <- merge(ssb, rec, by = "year")
-sr <- merge(sr, rps, by = "year")
-sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)
-
-plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
-                 col = cols[1])
-
-```
-
-### SSB x RPS
-
-```{r}
-p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
-                 col = cols[1])
-p
-```
-
-### log(SSB) x log(RPS)
-
-```{r}
-p %>% layout(yaxis = list(type = "log"),
-             xaxis = list(type = "log"))
-```
-
-```{r}
-
-message("Bits of the projection code is currently not working. Specifically the plots that require the frame argument in plotly.")
-
-```
-
-<!-- Assumptions  {data-navmenu="Projections" .storyboard} -->
-<!-- ================================================================================ -->
-
-<!-- ### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers -->
-
-<!-- ```{r} -->
-
-<!-- Cp <- fit$tmb.data$Cp -->
-<!-- Cp_string <- text_c(Cp) -->
-<!-- Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1)) -->
-<!-- Lp_string <- paste(text_c(Lp), "t") -->
-
-<!-- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>% -->
-<!--   filter(year >= max(years) - 4) %>% -->
-<!--   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   filter(year == max(years)) %>% -->
-<!--   add_lines(showlegend = FALSE, line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age")) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - The following catch multipliers were applied: `r #Cp_string` -->
-<!-- - This translates to `r #Lp_string` of catch in the projection years -->
-
-<!-- ### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994 -->
-
-<!-- ```{r} -->
-
-<!-- if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic") -->
-
-<!-- aveM <- tabs$log_aveM -->
-<!-- aveM <- exp(mean(aveM$est[!aveM$year %in% 1991:1994])) -->
-
-<!-- trans_est(tabs$log_aveM) %>% -->
-<!--   plot_trend(ylab = "Average M", col = cols[1]) %>% -->
-<!--   add_segments(x = 1995, xend = max(years) + 3, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_segments(x = 1990, xend = 1995, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), -->
-<!--                line = list(width = 1, dash = "dot")) %>% -->
-<!--   add_segments(x = min(years), xend = 1990, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_text(x = max(years) + 0.2, y = aveM, text = "long-term mean", textposition = "bottom right", -->
-<!--            color = I("black")) %>% -->
-<!--   layout(xaxis = list(range = c(min(years), max(years) + 3))) -->
-
-<!-- ``` -->
-
-<!-- ### Assume a relationship between recruits per spawner and number of spawners to project recruits in the future -->
-
-<!-- ```{r} -->
-
-<!-- if (fit$tmb.data$Rp_mode != 3) stop("TODO: make the recruitment mode documentation dynamic") -->
-
-<!-- rps_tab <- tabs$log_rps -->
-<!-- names(rps_tab) <- c("year", "rps", "rps_sd", "rps_lwr", "rps_upr") -->
-<!-- rps_tab <- rps_tab[rps_tab$year >= 1985, ] -->
-<!-- rps_tab$year <- rps_tab$year - 2 # scale back two year to allign with spawners that produced the recruits -->
-<!-- ssa_tab <- tabs$log_ssa -->
-<!-- names(ssa_tab) <- c("year", "ssa", "ssa_sd", "ssa_lwr", "ssa_upr") -->
-<!-- d <- merge(ssa_tab, rps_tab, by = "year") -->
-<!-- rec_tab <- tabs$log_recruitment -->
-<!-- names(rec_tab) <- c("year", "rec", "rec_sd", "rec_lwr", "rec_upr") -->
-<!-- rec_tab <- rec_tab[rec_tab$year >= 1985, ] -->
-<!-- rec_tab$year <- rec_tab$year - 2 # scale back two years -->
-<!-- d <- merge(d, rec_tab, by = "year") -->
-<!-- d$excluded <- d$year %in% 1991:1994 -->
-
-<!-- sub_d <- d[!d$year %in% 1991:1994, ] -->
-<!-- lin_rel <- lm(rps ~ ssa, data = sub_d) -->
-<!-- # coefficients(lin_rel) -->
-
-<!-- preds <- data.frame(year = NA, rps_sd = NA, ssa_sd = NA, ssa = seq(min(d$ssa), max(d$ssa), length.out = 100)) -->
-<!-- preds$fits <- predict(lin_rel, newdata = preds) -->
-<!-- preds$excluded <- FALSE -->
-
-<!-- plot_ly(data = d, x = ~ssa, y = ~rps, text = ~year, -->
-<!--         type = "scatter", mode = "markers", name = "est", -->
-<!--         color = ~excluded, colors = c(cols[2], "grey"), -->
-<!--         error_y = list(array = ~ qnorm(0.975) * rps_sd, -->
-<!--                        width = 0, thickness = 1), -->
-<!--         error_x = list(array = ~ qnorm(0.975) * ssa_sd, -->
-<!--                        width = 0, thickness = 1)) %>% -->
-<!--   add_text(text = ~year, textfont = list(color = toRGB("grey50")), textposition = "top right", -->
-<!--            error_y = list(visible = FALSE), -->
-<!--            error_x = list(visible = FALSE)) %>% -->
-<!--   layout(title = "Spawners vs Recruits / Spawner", -->
-<!--          xaxis = list(title = "log(Spawners)"), -->
-<!--          yaxis = list(title = "log(Recruits / Spawner)", zerolinecolor = toRGB("lightgrey")), -->
-<!--          showlegend = FALSE) %>% -->
-<!--   add_lines(data = preds, x = ~ssa, y = ~fits, name = "fit", color = I(cols[1])) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - Grey points indicate years excluded from the regressionn -->
-
-<!-- ### Assume averages weights and proportions mature from the last three years -->
-
-<!-- ```{r} -->
-
-<!-- mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(mat = mean(mat), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Proportion mature")) -->
-
-<!-- wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols), -->
-<!--           showlegend = FALSE) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(weight = mean(weight), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Weight (kg)")) -->
-
-<!-- subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- ### Catch multiplier projections for SSB / Blim from the last RAP -->
-
-<!-- ```{r, eval = !is.null(tmp$last)} -->
-
-<!-- d <- trans_est(last_tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- d <- d[d$year <= max(years) + 1, ] -->
-
-<!-- past_d <- trans_est(last_tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- current_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s)) -->
-<!-- current_d <- do.call(rbind, current_d) -->
-<!-- current_d$model <- "current RAP" -->
-<!-- d$model <- "last RAP" -->
-<!-- d <- rbind(d, current_d) -->
-
-<!-- plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model", -->
-<!--           vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Results {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~ -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_ratio) -->
-<!-- d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100 -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- xlim <- c(max(years) + 0.9, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%", -->
-<!-- #           vline = max(years) + 1, col = cols[1]) %>% -->
-<!-- #   layout(yaxis = list(tickvals = seq(-200, 200, by = 25))) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- ### SSB / Blim -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-
-<!-- past_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(years) - 2.1, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", -->
-<!-- #           vline = max(years) + 1, hline = 1, col = cols[1]) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_ratio -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-<!-- ### Prob(SSB~yp~ &ge; B~lim~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_lrp -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-Inputs {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
-```
-
-### Catch at age
-
-```{r}
-kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
-```
-
-<!-- ### Monthly landings -->
-
-<!-- ```{r} -->
-<!-- d <- inputs$catch.month -->
-<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
-<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
-<!-- ``` -->
-
-### RV index
-
-```{r}
-d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
-```
-
-### SN index
-
-```{r}
-d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
-mat <- d[, as.character(3:10)]
-rownames(mat) <- d$year
-kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
-```
-
-### SS index
-
-```{r}
-d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
-names(d) <- c("Year", "Month", "Biomass", "SE^2^")
-kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
-```
-
-### SS age composition
-
-```{r}
-kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
-```
-
-### Weight at age
-
-```{r}
-d <- inputs$pwts
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "weight")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
-```
-
-### Maturity at age
-
-```{r}
-d <- inputs$mats
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "mat")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div.
-2J+3KL from DFO RV autumn bottom trawl surveys.
-Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
-```
-
-### Tag summary
-
-```{r}
-d <- inputs$mtag1 %>% 
-  group_by(rel_year, exp) %>% 
-  summarise(NAFO = unique(div),
-            Nrel = sum(unique(released)), 
-            Ncap = sum(recaptured))
-names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
-kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
-```
-
-Settings {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Catch bounds
-
-```{r}
-d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
-kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
-```
-
-### Baseline M
-
-```{r}
-m <- fit$tmb.data$M_matrix
-rownames(m) <- ages
-colnames(m) <- years
-kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
-```
-
-Output {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Parameter estimates
-
-```{r}
-
-par_tab <- tidy_pars(fit)
-rownames(par_tab) <- NULL
-options(knitr.kable.NA = '')
-kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Parameter estimates (retro)
-
-```{r}
-
-par_tabs <- lapply(names(fit$retro), function(nm) {
-  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
-  d$id <- seq(nrow(d))
-  d
-})
-par_tabs <- do.call(rbind, par_tabs)
-final_par_tab <- data.frame(tyear = max(years), par_tab)
-final_par_tab$id <- seq(nrow(final_par_tab))
-par_tabs <- rbind(par_tabs, final_par_tab)
-par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-par_tabs$Quantity <- as.character(par_tabs$Quantity)
-par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
-par_tabs$Symbol <- as.character(par_tabs$Symbol)
-par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-rownames(par_tabs) <- NULL
-names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"
-
-kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Stock size
-
-```{r}
-
-tab_name <- c("log_abundance" = "Abundance (millions)", 
-              "log_recruitment" = "Recruits (age 2; millions)", 
-              "log_biomass" = "Biomass (kt)", 
-              "log_ssa" = "Spawners (millions)", 
-              "log_ssb" = "SSB (kt)", 
-              "log_ssb_lrp" = "SSB/Blim n(%)")
-
-nms <- names(tab_name)
-status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
-names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
-  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  status_tab <- merge(status_tab, tab, by = "Year")
-}
-
-knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Rates
-
-```{r}
-
-tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
-              "log_aveM" = "Average M (ages 5+)", 
-              "log_aveF" = "Average F (ages 5+)")
-
-nms <- names(tab_name)
-rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
-names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  rate_tab <- merge(rate_tab, tab, by = "Year")
-}
-
-knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Abundance-at-age
-
-```{r}
-N_matrix <- fit$rep$N_matrix
-dimnames(N_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Mature abundance-at-age
-
-```{r}
-ssa_matrix <- fit$rep$ssa_matrix
-dimnames(ssa_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Biomass-at-age
-
-```{r}
-B_matrix <- fit$rep$biomass_matrix
-dimnames(B_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Mature biomass-at-age
-
-```{r}
-ssb_matrix <- fit$rep$ssb_matrix
-dimnames(ssb_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Z-at-age
-
-```{r}
-Z_matrix <- fit$rep$Z_matrix
-dimnames(Z_matrix) <- list(age = ages, year = years)
-knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
-```
-
-### M-at-age
-
-```{r}
-M_matrix <- fit$rep$Mpe_matrix
-dimnames(M_matrix) <- list(age = ages, year = years)
-M_matrix <- M_matrix[as.character(2:8), ]
-rownames(M_matrix) <- c(as.character(2:7), "8+")
-knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
-```
-
-### F-at-age
-
-```{r}
-F_matrix <- fit$rep$F_matrix
-dimnames(F_matrix) <- list(age = ages, year = years)
-knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
-```

### 2020-03-14 inst/rmd/tables.Rmd
Revamped NCAM explorer

--- 2019-09-10 inst/rmd/vis_fit.Rmd
@@
----
-title: "NCAM Explorer"
-output: flexdashboard::flex_dashboard
-always_allow_html: yes
----
-
-<style>
-/* list */
-li {
-line-height: 30px;
-}
-</style>
-
-```{r setup, echo=FALSE}
-library(plotly)
-library(Matrix)
-library(knitr)
-library(data.table)
-library(dplyr)
-library(webshot)
-knitr::opts_knit$set(root.dir = tmp$output_dir)
-knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
-```
-
-```{r data, include=FALSE}
-
-cols <-  viridis::viridis(4)[c(2, 3, 1, 4)]
-cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
-fit <- tmp$fit # temp object from vis_fit function
-if (!is.null(tmp$comp)) { comp <- tmp$comp }
-if (!is.null(tmp$last)) { last <- tmp$last }
-est <- as.list(fit$sd.rep, "Est")
-sd <- as.list(fit$sd.rep, "Std")
-tabs <- tidy_model(fit)
-if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
-if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
-if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
-min_year <- min(fit$tmb.data$years)
-max_year <- max(fit$tmb.data$years)
-ages <- fit$tmb.data$ages
-years <- fit$tmb.data$years
-years_plus1 <- fit$tmb.data$years_plus1
-
-## helper functions for nice Rmarkdown text
-text_range <- function(x) paste(range(x), collapse = "-")
-text_c <- function(x) {
-  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
-}
-text_units <- function(x) {
-  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
-  if (x >= 1000) return(paste0(round(x / 1000), "k"))
-  if (x < 1000) return(as.character(round(x)))
-}
-
-```
-
-Background
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### NCAM
-
-<h1>NCAM: A state-space stock assessment model for Northern cod</h1>
-
-```{r}
-
-ax <- list(
-  title = "",
-  zeroline = FALSE,
-  showline = FALSE,
-  showticklabels = FALSE,
-  showgrid = FALSE
-)
-
-xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
-plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
-        fillcolor = 'rgba(49, 104, 142, 0.3)',
-        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
-  add_trace() %>%
-  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
-         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
-  plotly::config(displayModeBar = FALSE)
-
-```
-
-### General
-
-- An *integrated* *state-space* population dynamics model
-- *Integrated*
-    - Uses as much information as possible
-- *State-space*
-    - Includes process and observation equations + error
-    - Population processes are informed by observations from multiple sources
-    - Process errors are necessary for realistic stochastic projections
-
-### Data inputs
-
-- Reported landings (`r text_range(years)`)
-- Catch-at-age information (`r text_range(years)`)
-- Reported landings by month (`r text_range(years)`)
-- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
-- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
-- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
-- Tagging data (`r text_range(years)`)
-
-### Biological inputs
-
-- RV survey data were used fit cohort-based models of:
-   - Proportion mature at age
-   - Beginning of year and mid-year stock weights at age
-- Baseline natural mortality (M) values
-
-```{r}
-
-plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
-  add_surface() %>%
-  colorbar(title = "Mshift") %>%
-  layout(scene = list(
-    xaxis = list(title = "Year"),
-    yaxis = list(title = "Age"),
-    zaxis = list(title = "Mshift")
-  ))
-
-```
-
-### Partial catches
-
-- User sets lower and upper bounds on total catch weight
-- Bounds are used via a censored likelihood
-
-```{r}
-
-plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
-  add_lines(name = "Lower", color = I(cols[1])) %>%
-  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
-  add_lines(y = rep(1, length(years)), color = I("black"),
-            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Change in Q
-
-- A shift in Q was assumed for the offshore RV survey
-- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
-- Q for the inshore sentinel survey was assumed to follow a random walk
-    - Done to account for potential changes in migration patterns
-
-### Tagging data
-
-- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
-- Applies adjustments for tagging mortality, tag loss and reporting rates
-- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
-- The Baranov catch equation was applied to age-based tag survivors
-- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock
-
-### References
-
-Cadigan, N.
-2015.
-A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates.
-Canadian Journal of Fisheries and Aquatic Sciences 73 (2).
-NRC Research Press: 296-308.
-
-Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.
-
-Catch
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-
-plot_landings(fit)
-
-```
-
-> Y-axes in log scale
-
-### Numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = t(inputs$catch.num[as.character(years), as.character(ages)]),
-                    pred = fit$rep$pred_catch),
-               ages = ages, years = years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age", type = "log",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Predictions by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Numbers")
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Predictions by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = 1, end = 0.8),
-              ylab = "Numbers")
-```
-
-> Y-axes in log scale; also note that model fits to composition - this plot is included for illustration
-
-### Numbers at age surface
-
-```{r}
-plot_surface(mat = t(inputs$catch.num[as.character(years), as.character(ages)]), 
-             ages = ages, years = years, name = "N")
-```
-
-> Note: model fits to composition - this plot is included for illustration
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Proportions at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
-                    pred = fit$rep$pred_prop_catch),
-               ages = ages, years = years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "linear", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Proportion", stack_group = TRUE)
-```
-
-> Note: values are stacked
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "linear", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = 1, end = 0.8),
-              ylab = "Proportion", stack_group = TRUE)
-```
-
-> Note: values are stacked
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$CA_resid,
-                    fit = fit$rep$pred_crl),
-               ages = ages[-length(ages)], years = years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-> based on continuation ratio logits
-
-### Proportions at age surface
-
-```{r}
-plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
-```
-
-RV Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### RV survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
-pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
-obs[obs == 0] <- NA
-pred[pred == 0] <- NA
-d <- data.frame(year = years, obs = obs, pred = pred)
-
-# ## RV biomass
-# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# d <- data.frame(year = years, obs = obs, pred = pred)
-# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    zero = fit$tmb.data$ind_zero,
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d <- d[d$wt == 1, ]
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
-                    pred = exp(fit$rep$Elog_RV_index),
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d$obs[d$wt == 0] <- NA
-d$pred[d$wt == 0] <- NA
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Mean numbers per tow",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Mean numbers per tow")
-```
-
-> Y-axes in log scale
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE,
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = -1, end = 0.8),
-              ylab = "Mean numbers per tow")
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
-                    fit = fit$rep$Elog_RV_index),
-               ages = ages, years = years)
-d <- d[d$keep, ]
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SN Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$SN_GN_index)
-pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
-d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$log_SN_GN_resid,
-              ages = fit$tmb.data$SN_ages,
-              years = fit$tmb.data$SN_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
-                    pred = exp(fit$rep$Elog_SN_GN_index)),
-               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
-d$cohort <- factor(d$year - d$age)
-d$age <- factor(d$age)
-
-facet_obs_pred(d, by = "age",
-               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
-               nrows = 4, ylab = "Sentinel survey index",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Fits by age
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE, 
-              type = "log", group = "age",
-              group_cols = viridis::viridis(nlevels(d$age), 
-                                            direction = 1, end = 0.8),
-              ylab = "Sentinel survey index")
-```
-
-> Y-axes in log scale
-
-### Fits by cohort
-
-```{r}
-plot_obs_pred(data = d, showlegend = TRUE,
-              type = "log", group = "cohort",
-              group_cols = viridis::viridis(nlevels(d$cohort), 
-                                            direction = -1, end = 0.8),
-              ylab = "Sentinel survey index")
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
-                    fit = fit$rep$Elog_SN_GN_index),
-               ages = fit$tmb.data$SN_ages,
-               years = fit$tmb.data$SN_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SS Survey {data-navmenu="Surveys"}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Smith sound acoustic biomass estimates
-
-```{r}
-
-d <- data.frame(year = fit$tmb.data$SS_bms_years,
-                obs = fit$tmb.data$SSbms,
-                pred = fit$rep$ESSbms)
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
-              showlegend = TRUE)
-
-```
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
-              ages = fit$tmb.data$SS_ages,
-              years = fit$tmb.data$SS_age_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Smith Sound trawl numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
-                    pred = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
-               obs_col = cols[2], pred_col = cols[1])
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
-                    fit = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-Tagging {data-orientation=rows}
-================================================================================
-
-```{r}
-
-## Prep tag data and fits
-tag_dat <- data.frame(
-  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
-  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
-  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
-  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
-  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
-  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
-  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
-  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
-  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
-             rep("post 1996", length(fit$tmb.data$tag_exp)))
-)
-tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
-tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
-                      "age:", tag_dat$age, "</br>",
-                      "delay:", tag_dat$delay, "</br>")
-
-Udev <- data.frame(
-  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
-  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
-)
-
-Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
-names(Fay)[names(Fay) == "year"] <- "rel_year"
-
-tag_dat <- merge(tag_dat, Udev, by = "exp")
-tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
-tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
-
-## Share data across multiple plots
-d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% 
-  group_by(exp) %>% 
-  highlight(off = "plotly_relayout")
-
-```
-
-Row {data-height=700 .tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted tag recaptures by experimemnt
-
-```{r}
-
-y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
-y_tickvals <- sqrt(y_ticktext)
-
-rel_plot <- base %>%
-  filter(rel_year == cap_year) %>%
-  group_by(exp) %>%
-  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~jitter(rel_year, factor = 0.5), 
-              y = ~sqrt(tot_rel), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N released",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
-y_tickvals <- sqrt(y_ticktext)
-
-cap_plot <- base %>%
-  group_by(exp, rel_year, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
-            period = unique(period)) %>%
-  group_by(exp) %>%
-  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
-              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
-            alpha = 0.7, showlegend = FALSE, symbol = NULL,
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y2",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
-        nrows = 2, heights = c(0.3, 0.7))
-
-```
-
-> Y-axis is in square root scale
-
-### Animation
-
-```{r}
-
-sum_d <- tag_dat %>%
-  group_by(exp, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs))
-
-sum_d %>%
-  group_by(exp) %>%
-  plot_ly(x = ~cap_year, frame = ~exp, hoverinfo = "x+y+text",
-          showlegend = FALSE) %>%
-  add_markers(y = ~sqrt(sum_obs), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(y = ~sqrt(sum_pred), color = I(cols[2]),
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>%
-  layout(yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Tag reporting rate
-
-```{r}
-
-cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
-rr <- data.frame(year = fit$tmb.data$SRR_years,
-                 est = fit$tmb.data$log_SRR_est,
-                 sd = sqrt(diag(cov_mat)))
-rr$lwr <- rr$est - 1.96 * rr$sd
-rr$upr <- rr$est + 1.96 * rr$sd
-rr <- trans_est(rr)
-rr$model <- "Random Walk"
-
-ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
-                      est = est$log_SRR, sd = sd$log_SRR)
-ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
-ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
-ncam_rr <- trans_est(ncam_rr)
-ncam_rr$model <- "NCAM"
-
-rr <- rbind(rr, ncam_rr)
-
-plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)
-
-```
-
-### Initial F deviations for tagging experiments
-
-```{r}
-
-base %>%
-  filter(rel_year == cap_year) %>%
-  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
-              showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "Initial F deviation"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Observed vs. predicted recaptures
-
-```{r}
-
-x_ticktext <- c(5, pretty(tag_dat$pred, 10))
-y_ticktext <- c(5, pretty(tag_dat$obs, 10))
-x_tickvals <- sqrt(x_ticktext)
-y_tickvals <- sqrt(y_ticktext)
-
-base %>%
-  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
-              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       line = list(width = 1)),
-         xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Observed number of recaptures",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext))
-
-```
-
-> Axes are in square root scale
-
-### Standardized residuals vs. predicted recaptures
-
-```{r}
-
-y_ticktext <- c(5, pretty(tag_dat$res, 10))
-y_tickvals <- sqrt(x_ticktext)
-
-base %>%
-  add_markers( x = ~sqrt(pred), y = ~res,
-               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Standardized residuals"))
-
-```
-
-> X-axis is in square root scale
-
-Catchability
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Fishery selectivity
-
-```{r}
-
-S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
-plot_surface(S, ages = ages, years = years, name = "S")
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey catchability
-
-```{r}
-
-q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
-plot_surface(q, ages = ages, years = years, name = "q")
-
-```
-
-### Proportion of the stock in Smith Sound
-
-```{r}
-p <- matrix(0, nrow = length(ages), ncol = length(years),
-            dimnames = list(ages, years))
-ss_p <- fit$rep$SSD
-ss_years <- fit$tmb.data$SS_years
-ss_ages <- fit$tmb.data$SS_ages
-rownames(ss_p) <- ss_ages
-colnames(ss_p) <- ss_years
-p[as.character(ss_ages), as.character(ss_years)] <- ss_p
-plot_surface(p, ages = ages, years = years, name = "p")
-
-```
-
-### Sentinel gillnet survey catchability
-
-```{r}
-
-log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
-log_qa <- fit$rep$log_SN_GN_Qa
-q <- exp(log_qy + log_qa)
-plot_surface(q, ages = fit$tmb.data$SN_ages,
-             years = fit$tmb.data$SN_years,
-             name = "q")
-
-```
-
-Trends
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r}
-
-## Note: biomass is in tonnes and numbers are in thousands
-bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
-             data.frame(metric = "Ages 2+", tabs$log_biomass))
-abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
-             data.frame(metric = "Ages 2+", tabs$log_abundance))
-
-bms_plot <- trans_est(bms) %>%
-  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
-             annotate = TRUE)
-
-abn_plot <- trans_est(abn, scale = 0.001) %>%
-  plot_trend(ylab = "Abundance", group = "metric", col = cols,
-             annotate = TRUE)
-
-blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
-  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
-  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
-           color = I("black"), textposition = "right", hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Abundance at age
-
-```{r}
-plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
-```
-
-### Biomass at age
-
-```{r}
-plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
-```
-
-### SSB at age
-
-```{r}
-plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
-```
-
-<!-- ### Exploitable biomass at age -->
-
-<!-- ```{r} -->
-<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
-<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
-<!-- E <- B * S -->
-<!-- plot_surface(E, ages = ages, years = years, name = "Exploitable biomass") -->
-<!-- ``` -->
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r}
-
-f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
-z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
-m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
-
-z_plot <- trans_est(z) %>%
-  plot_trend(ylab = "Average Z", group = "metric",
-             annotate = TRUE, col = cols)
-
-m_plot <- trans_est(m) %>%
-  plot_trend(ylab = "Average M", group = "metric",
-             annotate = TRUE, col = cols)
-
-f_plot <- trans_est(f) %>%
-  plot_trend(ylab = "Average F", group = "metric",
-             annotate = TRUE, col = cols)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Z at age
-
-```{r}
-plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
-```
-
-### M at age
-
-```{r}
-plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
-```
-
-### F at age
-
-```{r}
-plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
-```
-
-### Z, M and F
-
-```{r}
-plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
-                 ages = ages, years = years)
-```
-
-> Z = purples, M = blues, F = reds
-
-### Process error
-
-```{r}
-
-pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], 
-             ages = fit$tmb.data$pe_ages, name = "pe", 
-             col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
-  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
-              color = I("black"), opacity = 0.8)
-
-```
-
-### Process error (bubble plot)
-
-```{r}
-pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = fit$tmb.data$pe_ages, value = "res")
-bubble_plot(pe)
-```
-
-Retro {`r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Mohn's rho
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_mohns(col = cols[1], ylab = "SSB")
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_mohns(ylab = "Recruitment (age 2)", col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_mohns(ylab = "SSB/Blim", col = cols[1])
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Relative difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
-
-### Difference
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- retro_tabs$log_ssb %>%
-  plot_retro_deltas(in_col = cols[1], ylab = "log(SSB)")
-
-abn_plot <- retro_tabs$log_recruitment %>%
-  plot_retro_deltas(ylab = "log(Recruitment)", in_col = cols[1])
-
-blim_plot <- retro_tabs$log_ssb_lrp %>%
-  plot_retro_deltas(ylab = "log(SSB/Blim)", in_col = cols[1])
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Mohn's rho
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_mohns(ylab = "Average Z (ages 5+)", col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_mohns(ylab = "Average M (ages 5+)", col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_mohns(ylab = "Average F (ages 5+)", col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Relative difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
-
-### Difference
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- retro_tabs$log_aveZ %>%
-  plot_retro_deltas(ylab = "log(Average Z)", in_col = cols[1])
-
-m_plot <- retro_tabs$log_aveM %>%
-  plot_retro_deltas(ylab = "log(Average M)", in_col = cols[1])
-
-f_plot <- retro_tabs$log_aveF %>%
-  plot_retro_deltas(ylab = "log(Average F)", in_col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
-  layout(annotations = left_label("Difference", offset = 50),
-         margin = list(l = 75))
-
-```
-
-> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
-
-Comps {`r if (is.null(tmp$comp)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(tmp$comp)}
-
-abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
-  plot_comps(ylab = "Recruitment (age 2)", col = cols)
-
-bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
-  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)
-
-blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
-  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(tmp$comp)}
-
-z_plot <- trans_est(comp_tabs$log_aveZ) %>%
-  plot_comps(ylab = "Average Z (ages 5+)", col = cols)
-
-m_plot <- trans_est(comp_tabs$log_aveM) %>%
-  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)
-
-f_plot <- trans_est(comp_tabs$log_aveF) %>%
-  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Productivity
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Recruitment
-
-```{r}
-trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1])
-```
-
-### SSB
-
-```{r}
-trans_est(tabs$log_ssb) %>%
-  plot_trend(ylab = "SSB (t)", col = cols[1])
-```
-
-### RPS
-
-```{r}
-trans_est(tabs$log_rps) %>%
-  filter(year >= min_year + min(ages)) %>% 
-  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
-  layout(xaxis = list(range = c(min_year, max_year + 1)))
-```
-
-### SSB x Recruitment
-
-```{r}
-
-ssb <- trans_est(tabs$log_ssb)
-rec <- trans_est(tabs$log_recruitment, scale = 0.001)
-rps <- trans_est(tabs$log_rps)
-names(ssb) <- paste0("ssb_", names(ssb))
-names(rec) <- paste0("rec_", names(rec))
-names(rps) <- paste0("rps_", names(rps))
-ssb$year <- ssb$ssb_year + min(ages)
-rec$year <- rec$rec_year
-rps$year <- rps$rps_year
-sr <- merge(ssb, rec, by = "year")
-sr <- merge(sr, rps, by = "year")
-sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)
-
-plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
-                 col = ~year, clab = "Year")
-
-```
-
-### SSB x RPS
-
-```{r}
-p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
-                 col = ~year, clab = "Year")
-p
-```
-
-### log(SSB) x log(RPS)
-
-```{r}
-p %>% layout(yaxis = list(type = "log"),
-             xaxis = list(type = "log"))
-```
-
-Assumptions  {data-navmenu="Projections" .storyboard}
-================================================================================
-
-### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers
-
-```{r}
-
-Cp <- fit$tmb.data$Cp
-Cp_string <- text_c(Cp)
-Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1))
-Lp_string <- paste(text_c(Lp), "t")
-
-tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>%
-  filter(year >= max(years) - 4) %>%
-  plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>%
-  add_lines(alpha = 0.5) %>%
-  filter(year == max(years)) %>%
-  add_lines(showlegend = FALSE, line = list(width = 3)) %>%
-  layout(xaxis = list(title = "Age"))
-
-```
-
-***
-- The following catch multipliers were applied: `r Cp_string`
-- This translates to `r Lp_string` of catch in the projection years
-
-### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994
-
-```{r}
-
-if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic")
-
-Mpe <- fit$rep$Mpe_matrix
-Mproj <- replicate(fit$tmb.data$Yp, fit$tmb.data$pnm)
-bpe <- fit$rep$bpe[, (fit$tmb.data$Y + 1):(fit$tmb.data$Y + fit$tmb.data$Yp)]
-Mproj <- Mproj * exp(bpe)
-Mproj <- cbind(Mpe, Mproj)
-proj_years <- fit$tmb.data$proj_years
-
-plot_surface(mat = Mproj, ages = ages, years = c(years, proj_years), name = "M")
-
-```
-
-### Assume mean recruitment from the last three years
-
-```{r}
-
-last_three <- tail(fit$tmb.data$years, 3)
-mean_rec <- exp(mean(tabs$log_recruitment$est[tabs$log_recruitment$year %in% last_three]))
-mean_rec <- data.frame(year = c(last_three, last_three[3] + fit$tmb.data$Yp), 
-                       mean_rec = mean_rec * 1000, grp = "mean")
-trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1]) %>% 
-  add_lines(data = mean_rec, x = ~year, y = ~mean_rec, name = "mean",
-            color = I("black"), line = list(width = 1)) %>% 
-  add_text(x = max(years) + 1.2, y = mean_rec$mean_rec[1], 
-           text = "mean recruitment", 
-           textposition = "bottom right",
-           color = I("black")) %>%
-  layout(xaxis = list(range = c(min(years), max(years) + fit$tmb.data$Yp)))
-
-```
-
-### Assume averages weights and proportions mature from the last three years
-
-```{r}
-
-mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>%
-  filter(year >= max(years) - 2) %>%
-  plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>%
-  add_lines(alpha = 0.5) %>%
-  group_by(age) %>%
-  summarise(mat = mean(mat), year = "3y mean") %>%
-  add_lines(line = list(width = 3)) %>%
-  layout(xaxis = list(title = "Age"),
-         yaxis = list(title = "Proportion mature"))
-
-wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>%
-  filter(year >= max(years) - 2) %>%
-  plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols),
-          showlegend = FALSE) %>%
-  add_lines(alpha = 0.5) %>%
-  group_by(age) %>%
-  summarise(weight = mean(weight), year = "3y mean") %>%
-  add_lines(line = list(width = 3)) %>%
-  layout(xaxis = list(title = "Age"),
-         yaxis = list(title = "Weight (kg)"))
-
-subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE)
-
-```
-
-Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`}
-================================================================================
-
-### Catch multiplier projections for SSB / Blim from the last RAP
-
-```{r, eval = !is.null(tmp$last)}
-
-d <- trans_est(last_tabs$Cproj_log_ssb_lrp)
-d$scenario <- factor(d$scenario)
-d$type <- "Projected"
-
-past_d <- trans_est(last_tabs$log_ssb_lrp)
-past_d <- past_d[past_d$year != max(past_d$year), ] # drop the last year here (it's in the projection output)
-past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
-past_d <- do.call(rbind, past_d)
-past_d$type <- "Estimated"
-d <- rbind(past_d, d)
-xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
-ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
-
-current_d <- trans_est(tabs$log_ssb_lrp)
-current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s))
-current_d <- do.call(rbind, current_d)
-current_d$model <- "current RAP"
-current_d$type <- "Estimated"
-d$model <- "last RAP"
-d <- rbind(d, current_d)
-
-plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model",
-          vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE,
-          darken_proj = TRUE)
-
-```
-
-Retro projections {data-navmenu="Projections" `r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-### Catch multiplier projections for SSB / Blim from retrospective fits
-
-```{r, eval = !is.null(fit$retro)}
-
-d <- trans_est(retro_tabs$Cproj_log_ssb_lrp)
-d$terminal_year <- as.numeric(as.character(d$terminal_year))
-d$scenario <- factor(d$scenario)
-d$type <- "Projected"
-
-past_d <- trans_est(retro_tabs$log_ssb_lrp)
-past_d$terminal_year <- as.numeric(as.character(past_d$terminal_year))
-past_d <- past_d[past_d$year <= past_d$terminal_year, ] # drop the last year (it's in the projection output)
-past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
-past_d <- do.call(rbind, past_d)
-past_d$type <- "Estimated"
-d <- rbind(past_d, d)
-d <- d[order(d$scenario, d$terminal_year, d$year), ]
-
-# xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1)
-# ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
-
-plot_proj(data = d, ylab = "SSB/Blim", group = "terminal_year",
-          hline = 1, col = rev(viridis::viridis(length(unique(d$terminal_year)))), 
-          showlegend = TRUE, markers = FALSE, darken_proj = TRUE)
-
-```
-
-Results {data-navmenu="Projections"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~
-
-```{r}
-
-d <- trans_est(tabs$Cproj_log_ssb_ratio)
-d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100
-d$scenario <- factor(d$scenario)
-xlim <- c(max(years) + 0.9, max(years) + 4.1)
-ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
-
-plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%",
-          vline = max(years) + 1, col = cols[1]) %>%
-  layout(yaxis = list(tickvals = seq(-200, 200, by = 25)))
-
-```
-
-### SSB / Blim
-
-```{r}
-
-d <- trans_est(tabs$Cproj_log_ssb_lrp)
-d <- d[d$year != max(years_plus1), ] # drop the first year (it's in the time series)
-d$type <- "Projected"
-d$scenario <- factor(d$scenario)
-
-## Append full time series to projections
-past_d <- trans_est(tabs$log_ssb_lrp)
-past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s))
-past_d <- do.call(rbind, past_d)
-past_d$type <- ifelse(past_d$year == max(years_plus1), "Projected", "Estimated")
-d <- rbind(past_d, d)
-xlim <- c(max(years) - 2.1, max(years) + 4.1)
-ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ])
-
-plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim",
-          vline = max(years) + 1, hline = 1, col = cols[1], darken_proj = TRUE)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~)
-
-```{r}
-
-growth <- tabs$Cproj_log_ssb_ratio
-growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)
-
-d <- growth[growth$year != max(years) + 1, ]
-d$scenario <- factor(d$scenario)
-plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
-  add_heatmap() %>%
-  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
-                  showarrow = FALSE, font = list(color = "white")) %>%
-  colorbar(title = "P", limits = c(0, 1)) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year", 
-                      tickvals = min(d$year):max(d$year)))
-
-```
-
-### Prob(SSB~yp~ &ge; B~lim~)
-
-```{r}
-
-growth <- tabs$Cproj_log_ssb_lrp
-growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd)
-
-d <- growth[growth$year != max(years) + 1, ]
-d$scenario <- factor(d$scenario)
-plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>%
-  add_heatmap() %>%
-  add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2),
-                  showarrow = FALSE, font = list(color = "white")) %>%
-  colorbar(title = "P", limits = c(0, 1)) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year",
-                      tickvals = min(d$year):max(d$year)))
-
-```
-
 Inputs {data-navmenu="Tables"}

### 2020-03-14 inst/rmd/tagging.Rmd
Revamped NCAM explorer

Tagging
================================================================================

Inputs {.sidebar}
--------------------------------------------------------------------------------

#### **Filters**

```{r}

## Prep tag data and fits
tag_dat <- data.frame(
  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
             rep("post 1996", length(fit$tmb.data$tag_exp))),
  theta = c(plogis(est$theta[fit$tmb.data$theta_map + 1]), rep(NA, length(fit$tmb.data$tag_exp))),
  theta_lwr = c(plogis(est_lwr$theta[fit$tmb.data$theta_map + 1]), rep(NA, length(fit$tmb.data$tag_exp))),
  theta_upr = c(plogis(est_upr$theta[fit$tmb.data$theta_map + 1]), rep(NA, length(fit$tmb.data$tag_exp)))
)
tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
                      "age:", tag_dat$age, "</br>",
                      "delay:", tag_dat$delay, "</br>")
tag_dat$exp_no <- ifelse(nchar(tag_dat$exp == 4), substr(tag_dat$exp, 3, 4), substr(tag_dat$exp, 5, 6)) %>% 
  as.numeric()
tag_dat$year_exp <- tag_dat$rel_year + tag_dat$exp_no / 100

Udev <- data.frame(
  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
  Udev = c(est$old_Udev, est$Udev),
  Udev_lwr = c(est_lwr$old_Udev, est_lwr$Udev),
  Udev_upr = c(est_upr$old_Udev, est_upr$Udev)
)

Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
names(Fay)[names(Fay) == "year"] <- "rel_year"

tag_dat <- merge(tag_dat, Udev, by = "exp")
tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
tag_dat$Fsay_lwr <- tag_dat$Fay * exp(tag_dat$Udev_lwr)
tag_dat$Fsay_upr <- tag_dat$Fay * exp(tag_dat$Udev_upr)

## Share data across multiple plots
d <- crosstalk::SharedData$new(tag_dat, ~exp)
base <- plot_ly(d, cliponaxis = FALSE) %>% 
  group_by(exp) %>% 
  highlight(off = "plotly_doubleclick")

crosstalk::filter_select("period", "Period", d, ~period)
crosstalk::filter_select("rel_year", "Release year", d, ~rel_year)
crosstalk::filter_select("exp", "Experiment", d, ~exp)

```

Row {.tabset}
--------------------------------------------------------------------------------

### Recaptures

```{r}

y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
y_tickvals <- sqrt(y_ticktext)

rel_plot <- base %>%
  filter(rel_year == cap_year) %>%
  group_by(exp) %>%
  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
  add_markers(x = ~jitter(rel_year, factor = 0.5), 
              y = ~sqrt(tot_rel), hoverinfo = "text",
              text = ~paste("Experiment:", exp, "<br>Release year:", rel_year, "<br>N:", tot_rel),
              stroke = I(obs_col), color = ~period, colors = c(obs_col, "white"),
              marker = list(opacity = 0.8, line = list(width = 1.5))) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N released",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year", tickformat = "d"))

y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
y_tickvals <- sqrt(y_ticktext)

cap_plot <- base %>%
  group_by(exp, rel_year, cap_year) %>%
  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
            period = unique(period)) %>%
  group_by(exp) %>%
  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), name = "",
              hoverinfo = "text",
              showlegend = FALSE, text = ~paste("Experiment:", exp, 
                                                "<br>Release year:", rel_year, 
                                                "<br>Recapture year:", cap_year, 
                                                "<br>Observed N:", sum_obs),
              stroke = I(obs_col), color = ~period, colors = c(obs_col, "white"),
              marker = list(opacity = 0.8, line = list(width = 1.5))) %>%
  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(pred_col), name = "Predicted",
            alpha = 0.7, showlegend = TRUE, symbol = NULL, hoverinfo = "text",
            text = ~paste("Experiment:", exp, 
                          "<br>Release year:", rel_year, 
                          "<br>Recapture year:", cap_year, 
                          "<br>Predicted N:", round(sum_pred))) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = 0, y1 = 1000, yref = "y2",
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "N recaptured",
                      tickvals = y_tickvals,
                      ticktext = y_ticktext,
                      range = range(y_tickvals)),
         xaxis = list(title = "Year", tickformat = "d"))

subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
        nrows = 2, heights = c(0.3, 0.7))

```

> Y-axis is in square root scale

### Residuals

```{r}

ticktext <- c(5, pretty(c(tag_dat$pred, tag_dat$obs), 10))
tickvals <- sqrt(ticktext)

max_caps <- sqrt(max(c(tag_dat$pred, tag_dat$obs)))

a <- base %>%
  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
              text = ~text, showlegend = FALSE,
              stroke = I(obs_col), color = ~period, colors = c(obs_col, "white"),
              marker = list(opacity = 0.8, line = list(width = 1.5))) %>%
  add_lines(x = ~sqrt(obs), y = ~sqrt(obs), split = ~period,
            line = list(color = "black", width = 1), # had to split by period to render line with WebGL (reason unknown)
            hoverinfo = "none", showlegend = FALSE) %>% 
  layout(xaxis = list(title = "Predicted number of recaptures",
                      tickvals = tickvals,
                      ticktext = ticktext),
         yaxis = list(title = "Observed number of recaptures",
                      tickvals = tickvals,
                      ticktext = ticktext,
                      scaleanchor = "x")) %>%
  toWebGL()

ticktext <- c(5, pretty(tag_dat$pred, 10))
tickvals <- sqrt(ticktext)

b <- base %>%
  add_markers(x = ~sqrt(pred), y = ~res,
              text = ~text, showlegend = FALSE,
              stroke = I(obs_col), color = ~period, colors = c(obs_col, "white"),
              marker = list(opacity = 0.8, line = list(width = 1.5))) %>%
  layout(xaxis = list(title = "Predicted number of recaptures",
                      tickvals = tickvals,
                      ticktext = ticktext),
         yaxis = list(title = "Standardized residuals")) %>% 
  toWebGL()

subplot(a, b, nrows = 1, shareX = FALSE, shareY = FALSE, 
        titleX = TRUE, titleY = TRUE, margin = 0.05)

```

> Observation and prediction axes are in square root scale

### Initial F deviations

```{r}

base %>%
  filter(rel_year == cap_year) %>%
  add_markers(x = ~jitter(rel_year, factor = 1.5), 
              y = ~Udev, text = ~paste("F deviation:", round(Udev, 2), "<br>",
                                       "Release year:", rel_year, "<br>",
                                       "Experiment:", exp),
              hoverinfo = "text",
              showlegend = FALSE, stroke = I(pred_col), 
              color = ~period, colors = c(pred_col, "white"),
              marker = list(opacity = 0.8, line = list(width = 1.5))) %>%
  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
                       line = list(width = 1, dash = "dot")),
         yaxis = list(title = "Initial F deviation"),
         xaxis = list(title = "Year", tickformat = "d"))

```

### Thetas

```{r}

base %>% 
  filter(!is.na(theta)) %>% 
  dplyr::distinct(exp, theta, theta_upr, theta_lwr, 
                  period, rel_year, year_exp) %>% 
  mutate(fexp = factor(exp)) %>% 
  add_segments(x = ~fexp, xend = ~fexp,
               y = ~theta_lwr, yend = ~theta_upr,
               color = I(cols[1]), showlegend = FALSE,
               text = ~exp, marker = list(color = "transparent")) %>%
  add_markers(x = ~fexp, y = ~theta, showlegend = FALSE,
              marker = list(symbol = "200", 
                            color = "white", size = 8,
                            line = list(color = toRGB(cols[1]),
                                        width = 1.5)),
              text = ~exp) %>% 
  layout(yaxis = list(title = "Theta"),
         xaxis = list(title = "Experiment"))

```

### Reporting rates

```{r}

cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
rr <- data.frame(year = fit$tmb.data$SRR_years,
                 est = fit$tmb.data$log_SRR_est,
                 sd = sqrt(diag(cov_mat)))
rr$lwr <- rr$est - 1.96 * rr$sd
rr$upr <- rr$est + 1.96 * rr$sd
rr <- trans_est(rr)
rr$model <- "Random Walk"

ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
                      est = est$log_SRR, sd = sd$log_SRR)
ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
ncam_rr <- trans_est(ncam_rr)
ncam_rr$model <- "NCAM"

rr <- rbind(rr, ncam_rr)

plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)

```

### 2020-03-14 inst/rmd/trends.Rmd
Revamped NCAM explorer

--- 2018-11-21 inst/rmd/vis_fit.Rmd
@@
----
-title: "NCAM Explorer"
-output: flexdashboard::flex_dashboard
-always_allow_html: yes
----
-
-<style>
-/* list */
-li {
-line-height: 30px;
-}
-</style>
-
-```{r setup, echo=FALSE}
-library(plotly)
-library(Matrix)
-library(knitr)
-library(data.table)
-library(webshot)
-knitr::opts_knit$set(root.dir = tmp$output_dir)
-knitr::opts_chunk$set(echo = FALSE, fig.height = 7, fig.width = 8)
-```
-
-```{r data, include=FALSE}
-
-cols <- viridis::viridis(4)[-1]
-cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
-fit <- tmp$fit # temp object from vis_fit function
-if (!is.null(tmp$comp)) { comp <- tmp$comp }
-if (!is.null(tmp$last)) { last <- tmp$last }
-est <- as.list(fit$sd.rep, "Est")
-sd <- as.list(fit$sd.rep, "Std")
-tabs <- tidy_model(fit)
-if (!is.null(tmp$comp)) { comp_tabs <- tidy_model(model_list = comp) }
-if (!is.null(tmp$last)) { last_tabs <- tidy_model(tmp$last) }
-if (!is.null(fit$retro)) { retro_tabs <- tidy_retro(fit) } 
-min_year <- min(fit$tmb.data$years)
-max_year <- max(fit$tmb.data$years)
-ages <- fit$tmb.data$ages
-years <- fit$tmb.data$years
-years_plus1 <- fit$tmb.data$years_plus1
-
-## helper functions for nice Rmarkdown text
-text_range <- function(x) paste(range(x), collapse = "-")
-text_c <- function(x) {
-  paste0(paste(x[-length(x)], collapse = ", "), ", and ", x[length(x)])
-}
-text_units <- function(x) {
-  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
-  if (x >= 1000) return(paste0(round(x / 1000), "k"))
-  if (x < 1000) return(as.character(round(x)))
-}
-
-```
-
-Background
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### NCAM
-
-<h1>NCAM: A state-space stock assessment model for Northern cod</h1>
-
-```{r}
-
-ax <- list(
-  title = "",
-  zeroline = FALSE,
-  showline = FALSE,
-  showticklabels = FALSE,
-  showgrid = FALSE
-)
-
-xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
-plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
-        fillcolor = 'rgba(49, 104, 142, 0.3)',
-        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
-  add_trace() %>%
-  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
-         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
-  plotly::config(displayModeBar = FALSE)
-
-```
-
-### General
-
-- An *integrated* *state-space* population dynamics model
-- *Integrated*
-    - Uses as much information as possible
-- *State-space*
-    - Includes process and observation equations + error
-    - Population processes are informed by observations from multiple sources
-    - Process errors are necessary for realistic stochastic projections
-
-### Data inputs
-
-- Reported landings (`r text_range(years)`)
-- Catch-at-age information (`r text_range(years)`)
-- Reported landings by month (`r text_range(years)`)
-- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
-- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
-- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
-- Tagging data (`r text_range(years)`)
-
-### Biological inputs
-
-- RV survey data were used fit cohort-based models of:
-   - Proportion mature at age
-   - Beginning of year and mid-year stock weights at age
-- Baseline natural mortality (M) values
-
-```{r}
-
-plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
-  add_surface() %>%
-  colorbar(title = "Mshift") %>%
-  layout(scene = list(
-    xaxis = list(title = "Year"),
-    yaxis = list(title = "Age"),
-    zaxis = list(title = "Mshift")
-  ))
-
-```
-
-### Partial catches
-
-- User sets lower and upper bounds on total catch weight
-- Bounds are used via a censored likelihood
-
-```{r}
-
-plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
-  add_lines(name = "Lower", color = I(cols[1])) %>%
-  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
-  add_lines(y = rep(1, length(years)), color = I("black"),
-            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Change in Q
-
-- A shift in Q was assumed for the offshore RV survey
-- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
-- Q for the inshore sentinel survey was assumed to follow a random walk
-    - Done to account for potential changes in migration patterns
-
-### Tagging data
-
-- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
-- Applies adjustments for tagging mortality, tag loss and reporting rates
-- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
-- The Baranov catch equation was applied to age-based tag survivors
-- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock
-
-### References
-
-Cadigan, N.
-2015.
-A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates.
-Canadian Journal of Fisheries and Aquatic Sciences 73 (2).
-NRC Research Press: 296-308.
-
-Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.
-
-Catch
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-
-plot_landings(fit)
-
-```
-
-> Y-axes in log scale
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted catch at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
-                    pred = fit$rep$pred_prop_catch),
-               ages = ages, years = years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$CA_resid,
-                    fit = fit$rep$pred_crl),
-               ages = ages[-length(ages)], years = years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-> based on continuation ratio logits
-
-### Catch at age surface (proportion)
-
-```{r}
-plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
-```
-
-<!-- ### Catch at age surface (numbers) -->
-
-<!-- ```{r} -->
-<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
-<!-- ``` -->
-
-<!-- > Note: model fits to composition - this plot is included for illustration -->
-
-RV Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### RV survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
-pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
-obs[obs == 0] <- NA
-pred[pred == 0] <- NA
-d <- data.frame(year = years, obs = obs, pred = pred)
-
-# ## RV biomass
-# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# d <- data.frame(year = years, obs = obs, pred = pred)
-# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    zero = fit$tmb.data$ind_zero,
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d <- d[d$wt == 1, ]
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
-                    pred = exp(fit$rep$Elog_RV_index),
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d$obs[d$wt == 0] <- NA
-d$pred[d$wt == 0] <- NA
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Mean numbers per tow",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
-                    fit = fit$rep$Elog_RV_index),
-               ages = ages, years = years)
-d <- d[d$keep, ]
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SN Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$SN_GN_index)
-pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
-d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$log_SN_GN_resid,
-              ages = fit$tmb.data$SN_ages,
-              years = fit$tmb.data$SN_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
-                    pred = exp(fit$rep$Elog_SN_GN_index)),
-               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
-facet_obs_pred(d, by = "age",
-               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
-               nrows = 4, ylab = "Sentinel survey index",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
-                    fit = fit$rep$Elog_SN_GN_index),
-               ages = fit$tmb.data$SN_ages,
-               years = fit$tmb.data$SN_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SS Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Smith sound acoustic biomass estimates
-
-```{r}
-
-d <- data.frame(year = fit$tmb.data$SS_bms_years,
-                obs = fit$tmb.data$SSbms,
-                pred = fit$rep$ESSbms)
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
-              showlegend = TRUE)
-
-```
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
-              ages = fit$tmb.data$SS_ages,
-              years = fit$tmb.data$SS_age_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Smith Sound trawl numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
-                    pred = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
-               obs_col = cols[2], pred_col = cols[1])
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
-                    fit = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-Tagging {data-orientation=rows}
-================================================================================
-
-```{r}
-
-## Prep tag data and fits
-tag_dat <- data.frame(
-  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
-  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
-  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
-  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
-  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
-  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
-  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
-  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
-  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
-             rep("post 1996", length(fit$tmb.data$tag_exp)))
-)
-tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
-tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
-                      "age:", tag_dat$age, "</br>",
-                      "delay:", tag_dat$delay, "</br>")
-
-Udev <- data.frame(
-  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
-  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
-)
-
-Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
-names(Fay)[names(Fay) == "year"] <- "rel_year"
-
-tag_dat <- merge(tag_dat, Udev, by = "exp")
-tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
-tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
-
-## Share data across multiple plots
-d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)
-
-```
-
-Row {data-height=700 .tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted tag recaptures by experimemnt
-
-```{r}
-
-y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
-y_tickvals <- sqrt(y_ticktext)
-
-rel_plot <- base %>%
-  filter(rel_year == cap_year) %>%
-  group_by(exp) %>%
-  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N released",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
-y_tickvals <- sqrt(y_ticktext)
-
-cap_plot <- base %>%
-  group_by(exp, rel_year, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
-            period = unique(period)) %>%
-  group_by(exp) %>%
-  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
-              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
-            alpha = 0.7, showlegend = FALSE, symbol = NULL,
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y2",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
-        nrows = 2, heights = c(0.3, 0.7))
-
-```
-
-> Y-axis is in square root scale
-
-<!-- ### Animation -->
-
-<!-- ```{r} -->
-
-<!-- sum_d <- tag_dat %>% -->
-<!--   group_by(exp, cap_year) %>% -->
-<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->
-
-<!-- sum_d %>%  -->
-<!--   group_by(exp) %>%  -->
-<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
-<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
-<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
-<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
-<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
-<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
-<!--   layout(yaxis = list(title = "N recaptured", -->
-<!--                       tickvals = y_tickvals, -->
-<!--                       ticktext = y_ticktext, -->
-<!--                       range = range(y_tickvals)), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-Row
---------------------------------------------------------------------------------
-
-### Tag reporting rate
-
-```{r}
-
-cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
-rr <- data.frame(year = fit$tmb.data$SRR_years,
-                 est = fit$tmb.data$log_SRR_est,
-                 sd = sqrt(diag(cov_mat)))
-rr$lwr <- rr$est - 1.96 * rr$sd
-rr$upr <- rr$est + 1.96 * rr$sd
-rr <- trans_est(rr)
-rr$model <- "Random Walk"
-
-ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
-                      est = est$log_SRR, sd = sd$log_SRR)
-ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
-ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
-ncam_rr <- trans_est(ncam_rr)
-ncam_rr$model <- "NCAM"
-
-rr <- rbind(rr, ncam_rr)
-
-plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)
-
-```
-
-### Initial F deviations for tagging experiments
-
-```{r}
-
-base %>%
-  distinct(rel_year, Udev, exp, period) %>%
-  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
-              showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "Initial F deviation"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Observed vs. predicted recaptures
-
-```{r}
-
-x_ticktext <- c(5, pretty(tag_dat$pred, 10))
-y_ticktext <- c(5, pretty(tag_dat$obs, 10))
-x_tickvals <- sqrt(x_ticktext)
-y_tickvals <- sqrt(y_ticktext)
-
-base %>%
-  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
-              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       line = list(width = 1)),
-         xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Observed number of recaptures",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext))
-
-```
-
-> Axes are in square root scale
-
-### Standardized residuals vs. predicted recaptures
-
-```{r}
-
-y_ticktext <- c(5, pretty(tag_dat$res, 10))
-y_tickvals <- sqrt(x_ticktext)
-
-base %>%
-  add_markers( x = ~sqrt(pred), y = ~res,
-               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Standardized residuals"))
-
-```
-
-> X-axis is in square root scale
-
-Catchability
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Fishery selectivity
-
-```{r}
-
-S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
-plot_surface(S, ages = ages, years = years, name = "S")
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey catchability
-
-```{r}
-
-q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
-plot_surface(q, ages = ages, years = years, name = "q")
-
-```
-
-### Proportion of the stock in Smith Sound
-
-```{r}
-p <- matrix(0, nrow = length(ages), ncol = length(years),
-            dimnames = list(ages, years))
-ss_p <- fit$rep$SSD
-ss_years <- fit$tmb.data$SS_years
-ss_ages <- fit$tmb.data$SS_ages
-rownames(ss_p) <- ss_ages
-colnames(ss_p) <- ss_years
-p[as.character(ss_ages), as.character(ss_years)] <- ss_p
-plot_surface(p, ages = ages, years = years, name = "p")
-
-```
-
-### Sentinel gillnet survey catchability
-
-```{r}
-
-log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
-log_qa <- fit$rep$log_SN_GN_Qa
-q <- exp(log_qy + log_qa)
-plot_surface(q, ages = fit$tmb.data$SN_ages,
-             years = fit$tmb.data$SN_years,
-             name = "q")
-
-```
-
-Trends
+Retro {data-navmenu="Trends" `r if (is.null(fit$retro)) ".hidden"`}
 ================================================================================
@@
 
-```{r}
-
-## Note: biomass is in tonnes and numbers are in thousands
-bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
-             data.frame(metric = "Ages 2+", tabs$log_biomass))
-abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
-             data.frame(metric = "Ages 2+", tabs$log_abundance))
-
-bms_plot <- trans_est(bms) %>%
-  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
-             annotate = TRUE)
-
-abn_plot <- trans_est(abn, scale = 0.001) %>%
-  plot_trend(ylab = "Abundance", group = "metric", col = cols,
-             annotate = TRUE)
-
-blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
-  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
-  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
-           color = I("black"), textposition = "right", hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Abundance at age
-
-```{r}
-plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
-```
-
-### Biomass at age
-
-```{r}
-plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
-```
-
-### SSB at age
-
-```{r}
-plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r}
-
-f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
-z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
-m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
-
-z_plot <- trans_est(z) %>%
-  plot_trend(ylab = "Average Z", group = "metric",
-             annotate = TRUE, col = cols)
-
-m_plot <- trans_est(m) %>%
-  plot_trend(ylab = "Average M", group = "metric",
-             annotate = TRUE, col = cols)
-
-f_plot <- trans_est(f) %>%
-  plot_trend(ylab = "Average F", group = "metric",
-             annotate = TRUE, col = cols)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Z at age
-
-```{r}
-plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
-```
-
-### M at age
-
-```{r}
-plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
-```
-
-### F at age
-
-```{r}
-plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
-```
-
-### Z, M and F
-
-```{r}
-plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
-                 ages = ages, years = years)
-```
-
-> Z = purples, M = blues, F = reds
-
-### Process error
-
-```{r}
-
-pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
-             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
-  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
-              color = I("black"), opacity = 0.8)
-
-```
-
-### Process error (bubble plot)
-
-```{r}
-pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = 2:8, value = "res")
-bubble_plot(pe)
-```
-
-Retro {`r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
 ```{r, eval = !is.null(fit$retro)}
@@
 bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])
+  plot_retro(ylab = "SSB (t)", lwd = 1, col = pred_col)
 
 abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])
+  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = pred_col)
 
 blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
+  plot_retro(ylab = "SSB/Blim", lwd = 1, col = pred_col) %>%
   add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
@@
 
-Row
+### Mohn's rho
+
+```{r, eval = !is.null(fit$retro)}
+
+bms_plot <- trans_est(retro_tabs$log_ssb) %>%
+  plot_mohns(col = pred_col, ylab = "SSB")
+
+abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
+  plot_mohns(ylab = "Recruitment (age 2)", col = pred_col)
+
+blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
+  plot_mohns(ylab = "SSB/Blim", col = pred_col)
+
+subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Relative difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
+
+### Difference
+
+```{r, eval = !is.null(fit$retro)}
+
+bms_plot <- retro_tabs$log_ssb %>%
+  plot_retro_deltas(in_col = pred_col, ylab = "log(SSB)")
+
+abn_plot <- retro_tabs$log_recruitment %>%
+  plot_retro_deltas(ylab = "log(Recruitment)", in_col = pred_col)
+
+blim_plot <- retro_tabs$log_ssb_lrp %>%
+  plot_retro_deltas(ylab = "log(SSB/Blim)", in_col = pred_col)
+
+subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
+
+Row {.tabset}
 --------------------------------------------------------------------------------
@@
 z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])
+  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = pred_col)
 
 m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])
+  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = pred_col)
 
 f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])
+  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = pred_col)
 
@@
 
-Comps {`r if (is.null(tmp$comp)) ".hidden"`}
+### Mohn's rho
+
+```{r, eval = !is.null(fit$retro)}
+
+z_plot <- trans_est(retro_tabs$log_aveZ) %>%
+  plot_mohns(ylab = "Average Z (ages 5+)", col = pred_col)
+
+m_plot <- trans_est(retro_tabs$log_aveM) %>%
+  plot_mohns(ylab = "Average M (ages 5+)", col = pred_col)
+
+f_plot <- trans_est(retro_tabs$log_aveF) %>%
+  plot_mohns(ylab = "Average F (ages 5+)", col = pred_col)
+
+subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Relative difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Relative difference = (retro estimate - current estimate) / current estimate; Mohn's rho = mean(relative difference)
+
+### Difference
+
+```{r, eval = !is.null(fit$retro)}
+
+z_plot <- retro_tabs$log_aveZ %>%
+  plot_retro_deltas(ylab = "log(Average Z)", in_col = pred_col)
+
+m_plot <- retro_tabs$log_aveM %>%
+  plot_retro_deltas(ylab = "log(Average M)", in_col = pred_col)
+
+f_plot <- retro_tabs$log_aveF %>%
+  plot_retro_deltas(ylab = "log(Average F)", in_col = pred_col)
+
+subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE) %>% 
+  layout(annotations = left_label("Difference", offset = 50),
+         margin = list(l = 75))
+
+```
+
+> Difference = retro estimate - current estimate; approximate 95% intervals based on retro standard deviation + current standard deviation
+
+Comps {data-navmenu="Trends" `r if (is.null(tmp$comp)) ".hidden"`}
 ================================================================================
@@
 
-Productivity
+Status and rates {data-navmenu="Trends"}
+================================================================================
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Stock status
+
+```{r}
+
+## Note: biomass is in tonnes and numbers are in thousands
+bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
+             data.frame(metric = "Ages 2+", tabs$log_biomass))
+abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
+             data.frame(metric = "Ages 2+", tabs$log_abundance))
+
+bms_plot <- trans_est(bms) %>%
+  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
+             annotate = TRUE)
+
+abn_plot <- trans_est(abn, scale = 0.001) %>%
+  plot_trend(ylab = "Abundance", group = "metric", col = cols,
+             annotate = TRUE)
+
+blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
+  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
+  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
+               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
+  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
+           color = I("black"), textposition = "right", hoverinfo = "none")
+
+subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
+
+```
+
+### Abundance at age
+
+```{r}
+plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
+```
+
+<!-- ### Biomass at age -->
+
+<!-- ```{r} -->
+<!-- plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass") -->
+<!-- ``` -->
+
+### SSB at age
+
+```{r}
+plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
+```
+
+<!-- ### Exploitable biomass at age -->
+
+<!-- ```{r} -->
+<!-- S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x)) -->
+<!-- B <- fit$rep$N_matrix[, seq(ncol(S))] * fit$tmb.data$midyw -->
+<!-- E <- B * S -->
+<!-- plot_surface(E, ages = ages, years = years, name = "Exploitable biomass") -->
+<!-- ``` -->
+
+### Demographics
+
+```{r}
+
+a <- tidy_mat(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, value = "abundance") %>% 
+  dplyr::left_join(trans_est(tabs$log_abundance, scale = 0.001), by = "year") %>% 
+  rename(tot_abundance = est, tot_abundance_lwr = lwr, tot_abundance_upr = upr,
+         tot_abundance_sd = sd)
+
+b <- tidy_mat(fit$rep$biomass_matrix, ages = ages, years = years_plus1, value = "biomass") %>% 
+  dplyr::left_join(trans_est(tabs$log_biomass), by = "year") %>% 
+  rename(tot_biomass = est, tot_biomass_lwr = lwr, tot_biomass_upr = upr,
+         tot_biomass_sd = sd)
+
+c <- tidy_mat(fit$rep$ssb_matrix, ages = ages, years = years_plus1, value = "ssb") %>% 
+  dplyr::left_join(trans_est(tabs$log_ssb), by = "year") %>% 
+  dplyr::rename(tot_ssb = est, tot_ssb_lwr = lwr, tot_ssb_upr = upr,
+                tot_ssb_sd = sd)
+
+d <- dplyr::left_join(a, b, by = c("year", "age")) %>%
+  dplyr::left_join(c, by = c("year", "age")) %>%
+  mutate(cohort = year - age,
+         abundance_prop = abundance / tot_abundance,
+         abundance_lab = paste0("Abundance<br>", 
+                                ifelse(gsub("[[:digit:]]", "", text_units(tot_abundance_lwr)) ==
+                                         gsub("[[:digit:]]", "", text_units(tot_abundance_upr)),
+                                       gsub("[[:alpha:]]", "", text_units(tot_abundance_lwr)),
+                                       text_units(tot_abundance_lwr)), "-", 
+                                text_units(tot_abundance_upr)),
+         biomass_prop = -(biomass / tot_biomass),
+         biomass_lab = paste0("Biomass<br>", 
+                              ifelse(gsub("[[:digit:]]", "", text_units(tot_biomass_lwr)) ==
+                                       gsub("[[:digit:]]", "", text_units(tot_biomass_upr)),
+                                     gsub("[[:alpha:]]", "", text_units(tot_biomass_lwr)),
+                                     paste0(text_units(tot_biomass_lwr), "t")), "-", 
+                              text_units(tot_biomass_upr), "t"),
+         ssb_prop = -(ssb / tot_ssb),
+         ssb_lab = paste0("SSB<br>", 
+                          ifelse(gsub("[[:digit:]]", "", text_units(tot_ssb_lwr)) ==
+                                   gsub("[[:digit:]]", "", text_units(tot_ssb_upr)),
+                                 gsub("[[:alpha:]]", "", text_units(tot_ssb_lwr)),
+                                 paste0(text_units(tot_ssb_lwr), "t")), "-", 
+                          text_units(tot_ssb_upr), "t")) %>%
+  arrange(year, age)
+
+max_x <- max(c(d$abundance_prop, d$biomass_prop)) * 1.2
+xlabs <- pretty(c(-max_x, max_x), n = 10)
+
+rec <- fit$rep$N_matrix[1, ]
+rec_spikes <- find_peaks(rec, m = 1)
+rec_spikes <- years[rec_spikes] - min(ages)
+
+d$spike <- d$cohort
+d$spike_start <- d$biomass_prop
+d$spike_end <- d$abundance_prop
+d$spike[!(d$spike %in% rec_spikes)] <- " "
+d$spike_start[!(d$spike %in% rec_spikes)] <- 0
+d$spike_end[!(d$spike %in% rec_spikes)] <- 0
+
+d %>%
+  plot_ly(y = ~age, frame = ~year, showlegend = FALSE,
+          hoverinfo = "none", cliponaxis = FALSE) %>%
+  add_paths(x = ~abundance_prop, fill = "tozerox",
+            color = I(cols[1]), line = list(width = 1)) %>%
+  # add_paths(x = ~biomass_prop, fill = "tozerox",
+  #           color = I(cols[2]), line = list(width = 1)) %>%
+  add_paths(x = ~ssb_prop, fill = "tozerox",
+            color = I(cols[2]), line = list(width = 1)) %>%
+  add_text(x = max_x, text = ~cohort, textposition = "right") %>%
+  add_text(x = ~abundance_prop + 0.01, color = I("grey50"),
+           text = ~spike, textposition = "center right") %>%
+  add_text(x = ~ssb_prop - 0.01, color = I("grey50"),
+           text = ~spike, textposition = "center left") %>%
+  add_text(x = 0, y = ~max(age), text = ~paste0(year, "<br>"),
+           textposition = "center top", size = I(18)) %>%
+  # add_text(x = max_x / 2, y = ~max(age), text = "Abundance<br>",
+  #          textposition = "center top", color = I(cols[1]),
+  #          size = I(18)) %>%
+  add_text(x = max_x / 2, y = ~max(age), text = ~abundance_lab,
+           textposition = "center top", color = I(cols[1]),
+           size = I(18)) %>%
+  # add_text(x = -max_x / 2, y = ~max(age), text = ~biomass_lab,
+  #          textposition = "center top", color = I(cols[2]),
+  #          size = I(18)) %>%
+  add_text(x = -max_x / 2, y = ~max(age), text = ~ssb_lab,
+           textposition = "center top", color = I(cols[2]),
+           size = I(18)) %>%
+  add_annotations(x = 0, y = 1, yshift = 35, text = "Age",
+                  xanchor = "right", yanchor = "top",
+                  font = list(size = 14),
+                  xref = "paper", yref = "paper", showarrow = FALSE) %>%
+  add_annotations(x = 1, y = 1, yshift = 35, xshift = -5,
+                  text = "Cohort",
+                  xanchor = "left", yanchor = "top",
+                  font = list(size = 14),
+                  xref = "paper", yref = "paper", showarrow = FALSE) %>%
+  layout(xaxis = list(range = c(-max_x, max_x),
+                      title = "Percent",
+                      tickmode = "array",
+                      tickvals = xlabs,
+                      ticktext = paste0(abs(xlabs) * 100, "%")),
+         yaxis = list(title = ""),
+         margin = list(l = 50, r = 70, t = 50, b = 50)) %>%
+  animation_opts(frame = 1000, transition = 0) %>%
+  animation_slider(currentvalue = list(prefix = "Year: ",
+                                       font = list(color = "transparent")))
+
+```
+
+> Totals are lower and upper 95% confidence intervals
+
+Row {.tabset}
+--------------------------------------------------------------------------------
+
+### Mortality Rates
+
+```{r}
+
+f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
+z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
+m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
+
+z_plot <- trans_est(z) %>%
+  plot_trend(ylab = "Average Z", group = "metric",
+             annotate = TRUE, col = cols)
+
+m_plot <- trans_est(m) %>%
+  plot_trend(ylab = "Average M", group = "metric",
+             annotate = TRUE, col = cols)
+
+f_plot <- trans_est(f) %>%
+  plot_trend(ylab = "Average F", group = "metric",
+             annotate = TRUE, col = cols)
+
+subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
+
+```
+
+<!-- ### Z at age -->
+
+<!-- ```{r} -->
+<!-- plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z") -->
+<!-- ``` -->
+
+### M at age
+
+```{r}
+plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
+```
+
+### F at age
+
+```{r}
+plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
+```
+
+### Z, M and F
+
+```{r}
+plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
+                 ages = ages, years = years)
+```
+
+> Z = purples, M = blues, F = reds
+
+### Process error
+
+```{r}
+
+pe <- fit$rep$pe
+plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], 
+             ages = fit$tmb.data$pe_ages, name = "pe", 
+             col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
+  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
+              color = I("black"), opacity = 0.8)
+
+```
+
+<!-- ### Process error (bubble) -->
+
+<!-- ```{r} -->
+<!-- pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1], -->
+<!--                ages = fit$tmb.data$pe_ages, value = "res") -->
+<!-- bubble_plot(pe, pos_col = pos_col, neg_col = neg_col) -->
+<!-- ``` -->
+
+Productivity  {data-navmenu="Trends"}
 ================================================================================
@@
 trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1])
+  plot_trend(ylab = "Recruitment", col = pred_col)
 ```
@@
 trans_est(tabs$log_ssb) %>%
-  plot_trend(ylab = "SSB (t)", col = cols[1])
+  plot_trend(ylab = "SSB (t)", col = pred_col)
 ```
@@
   filter(year >= min_year + min(ages)) %>% 
-  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
+  plot_trend(ylab = "RPS (thousand / tonne)", col = pred_col) %>% 
   layout(xaxis = list(range = c(min_year, max_year + 1)))
@@
 ```
-
-```{r}
-
-message("Bits of the projection code is currently not working. Specifically the plots that require the frame argument in plotly.")
-
-```
-
-<!-- Assumptions  {data-navmenu="Projections" .storyboard} -->
-<!-- ================================================================================ -->
-
-<!-- ### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers -->
-
-<!-- ```{r} -->
-
-<!-- Cp <- fit$tmb.data$Cp -->
-<!-- Cp_string <- text_c(Cp) -->
-<!-- Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1)) -->
-<!-- Lp_string <- paste(text_c(Lp), "t") -->
-
-<!-- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>% -->
-<!--   filter(year >= max(years) - 4) %>% -->
-<!--   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   filter(year == max(years)) %>% -->
-<!--   add_lines(showlegend = FALSE, line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age")) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - The following catch multipliers were applied: `r #Cp_string` -->
-<!-- - This translates to `r #Lp_string` of catch in the projection years -->
-
-<!-- ### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994 -->
-
-<!-- ```{r} -->
-
-<!-- if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic") -->
-
-<!-- aveM <- tabs$log_aveM -->
-<!-- aveM <- exp(mean(aveM$est[!aveM$year %in% 1991:1994])) -->
-
-<!-- trans_est(tabs$log_aveM) %>% -->
-<!--   plot_trend(ylab = "Average M", col = cols[1]) %>% -->
-<!--   add_segments(x = 1995, xend = max(years) + 3, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_segments(x = 1990, xend = 1995, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), -->
-<!--                line = list(width = 1, dash = "dot")) %>% -->
-<!--   add_segments(x = min(years), xend = 1990, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_text(x = max(years) + 0.2, y = aveM, text = "long-term mean", textposition = "bottom right", -->
-<!--            color = I("black")) %>% -->
-<!--   layout(xaxis = list(range = c(min(years), max(years) + 3))) -->
-
-<!-- ``` -->
-
-<!-- ### Assume a relationship between recruits per spawner and number of spawners to project recruits in the future -->
-
-<!-- ```{r} -->
-
-<!-- if (fit$tmb.data$Rp_mode != 3) stop("TODO: make the recruitment mode documentation dynamic") -->
-
-<!-- rps_tab <- tabs$log_rps -->
-<!-- names(rps_tab) <- c("year", "rps", "rps_sd", "rps_lwr", "rps_upr") -->
-<!-- rps_tab <- rps_tab[rps_tab$year >= 1985, ] -->
-<!-- rps_tab$year <- rps_tab$year - 2 # scale back two year to allign with spawners that produced the recruits -->
-<!-- ssa_tab <- tabs$log_ssa -->
-<!-- names(ssa_tab) <- c("year", "ssa", "ssa_sd", "ssa_lwr", "ssa_upr") -->
-<!-- d <- merge(ssa_tab, rps_tab, by = "year") -->
-<!-- rec_tab <- tabs$log_recruitment -->
-<!-- names(rec_tab) <- c("year", "rec", "rec_sd", "rec_lwr", "rec_upr") -->
-<!-- rec_tab <- rec_tab[rec_tab$year >= 1985, ] -->
-<!-- rec_tab$year <- rec_tab$year - 2 # scale back two years -->
-<!-- d <- merge(d, rec_tab, by = "year") -->
-<!-- d$excluded <- d$year %in% 1991:1994 -->
-
-<!-- sub_d <- d[!d$year %in% 1991:1994, ] -->
-<!-- lin_rel <- lm(rps ~ ssa, data = sub_d) -->
-<!-- # coefficients(lin_rel) -->
-
-<!-- preds <- data.frame(year = NA, rps_sd = NA, ssa_sd = NA, ssa = seq(min(d$ssa), max(d$ssa), length.out = 100)) -->
-<!-- preds$fits <- predict(lin_rel, newdata = preds) -->
-<!-- preds$excluded <- FALSE -->
-
-<!-- plot_ly(data = d, x = ~ssa, y = ~rps, text = ~year, -->
-<!--         type = "scatter", mode = "markers", name = "est", -->
-<!--         color = ~excluded, colors = c(cols[2], "grey"), -->
-<!--         error_y = list(array = ~ qnorm(0.975) * rps_sd, -->
-<!--                        width = 0, thickness = 1), -->
-<!--         error_x = list(array = ~ qnorm(0.975) * ssa_sd, -->
-<!--                        width = 0, thickness = 1)) %>% -->
-<!--   add_text(text = ~year, textfont = list(color = toRGB("grey50")), textposition = "top right", -->
-<!--            error_y = list(visible = FALSE), -->
-<!--            error_x = list(visible = FALSE)) %>% -->
-<!--   layout(title = "Spawners vs Recruits / Spawner", -->
-<!--          xaxis = list(title = "log(Spawners)"), -->
-<!--          yaxis = list(title = "log(Recruits / Spawner)", zerolinecolor = toRGB("lightgrey")), -->
-<!--          showlegend = FALSE) %>% -->
-<!--   add_lines(data = preds, x = ~ssa, y = ~fits, name = "fit", color = I(cols[1])) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - Grey points indicate years excluded from the regressionn -->
-
-<!-- ### Assume averages weights and proportions mature from the last three years -->
-
-<!-- ```{r} -->
-
-<!-- mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(mat = mean(mat), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Proportion mature")) -->
-
-<!-- wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols), -->
-<!--           showlegend = FALSE) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(weight = mean(weight), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Weight (kg)")) -->
-
-<!-- subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- ### Catch multiplier projections for SSB / Blim from the last RAP -->
-
-<!-- ```{r, eval = !is.null(tmp$last)} -->
-
-<!-- d <- trans_est(last_tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- d <- d[d$year <= max(years) + 1, ] -->
-
-<!-- past_d <- trans_est(last_tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- current_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s)) -->
-<!-- current_d <- do.call(rbind, current_d) -->
-<!-- current_d$model <- "current RAP" -->
-<!-- d$model <- "last RAP" -->
-<!-- d <- rbind(d, current_d) -->
-
-<!-- plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model", -->
-<!--           vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Results {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~ -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_ratio) -->
-<!-- d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100 -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- xlim <- c(max(years) + 0.9, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%", -->
-<!-- #           vline = max(years) + 1, col = cols[1]) %>% -->
-<!-- #   layout(yaxis = list(tickvals = seq(-200, 200, by = 25))) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- ### SSB / Blim -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-
-<!-- past_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(years) - 2.1, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", -->
-<!-- #           vline = max(years) + 1, hline = 1, col = cols[1]) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_ratio -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-<!-- ### Prob(SSB~yp~ &ge; B~lim~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_lrp -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-Inputs {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
-```
-
-### Catch at age
-
-```{r}
-kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
-```
-
-<!-- ### Monthly landings -->
-
-<!-- ```{r} -->
-<!-- d <- inputs$catch.month -->
-<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
-<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
-<!-- ``` -->
-
-### RV index
-
-```{r}
-d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
-```
-
-### SN index
-
-```{r}
-d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
-mat <- d[, as.character(3:10)]
-rownames(mat) <- d$year
-kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
-```
-
-### SS index
-
-```{r}
-d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
-names(d) <- c("Year", "Month", "Biomass", "SE^2^")
-kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
-```
-
-### SS age composition
-
-```{r}
-kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
-```
-
-### Weight at age
-
-```{r}
-d <- inputs$pwts
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "weight")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
-```
-
-### Maturity at age
-
-```{r}
-d <- inputs$mats
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "mat")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div.
-2J+3KL from DFO RV autumn bottom trawl surveys.
-Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
-```
-
-### Tag summary
-
-```{r}
-d <- inputs$mtag1 %>% 
-  group_by(rel_year, exp) %>% 
-  summarise(NAFO = unique(div),
-            Nrel = sum(unique(released)), 
-            Ncap = sum(recaptured))
-names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
-kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
-```
-
-Settings {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Catch bounds
-
-```{r}
-d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
-kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
-```
-
-### Baseline M
-
-```{r}
-m <- fit$tmb.data$M_matrix
-rownames(m) <- ages
-colnames(m) <- years
-kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
-```
-
-Output {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Parameter estimates
-
-```{r}
-
-par_tab <- tidy_pars(fit)
-rownames(par_tab) <- NULL
-options(knitr.kable.NA = '')
-kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Parameter estimates (retro)
-
-```{r}
-
-par_tabs <- lapply(names(fit$retro), function(nm) {
-  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
-  d$id <- seq(nrow(d))
-  d
-})
-par_tabs <- do.call(rbind, par_tabs)
-final_par_tab <- data.frame(tyear = max(years), par_tab)
-final_par_tab$id <- seq(nrow(final_par_tab))
-par_tabs <- rbind(par_tabs, final_par_tab)
-par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-par_tabs$Quantity <- as.character(par_tabs$Quantity)
-par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
-par_tabs$Symbol <- as.character(par_tabs$Symbol)
-par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-rownames(par_tabs) <- NULL
-names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"
-
-kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Stock size
-
-```{r}
-
-tab_name <- c("log_abundance" = "Abundance (millions)", 
-              "log_recruitment" = "Recruits (age 2; millions)", 
-              "log_biomass" = "Biomass (kt)", 
-              "log_ssa" = "Spawners (millions)", 
-              "log_ssb" = "SSB (kt)", 
-              "log_ssb_lrp" = "SSB/Blim n(%)")
-
-nms <- names(tab_name)
-status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
-names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
-  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  status_tab <- merge(status_tab, tab, by = "Year")
-}
-
-knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Rates
-
-```{r}
-
-tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
-              "log_aveM" = "Average M (ages 5+)", 
-              "log_aveF" = "Average F (ages 5+)")
-
-nms <- names(tab_name)
-rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
-names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  rate_tab <- merge(rate_tab, tab, by = "Year")
-}
-
-knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Abundance-at-age
-
-```{r}
-N_matrix <- fit$rep$N_matrix
-dimnames(N_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Mature abundance-at-age
-
-```{r}
-ssa_matrix <- fit$rep$ssa_matrix
-dimnames(ssa_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Biomass-at-age
-
-```{r}
-B_matrix <- fit$rep$biomass_matrix
-dimnames(B_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Mature biomass-at-age
-
-```{r}
-ssb_matrix <- fit$rep$ssb_matrix
-dimnames(ssb_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Z-at-age
-
-```{r}
-Z_matrix <- fit$rep$Z_matrix
-dimnames(Z_matrix) <- list(age = ages, year = years)
-knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
-```
-
-### M-at-age
-
-```{r}
-M_matrix <- fit$rep$Mpe_matrix
-dimnames(M_matrix) <- list(age = ages, year = years)
-M_matrix <- M_matrix[as.character(2:8), ]
-rownames(M_matrix) <- c(as.character(2:7), "8+")
-knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
-```
-
-### F-at-age
-
-```{r}
-F_matrix <- fit$rep$F_matrix
-dimnames(F_matrix) <- list(age = ages, year = years)
-knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
-```

### 2020-03-14 inst/rmd/vis_fit.Rmd
Revamped NCAM explorer

--- 2018-11-21 inst/rmd/vis_fit.Rmd
@@
 title: "NCAM Explorer"
-output: flexdashboard::flex_dashboard
-always_allow_html: yes
+output: 
+  flexdashboard::flex_dashboard:
+    css: style.css
 ---
-
-<style>
-/* list */
-li {
-line-height: 30px;
-}
-</style>
 
@@
 
-cols <- viridis::viridis(4)[-1]
+## https://projects.susielu.com/viz-palette?colors=[%22#243a68%22,%22#6a8b3e%22,%22#ff9800%22,%22#f44336%22,%22#9e9e9e%22]&backgroundColor=%22white%22&fontColor=%22black%22&mode=%22normal%22
+## https://material.io/resources/color/#!/?view.left=0&view.right=0&primary.color=243A68&secondary.color=9E9E9E
+
+cols <- c("#243a68", "#6a8b3e", "#ff9800", "#f44336", "#9e9e9e")
 cols_rgb <-  paste0("rgb(", apply(col2rgb(cols), 2, paste, collapse = ","), ")")
+obs_col <- cols[5]
+pred_col <- pos_col <- cols[1]
+neg_col <- cols[4]
+
 fit <- tmp$fit # temp object from vis_fit function
@@
 sd <- as.list(fit$sd.rep, "Std")
+est_lwr <- lapply(seq_along(est), function(i) est[[i]] - qnorm(0.975) * sd[[i]])
+est_upr <- lapply(seq_along(est), function(i) est[[i]] + qnorm(0.975) * sd[[i]])
+names(est_lwr) <- names(est_upr) <- names(est)
 tabs <- tidy_model(fit)
@@
 text_units <- function(x) {
-  if (x >= 1000000) return(paste0(round(x / 1000000), "M"))
-  if (x >= 1000) return(paste0(round(x / 1000), "k"))
-  if (x < 1000) return(as.character(round(x)))
+  ifelse(x >= 1000000000, paste0(round(x / 1000000000), "B"),
+         ifelse(x >= 1000000, paste0(round(x / 1000000), "M"),
+         ifelse(x >= 1000, paste0(round(x / 1000), "k"), 
+         as.character(round(x)))))
+}
+
+## Function for finding local maxima
+## from https://github.com/stas-g/findPeaks
+find_peaks <- function (x, m = 3){
+  shape <- diff(sign(diff(x, na.pad = FALSE)))
+  pks <- sapply(which(shape < 0), FUN = function(i){
+    z <- i - m + 1
+    z <- ifelse(z > 0, z, 1)
+    w <- i + m + 1
+    w <- ifelse(w < length(x), w, length(x))
+    if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
+  })
+  pks <- unlist(pks)
+  pks
 }
@@
 
-Background
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### NCAM
-
-<h1>NCAM: A state-space stock assessment model for Northern cod</h1>
-
-```{r}
-
-ax <- list(
-  title = "",
-  zeroline = FALSE,
-  showline = FALSE,
-  showticklabels = FALSE,
-  showgrid = FALSE
-)
-
-xy <- approx(x = years_plus1, y = fit$rep$abundance + 1000000, n = 50)
-plot_ly(x = xy$x, y = xy$y, hoverinfo = "none",
-        fillcolor = 'rgba(49, 104, 142, 0.3)',
-        type = 'scatter', mode = 'none', fill = 'tozeroy', showlegend = FALSE) %>%
-  add_trace() %>%
-  layout(xaxis = ax, yaxis = c(list(range = c(0, max(xy$y))), ax),
-         margin = list(l = 0, r = 0, b = 0, t = 0, pad = 0)) %>% 
-  plotly::config(displayModeBar = FALSE)
+```{r background, child = "background.Rmd"}
 
@@
 
-### General
-
-- An *integrated* *state-space* population dynamics model
-- *Integrated*
-    - Uses as much information as possible
-- *State-space*
-    - Includes process and observation equations + error
-    - Population processes are informed by observations from multiple sources
-    - Process errors are necessary for realistic stochastic projections
-
-### Data inputs
-
-- Reported landings (`r text_range(years)`)
-- Catch-at-age information (`r text_range(years)`)
-- Reported landings by month (`r text_range(years)`)
-- Offshore research vessel (RV) survey indices (`r text_range(c(1983, max(years)))`)
-- Inshore sentinel (SN) gillnet survey indices (`r text_range(fit$tmb.data$SN_years)`)
-- Inshore acoustic biomass surveys and age composition data from Smith Sound (SS; `r text_range(fit$tmb.data$SS_years)`)
-- Tagging data (`r text_range(years)`)
-
-### Biological inputs
-
-- RV survey data were used fit cohort-based models of:
-   - Proportion mature at age
-   - Beginning of year and mid-year stock weights at age
-- Baseline natural mortality (M) values
-
-```{r}
-
-plot_ly(x = years, y = ages, z = fit$rep$M_matrix, height = 400, width = 500) %>%
-  add_surface() %>%
-  colorbar(title = "Mshift") %>%
-  layout(scene = list(
-    xaxis = list(title = "Year"),
-    yaxis = list(title = "Age"),
-    zaxis = list(title = "Mshift")
-  ))
+```{r fishery, child = "fishery.Rmd"}
 
@@
 
-### Partial catches
-
-- User sets lower and upper bounds on total catch weight
-- Bounds are used via a censored likelihood
-
-```{r}
-
-plot_ly(x = years, y = fit$tmb.data$LRL, height = 300, width = 500) %>%
-  add_lines(name = "Lower", color = I(cols[1])) %>%
-  add_lines(y = fit$tmb.data$LRU, name = "Upper", color = I(cols[2])) %>%
-  add_lines(y = rep(1, length(years)), color = I("black"),
-            line = list(width = 1, dash = "dot"), showlegend = FALSE) %>%
-  layout(yaxis = list(title = "Catch multiplier"),
-         xaxis = list(title = "Year"))
+```{r surveys, child = "surveys.Rmd"}
 
@@
 
-### Change in Q
-
-- A shift in Q was assumed for the offshore RV survey
-- A variable portion of the stock is assumed to occupy Smith Sound between `r text_range(fit$tmb.data$SS_years)`
-- Q for the inshore sentinel survey was assumed to follow a random walk
-    - Done to account for potential changes in migration patterns
-
-### Tagging data
-
-- Age of tagged fish estimated from length-at-release using a spatiotemporal Von Bertalanffy growth model
-- Applies adjustments for tagging mortality, tag loss and reporting rates
-- Reporting rates estimated by an external model using the high-reward tag data collected since 1997
-- The Baranov catch equation was applied to age-based tag survivors
-- Assumes that initial fishing mortality experienced by fish in each experiment are random deviations from the fishing mortality experienced by the whole stock
-
-### References
-
-Cadigan, N.
-2015.
-A state-space stock assessment model for Northern cod, including under-reported catches and variable natural mortality rates.
-Canadian Journal of Fisheries and Aquatic Sciences 73 (2).
-NRC Research Press: 296-308.
-
-Cadigan, N. 2016. Updates to a Northern cod (*Gadus morhua*) state-space integrated assessment model. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/022. v + 58 p.
-
-Catch
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-
-plot_landings(fit)
+```{r tagging, child = "tagging.Rmd"}
 
@@
 
-> Y-axes in log scale
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted catch at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$catch_prop,
-                    pred = fit$rep$pred_prop_catch),
-               ages = ages, years = years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Proportion",
-               obs_col = cols[2], pred_col = cols[1]) %>%
-  layout(yaxis = list(rangemode = "tozero"))
+```{r trends, child = "trends.Rmd"}
 
@@
 
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$CA_resid,
-                    fit = fit$rep$pred_crl),
-               ages = ages[-length(ages)], years = years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
+```{r projections, child = "projections.Rmd"}
 
@@
 
-> based on continuation ratio logits
-
-### Catch at age surface (proportion)
-
-```{r}
-plot_surface(mat = fit$rep$pred_prop_catch, ages = ages, years = years, name = "P")
-```
-
-<!-- ### Catch at age surface (numbers) -->
-
-<!-- ```{r} -->
-<!-- plot_surface(mat = t(inputs$catch.num), ages = ages, years = years, name = "N") -->
-<!-- ``` -->
-
-<!-- > Note: model fits to composition - this plot is included for illustration -->
-
-RV Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### RV survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$est_wt * fit$tmb.data$RV_index * fit$tmb.data$trawl_units)
-pred <- colSums(fit$tmb.data$est_wt * exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units)
-obs[obs == 0] <- NA
-pred[pred == 0] <- NA
-d <- data.frame(year = years, obs = obs, pred = pred)
-
-# ## RV biomass
-# obs <- colSums(fit$tmb.data$RV_index * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# pred <- colSums(exp(fit$rep$Elog_RV_index) * fit$tmb.data$trawl_units * fit$tmb.data$weight)
-# d <- data.frame(year = years, obs = obs, pred = pred)
-# write.table(d, file = "clipboard", sep = "\t", row.names = FALSE, quote = FALSE)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "RV survey index",
-              showlegend = TRUE, type = "log")
+```{r tables, child = "tables.Rmd"}
 
 ```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    zero = fit$tmb.data$ind_zero,
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d <- d[d$wt == 1, ]
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$RV_index,
-                    pred = exp(fit$rep$Elog_RV_index),
-                    wt = fit$tmb.data$est_wt),
-               ages = ages, years = years)
-d$obs[d$wt == 0] <- NA
-d$pred[d$wt == 0] <- NA
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Mean numbers per tow",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$RV_resid,
-                    keep = fit$tmb.data$ind_zero == 0 & fit$tmb.data$est_wt == 1,
-                    fit = fit$rep$Elog_RV_index),
-               ages = ages, years = years)
-d <- d[d$keep, ]
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SN Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index
-
-```{r}
-
-obs <- colSums(fit$tmb.data$SN_GN_index)
-pred <- colSums(exp(fit$rep$Elog_SN_GN_index))
-d <- data.frame(year = fit$tmb.data$SN_years, obs = obs, pred = pred)
-
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Sentinel survey index",
-              showlegend = TRUE, type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$log_SN_GN_resid,
-              ages = fit$tmb.data$SN_ages,
-              years = fit$tmb.data$SN_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Sentinel gillnet survey index at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SN_GN_index,
-                    pred = exp(fit$rep$Elog_SN_GN_index)),
-               ages = fit$tmb.data$SN_ages, years = fit$tmb.data$SN_years)
-facet_obs_pred(d, by = "age",
-               which = c(3, 7, 4, 8, 5, 9, 6, 10), # done to order by column
-               nrows = 4, ylab = "Sentinel survey index",
-               obs_col = cols[2], pred_col = cols[1], type = "log")
-
-```
-
-> Y-axes in log scale
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$log_SN_GN_resid,
-                    fit = fit$rep$Elog_SN_GN_index),
-               ages = fit$tmb.data$SN_ages,
-               years = fit$tmb.data$SN_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-SS Survey
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Smith sound acoustic biomass estimates
-
-```{r}
-
-d <- data.frame(year = fit$tmb.data$SS_bms_years,
-                obs = fit$tmb.data$SSbms,
-                pred = fit$rep$ESSbms)
-plot_obs_pred(d, obs_col = cols[2], pred_col = cols[1], ylab = "Biomass (t)",
-              showlegend = TRUE)
-
-```
-
-### Standardized log residuals
-
-```{r}
-
-d <- tidy_mat(fit$rep$SS_agecomps_resid_std,
-              ages = fit$tmb.data$SS_ages,
-              years = fit$tmb.data$SS_age_years, value = "res")
-bubble_plot(d)
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Smith Sound trawl numbers at age
-
-```{r}
-
-d <- tidy_mats(list(obs = fit$tmb.data$SS_agecomps,
-                    pred = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-facet_obs_pred(d, by = "age",
-               which = c(2, 8, 3, 9, 4, 10, 5, 11, 6, 12, 7, 13), # done to order by column
-               nrows = 6, ylab = "Numbers caught", rangemode = "tozero",
-               obs_col = cols[2], pred_col = cols[1])
-
-```
-
-### Residuals by year, cohort, age and expected values
-
-```{r}
-
-d <- tidy_mats(list(res = fit$rep$SS_agecomps_resid_std,
-                    fit = fit$rep$ESS_agecomps),
-               ages = fit$tmb.data$SS_ages,
-               years = fit$tmb.data$SS_age_years)
-d$cohort <- d$year - d$age
-d$rec <- seq(nrow(d))
-plot_resid(d, col = cols[1])
-
-```
-
-Tagging {data-orientation=rows}
-================================================================================
-
-```{r}
-
-## Prep tag data and fits
-tag_dat <- data.frame(
-  exp = c(fit$tmb.data$old_tag_exp, fit$tmb.data$tag_exp),
-  n_rel = c(fit$tmb.data$old_N_rel, fit$tmb.data$N_rel),
-  rel_year = c(fit$tmb.data$old_rel_year, fit$tmb.data$rel_year) + min(years),
-  cap_year = c(fit$tmb.data$old_cap_year, fit$tmb.data$cap_year) + min(years),
-  age = c(fit$tmb.data$old_tag_age, fit$tmb.data$tag_age),
-  pred = c(fit$rep$old_tag_report, fit$rep$tag_report),
-  obs = c(fit$tmb.data$old_N_cap, fit$tmb.data$N_cap),
-  res = c(fit$rep$old_tag_std_resid, fit$rep$tag_std_resid),
-  period = c(rep("pre 1996", length(fit$tmb.data$old_tag_exp)),
-             rep("post 1996", length(fit$tmb.data$tag_exp)))
-)
-tag_dat$delay <- tag_dat$cap_year - tag_dat$rel_year
-tag_dat$text <- paste("exp:", tag_dat$exp, "</br>",
-                      "age:", tag_dat$age, "</br>",
-                      "delay:", tag_dat$delay, "</br>")
-
-Udev <- data.frame(
-  exp = c(unique(fit$tmb.data$old_tag_exp), unique(fit$tmb.data$tag_exp)),
-  Udev = c(fit$rep$old_Udev, fit$rep$Udev)
-)
-
-Fay <- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "Fay")
-names(Fay)[names(Fay) == "year"] <- "rel_year"
-
-tag_dat <- merge(tag_dat, Udev, by = "exp")
-tag_dat <- merge(tag_dat, Fay, by = c("rel_year", "age"))
-tag_dat$Fsay <- tag_dat$Fay * exp(tag_dat$Udev)
-
-## Share data across multiple plots
-d <- crosstalk::SharedData$new(tag_dat, ~exp, "Select an experiment")
-base <- plot_ly(d, symbol = ~period, colors = cols[1:2], symbols = c(16, 1)) %>% group_by(exp)
-
-```
-
-Row {data-height=700 .tabset}
---------------------------------------------------------------------------------
-
-### Observed and predicted tag recaptures by experimemnt
-
-```{r}
-
-y_ticktext <- c(50, 500, pretty(c(0, 6000), 4))
-y_tickvals <- sqrt(y_ticktext)
-
-rel_plot <- base %>%
-  filter(rel_year == cap_year) %>%
-  group_by(exp) %>%
-  summarise(rel_year = unique(rel_year), tot_rel = sum(n_rel), period = unique(period)) %>%
-  add_markers(x = ~rel_year, y = ~sqrt(tot_rel), color = I(cols[1]),
-              text = ~paste("exp:", exp, "</br>", "N:", tot_rel)) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N released",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-y_ticktext <- c(5, 20, pretty(c(0, 500), 10))
-y_tickvals <- sqrt(y_ticktext)
-
-cap_plot <- base %>%
-  group_by(exp, rel_year, cap_year) %>%
-  summarise(sum_pred = sum(pred), sum_obs = sum(obs),
-            period = unique(period)) %>%
-  group_by(exp) %>%
-  add_markers(x = ~cap_year, y = ~sqrt(sum_obs), color = I(cols[1]), name = "",
-              showlegend = FALSE, text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>%
-  add_lines(x = ~cap_year, y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted",
-            alpha = 0.7, showlegend = FALSE, symbol = NULL,
-            text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = 0, y1 = 1000, yref = "y2",
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "N recaptured",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext,
-                      range = range(y_tickvals)),
-         xaxis = list(title = "Year"))
-
-subplot(rel_plot, cap_plot, shareX = TRUE, titleX = TRUE, titleY = TRUE,
-        nrows = 2, heights = c(0.3, 0.7))
-
-```
-
-> Y-axis is in square root scale
-
-<!-- ### Animation -->
-
-<!-- ```{r} -->
-
-<!-- sum_d <- tag_dat %>% -->
-<!--   group_by(exp, cap_year) %>% -->
-<!--   summarise(sum_pred = sum(pred), sum_obs = sum(obs)) -->
-
-<!-- sum_d %>%  -->
-<!--   group_by(exp) %>%  -->
-<!--   plot_ly(x = ~cap_year, frame = ~exp) %>% -->
-<!--   add_markers(y = ~sqrt(sum_obs), color = I(cols[1]), name = "Observed", -->
-<!--               text = ~paste("exp:", exp, "</br>", "observed N:", sum_obs)) %>% -->
-<!--   add_lines(y = ~sqrt(sum_pred), color = I(cols[2]), name = "Predicted", -->
-<!--             text = ~paste("exp:", exp, "</br>", "predicted N:", round(sum_pred))) %>% -->
-<!--   animation_opts(1000, transition = 500, easing = "exp", redraw = FALSE) %>% -->
-<!--   layout(yaxis = list(title = "N recaptured", -->
-<!--                       tickvals = y_tickvals, -->
-<!--                       ticktext = y_ticktext, -->
-<!--                       range = range(y_tickvals)), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-Row
---------------------------------------------------------------------------------
-
-### Tag reporting rate
-
-```{r}
-
-cov_mat <- Matrix::as.matrix(fit$tmb.data$SRR_cov_matrix)
-rr <- data.frame(year = fit$tmb.data$SRR_years,
-                 est = fit$tmb.data$log_SRR_est,
-                 sd = sqrt(diag(cov_mat)))
-rr$lwr <- rr$est - 1.96 * rr$sd
-rr$upr <- rr$est + 1.96 * rr$sd
-rr <- trans_est(rr)
-rr$model <- "Random Walk"
-
-ncam_rr <- data.frame(year = fit$tmb.data$SRR_years,
-                      est = est$log_SRR, sd = sd$log_SRR)
-ncam_rr$lwr <- ncam_rr$est - 1.96 * ncam_rr$sd
-ncam_rr$upr <- ncam_rr$est + 1.96 * ncam_rr$sd
-ncam_rr <- trans_est(ncam_rr)
-ncam_rr$model <- "NCAM"
-
-rr <- rbind(rr, ncam_rr)
-
-plot_trend(rr, group = "model", col = cols, ylab = "Reporting rates", annotate = TRUE)
-
-```
-
-### Initial F deviations for tagging experiments
-
-```{r}
-
-base %>%
-  distinct(rel_year, Udev, exp, period) %>%
-  add_markers(x = ~rel_year, y = ~Udev, text = ~exp, color = I(cols[1]),
-              showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 1996, x1 = 1996,
-                       y0 = min(tag_dat$Udev), y1 = max(tag_dat$Udev),
-                       line = list(width = 1, dash = "dot")),
-         yaxis = list(title = "Initial F deviation"),
-         xaxis = list(title = "Year"))
-
-```
-
-### Observed vs. predicted recaptures
-
-```{r}
-
-x_ticktext <- c(5, pretty(tag_dat$pred, 10))
-y_ticktext <- c(5, pretty(tag_dat$obs, 10))
-x_tickvals <- sqrt(x_ticktext)
-y_tickvals <- sqrt(y_ticktext)
-
-base %>%
-  add_markers(x = ~sqrt(pred), y = ~sqrt(obs),
-              text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(shapes = list(type = "line", x0 = 0, x1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       y0 = 0, y1 = max(sqrt(c(tag_dat$pred, tag_dat$obs))),
-                       line = list(width = 1)),
-         xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Observed number of recaptures",
-                      tickvals = y_tickvals,
-                      ticktext = y_ticktext))
-
-```
-
-> Axes are in square root scale
-
-### Standardized residuals vs. predicted recaptures
-
-```{r}
-
-y_ticktext <- c(5, pretty(tag_dat$res, 10))
-y_tickvals <- sqrt(x_ticktext)
-
-base %>%
-  add_markers( x = ~sqrt(pred), y = ~res,
-               text = ~text, color = I(cols[1]), showlegend = FALSE) %>%
-  layout(xaxis = list(title = "Predicted number of recaptures",
-                      tickvals = x_tickvals,
-                      ticktext = x_ticktext),
-         yaxis = list(title = "Standardized residuals"))
-
-```
-
-> X-axis is in square root scale
-
-Catchability
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Fishery selectivity
-
-```{r}
-
-S <- apply(fit$rep$F_matrix, 2, function(x) x / max(x))
-plot_surface(S, ages = ages, years = years, name = "S")
-
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### RV survey catchability
-
-```{r}
-
-q <- exp(fit$rep$log_RV_Qa + log(fit$tmb.data$trawl_units))
-plot_surface(q, ages = ages, years = years, name = "q")
-
-```
-
-### Proportion of the stock in Smith Sound
-
-```{r}
-p <- matrix(0, nrow = length(ages), ncol = length(years),
-            dimnames = list(ages, years))
-ss_p <- fit$rep$SSD
-ss_years <- fit$tmb.data$SS_years
-ss_ages <- fit$tmb.data$SS_ages
-rownames(ss_p) <- ss_ages
-colnames(ss_p) <- ss_years
-p[as.character(ss_ages), as.character(ss_years)] <- ss_p
-plot_surface(p, ages = ages, years = years, name = "p")
-
-```
-
-### Sentinel gillnet survey catchability
-
-```{r}
-
-log_qy <- t(replicate(length(fit$tmb.data$SN_ages), fit$rep$SN_GN_Qy))
-log_qa <- fit$rep$log_SN_GN_Qa
-q <- exp(log_qy + log_qa)
-plot_surface(q, ages = fit$tmb.data$SN_ages,
-             years = fit$tmb.data$SN_years,
-             name = "q")
-
-```
-
-Trends
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r}
-
-## Note: biomass is in tonnes and numbers are in thousands
-bms <- rbind(data.frame(metric = "SSB", tabs$log_ssb),
-             data.frame(metric = "Ages 2+", tabs$log_biomass))
-abn <- rbind(data.frame(metric = "Age 2", tabs$log_recruitment),
-             data.frame(metric = "Ages 2+", tabs$log_abundance))
-
-bms_plot <- trans_est(bms) %>%
-  plot_trend(ylab = "Biomass (t)", group = "metric", col = cols,
-             annotate = TRUE)
-
-abn_plot <- trans_est(abn, scale = 0.001) %>%
-  plot_trend(ylab = "Abundance", group = "metric", col = cols,
-             annotate = TRUE)
-
-blim_plot <- trans_est(tabs$log_ssb_lrp) %>%
-  plot_trend(ylab = "SSB/Blim", annotate = TRUE, col = cols) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none") %>%
-  add_text(x = max_year + 1.2, y = 1, text = paste(round(fit$rep$blim / 1000), "kt"),
-           color = I("black"), textposition = "right", hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Abundance at age
-
-```{r}
-plot_surface(fit$rep$N_matrix * 1000, ages = ages, years = years_plus1, name = "N")
-```
-
-### Biomass at age
-
-```{r}
-plot_surface(fit$rep$biomass_matrix, ages = ages, years = years_plus1, name = "Biomass")
-```
-
-### SSB at age
-
-```{r}
-plot_surface(fit$rep$ssb_matrix, ages = ages, years = years_plus1, name = "SSB")
-```
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r}
-
-f <- data.frame(metric = "Ages 5+", tabs$log_aveF)
-z <- data.frame(metric = "Ages 5+", tabs$log_aveZ)
-m <- data.frame(metric = "Ages 5+", tabs$log_aveM)
-
-z_plot <- trans_est(z) %>%
-  plot_trend(ylab = "Average Z", group = "metric",
-             annotate = TRUE, col = cols)
-
-m_plot <- trans_est(m) %>%
-  plot_trend(ylab = "Average M", group = "metric",
-             annotate = TRUE, col = cols)
-
-f_plot <- trans_est(f) %>%
-  plot_trend(ylab = "Average F", group = "metric",
-             annotate = TRUE, col = cols)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-### Z at age
-
-```{r}
-plot_surface(fit$rep$Z_matrix, ages = ages, years = years, name = "Z")
-```
-
-### M at age
-
-```{r}
-plot_surface(fit$rep$Mpe_matrix, ages = ages, years = years, name = "M")
-```
-
-### F at age
-
-```{r}
-plot_surface(fit$rep$F_matrix, ages = ages, years = years, name = "F")
-```
-
-### Z, M and F
-
-```{r}
-plot_zmf_surface(z = fit$rep$Z_matrix, m = fit$rep$Mpe_matrix, f = fit$rep$F_matrix,
-                 ages = ages, years = years)
-```
-
-> Z = purples, M = blues, F = reds
-
-### Process error
-
-```{r}
-
-pe <- fit$rep$pe
-plot_surface(pe, years = c(years, fit$tmb.data$proj_years)[-1], ages = 2:8,
-             name = "pe", col = RColorBrewer::brewer.pal(9, "RdBu")) %>%
-  add_surface(matrix(0, nrow = nrow(pe), ncol = ncol(pe)), showscale = FALSE,
-              color = I("black"), opacity = 0.8)
-
-```
-
-### Process error (bubble plot)
-
-```{r}
-pe <- tidy_mat(fit$rep$pe, years = c(years, fit$tmb.data$proj_years)[-1],
-               ages = 2:8, value = "res")
-bubble_plot(pe)
-```
-
-Retro {`r if (is.null(fit$retro)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(fit$retro)}
-
-bms_plot <- trans_est(retro_tabs$log_ssb) %>%
-  plot_retro(ylab = "SSB (t)", lwd = 1, col = cols[1])
-
-abn_plot <- trans_est(retro_tabs$log_recruitment, scale = 0.001) %>%
-  plot_retro(ylab = "Recruitment (age 2)", lwd = 1, col = cols[1])
-
-blim_plot <- trans_est(retro_tabs$log_ssb_lrp) %>%
-  plot_retro(ylab = "SSB/Blim", lwd = 1, col = cols[1]) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none")
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(fit$retro)}
-
-z_plot <- trans_est(retro_tabs$log_aveZ) %>%
-  plot_retro(ylab = "Average Z (ages 5+)", lwd = 1, col = cols[1])
-
-m_plot <- trans_est(retro_tabs$log_aveM) %>%
-  plot_retro(ylab = "Average M (ages 5+)", lwd = 1, col = cols[1])
-
-f_plot <- trans_est(retro_tabs$log_aveF) %>%
-  plot_retro(ylab = "Average F (ages 5+)", lwd = 1, col = cols[1])
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Comps {`r if (is.null(tmp$comp)) ".hidden"`}
-================================================================================
-
-Row
---------------------------------------------------------------------------------
-
-### Stock status
-
-```{r, eval = !is.null(tmp$comp)}
-
-abn_plot <- trans_est(comp_tabs$log_recruitment, scale = 0.001) %>% 
-  plot_comps(ylab = "Recruitment (age 2)", col = cols)
-
-bms_plot <- trans_est(comp_tabs$log_ssb) %>% 
-  plot_comps(ylab = "SSB (t)", col = cols, showlegend = FALSE)
-
-blim_plot <- trans_est(comp_tabs$log_ssb_lrp) %>%
-  plot_comps(ylab = "SSB/Blim", col = cols, showlegend = FALSE) %>%
-  add_segments(x = min_year, xend = max_year + 1, y = 1, yend = 1, color = I("black"),
-               linetype = I(3), name = "Blim", text = NULL, hoverinfo = "none", showlegend = FALSE)
-
-subplot(abn_plot, bms_plot, blim_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Row
---------------------------------------------------------------------------------
-
-### Mortality Rates
-
-```{r, eval = !is.null(tmp$comp)}
-
-z_plot <- trans_est(comp_tabs$log_aveZ) %>%
-  plot_comps(ylab = "Average Z (ages 5+)", col = cols)
-
-m_plot <- trans_est(comp_tabs$log_aveM) %>%
-  plot_comps(ylab = "Average M (ages 5+)", col = cols, showlegend = FALSE)
-
-f_plot <- trans_est(comp_tabs$log_aveF) %>%
-  plot_comps(ylab = "Average F (ages 5+)", col = cols, showlegend = FALSE)
-
-subplot(z_plot, m_plot, f_plot, nrows = 3, shareX = TRUE, titleY = TRUE)
-
-```
-
-Productivity
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Recruitment
-
-```{r}
-trans_est(tabs$log_recruitment, scale = 0.001) %>%
-  plot_trend(ylab = "Recruitment", col = cols[1])
-```
-
-### SSB
-
-```{r}
-trans_est(tabs$log_ssb) %>%
-  plot_trend(ylab = "SSB (t)", col = cols[1])
-```
-
-### RPS
-
-```{r}
-trans_est(tabs$log_rps) %>%
-  filter(year >= min_year + min(ages)) %>% 
-  plot_trend(ylab = "RPS (thousand / tonne)", col = cols[1]) %>% 
-  layout(xaxis = list(range = c(min_year, max_year + 1)))
-```
-
-### SSB x Recruitment
-
-```{r}
-
-ssb <- trans_est(tabs$log_ssb)
-rec <- trans_est(tabs$log_recruitment, scale = 0.001)
-rps <- trans_est(tabs$log_rps)
-names(ssb) <- paste0("ssb_", names(ssb))
-names(rec) <- paste0("rec_", names(rec))
-names(rps) <- paste0("rps_", names(rps))
-ssb$year <- ssb$ssb_year + min(ages)
-rec$year <- rec$rec_year
-rps$year <- rps$rps_year
-sr <- merge(ssb, rec, by = "year")
-sr <- merge(sr, rps, by = "year")
-sr$lab <- paste0("Recruitment year: ", sr$rec_year, "<br>SSB year: ", sr$ssb_year)
-
-plot_xy_errorbar(sr, x = ~ssb_est, y = ~rec_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rec_lwr, y_upr = ~rec_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "Recruitment",
-                 col = cols[1])
-
-```
-
-### SSB x RPS
-
-```{r}
-p <- plot_xy_errorbar(sr, x = ~ssb_est, y = ~rps_est,
-                 x_lwr = ~ssb_lwr, x_upr = ~ssb_upr,
-                 y_lwr = ~rps_lwr, y_upr = ~rps_upr,
-                 text = ~lab, xlab = "SSB (t)", ylab = "RPS (thousand / tonne)",
-                 col = cols[1])
-p
-```
-
-### log(SSB) x log(RPS)
-
-```{r}
-p %>% layout(yaxis = list(type = "log"),
-             xaxis = list(type = "log"))
-```
-
-```{r}
-
-message("Bits of the projection code is currently not working. Specifically the plots that require the frame argument in plotly.")
-
-```
-
-<!-- Assumptions  {data-navmenu="Projections" .storyboard} -->
-<!-- ================================================================================ -->
-
-<!-- ### Assume the same selectivity curve as estimated by NCAM in the terminal year and apply catch multipliers -->
-
-<!-- ```{r} -->
-
-<!-- Cp <- fit$tmb.data$Cp -->
-<!-- Cp_string <- text_c(Cp) -->
-<!-- Lp <- round(Cp * tail(fit$rep$pred_catch_wt, 1)) -->
-<!-- Lp_string <- paste(text_c(Lp), "t") -->
-
-<!-- tidy_mat(fit$rep$F_matrix, ages = ages, years = years, value = "F") %>% -->
-<!--   filter(year >= max(years) - 4) %>% -->
-<!--   plot_ly(x = ~age, y = ~F, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   filter(year == max(years)) %>% -->
-<!--   add_lines(showlegend = FALSE, line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age")) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - The following catch multipliers were applied: `r #Cp_string` -->
-<!-- - This translates to `r #Lp_string` of catch in the projection years -->
-
-<!-- ### Assume natural mortality converges on the long-term mean, excluding the M spike between 1991 - 1994 -->
-
-<!-- ```{r} -->
-
-<!-- if (all(fit$tmb.data$Mp_mult != 1)) stop("TODO: make the mortality multiplier projection documentation dynamic") -->
-
-<!-- aveM <- tabs$log_aveM -->
-<!-- aveM <- exp(mean(aveM$est[!aveM$year %in% 1991:1994])) -->
-
-<!-- trans_est(tabs$log_aveM) %>% -->
-<!--   plot_trend(ylab = "Average M", col = cols[1]) %>% -->
-<!--   add_segments(x = 1995, xend = max(years) + 3, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_segments(x = 1990, xend = 1995, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), -->
-<!--                line = list(width = 1, dash = "dot")) %>% -->
-<!--   add_segments(x = min(years), xend = 1990, y = aveM, yend = aveM, -->
-<!--                name = "long-term mean", color = I("black"), line = list(width = 1)) %>% -->
-<!--   add_text(x = max(years) + 0.2, y = aveM, text = "long-term mean", textposition = "bottom right", -->
-<!--            color = I("black")) %>% -->
-<!--   layout(xaxis = list(range = c(min(years), max(years) + 3))) -->
-
-<!-- ``` -->
-
-<!-- ### Assume a relationship between recruits per spawner and number of spawners to project recruits in the future -->
-
-<!-- ```{r} -->
-
-<!-- if (fit$tmb.data$Rp_mode != 3) stop("TODO: make the recruitment mode documentation dynamic") -->
-
-<!-- rps_tab <- tabs$log_rps -->
-<!-- names(rps_tab) <- c("year", "rps", "rps_sd", "rps_lwr", "rps_upr") -->
-<!-- rps_tab <- rps_tab[rps_tab$year >= 1985, ] -->
-<!-- rps_tab$year <- rps_tab$year - 2 # scale back two year to allign with spawners that produced the recruits -->
-<!-- ssa_tab <- tabs$log_ssa -->
-<!-- names(ssa_tab) <- c("year", "ssa", "ssa_sd", "ssa_lwr", "ssa_upr") -->
-<!-- d <- merge(ssa_tab, rps_tab, by = "year") -->
-<!-- rec_tab <- tabs$log_recruitment -->
-<!-- names(rec_tab) <- c("year", "rec", "rec_sd", "rec_lwr", "rec_upr") -->
-<!-- rec_tab <- rec_tab[rec_tab$year >= 1985, ] -->
-<!-- rec_tab$year <- rec_tab$year - 2 # scale back two years -->
-<!-- d <- merge(d, rec_tab, by = "year") -->
-<!-- d$excluded <- d$year %in% 1991:1994 -->
-
-<!-- sub_d <- d[!d$year %in% 1991:1994, ] -->
-<!-- lin_rel <- lm(rps ~ ssa, data = sub_d) -->
-<!-- # coefficients(lin_rel) -->
-
-<!-- preds <- data.frame(year = NA, rps_sd = NA, ssa_sd = NA, ssa = seq(min(d$ssa), max(d$ssa), length.out = 100)) -->
-<!-- preds$fits <- predict(lin_rel, newdata = preds) -->
-<!-- preds$excluded <- FALSE -->
-
-<!-- plot_ly(data = d, x = ~ssa, y = ~rps, text = ~year, -->
-<!--         type = "scatter", mode = "markers", name = "est", -->
-<!--         color = ~excluded, colors = c(cols[2], "grey"), -->
-<!--         error_y = list(array = ~ qnorm(0.975) * rps_sd, -->
-<!--                        width = 0, thickness = 1), -->
-<!--         error_x = list(array = ~ qnorm(0.975) * ssa_sd, -->
-<!--                        width = 0, thickness = 1)) %>% -->
-<!--   add_text(text = ~year, textfont = list(color = toRGB("grey50")), textposition = "top right", -->
-<!--            error_y = list(visible = FALSE), -->
-<!--            error_x = list(visible = FALSE)) %>% -->
-<!--   layout(title = "Spawners vs Recruits / Spawner", -->
-<!--          xaxis = list(title = "log(Spawners)"), -->
-<!--          yaxis = list(title = "log(Recruits / Spawner)", zerolinecolor = toRGB("lightgrey")), -->
-<!--          showlegend = FALSE) %>% -->
-<!--   add_lines(data = preds, x = ~ssa, y = ~fits, name = "fit", color = I(cols[1])) -->
-
-<!-- ``` -->
-
-<!-- *** -->
-<!-- - Grey points indicate years excluded from the regressionn -->
-
-<!-- ### Assume averages weights and proportions mature from the last three years -->
-
-<!-- ```{r} -->
-
-<!-- mat_p <- tidy_mat(fit$tmb.data$mat, ages = factor(ages), years = years, value = "mat") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~mat, color = ~factor(year), colors = rev(cols)) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(mat = mean(mat), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Proportion mature")) -->
-
-<!-- wt_p <- tidy_mat(fit$tmb.data$weight, ages = factor(ages), years = years, value = "weight") %>% -->
-<!--   filter(year >= max(years) - 2) %>% -->
-<!--   plot_ly(x = ~age, y = ~weight, color = ~factor(year), colors = rev(cols), -->
-<!--           showlegend = FALSE) %>% -->
-<!--   add_lines(alpha = 0.5) %>% -->
-<!--   group_by(age) %>% -->
-<!--   summarise(weight = mean(weight), year = "3y mean") %>% -->
-<!--   add_lines(line = list(width = 3)) %>% -->
-<!--   layout(xaxis = list(title = "Age"), -->
-<!--          yaxis = list(title = "Weight (kg)")) -->
-
-<!-- subplot(mat_p, wt_p, titleY = TRUE, titleX = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Past projections {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- ### Catch multiplier projections for SSB / Blim from the last RAP -->
-
-<!-- ```{r, eval = !is.null(tmp$last)} -->
-
-<!-- d <- trans_est(last_tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- d <- d[d$year <= max(years) + 1, ] -->
-
-<!-- past_d <- trans_est(last_tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(last$tmb.data$years) - 5.1, max(years) + 1.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- current_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- current_d <- lapply(levels(d$scenario), function(s) data.frame(current_d, scenario = s)) -->
-<!-- current_d <- do.call(rbind, current_d) -->
-<!-- current_d$model <- "current RAP" -->
-<!-- d$model <- "last RAP" -->
-<!-- d <- rbind(d, current_d) -->
-
-<!-- plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", group = "model", -->
-<!--           vline = max(last$tmb.data$years) + 1, hline = 1, col = cols, showlegend = TRUE) -->
-
-<!-- ``` -->
-
-<!-- Results {data-navmenu="Projections" `r if (is.null(tmp$last)) ".hidden"`} -->
-<!-- ================================================================================ -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Catch multiplier projections for SSB~yp~ / SSB~`r max(years) + 1`~ -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_ratio) -->
-<!-- d[, c("est", "lwr", "upr")] <- (d[, c("est", "lwr", "upr")] - 1) * 100 -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- xlim <- c(max(years) + 0.9, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "Percent growth", ticksuffix = "%", -->
-<!-- #           vline = max(years) + 1, col = cols[1]) %>% -->
-<!-- #   layout(yaxis = list(tickvals = seq(-200, 200, by = 25))) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- ### SSB / Blim -->
-
-<!-- ```{r} -->
-
-<!-- d <- trans_est(tabs$Cproj_log_ssb_lrp) -->
-<!-- d$scenario <- factor(d$scenario) -->
-
-<!-- past_d <- trans_est(tabs$log_ssb_lrp) -->
-<!-- past_d <- lapply(levels(d$scenario), function(s) data.frame(past_d, scenario = s)) -->
-<!-- past_d <- do.call(rbind, past_d) -->
-<!-- d <- rbind(past_d, d) -->
-<!-- xlim <- c(max(years) - 2.1, max(years) + 4.1) -->
-<!-- ylim <- range(d[d$year >= xlim[1] & d$year <= xlim[2], c("est", "lwr", "upr"), ]) -->
-
-<!-- # plot_proj(data = d, xlim = xlim, ylim = ylim, ylab = "SSB/Blim", -->
-<!-- #           vline = max(years) + 1, hline = 1, col = cols[1]) -->
-<!-- message("plot_proj not working properly at the moment") -->
-
-<!-- ``` -->
-
-<!-- Row {.tabset} -->
-<!-- -------------------------------------------------------------------------------- -->
-
-<!-- ### Prob(SSB~yp~ &ge; SSB~`r max(years) + 1`~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_ratio -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-<!-- ### Prob(SSB~yp~ &ge; B~lim~) -->
-
-<!-- ```{r} -->
-
-<!-- growth <- tabs$Cproj_log_ssb_lrp -->
-<!-- growth$p0 <- 1 - pnorm(0, mean = growth$est, sd = growth$sd) -->
-
-<!-- d <- growth[growth$year != max(years) + 1, ] -->
-<!-- d$scenario <- factor(d$scenario) -->
-<!-- plot_ly(data = d, x = ~year, y = ~scenario, z = ~p0) %>% -->
-<!--   add_heatmap() %>% -->
-<!--   add_annotations(x = ~year, y = ~scenario, text = ~round(p0, 2), -->
-<!--                   showarrow = FALSE, font = list(color = "white")) %>% -->
-<!--   colorbar(title = "P", limits = c(0, 1)) %>% -->
-<!--   layout(yaxis = list(title = "Catch multiplier"), -->
-<!--          xaxis = list(title = "Year")) -->
-
-<!-- ``` -->
-
-Inputs {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Landings
-
-```{r}
-kable(inputs$landings, caption = "Reported landings (t) of cod from NAFO Divs. 2J+3KL from 1959 onward")
-```
-
-### Catch at age
-
-```{r}
-kable(inputs$catch.num, caption = "Catch numbers-at-age (thousands) for the commercial cod fishery in NAFO Divs. 2J+3KL.")
-```
-
-<!-- ### Monthly landings -->
-
-<!-- ```{r} -->
-<!-- d <- inputs$catch.month -->
-<!-- dimnames(d) <- list(year = rownames(d), month = month.name) -->
-<!-- kable(d, caption = "Monthly reported landings of cod from NAFO Divs. 2J+3KL.") -->
-<!-- ``` -->
-
-### RV index
-
-```{r}
-d <- data.table::dcast(inputs$index, year ~ age, value.var = "index")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Mean number of cod per tow at age in the index strata for the autumn DFO RV bottom-trawl surveys of NAFO Divs. 2J+3KL.", digits = 3)
-```
-
-### SN index
-
-```{r}
-d <- data.table::dcast(inputs$sentinel, year ~ AGE, value.var = "EST")
-mat <- d[, as.character(3:10)]
-rownames(mat) <- d$year
-kable(mat, caption = 'Standardized age-disaggregated sentinel gill-net (5 1/2" mesh) catch rates (fish / net) of cod in NAFO Divs. 2J+3KL.', digits = 3)
-```
-
-### SS index
-
-```{r}
-d <- inputs$SS.bms[, c("Year", "month", "Biomass", "se2")]
-names(d) <- c("Year", "Month", "Biomass", "SE^2^")
-kable(d, caption = "Acoustic biomass estimates (t) for Smith Sound, Trinity Bay, from Rose et al. (2011).")
-```
-
-### SS age composition
-
-```{r}
-kable(inputs$SS_agecomps, caption = "Age composition of cod sampled during acoustic surveys in Smith Sound during 1995-2003. For details see Rose et al. 2011.")
-```
-
-### Weight at age
-
-```{r}
-d <- inputs$pwts
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "weight")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Beginning of year weight-at-age estimates (kg) from a generalized Von Bertalanffy (VonB2) growth model described in Cadigan (2016) fitted by cohort to average weight at age data for cod from autumn bottom-trawl surveys in Division 2J+3KL.", digits = 3)
-```
-
-### Maturity at age
-
-```{r}
-d <- inputs$mats
-d <- d[d$year <= max(inputs$index$year), ]
-d <- data.table::dcast(d, year ~ age, value.var = "mat")
-mat <- d[, as.character(2:14)]
-rownames(mat) <- d$year
-kable(mat, caption = "Estimated proportions mature for female cod from NAFO Div.
-2J+3KL from DFO RV autumn bottom trawl surveys.
-Estimates were obtained from a cohort-specific binomial logistic regression model fitted to observed proportions mature at age.", digits = 3)
-```
-
-### Tag summary
-
-```{r}
-d <- inputs$mtag1 %>% 
-  group_by(rel_year, exp) %>% 
-  summarise(NAFO = unique(div),
-            Nrel = sum(unique(released)), 
-            Ncap = sum(recaptured))
-names(d) <- c("Year", "Experiment", "NAFO division", "N~released~", "N~recaptured~")
-kable(d, caption = "Number of tagged cod released and recaptured by experiment.")
-```
-
-Settings {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Catch bounds
-
-```{r}
-d <- data.frame(Year = fit$tmb.data$years, Lower = fit$tmb.data$LRL, Upper = fit$tmb.data$LRU)
-kable(d, caption = "Lower and upper catch multipliers used for generating catch bounds.")
-```
-
-### Baseline M
-
-```{r}
-m <- fit$tmb.data$M_matrix
-rownames(m) <- ages
-colnames(m) <- years
-kable(t(m), caption = "Baseline levels of natural mortality (M) at age.")
-```
-
-Output {data-navmenu="Tables"}
-================================================================================
-
-Row {.tabset}
---------------------------------------------------------------------------------
-
-### Parameter estimates
-
-```{r}
-
-par_tab <- tidy_pars(fit)
-rownames(par_tab) <- NULL
-options(knitr.kable.NA = '')
-kable(par_tab, digits = 3, caption = "Estimates of model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Parameter estimates (retro)
-
-```{r}
-
-par_tabs <- lapply(names(fit$retro), function(nm) {
-  d <- data.frame(tyear = as.numeric(nm), tidy_pars(fit$retro[[nm]]))
-  d$id <- seq(nrow(d))
-  d
-})
-par_tabs <- do.call(rbind, par_tabs)
-final_par_tab <- data.frame(tyear = max(years), par_tab)
-final_par_tab$id <- seq(nrow(final_par_tab))
-par_tabs <- rbind(par_tabs, final_par_tab)
-par_tabs <- par_tabs[order(par_tabs$id, par_tabs$tyear), ]
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-par_tabs$Quantity <- as.character(par_tabs$Quantity)
-par_tabs$Quantity[duplicated(par_tabs$Quantity)] <- ""
-par_tabs$Symbol <- as.character(par_tabs$Symbol)
-par_tabs$Symbol[duplicated(par_tabs$Symbol)] <- ""
-par_tabs <- par_tabs[, c("Quantity", "Symbol", "tyear", "Estimate", "CV")]
-rownames(par_tabs) <- NULL
-names(par_tabs)[names(par_tabs) == "tyear"] <- "Terminal Year"
-
-kable(par_tabs, digits = 3, caption = "Estimates of retrospective model fits, population parameters and variance parameters, with coefficients of variation.")
-
-```
-
-> Note: RV = DFO Research Vessel survey; SN = Sentinel survey; SS = Smith Sound; D = proportion of stock in Smith Sound; NB = negative binomial
-
-### Stock size
-
-```{r}
-
-tab_name <- c("log_abundance" = "Abundance (millions)", 
-              "log_recruitment" = "Recruits (age 2; millions)", 
-              "log_biomass" = "Biomass (kt)", 
-              "log_ssa" = "Spawners (millions)", 
-              "log_ssb" = "SSB (kt)", 
-              "log_ssb_lrp" = "SSB/Blim n(%)")
-
-nms <- names(tab_name)
-status_tab <- trans_est(tabs[[nms[1]]], scale = 1000)[, c("year", "est", "lwr", "upr")]
-names(status_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  s <- ifelse(n == "log_ssb_lrp", 0.01, 1000)
-  tab <- trans_est(tabs[[n]], scale = s)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  status_tab <- merge(status_tab, tab, by = "Year")
-}
-
-knitr::kable(status_tab, digits = 0, caption = "Northern cod stock size estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Rates
-
-```{r}
-
-tab_name <- c("log_aveZ" = "Average Z (ages 5+)", 
-              "log_aveM" = "Average M (ages 5+)", 
-              "log_aveF" = "Average F (ages 5+)")
-
-nms <- names(tab_name)
-rate_tab <- trans_est(tabs[[nms[1]]])[, c("year", "est", "lwr", "upr")]
-names(rate_tab) <- c("Year", paste(tab_name[[nms[1]]], c("", "L 95% CI", "U 95% CI")))
-
-for (n in nms[-1]) {
-  tab <- trans_est(tabs[[n]], scale = 1)[, c("year", "est", "lwr", "upr")]
-  names(tab) <- c("Year", paste(tab_name[[n]], c("", "L 95% CI", "U 95% CI")))
-  rate_tab <- merge(rate_tab, tab, by = "Year")
-}
-
-knitr::kable(rate_tab, digits = 3, caption = "Northern cod stock mortality rate estimates with lower (L) and upper (U) 95% confidence intervals (CI)")
-
-```
-
-### Abundance-at-age
-
-```{r}
-N_matrix <- fit$rep$N_matrix
-dimnames(N_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(N_matrix) / 1000, caption = "Abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Mature abundance-at-age
-
-```{r}
-ssa_matrix <- fit$rep$ssa_matrix
-dimnames(ssa_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssa_matrix) / 1000, caption = "Mature abundance-at-age estimates (millions).", digits = 2)
-```
-
-### Biomass-at-age
-
-```{r}
-B_matrix <- fit$rep$biomass_matrix
-dimnames(B_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(B_matrix) / 1000, caption = "Biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Mature biomass-at-age
-
-```{r}
-ssb_matrix <- fit$rep$ssb_matrix
-dimnames(ssb_matrix) <- list(age = ages, year = years_plus1)
-knitr::kable(t(ssb_matrix) / 1000, caption = "Mature biomass-at-age estimates (kt).", digits = 2)
-```
-
-### Z-at-age
-
-```{r}
-Z_matrix <- fit$rep$Z_matrix
-dimnames(Z_matrix) <- list(age = ages, year = years)
-knitr::kable(t(Z_matrix), caption = "Z-at-age estimates.", digits = 3)
-```
-
-### M-at-age
-
-```{r}
-M_matrix <- fit$rep$Mpe_matrix
-dimnames(M_matrix) <- list(age = ages, year = years)
-M_matrix <- M_matrix[as.character(2:8), ]
-rownames(M_matrix) <- c(as.character(2:7), "8+")
-knitr::kable(t(M_matrix), caption = "M-at-age estimates.", digits = 3)
-```
-
-### F-at-age
-
-```{r}
-F_matrix <- fit$rep$F_matrix
-dimnames(F_matrix) <- list(age = ages, year = years)
-knitr::kable(t(F_matrix), caption = "F-at-age estimates.", digits = 3)
-```

### 2020-03-14 R/plot.R

@@ -157,7 +157,8 @@ plot_trend <- function(data, group = NULL, ylab = "", title = "", ci = "ribbon",
   p <- p %>%
     add_lines(y = ~est, name = trace_names[2], line = list(width = lwd)) %>%
     layout(title = title,
-           xaxis = list(title = "Year", range = c(min_year, max_year)),
+           xaxis = list(title = "Year", range = c(min_year, max_year),
+                        tickformat = "d"),
            yaxis = list(title = ylab))
   p
 
@@ -184,7 +185,7 @@ plot_retro <- function(data, col = '#1f77b4', ...) {
   data[data$terminal_year != max_year, c("lwr", "upr")] <- NA
   p <- plot_trend(data, group = "terminal_year", col = col, ...)
   data$grp <- data$terminal_year
-  marker_data <- data %>% group_by(terminal_year) %>% filter(dplyr::row_number() == n())
+  marker_data <- data %>% group_by(terminal_year) %>% filter(dplyr::row_number() == dplyr::n())
   p %>% add_markers(data = marker_data, x = ~year, y = ~est)
 
 }
@@ -224,7 +225,7 @@ plot_mohns <- function(data = NULL, ylab = "Relative difference", col = '#1f77b4
           color = I(col)) %>%
     add_markers() %>%
     add_segments(x = ~year, xend = ~year, y = 0, yend = ~pdiff) %>%
-    layout(xaxis = list(title = "Year"),
+    layout(xaxis = list(title = "Year", tickformat = "d"),
            yaxis = list(title = ylab),
            annotations = top_label(paste("Mohn's rho =", round(mohns, 3))))
 
@@ -272,7 +273,7 @@ plot_retro_deltas <- function(data = NULL, ylab = "Difference",
     add_markers(name = "Difference") %>%
     add_segments(x = ~year, xend = ~year, y = ~lwr, yend = ~upr,
                  name = "95% CI") %>%
-    layout(xaxis = list(title = "Year"),
+    layout(xaxis = list(title = "Year", tickformat = "d"),
            yaxis = list(title = ylab))
 
 }
@@ -353,7 +354,8 @@ plot_proj <- function(data = NULL, group = NULL, xlim = NULL, ylim = NULL, ylab 
                                          font = list(color = "black", size = 13))) %>%
     layout(shapes = lines,
            xaxis = list(title = "Year", range = xlim,
[…]

### 2020-03-14 R/tidy.R

@@ -58,6 +58,14 @@ tidy_sdreport <- function(model) {
       if (length(split_values[[nm]]) == tmb.data$Y + 1) {
         vals <- cbind(data.frame(year = tmb.data$years_plus1), vals)
       }
+      if (length(split_values[[nm]]) == tmb.data$Y * tmb.data$A) {
+        vals <- cbind(expand.grid(age = tmb.data$ages, year = tmb.data$years), vals)
+      }
+      if (length(split_values[[nm]]) == tmb.data$Y * length(tmb.data$M_breaks_start)) {
+        grp <- paste0(tmb.data$M_breaks_start + min(tmb.data$ages), "-",
+                      tmb.data$M_breaks_end + min(tmb.data$ages))
+        vals <- cbind(expand.grid(age_group = grp, year = tmb.data$years), vals)
+      }
     }
     l[[nm]] <- vals
   }

### 2020-03-14 R/vis_fit.R

@@ -20,6 +20,17 @@ vis_fit <- function(fit = NULL, comp = NULL, last = NULL, output_file = NULL, ou
       stop(paste(p, "is needed for vis_fit to work. Please install it."), call. = FALSE)
     }
   }
+  if (packageVersion("plotly") > "4.8") {
+    message("Warning: plots will not render correctly in the dashboard using a version of plotly > 4.8.")
+    ans <- readline(prompt = "Press [enter] to revert to version 4.8 or [esc] to abort. ")
+    ans <- readline(prompt = "are you sure you want to revert to plotly version 4.8? ")
+    if (substr(ans, 1, 1) == "y") {
+      remove.packages("plotly")
+      devtools::install_version("plotly", version = "4.8.0", repos = "http://cran.us.r-project.org")
+    } else {
+      stop("plotly version 4.8 is required.")
+    }
+  }
 
   ## simplify file and dir specification
   if (!is.null(output_file)) {

### 2020-03-14 analysis/paper/figures.R

@@ -1,56 +1,0 @@
-
-library(webshot)
-library(magick)
-
-## helper function for exporting dashboard tabs using webshot
-dashboard_webshot <- function(in_file = NULL, out_file = NULL, 
-                              vwidth = 1200, vheight = 800, zoom = 3, delay = 10,
-                              ...) {
-    
-    webshot(url = file.path("file://", getwd(), in_file),
-            file = out_file,
-            delay = delay, vwidth = vwidth, vheight = vheight, zoom = zoom)
-}
-
-## take screenshots of NCAM dashboard
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#background",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_background_tab.png")
-
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#catch",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_catch_tab.png")
-
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#rv-survey",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png")
-
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#tagging",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_tagging_tab.png")
-
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#trends",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_trends_tab.png")
-
-dashboard_webshot(in_file = "analysis/dashboards/NCAM_dashboard.html#output",
-                  out_file = "analysis/paper/figures/NCAM_dashboard_tables_tab.png")
-
-## combine screenshots into one image
-img_a <- image_read("analysis/paper/figures/NCAM_dashboard_background_tab.png") %>% 
-    image_border("white", "300x50") %>% 
-    image_annotate("a)", color = "black", size = 300)
-img_b <- image_read("analysis/paper/figures/NCAM_dashboard_rv-survey_tab.png") %>% 
-    image_border("white", "300x50") %>% 
-    image_annotate("b)", color = "black", size = 300)
-img_c <- image_read("analysis/paper/figures/NCAM_dashboard_tagging_tab.png") %>% 
-    image_border("white", "300x50") %>% 
-    image_annotate("c)", color = "black", size = 300)
[…]

### 2020-03-15 analysis/paper/manuscript.Rmd
Address Greg and Keith's suggestions

@@
 
-Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science.
+Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools available that facilitate the exploration and communication of fisheries science.
 For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples).
-While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore data and models in ways not amenable to static formats and, in the process, these heuristic improve accessibility and augment communication efforts.
-These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.
+Such tools complement static documentation as they allow audiences to explore data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts.
+These will be helpful tools to utilize as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid in the peer review of stock assessments.
 
@@
 The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
-These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).
+These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table \@ref(tab:pkg).
 See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.
@@
 For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
-All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
+All of the above-mentioned data sets are complex and, as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
 It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988].
@@
 These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981].
-Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
-In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)).
+Both the inputs and outputs from RStrap analyses are very large and, depending on the species, may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
+In order for fisheries scientists to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)).
 RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results.
@@
 Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on.
-This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough the data checking and analysis required for formal stock assessment begins.
+This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough data checking and analysis required for formal stock assessment begins.
 
@@
 The intent of this tool was not to replace detailed documentation on the model [i.e.
-@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.
+@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the stock and model.
+Explanatory text and detail were therefore kept to a minimum, allowing the audience to explore the figures and data unencumbered by verbose descriptions.
 
@@
 The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents.
-However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.
+However, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides.
 First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
@@
 These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
-Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieve concerns that something is hidden.
+Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieves any potential concerns that something is hidden.
 
@@
 
-The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations.
-Consequently, there exists the possibility of generalizing these tools to simplify their application to other regions.
-Region specific data management and modelling approaches, however, will surely complicate matters as tools built for one region may be difficult to apply in another region.
-Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications to be applied to another model being developed.
-Nevertheless, considerable chunks of code could be recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard.
-This effort was one step forward in improving the generality of the tool and another step, which is actively being worked on by a colleague in IFREMER, is to modify the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014].
-Though more time is required to make the dashboard code more general as the developer must ensure the code is robust to a wider range of scenarios, substantial value could be added by making it more widely applicable.
+The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations but the solutions presented are specific to the Northwest Atlantic Fisheries Centre.
+The concept is therefore broadly applicable, however, the underlying code will require modifications to accommodate region and stock specific data management and modelling approaches.
+Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications when applied to a model developed for another cod stock.
+Nevertheless, considerable chunks of code were recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard.
+Additionally, a French colleague is currently modifying the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014].
 Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g.
@@
 These are, of course, early days in the development and application of interactive tools to stock assessment.
-Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards.
-Regardless of the approach, one should strive to make the path taken transparent and reproducible.
+Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards depending on user needs and preferences.
+Regardless of the approach, a common goal is to make the workflow more efficient, reproducible and transparent.
 

### 2020-03-15 analysis/paper/manuscript.Rmd
Modify text to reflect changes to NCAM explorer

@@
 The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background").
-Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data.
+Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Landings"; accessed from the "Fishery" drop-down menu), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data.
 For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)).
-The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity").
-Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
+The dashboard also includes pages focused on model estimates such as selectivity ("Selectivity"; accessed from the "Fishery" drop-down menu) and catchability ("Catchability"; accessed from the "Surveys" drop-down menu).
+Pages for viewing trends in various population parameters are included under the "Trends" drop-down menu, where the "Retro" page displays results from a retrospective analysis, the "Comps" page displays population trends from different models, the "Status and rates" page displays trends in stock size and vital rates (Figure \@ref(fig:trends)), and the "Productivity" page focuses on trends in SSB and recruitment.
+Finally, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
 The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents.

### 2020-03-15 analysis/paper/response_letter.Rmd
Address Greg and Keith's suggestions for the response letter and prep files for re-submission

@@
 We have also clarified the intent and audience for the tools highlighted in our manuscript and these changes should curtail impressions that these tools are supposed to be stand-alone documents.
-These changes should also alleviate impressions that interactive tools serve as a complete replacement for static documents when, in actual fact, they are most often used to complement static documentation.
+These changes should also alleviate impressions that interactive tools serve as a complete replacement for static documents when, in fact, they are best used to complement static documentation.
 Please see below for more details on the changes we made in response to the reviews.
@@
 In the following section, "Towards open stock assessment", we note that versioning can be traced using GitHub.
-The only item we did not cover was ensuring compatibility; this is where automated testing via TravisCI and the testthat package are sure to help, however, we did not add this to the discussion because 1) these are programming tools the authors have limited experience with, and 2) it is potentially superfluous to the main message of the paper.*
+The only item we did not cover was ensuring compatibility; this is where automated testing via TravisCI and the testthat package are sure to help, however, we did not add this to the discussion because 1) these are programming tools the authors have limited experience with, and 2) moving into technical areas beyond the scope of this paper.*
 
@@
 
-*This is a fair point and, in hindsight, we see why the paper leaves this impression.
+*This is a very good point and, in hindsight, we see why the paper leaves this impression.
 To try and rectify this, we have clarified the intent of interactive tools in several places in the paper.
 We now explicitly note in the introduction that these tools are not meant to replace static modes of communication, rather they are often intended to supplement standard documents.
-This point is reiterated in the  "Towards open stock assessment" section, where we also highlight an example from Appendix A to make the point that advisory documents can be interactive, and these tend to be linear in structure.
+This point is reiterated in the  "Towards open stock assessment" section, where we also highlight an example from Appendix A to make the point that advisory documents (i.e.
+a linear story) can be interactive.
 As with the previous iteration, we contend that a non-linear format can be useful for evaluating a stock assessment model.
@@
 This format is possible with interactive documents as well because dashboards can be nested within html slide-shows, or dashboards can include "story-boards".
-These are very flexible tools and in this iteration of the manuscript we hope we have clarified their flexibility without discounting static formats.
+These are very flexible tools and in this iteration of the manuscript we hope we have clarified their flexibility without discounting static or linear formats.
 In short, we want to make the point that interactive tools are a potentially powerful tool to add to a biologists’ toolkit.*
@@
 
-*See "stand alone" comment above.*
+*See "stand alone" comment above. Nevertheless, we have added equations and an additional figure to provide more context. We have also added more text and equations to subsequent tabs.*
 
@@
 
-*These are the key references for the accepted assessment model.*
+*More references have been added (all of the key references for the accepted assessment model).*
 
@@
 
-*TODO: Paul*
+*Agreed, the colors were not ideal. We have made the colour scheme more consistent throughout the tool.*
 
@@
 
-*TODO: Paul*
+*True, the fits by age plots were confusing and only useful if the user selected a specific age using the legend.
+These plots have been removed and replaced by plots with a slider for selecting specific ages.
+Now the facet plot allows a user to assess the fits at a glance with one plot and, if interested in the details, shift to the frame plot to take a closer look at the fits to specific ages.*
 
@@
 
-*Undoubtedly, there are many ways to organize this information and we have done our best to organize it in a way that makes sense to us.
-That is, the pages follow the general order in which our stock assessments tend to flow: 1) assess fits and residuals, 2) assess retrospective plots and derived quantities, 3) if 1 and 2 pass peer review, assess the projections.
-We know the tool is far from perfect and agree that it would benefit from input from programmers and/or graphic design experts but, unfortunately, we lack funding to pursue such consultations.
+*Undoubtedly, there are many ways to organize this information and, given this comment, we have tried to improve the layout of the tool.
+Overall, we have done our best to organize it in a way that makes sense to the aforementioned experts at the stock assessment meeting.
+That is, the pages follow the general order in which our stock assessments tend to flow: 1) assess fits and residuals, 2) assess retrospective plots and derived quantities, 3) if 1 and 2 pass peer review, assess the projections.*
+
+*We know the tool is far from perfect and agree that it would benefit from input from programmers and/or graphic design experts but, unfortunately, we lack funding to pursue such consultations.
 This is a reality that many biologists face and a key point of our paper is that this limitation should not preclude the use of these tools.*
@@
 
-*TODO?: Greg & Bob*
+*All excellent suggestions, which will improve future iterations of the tools.
+These two tools, unlike the NCAM dashboard, were mainly built for use by the developer themselves.
+When these tools move to being used by larger communities, further documentation will be essential.
+Our intent in showing these two ‘in-house’ tools was to demonstrate that interactive visualization tools can be a valuable addition to an analysts tool box.*
 
@@
 
-*TODO?: Keith*
+*We have expanded the "Torwards open stock-assessment" section and, in the process, we hope we have clarified how these tools improve accessability and transparency by making it easy for others to access the details.*
 
@@
 
-*TODO?: Bob*
+*This is a very astute observation and it is most likely related to changes in how Plaice length frequencies were recorded (3 to 1 cm scale).
+While this is an important detail to be cognisant of when analyzing these data, we have opted not to get into the minutiae of the sampling as it is outside the scope of this paper.*
 

### 2020-04-17 analysis/paper/response_letter.Rmd
Revisions following latest CJFAS review

--- 2020-03-06 analysis/paper/response_letter.Rmd
@@
 
-Thank you for considering another revision of manuscript cjfas-2019-0424,  **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** by Paul M.
+Thank you for considering another revision of manuscript cjfas-2019-0424.R1,  **"Improving the communication and accessibility of stock assessment using interactive visualization tools"** by Paul M.
 Regular, Gregory J.
@@
 Lewis.
-We are very grateful for the constructive and detailed reviews provided and we have made every effort to do justice to the reviewer's suggestions.
-In the process, we believe our manuscript has been greatly improved.
-Most notably we have expanded our introduction and discussion to include a more through exploration of the literature, emphasizing advances in fisheries science.
-We have also clarified the intent and audience for the tools highlighted in our manuscript and these changes should curtail impressions that these tools are supposed to be stand-alone documents.
-These changes should also alleviate impressions that interactive tools serve as a complete replacement for static documents when, in actual fact, they are most often used to complement static documentation.
-Please see below for more details on the changes we made in response to the reviews.
+We have carefully considered the latest round of suggestions and below we outline how we have addressed each point.
 
@@
 
-Associate Editor:
+Comments to the Author:
 
-Dr. Regular,
+The revised manuscript has undergone review by two of the original reviewers, both of whom have noted that all of their concerns had been addressed.
+I agree with the reviewers and have noted only a few minor issues and typos (outlined below).
+The article is well-written and presented in a straightforward way that, through the use of three excellent examples, demonstrates the power of interactive graphics/documents for presenting fisheries data and stock assessment results.
+I think the article will be of wide interest across the full complement of CJFAS readers including researchers, managers, and even fisheries stakeholders, which is further aided by its easily approached format and word-choice (i.e., general minimization of scientific/assessment jargon).
 
-Thank you for your submission to the Canadian Journal of Fisheries and Aquatic Sciences.
-The submitted manuscript has now undergone extensive and thorough review by three independent reviewers.
-Each of the reviewers have indicated that the manuscript addresses a timely topic worthy of publication in CJFAS.
-However, a number of concerns were raised regarding overlooked issues related to communication and visualization.
-After reviewing the article myself, I agree with the reviewers that this perspectives article addresses an important subject that would be of interest to the broad readership of CJFAS, but it is slightly limited in its review of the existing literature on data visualization in fisheries applications and does not tackle some critical problems with utilizing interactive visualization tools.
-I have summarized the reviewers’ comments along with my own below.
-Given the concerns raised, I have suggested that the article would be suitable for future publication after major revision and further referee assignment.
+It is worth noting that both the supplemental files are in html format.
+I found these files, particularly supplement 1 (interactive tutorial on how to develop your own dashboard for presenting interactive fisheries, or any other type, data), extremely useful and added greatly to the article itself.
+It will be important to provide readers access to these files, but I am not sure the best way to do so.
+The CJFAS editorial staff may have to work with the authors to determine the best approach.
+I was thinking perhaps a link on the abstract page, as well as, in the final document (maybe under the acknowledgements section).
+Hopefully, we can figure that out easily, but these files should definitely be accessible to readers.
 
-Although this is a perspective piece and I appreciated the succinctness of the manuscript, I think that a more thorough exploration of the literature on data visualization and stakeholder communication is warranted.
-As reviewer 1 points out, there have been a number of papers on this subject in recent years including Miller et al.
-(2019), which was part of a larger MSE special issue with a strong focus on stakeholder engagement (https://www.nrcresearchpress.com/toc/cjfas-mse/01/01).
-Other articles discussing communication in MSEs (e.g., Punt et al., 2017; https://academic.oup.com/icesjms/article/74/2/499/2907904) have also highlighted the importance of innovative and interactive visualization techniques.
-Similarly, the marine policy and socioeconomic literature has many examples of how data visualization can improve stakeholder interactions within participatory modeling initiatives (e.g., Levontin et al.
-2017; https://www.sciencedirect.com/science/article/pii/S0308597X1630793X?via%3Dihub).
-I would suggest exploring the literature a bit more in depth and using these to expand the introduction and discussion.
+Once you address the minor issues outlined below, the revised manuscript should be ready for publication in CJFAS.
 
-*We thank you for alerting us to these papers and have incorporated them where appropriate.*
+*We are thankful for the kind words and the helpful suggestions.*
 
-I think the introduction could include further discussion (with citations) of why improved data visualization techniques are warranted and how they can improve stakeholder engagement.
-
-*We now reference Levontin et al.
-2017 in our introduction and we have added a general paragraph to the that highlights the diversity of existing interactive applications in fisheries research.
-With these changes, we have tried maintain the succinctness and generality of the introduction while adding more discussion of the fisheries research literature.*
-
-Similarly, the discussion could be expanded to better highlight some of the issues with developing these type of graphic tools.
-For instance, as noted by reviewer 1, scientists are not necessarily the best equipped to develop visualizations due to lack of expertise and time constraints.
-Ideally, graphic designers could be consulted to help design visuals that take into account the needs of all stakeholders, but the cost of hiring graphic designers may be prohibitive in some circumstances.
-
-*We agree, and we have addressed this oversight in the _Conclusion_ section by noting that professional programmers and graphic designers would, ideally, be hired to construct interactive tools, however, it is rare for RFMOs to fund such initiatives.
-Most often, the task of communicating stock assessment results is left up to scientists without formal training in programming or graphic design.
-In our case, for instance, we would happily consult graphic design experts if we had funding to do so; alas, this is not the case and improvements must come from self-critical judgement, based on what we have gleaned from books on data visualization (e.g.
-The Visual Display of Quantitative Information by Edward Tufte, Data Visualization by Kieran Healy), along with constructive feedback from peers (e.g.
-Reviewer 1).
-In reality, all scientists have to communicate their research to the best of their ability, and the point of the paper is that interactive tools can be created by scientists without formal training in programming or graphic design.
-To use an analogy, consulting a chef will improve ones cooking, but one does not need to be a chef to make a meal.
-We have modified our conclusion to bolster this point.*
-
-*Regarding time constraints, we have expanded our discussion on the learning curve and emphasize that an R user can start producing basic dashboards within a day using Supplement 1 as a guide.
-We also contend that, once over the learning curve, the time investment may not be significantly different than the investment one would make constructing a static presentation.*
-
-Reviewer 1 notes that some of the language may need to be toned down and further supported by the literature.
-Interactive visualization tools are clearly useful and helpful for presenting the results of complex modeling approaches.
-However, further implementation is probably needed before any categorical statements can be made along the lines of transparency and accessibility of assessment models.
-Overall, I would also like to see more discussion of similar approaches and how they have been received by stakeholders.
-You provide a table in the appendix, but these were not really discussed in-depth.
-Using the case studies to support your points was very helpful and gave the manuscript a good flow, but I think highlighting a few other examples (spanning a variety of geographic areas and management agencies) within each sub-section (along with any difficulties encountered) would solidify the paper.
-
-*We have gone through our manuscript and we have tempered several conclusions.
-Some remain as our perspective is based on extensive verbal feedback at meetings from stakeholders and following tutorials the primary author has provided on this topic.
-In the "Towards open stock assessment" section, we also clarify why we think these interactive tools improve accessibility.*
-
-*We acknowledge that a formal survey would be more scientific and it would offer insights into ways of improving the dashboards, however, such a formal approach was outside the scope of this work.
-Likewise, a jurisdictional survey to learn about and speak to experiences outside our region would be an excellent addition to the paper, however, this too was outside the scope of this work.
-If we implemented such surveys we would likely be submitting this paper as a research article with stronger categorical statements.*
-
-Similarly, there are a number of discussion points raised by the reviewers that should be incorporated into the manuscript.
-For instance, reviewer 3 suggests discussion on how to incorporate graphics that compare results across multiple model runs, how to streamline versioning and code updates, how to ensure compatibility among graphics code and assessment outputs as either are updated or reparametrized, and what the benefits/difficulties are of developing generic versus study-specific visualization tools.
-
-*These are important points, several of which the first author coincidentally discussed with colleagues while this paper was in review.
-We have added a new section, "From bespoke to general tools", where we discuss the challenges and potential benefits of developing generic tools.
-We also acknowledge that the tools can be modified to visualize multiple model runs.
-In the following section, "Towards open stock assessment", we note that versioning can be traced using GitHub.
-The only item we did not cover was ensuring compatibility; this is where automated testing via TravisCI and the testthat package are sure to help, however, we did not add this to the discussion because 1) these are programming tools the authors have limited experience with, and 2) it is potentially superfluous to the main message of the paper.*
-
-I also agree that the link between interactive and static reports or presentations is unclear.
-It would seem that there needs to be a mix of approaches, but this is not really discussed.
-Similarly, what tips would the authors give for presenting results using interactive graphics?
-One element that is often useful with static presentations is the building of a ‘story or progress line’ (especially in model building or continuity exercises), which might be difficult to replicate if only using interactive graphics.
-It does not seem like an all or nothing approach is warranted, but that is the impression that I received from the article.
-I would suggest carefully reworking the discussion of how interactive graphics can be used when presenting to stakeholders and how static and interactive methods can be used synergistically.
-
-*This is a fair point and, in hindsight, we see why the paper leaves this impression.
-To try and rectify this, we have clarified the intent of interactive tools in several places in the paper.
-We now explicitly note in the introduction that these tools are not meant to replace static modes of communication, rather they are often intended to supplement standard documents.
-This point is reiterated in the  "Towards open stock assessment" section, where we also highlight an example from Appendix A to make the point that advisory documents can be interactive, and these tend to be linear in structure.
-As with the previous iteration, we contend that a non-linear format can be useful for evaluating a stock assessment model.
-That being said, we take your point that a story may be more suitable in some cases.
-This format is possible with interactive documents as well because dashboards can be nested within html slide-shows, or dashboards can include "story-boards".
-These are very flexible tools and in this iteration of the manuscript we hope we have clarified their flexibility without discounting static formats.
-In short, we want to make the point that interactive tools are a potentially powerful tool to add to a biologists’ toolkit.*
-
-Finally, all of the reviewers have provided useful suggestions for improving the graphics, especially those utilized within the manuscript. Please carefully consider these and revise as warranted.
-
-*We have made every effort to improve the graphics as suggested and while not compromising the readability and utility of the dashboard tool.  Please see responses to individual Reviewers below.*
-
-Similarly, please provide access to the source code (r code) via an online database (e.g., GitHub) as requested by reviewer 1.
-
-*We completely agree with the Reviewer.  The data and code used to produce the manuscript, supplement 1, and supplement 2 are available on GitHub.*
-
-If the authors are able to revise the manuscript to adequately address the reviewers’ concerns, then it should be suitable for future publication.
-
-Thank you,  
+Thank you for you submission,
 Daniel Goethel
 
-Comments to the Author: See attached.
+Note from Editorial Assistant: Supplementary files are provided to readers in the exact format you provide them so I suggest you upload the supplementary files as html but make sure you use the supplementary file designation.
+
+Line 30: 'to stock assessments'
+
+*We have re-phrased this line to state "...these advances have surely improved the advice provided on stock status."*
+
+Lines 31-34: Bit of a long sentence.  I would suggest ending the first before the 'however' and beginning the second with it.
+
+*Agreed, we have split this sentence as suggested.*
+
+Line 148: Should start "A screen shot of the cod..."
+
+*Done.*
+
+Line 150: Should say "the recovery positions".
+
+*Done.*
+
+Lines 164-167: This statement is a bit too close in nature to the exact wording used throughout the Maunder and Punt (2013) reference.
+For instance from their abstract: "Contemporary use of integrated analysis involves using all available data, in as raw a form as appropriate, in a single analysis." And from the current paper: "Such issues have largely been curtailed in contemporary stock assessments thanks to advances in statistical computing that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis." Please just reword the sentence a bit or use an exact quote from the reference.
+
+*Agreed, the wording was very similar.
+The sentence now reads "Such issues have largely been curtailed in contemporary stock assessments as advances in statistical computing have enabled integrated analyses of data from multiple monitoring programs (Maunder and Punt 2013)."*
+
+Line 182: What are 'partial fishery landings'? It might be worth explaining or just say 'fishery landings' to avoid unnecessary confusion.
+
+*What we did not outline in the text is that the recreational fishery is not accounted for in the reported landings so we call the landings partial.
+Rather than get into these details, we have replaced "partial" with "reported" to avoid confusion and to keep things simple.*
+
+Line 217: It is probably not worth going back and changing the NCAM explorer, but the use of 'comps' here is a bit confusing.
+I was expecting this to be an age/length composition tab not a model comparison tab.
+
+*We understand this expectation and we changed the page name to "Model comparison".*
+
+Line 218:  I think saying explicitly 'from different model runs' would avoid confusion, because this could be interpreted as completely different assessment models were applied.
+
+*Done.*
+
+Figure 5: I think the figures were updated without refining the caption text.
+The caption seems to focus only on the right hand part of the figure and ignores the larger figure on the left.
+Also, I do not see any red/blue coloring defining pos/neg values as stated in the caption.
+
+*That was an oversight after the last revision; we have now revised this caption to reflect the revised dashboard.
+The caption now reads "Screenshot of the 'RV survey' tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the left panel and standardized log residuals by year, cohort, age, and expected value (dots), with lowess curves (lines), are shown in the right panel."*
+
+Figure 6: Same as above.
+Left hand figure is abund, bio, and b/blim...recruitment is not shown.
+Also, it might be worth just stating the basis of Blim (e.g., B30...) in parenthesis, while also noting what the dashed line at 823kt represents.
+
+*Done, the caption now reads "Screenshot of the “Trends” tab which displays stock status estimates (abundance, biomass, and biomass relative to Blim [average biomass through the 1980s]) on the left panel, and mortality rates (average F, M, and Z) on the right panel."*
+
+292: Delete 'along'.
+
+*Deleted.*
+
+Line 302: Should this read 'to document in a pipeline'?
+
+*We have revised the sentence to read "...commonplace for the full data pipeline to be open and accessible."*
+
+Line 315: Delete the 'a'.
+
+*Deleted.*
+
+Lines 316-318: I don't really understand what this sentence is saying.
+It appears to be somewhat contradictory.
+I imagine the idea is along the lines of 'developing dashboards that document the full pipeline allow better access to details regarding model development decisions compared to current stock assessment practices that might be fully open but effectively inaccessible'.
+I'm not sure if this is correct or how exactly to word it, but please clarify the statement.
+
+*In hindsight we see that this sentence was not clear.
+We hope this revision does a better job of conveying what we are trying to say: "Although these tools do not reveal the entire data pipeline, a wide range of stakeholders and scientists can easily access and review details of an analysis without running a line of code.
+For most participants of a stock assessment, even if the entire pipeline was completely open, the details would remain inaccessible simply because most participants are already overcommitted (Banks 2011) and do not have the time to fully review and recreate a particular analysis in its entirety."*
+
+Line 325: Delete 'of'.
+
+*Deleted.*
+
+Line 329: I might suggest replacing 'their' with something along the lines of 'extend the utility of data visualizations or graphics by....'
+
+*Done.*
+
+Line 349-351: Just a wording suggestion, could say "...that the upfront time investment needed to develop interactive visualization tools has already improved efficiency in delivering....'
+
+*We have rephrased the sentence to read "...that the upfront time investment has been worthwhile considering the improved efficiency of delivering...*
 
@@
 Comments to the Author
-
-Summary
-
-This paper states a hard to disagree with proposition that fisheries science can benefit from data visualisation tools that are widely available and applied in other fields.
-It correctly situates data visualisation within the ambition to move stock assessment towards a model of reproducible research that is now commonly taught in beginner data science courses.
-The authors describe various efforts they have made in developing visualisation tools, such as interactive dashboards, and contextualise these as steps towards stock assessments meeting the standards of reproducible research.
-
-The paper is clearly written, and is important for its promotion of timely developments within fisheries science.
-
-*We thank the Reviewer for the kind comments.*
-
-Major problems
-
-The main shortcoming of the paper is that the authors don’t have sufficient relevant expertise to discuss matters related to design and lack skills necessary to implement effective visualisations.
-Yet this is the crux of the paper and visualisations are offered as the main results or innovations being presented.
-Typography, the use of colour and text, the layout, aesthetics, ergonomic organisation of information and the use of interactivity in the tools presented – all fall below professional standards, preventing these tools from being effective at communication.
-This is potentially damaging to the author’s cause.
-What the authors need is to work with a graphic designer and address these shortcomings before the paper can be published.
-The authors’ aim is to enable ‘a broader audience to conduct detailed exploration of the results’ so that they can arrive at a ‘deeper and collective understanding’ of stock assessments – programming and science knowledge are insufficient for this task, the authors need to work alongside experts on visual communication.
-
-*Please see the response to these comments in the AE section above.*
-
-Another issue, is accessibility to the code which is essential for reproducible research.
-Only one of the tools was available for review as a website, no source code was provided along with the paper and it seems that the code is not intended for release alongside the paper, e.g.
-via a link to GitHub.
-The publication of this paper should be made contingent on the availability of the code.
-The data provided along with the code can be simulated if there are privacy concerns.
-
-*Please see the response to these comments in the AE section above.*
-
-There is no discussion of data conflicts, or data weighting issues, that plague many integrated stock assessments.
-
-*While we have not gone into a detailed discussion of data conflicts, as that has been covered quite nicely by Maunder and Piner (Maunder, M.N.
-and Piner, K.R., 2017.
-Dealing with data conflicts in statistical inference of population assessment models that integrate information from multiple diverse data sets.
-Fisheries Research, 192, pp.16-27), we have noted that NCAM explorer helped us efficiently identify conflicts between two of the surveys included in the model.
-Data weighting issues, however, are not mentioned in the context of this paper as NCAM is self-weighting.*
-
-Ability to examine a variety of graphs in pairs would be an important feature enabling the user to examine contribution of individual data sources and detect data conflicts of model misspecifications.
-Features to examine convergence, boundary behaviour, comparisons with simpler models or external data might be good to consider as additional functionality.
-
-*We agree with these points.
-These tools are a work in progress and we continue to add functionality to them.
-We have now added a new section ("From bespoke to general tools") where we discuss some of our experiences with making these tools more general.
-We acknowledge that these are early days and much work has yet to be done; still, we are hopeful that we have included enough to spur discussions among the broader community and catalyze the development of more features.*
-
-Minor problems
-
-The authors need to be more tentative in some of their conclusions, such as ‘peer review process is more open and accessible’ as a result.
-Further testing with more and varied stakeholders is necessary to demonstrate that these tools are indeed achieving the aims to make stock assessments more reproducible, easier to critique by a wider range of stakeholders, and as a result some systemic changes in the processes of performing stock assessments are occurring.
-
-*Please see previous comments to the AE.*
-
-The authors are missing some important references on using visualisation tools in fisheries, e.g.
-Miller.
-S.K., Anganuzzi, A., Butterworth, D.S., Davies, C.R., Donovan, G.P., Nickson, A., Rademeyer, R.A., Restrepo, V.
-2018.
-Improving communication: the key to more effective MSE processes.
-Canadian Journal of Fisheries and Aquatic Sciences, https://doi.org/10.1139/cjfas-2018-0134.
-
-*We thank the Reviewer for alerting us to this paper and have added where appropriate.*
-
-Comments on the NCAM tool It’s great to see more interactive visualisation tools being developed for fisheries science.
-I greatly enjoyed playing with the tool.
-Please find below some comments and suggestions for how to improve it, but I would strongly recommend getting professional help from a designer.
-Please upload the full code to a GitHub directory and provide a link, use simulated data if necessary to protect confidentiality of any of the data that is not publicly available.
-
-*We thank the Reviewer for the kinds words.  Please see the response to these comments in the AE section above.*
-
-The tool should be made easily transferrable to other case studies otherwise the contribution made by the paper is diminished.
-
-*The code for the tools are available for other researchers and stock assessment biologists to adapt to their own work flows.
-Further, the tutorial that is provided should ensure that biologists can learn this approach in a short period of time.
-Finally, these tools are just examples of a more general approach – they are not meant to be the only possible solution.*
-
-*In the comments below, the Reviewer makes a number of good suggestions about the layout of the dashboard.  We will deal with these individually. *
-
-*However, we argue that there may be a misunderstanding that we hope to rectify.
-Many of the below suggestions seem to be based on the premise that the dashboard is a stand alone piece of work like a paper or book when, in reality, it is better to think of it as a PowerPoint replacement.
-There are DFO research papers, stock assessment reports, and primary publications that support the dashboard.
-Reviewers, stakeholders and participants generally have a working knowledge of this documentation and, as such, excessive texts and guides are not required.
-Moreover, excessive texts may be a distraction as the presenter actively walks participants through the latest model results (i.e.
-wordy PowerPoints are generally discouraged).
-Therefore, although these suggestions are greatly appreciated, not all of them are relevant and we have tried to make this clear in the text by stating that the tool was "built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model."  We also state the tool was not built to replace detailed documentation on the model.
-We refer to this paragraph as the "stand alone comment" below.*
-
-Background tab:
-
--NCAM tab needs more information about the tool overall and a better image.
-
-*See "stand alone" comment with regards to the more information suggestion.*
-
-*As for a better picture, we agree that it would be nice to have, however, for the time being, we have opted to focus on changes that improve the functionality of the tool.*
-
--General tab similarly should have more text and an overall guide with images and links to the rest of the tool
-
-*See "stand alone" comment above.*
-
-Data inputs tab:
-
--Should have information about the data, how it is collected and processed with links to raw data and a code for processing if any was used
-
-*See "stand alone" comment above.*
-
--Something about data quality, uncertainties, and potential conflicts in data should be mentioned here
-
-*See "stand alone" comment above.*
-
-Biological inputs tab:
-
--Confusing graphics perhaps because of a lack of explanatory text – this is a problem for most of the pages, two dimensional graphics should be added for clarity.
-
-*See "stand alone" comment above.*
-
-Partial catches:
-
--Need explanation as well as context, the tool overall should be able to tell a story by itself but it feels like a somewhat chaotic assembly of cut out figures from a stock assessment report.
-
-*See "stand alone" comment above.*
-
-Change in Q:
-
--Similarly you have an entire page to explain the context related to catchability, but there is insufficient information to be able to critically engage with the visualisation
-
-*See "stand alone" comment above.*
-
-Tagging:
-
--Likewise, there is a whole page to explain the issues of using tagging data in stock assessment – it seems this would be beneficial
-
-*See "stand alone" comment above.*
-
-References:
-
--Should be more references and they should be linked to the documents.
-
-*These are the key references for the accepted assessment model.*
-
-Catch tab:
-
--Everywhere in the tool there is insufficient labelling and cryptic legends, for example it is not clear what is meant by ‘bounds’ or what is represented by the shaded area -the choices of colour and line types are questionable and inconsistent throughout the tool -there are two shaded areas in the bottom graph (absolute landings) without a key
-
-*TODO: Paul*
-
-On the right side:
-
--there should be consistency in how observations and estimation are signified, going from a dashed line to solid circles adds unnecessary confusion -the graphs can have better titles such as ‘Age 2’ rather than ‘2’, for instance -other graphs, such as ‘fits by age’ are unwieldly and confusing, this, like a majority of graphs, does not represent a good use of data visualisation principles and betrays the authors’ lack of experience with graphic design.
-When there is a lot of data in one graph, making sure that interactivity works well is paramount, currently selecting a specific line is not so clear.
-Use different colour scheme.
-
-*TODO: Paul*
-
-Overall there are too many graphs, offered in such a way as it is difficult to prioritise information.
-The design can be far more ergonomic, with graphs organised more helpfully and with a sense of hierarchy and a narrative, encouraging to explore key results first, while offering extra levels of details as a choice to those who want to pursue a particular matter further.
-There as an inconsistent and suboptimal from a design point of view use of colour.
-A data visualisation professional should be able to help with making a good use of colour to encode information and to help the user navigate through results without getting overwhelmed.
-I would particularly recommend enabling the user to make side by side comparisons of different combination of graphs that are currently in fixed/distinct places in the tool.
-
-*Undoubtedly, there are many ways to organize this information and we have done our best to organize it in a way that makes sense to us.
-That is, the pages follow the general order in which our stock assessments tend to flow: 1) assess fits and residuals, 2) assess retrospective plots and derived quantities, 3) if 1 and 2 pass peer review, assess the projections.
-We know the tool is far from perfect and agree that it would benefit from input from programmers and/or graphic design experts but, unfortunately, we lack funding to pursue such consultations.
-This is a reality that many biologists face and a key point of our paper is that this limitation should not preclude the use of these tools.*
-
-*See also previous comments to the AE.*
+I agree with the authors that the paper has substantially improved and am now recommending it for publication without further changes.
 
@@
 Comments to the Author
-
-This is a clearly written and organized manuscript that presents an application of novel, open-source tools to disseminate stock assessment inputs and results.
-The interactive visualization tools were developed by others but modified by the authors to suit their needs; the original developers are credited appropriately.
-The authors make a strong case for using these tools for visualization of stock assessment results, and in that way, this paper is fitting for a perspective since it promotes a new paradigm for dissemination of stock assessments.
-There are no new research results, nor is there anything controversial in this piece.
-An important and laudable message of the paper is that these tools can help the transition to "a culture of open science and reproducible research."
-
-*We thank the Reviewer for the kind comments.*
-
-The manuscript repeatedly mentions the usefulness of these tools for stakeholders, both for data exploration and data modeling.
-In the data exploration figures (Figures 1 through 3), the interfaces would be more digestible to stakeholders if they included an annotation of what the user is viewing.
-For example, the Figure 1 view could include a note specifying that the error bars represent 95% confidence intervals, and species names should be used rather than, or in addition to, codes.
-Figure 2 could be improved by including units for the length bins, and Figure 3 should have a legend for the kernals.
-Each could also have a text field that gives an overview of the what the figure is showing and how a user can interact with the figure.
-
-*TODO?: Greg & Bob*
-
-More broadly, the NCAM Explorer (Supplement 2) would also be more functional with increased annotation.
-The background tab provides important context but should be expanded beyond the "terse point-form" to serve as a user guide for each tab.
-These improvements apply to the tools themselves rather than to the manuscript.
-
-*We agree with these comments and have added text to the Background tab and added references for background.
-See "stand alone" comment above.
-We have tried to strike a balance between addressing the spirit of the Reviewers comments and unnecessarily cluttering the dashboard.*
-
-The one area where I think the manuscript should be expanded is in the "Towards open stock-assessment" section.
-The main reason given for increasing the transparency of stock assessment through tools such as these is to improve peer review.
-However, as mentioned at several other points in the manuscript, these tools are also valuable for increasing stakeholder engagement in and understanding of stock assessment.
-That function warrants equal attention to peer review in this section and could present ideas for how to introduce these tools to stakeholders and how to help them sift through the large volume of data to find what is most relevant to them.
-
-*TODO?: Keith*
-
-Lastly, Supplement 1 (getting_started.html) isn’t referenced anywhere in the paper. It can be removed from the supplementary files.
-
-*We have now added a reference to Supplement 1 to the introduction and conclusion.*
-
-Reviewer: 3
-
-Comments to the Author
-
-This manuscript illustrates a novel approach to exploring and communicating data and model results related to fisheries stock assessments.
-Communication methods for complex patterns in data or stock assessment results is not a frequent topic covered in fisheries science journals but it is an area where new ideas are rapidly developing, much further progress is needed, and greater prominence in the literature is deserved.
-I highly recommend publishing this manuscript.
-
-*We thank the Reviewer for the kind comments.*
-
-The availability of the NCAM Explorer file NCAM_dashboard.html (referenced in the manuscript as Supplement 2) is vital for the understanding of the interactive nature of the tools developed.
-It would be useful to reference the other supplemental file getting_started.html in the text and include as an additional supplement an example of the RStrap Explorer as well in addition to the static figures in the text.
-
-*We agree with the Reviewers comments and we have included a reference to Supplement 1 in the introduction and conclusion.
-RStrap Explorer is harder to include as a supplement as it is Shiny based and must be hosted on a server.
-At the moment, the use of this tool has been limited to local computers; we have yet to host the tool on an online server.*
-
-I raise a few minor specific issues below, but the area where I would suggest the greatest change is adding to the introduction and conclusion further discussion of a few larger questions that this new approach raises.
-First, the visualization and communication needs are very broad within the stock assessment community, but the example tools presented are specific to analyses used in the region where the authors work: the RStrap analysis of survey data and the NCAM assessment model.
-I would have like to see some discussion about the trade-offs associated with developing generic dashboards that could be applied to other survey or assessment tools vs.
-developing separate dashboard applications for each database or stock assessment model.
-Comparing results from multiple stock assessment models which might be applied to the same stock is another challenge that might be discussed.
-
-*Please see previous comments to the AE.*
-
-Second, I would have liked more discussion of version control and archiving.
-The supplemental file NCAM_dashboard.html includes a comparison of two models "NCAM 2018" and "NCAM 2019" but in the assessment process there is typically a proliferation of numerous alternative models being considered in any given year.
-Exploration of these alternative models would presumably lead to a proliferation of these html dashboard files.
-Keeping track of which files show the results from which models is a challenge not addressed in the manuscript (and not obvious in the supplemental file).
-Furthermore, keeping track of the software version for the dashboard software (and R packages on which it depends) and format of the database or assessment model and maintaining compatibility between the two is a challenge not discussed here.
-
-*This is an excellent point, however, we do discuss this issue as this approach does not offer a novel solution to file organization.
-The fact is that there is little difference between the file management with this approach versus a more standard approach.
-Through pre-dashboard days, different model runs were nested within independent folders and within these folders were figure and table folders.
-The only difference with the dashboard approach is that key figures and tables are held in the html file (i.e.
-there is no longer figure and table folders).
-A convention we apply to help keep track of things is to use a consistent name for the folder, the RData file containing all the inputs and outputs, and the html file.
-Though an interesting and important challenge, we did not cover this as file management and project conventions is a potential tangent that may add little to the paper as potential users are likely going to apply their own personal preferences.*
-
-*Regarding software version, we now note in the conclusion that these files can be traceable and accessible via GitHub.*
-
-Third, the link between this tool and more traditional static reports could be also addressed.
-The manuscript presents a compelling argument in favor of interactive dashboards as a tool for the rich volume of information needing to be explored in fisheries models, but it is not suggested that static reports will go away.
-It is noted that "The user also has the ability to hover over the points to compare estimates and export the figure as a png file" (line 110) but this would not be an efficient way to produce figures for a static report.
-Further discussion of how the same tools used for the dashboard view could used to make the development of static reports more efficient may be worth including in the discussion.
-
-*We agree, it would not be very efficient to manually export individual plots.
-We have removed this note and have expanded our introductory text to clarify how synergies can be created between static and interactive documents.*
-
-Specific issues: Line 29: "For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way." I think this is a mischaracterization of the state of the science.
-Perhaps Northern Cod is an exception, but I believe the challenge of having adequate information to provide sound advice remains as present as ever, even for relatively well-monitored stocks, while the second of meaningfully presenting the resulting information has grown beside it.
-
-*Fair point.
-We have rephrased the sentence to read: "While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way."*
-
-Line 81: the reference to a 1981 report for information on the "locally developed R package called RStrap" seems inadequate.
-Perhaps "locally developed" means only locally available and thus no citation for the R package is available (in contrast to the other packages cited in the manuscript).
-
-*This is correct.  We now simply state that it is a local package.*
-
-Line 121: "data base" should be one word for consistency with earlier occurrence on the same line.
-
-*We have made the suggested change.*
-
-Lines 262: The statement "Although these tools do not reveal the entire data pipeline" may be contradicted by the ICES Transparent Assessment Framework (https://taf.ices.dk/app/about) which archives the full process of data processing to running assessment models in a reproducible way and should be added to Appendix A.
-
-*We thank the Reviewer for drawing our attention to the TAF.
-We have now re-structured our "Towards open stock assessment" section to highlight TAF and several other open science initiatives in fisheries research.*
-
-Figure 2: I realize this figure is only for illustrative purposes, but it might be good to add a note about the strange pattern in the length compositions of American plaice in the 1980s.
-The spiky pattern with peaks in the same bins each year seems like an artifact of rounding error or conversion from some other type of length measurement.
-
-*TODO?: Bob*
-
-Again, I think this manuscript is a valuable contribution to the field, is of broad interest, and I encourage it to be published with or without the refinements proposed above.
-
-*We thank the Reviewer for the kind comments.*
+The changes improve the manuscript and address the issues that I raised in previous review (in which I already only suggested minor changes).

### 2020-04-17 analysis/paper/manuscript.Rmd
Revisions following latest CJFAS review

@@
 Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013].
-While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way.
-Visualizations play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g.
+While these advances have surely improved the advice provided on stock status, it has become more difficult to communicate the basis of this advice in a meaningful way.
+Visualizations play a critical role in communicating this information to stakeholders [@levontin2017].
+However, traditional formats (e.g.
 slides) are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
@@
 We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
-Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model.
+Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose, and communicate results from an integrated assessment model.
 This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
@@
 
-Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day.
+Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting as a dedicated program in 1954 [@taggart1995] and continuing to this day.
 The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs.
 This database has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
-The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.
+The tagging and recovery data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.
 To begin to explore and understand this large data set, we built a simple application using shiny and leaflet (Table \@ref(tab:pkg)) that allows a user to quickly and dynamically subset the data (e.g.
-ranges of release and capture years, specific geographic locations).
+ranges of tag release and tag recovery years, specific geographic locations).
 As the application developed, it was incorporated into shinydashboard, to take advantage of the number of layout options available in shinydashboard (Table \@ref(tab:pkg)).
@@
 
-```{r map, echo = FALSE, fig.cap = 'Screen shot of cod tag mapping tool using shinydashboard and leaflet.
-The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
-Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date captured), which is particularly useful when error checking.'}
+```{r map, echo = FALSE, fig.cap = 'A screen shot of cod tag mapping tool using shinydashboard and leaflet.
+The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recovery positions (blue) are much fewer, and are left to be plotted individually at all scales.
+Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date recovered), which is particularly useful when error checking.'}
 knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
@@
 ```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
-In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
+In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the spatial distribution of the selected tag recoveries.'}
 knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")
@@
 This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
-Such issues have largely been curtailed in contemporary stock assessments thanks to advances in statistical computing that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
-Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
+).
+Such issues have largely been curtailed in contemporary stock assessments as advances in statistical computing have enabled integrated analyses of data from multiple monitoring programs [@maunder2013].
+Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012], and Template Model Builder [TMB; @kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
 From a computational perspective, analyses of a variety of large data sets has never been easier.
@@
 As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016].
-The model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
-Using a series of observation equations, this TMB based model simplifies thousands of historical data points into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
+The model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions, and reported fishery landings (1983-present), and tagging (1983-present).
+Using a series of observation equations, this TMB based model simplifies thousands of data points into quantities such as annual recruitment, spawning stock biomass, fishing mortality, and natural mortality.
 Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
 The traditional approach would involve producing static presentations and documents with a series of figures and tables.
-However, with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved for at least two reasons.
+However, with numerous model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved for at least two reasons.
 First, it is no longer feasible for the analyst to include and describe every figure and table produced in a single document.
@@
 The intent of this tool was not to replace detailed documentation on the model [i.e.
-@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the stock and model.
+@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the stock and assessment model.
 Explanatory text and detail were therefore kept to a minimum, allowing the audience to explore the figures and data unencumbered by verbose descriptions.
@@
 The dashboard also includes pages focused on model estimates such as selectivity ("Selectivity"; accessed from the "Fishery" drop-down menu) and catchability ("Catchability"; accessed from the "Surveys" drop-down menu).
-Pages for viewing trends in various population parameters are included under the "Trends" drop-down menu, where the "Retro" page displays results from a retrospective analysis, the "Comps" page displays population trends from different models, the "Status and rates" page displays trends in stock size and vital rates (Figure \@ref(fig:trends)), and the "Productivity" page focuses on trends in SSB and recruitment.
+Pages for viewing trends in various population parameters are included under the "Trends" drop-down menu, where the "Retrospective analysis" page displays results from a retrospective analysis, the "Model comparison" page displays population trends from different model runs, the "Status and rates" page displays trends in stock size and vital rates (Figure \@ref(fig:trends)), and the "Productivity" page focuses on trends in Spawning Stock Biomass (SSB) and recruitment.
 Finally, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
@@
 However, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides.
-First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
+First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across a smaller number of dynamic pages holding data-rich illustrations (i.e.
 both broader patterns and finer details in the data are accessible via zooming and tooltips).
-Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g.
+Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables, and values into documents and slides, although, we note that this issue can be addressed through other approaches (e.g.
 slide shows can easily be made using R markdown).
@@
 For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model.
-These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
+These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers, and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
 Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieves any potential concerns that something is hidden.
 
-```{r rv, echo = FALSE, fig.cap = 'Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
-Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
+```{r rv, echo = FALSE, fig.cap = 'Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the left panel and standardized log residuals by year, cohort, age, and expected value (dots), with lowess curves (lines), are shown in the right panel.'}
 knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
@@
 
-```{r trends, echo = FALSE, fig.cap = 'Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
+```{r trends, echo = FALSE, fig.cap = 'Screenshot of the "Trends" tab which displays stock status estimates (abundance, biomass, and biomass relative to B~lim~ [average biomass through the 1980s]) on the left panel, and mortality rates (average F, M, and Z) on the right panel.'}
 knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
@@
 These are, of course, early days in the development and application of interactive tools to stock assessment.
-Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards depending on user needs and preferences.
-Regardless of the approach, a common goal is to make the workflow more efficient, reproducible and transparent.
+Because the tools we highlight are themselves general, flexible, and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards depending on user needs and preferences.
+Regardless of the approach, a common goal is to make the workflow more efficient, reproducible, and transparent.
 
@@
 However, decisions made along the data pipeline increasingly influence the outcome of the study.
-Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis.
+Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made during different stages of the analysis.
 Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible.
@@
 
-A culture of openness is actively evolving in the field of stock assessment and, at this point, it is not hard to picture a future where it is commonplace for the the full data to document pipeline to be open and accessible.
+A culture of openness is actively evolving in the field of stock assessment and, at this point, it is not hard to picture a future where it is commonplace for the full data pipeline to be open and accessible.
 Open resources are actively being developed and used to improve accessibility to all the data, methods and results that feed into stock assessments [e.g.
@@
 Providing open access to code and data, however, does not necessarily make a stock assessment more accessible.
-We believe that interactive tools, such as those presented here and in Appendix A, represent an important a step forward in terms of open science as they make the peer review process more efficient.
-Although these tools do not reveal the entire data pipeline, a wide range of peers can easily access and review details that may be inaccessible even if the pipeline is open.
-For example, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
+We believe that interactive tools, such as those presented here and in Appendix A, represent an important step forward in terms of open science as they make the peer review process more efficient.
+Although these tools do not reveal the entire data pipeline, a wide range of stakeholders and scientists can easily access and review details of an analysis without running a line of code.
+For most participants of a stock assessment, even if the entire pipeline was completely open, the details would remain inaccessible simply because most participants are already overcommitted [@banks2011] and do not have the time to fully review and recreate a particular analysis in its entirety.
 Further, in the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
@@
 As such, participants can assess the results and diagnostics at their leisure.
-Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
+Specific experience with the tools used to generate the results are not required to constructively and critically review the results presented in the dashboard.
 
@@
 
-Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information.
+Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend the utility of data visualizations by providing a richer approach to communicating scientific information.
 However, we must acknowledge there are costs to adopting the use of interactive tools.
@@
 We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents.
-It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
+It is our experience that the upfront time investment has been worthwhile considering the improved efficiency of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
 

### 2020-04-17 inst/rmd/trends.Rmd
Revisions following latest CJFAS review

@@
-Retro {data-navmenu="Trends" `r if (is.null(fit$retro)) ".hidden"`}
+Retrospective analysis {data-navmenu="Trends" `r if (is.null(fit$retro)) ".hidden"`}
 ================================================================================
@@
 
-Comps {data-navmenu="Trends" `r if (is.null(tmp$comp)) ".hidden"`}
+Model comparison {data-navmenu="Trends" `r if (is.null(tmp$comp)) ".hidden"`}
 ================================================================================

### 2020-04-21 analysis/paper/manuscript.Rmd
Tweaks too the MS for CJFAS "First Look" steps

@@
 As a consequence, the peer review process is more open and accessible.
+
+  **Keywords**: dashboard, data exploration, integrated analysis, model validation, open stock assessment
 
@@
 
-```{r map, echo = FALSE, fig.cap = 'A screen shot of cod tag mapping tool using shinydashboard and leaflet.
+```{r map, echo = FALSE, fig.cap = "A screen shot of cod tag mapping tool.
 The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recovery positions (blue) are much fewer, and are left to be plotted individually at all scales.
-Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date recovered), which is particularly useful when error checking.'}
+Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date recovered), which is particularly useful when error checking.
+Base map created with leaflet [@cheng2018] using ESRI's Ocean BaseMap (Sources: Esri, GEBCO, NOAA, National Geographic, DeLorme, HERE, Geonames.org, and other contributors)."}
 knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
@@
 
-```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
-In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the spatial distribution of the selected tag recoveries.'}
+```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the cod tag mapping tool.
+In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the spatial distribution of the selected tag recoveries.
+Base map created in R with package maps [@becker2018a] and mapdata [@becker2018b].'}
 knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")

### 2020-07-23 analysis/site/index.Rmd
Use R Markdown Websites to construct a site for this project

--- 2020-03-15 analysis/paper/manuscript.Rmd
@@
 ---
-title: |
-  Improving the communication and accessibility of stock assessment using interactive visualization tools
-author: "Paul M. Regular, Gregory J. Robertson, Robert Rogers, Keith P. Lewis"
-date: |
-  Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
-output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
-bibliography: references.bib
-csl: cjfas.csl
-abstract: |
+title: Improving the communication and accessibility of stock assessment using interactive visualization tools
+---
 
-  **Corresponding author**: Paul M.
-Regular, Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St.
-John's, Newfoundland and Labrador, A1C 5X1, Canada, E-mail: Paul.Regular@dfo-mpo.gc.ca, Phone: 1-709-772-2067
+## Abstract
 
-  **Abstract**: Scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
+Scientists across many fields are faced with the challenge of synthesizing and communicating information from large and complex data sets.
 The field of stock assessment is no exception as the volume and variety of the data has grown alongside the computational methods used to integrate them.
@@
 
----
+## Documentation
 
-*The greatest value of a picture is when it forces us to notice what we never expected to see.*  
-- Tukey [-@tukey1977]
+The paper has been published in the Canadian Journal of Fisheries and Aquatic Sciences, https://doi.org/10.1139/cjfas-2019-0424.
 
-Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992].
-Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments.
-These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series.
-Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013].
-While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way.
-Visualizations play a critical role in communicating this information to stakeholders [@levontin2017], however, traditional formats (e.g.
-slides) are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions.
-A solution to this problem is the effective use of interactive visualization tools [@keena2016].
-These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018].
-As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g.
-@jones2016; @yeatman2018; @letcher2018].
+See [Supplement 1](getting_started.html) for a “getting started” tutorial that highlights one approach to constructing a dashboard using R
 
-Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools available that facilitate the exploration and communication of fisheries science.
-For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples).
-Such tools complement static documentation as they allow audiences to explore data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts.
-These will be helpful tools to utilize as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid in the peer review of stock assessments.
-
-We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs.
-Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model.
-This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling.
-In both cases interactive visuals are nested inside dynamic documents called "dashboards".
-The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators.
-These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table \@ref(tab:pkg).
-See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.
-
-```{r pkg, results = "asis", echo = FALSE}
-
-## Package name and description
-pkg <- c(shiny = 'Makes it incredibly easy to build interactive web applications with R.
-Automatic "reactive" binding between inputs and outputs and extensive prebuilt widgets make it possible to build beautiful, responsive, and powerful applications with minimal effort [@chang2018a]',
-         shinydashboard = "Create dashboards with 'Shiny'. This package provides a theme on top of 'Shiny', making it easy to create attractive dashboards [@chang2018b].",
-         rmarkdown = "Convert R Markdown documents into a variety of formats [@allaire2018].",
-         flexdashboard = "Format for converting an R Markdown document to a grid oriented dashboard.
-The dashboard flexibly adapts the size of its components to the containing web page [@allaire2017].",
-         plotly = "Easily translate 'ggplot2' graphs to an interactive web-based version and/or create custom web-based visualizations directly from R [@sievert2018].",
-         leaflet = "Create and customize interactive maps using the 'Leaflet' JavaScript library and the 'htmlwidgets' package.
-These maps can be used directly from the R console, from 'RStudio', in Shiny applications and R [@cheng2018].",
-         crosstalk = 'Provides building blocks for allowing HTML widgets to communicate with each other, with Shiny or without (i.e. static .html files) [@cheng2016].')
-
-tab <- data.frame(names(pkg), unname(pkg))
-names(tab) <- NULL
-# names(tab) <- c("**Package**", "**Description**")
-tab[, 1] <- paste0("**`", tab[, 1], "`**")
-
-knitr::kable(tab, caption = "Names and descriptions of the key packages used to build the interactive applications presented in this paper.
-Descriptions are directly from the `DESCRIPTION` file of each package.")
-
-```
-
-# Data exploration
-
-The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both.
-As such, stock assessment biologists need to manage large volumes of data from a variety of sources.
-Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in relative abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages.
-For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality.
-All of the above-mentioned data sets are complex and, as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each.
-It has long been recognized that this challenge can be mediated, to a degree, by the application of interactive visualization tools as they allow detailed exploration of the data behind a plot [@tukey1985; @fisherkeller1988].
-For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data.
-These abilities are illustrated  by two recently developed interactive tools: 1) a tool designed to quickly examine fishery-independent survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.
-
-## Survey data
-
-Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018].
-These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981].
-Both the inputs and outputs from RStrap analyses are very large and, depending on the species, may include > 40 years and span the majority of the Newfoundland and Labrador shelf.
-In order for fisheries scientists to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)).
-RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results.
-The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session.
-Both crosstalk and plotly were incorporated to allow the user to interact with the data.
-
-RStrap Explorer contains four primary tabs: 1) "Survey Indices" contains stock level estimates of biomass and abundance (Figure \@ref(fig:biomass)), 2) "Age & Length Distributions" contains length and age frequency plots (Figure \@ref(fig:freq)), 3) "Recruitment" displays recruitment indices, and 4) "Help" provides additional context to the survey data and analysis.
-The input interface for RStrap Explorer is divided into two parts: basic and advanced inputs.
-Basic inputs are those like species, season, and survey year(s).
-Advanced inputs allow the user to define the length or age of recruitment, select whether they would like analyses conducted by length and/or age, and so on.
-This application was built for data exploration and allow the user to obtain quick and easy visualization of survey data, before the thorough data checking and analysis required for formal stock assessment begins.
-
-```{r biomass, echo = FALSE, fig.cap = '- Screen shot of the "Survey Indices" page of the RStrap Explorer application.
-The "Basic Inputs" sidebar is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest.
-The "Biomass Index" sub-tab is illustrated here.
-This sub-tab shows a time-series of biomass data with 95% confidence intervals, with a break in the lines when the survey gear changed from an Engle to Campelen trawl, for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889).
-All trend lines can be toggled off using plotly’s dynamic user interface.'}
-knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
-```
-
-```{r freq, echo = FALSE, fig.cap = '- "Age and Length Frequencies" tab of the RStrap Explorer tool.
-This tab illustrates the length frequency distributions of both cod (*Gadus morhua*; species code 438) and American plaice (*Hippoglossoides platessoides*; species code 889) binned into 2 cm length bins.
-The binning is easily modified using the "Length Group" field in the "Advanced Options" drop down menu (shown on left side).'}
-knitr::include_graphics("figures/RStrap_dashboard_freq.png")
-```
-
-## Tagging data
-
-Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day.
-The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs.
-This database has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades.
-The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective.
-To begin to explore and understand this large data set, we built a simple application using shiny and leaflet (Table \@ref(tab:pkg)) that allows a user to quickly and dynamically subset the data (e.g.
-ranges of release and capture years, specific geographic locations).
-As the application developed, it was incorporated into shinydashboard, to take advantage of the number of layout options available in shinydashboard (Table \@ref(tab:pkg)).
-Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging.
-The markercluster (https://github.com/Leaflet/Leaflet.markercluster) function available in leaflet was particularly useful as a means to dynamically scale the level of pooling of spatial points (Figure \@ref(fig:map)).
-This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, as well as explore options on how to spatially pool the data for subsequent demographic analysis.
-Further tabs were added to provide basic summaries of the selected data (Figure \@ref(fig:summary)).
-
-```{r map, echo = FALSE, fig.cap = 'Screen shot of cod tag mapping tool using shinydashboard and leaflet.
-The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales.
-Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date captured), which is particularly useful when error checking.'}
-knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
-```
-
-```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool.
-In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
-knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")
-```
-
-# Data modeling
-
-Synthesizing data from multiple sources presents a key challenge to stock assessment.
-Analyses of different data sources were traditionally carried out independently and the summaries or parameters from these analyses were used in the assessment model.
-This approach, however, is less than ideal because information may be lost and uncertainty may be unaccounted for when we "do statistics on the statistics" [@link1999; @maunder2013].
-Such issues have largely been curtailed in contemporary stock assessments thanks to advances in statistical computing that have facilitated the analysis of all available data, in as raw a form as appropriate, in a single integrated analysis [@maunder2013].
-Specifically, statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015] allow the construction of a joint likelihood for an array of observations to, in theory, extract as much information as possible about the biological and fishery processes.
-From a computational perspective, analyses of a variety of large data sets has never been easier.
-However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of understanding and integrating data from multiple sources into a single model and communicating the methods and results to stakeholders and fisheries managers.
-This challenge was palpable for the data-rich case of Northern cod.
-
-The Northern cod stock off southern Labrador and eastern Newfoundland is one of the most well studied stocks in eastern Canada.
-As such, there are multiple monitoring programs that help inform the status of the stock and data from most of these programs have been integrated into a state-space stock assessment model, called NCAM [@cadigan2016].
-The model includes information from research vessel autumn trawl surveys (1983-present), Sentinel fishery surveys (1995-present), inshore acoustic surveys (1995-2009), fishery catch-at-age compositions and partial fishery landings (1983-present), and tagging (1983-present).
-Using a series of observation equations, this TMB based model simplifies thousands of historical data points into quantities such as recruitment, spawning stock biomass, fishing mortality and natural mortality.
-Once the model is fit to the data, the next step is to produce visual representations of the data and model output.
-The traditional approach would involve producing static presentations and documents with a series of figures and tables.
-However, with large amounts of model inputs and outputs, this approach quickly becomes overwhelming for both the analyst and the stakeholders involved for at least two reasons.
-First, it is no longer feasible for the analyst to include and describe every figure and table produced in a single document.
-Second, it is difficult for stakeholders to efficiently digest the information that has been compressed into a series of static slides or pages.
-Interactive documents provide a potential solution to this problem as they allow much more information to be contained and accessible on a single screen.
-
-In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019].
-We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document.
-We also used the crosstalk package to link the data displayed across multiple plots.
-Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.
-The intent of this tool was not to replace detailed documentation on the model [i.e.
-@cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the stock and model.
-Explanatory text and detail were therefore kept to a minimum, allowing the audience to explore the figures and data unencumbered by verbose descriptions.
-
-The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background").
-Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Landings"; accessed from the "Fishery" drop-down menu), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data.
-For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)).
-The dashboard also includes pages focused on model estimates such as selectivity ("Selectivity"; accessed from the "Fishery" drop-down menu) and catchability ("Catchability"; accessed from the "Surveys" drop-down menu).
-Pages for viewing trends in various population parameters are included under the "Trends" drop-down menu, where the "Retro" page displays results from a retrospective analysis, the "Comps" page displays population trends from different models, the "Status and rates" page displays trends in stock size and vital rates (Figure \@ref(fig:trends)), and the "Productivity" page focuses on trends in SSB and recruitment.
-Finally, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages).
-The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents.
-However, there are two key benefits of this approach over the *modus operandi* of producing static documents and slides.
-First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e.
-both broader patterns and finer details in the data are accessible via zooming and tooltips).
-Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g.
-slide shows can easily be made using R markdown).
-Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.
-For instance, explorations of the residual plots in the dashboard allowed the authors to quickly identify conflicts between two surveys used in the model.
-Potential solutions to this problem were then explored by modifying the configuration of the model and these changes were quickly evaluated using a dashboard that is produced in seconds.
-
-The dashboard also makes it easy to share the results with colleagues and stakeholders as it is rendered into a self-contained html file.
-This allows others to independently scrutinize details of both data and the model that are typically only accessible to the analyst.
-Such access improves the transparency of the stock assessment model which, in turn, leads to richer discussion and scrutiny of the biological and statistical rigor of the model.
-For instance, visualizations depicting the model assumptions, process errors and confidence intervals around the projections raised important questions on the impact of the assumptions on the projections from the model.
-These questions were raised during the first assessment meeting in which this tool was used and the ensuing discussions left us with the impression that this tool helped the assessment biologists, academics, resource managers and fisheries representatives in the room gain a deeper understanding of the uncertainties and the patterns in the risk tables.
-Feedback and personal testimony from multiple stakeholders who have used the dashboard reaffirmed our impression that the tool is helpful as there was general agreement that access to the details eases communication and relieves any potential concerns that something is hidden.
-
-```{r rv, echo = FALSE, fig.cap = 'Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero).
-Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
-knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
-```
-
-```{r trends, echo = FALSE, fig.cap = 'Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
-knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
-```
-
-# From bespoke to general tools
-
-The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations but the solutions presented are specific to the Northwest Atlantic Fisheries Centre.
-The concept is therefore broadly applicable, however, the underlying code will require modifications to accommodate region and stock specific data management and modelling approaches.
-Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications when applied to a model developed for another cod stock.
-Nevertheless, considerable chunks of code were recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard.
-Additionally, a French colleague is currently modifying the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014].
-Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g.
-StoX software; @johnsen2019].
-These are, of course, early days in the development and application of interactive tools to stock assessment.
-Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards depending on user needs and preferences.
-Regardless of the approach, a common goal is to make the workflow more efficient, reproducible and transparent.
-
-# Towards open stock assessment
-
-The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018].
-The majority of this sequence of events, sometimes called "the data pipeline" [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis.
-However, decisions made along the data pipeline increasingly influence the outcome of the study.
-Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis.
-Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible.
-A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015].
-Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009; @lowndes2017].
-
-A culture of openness is actively evolving in the field of stock assessment and, at this point, it is not hard to picture a future where it is commonplace for the the full data to document pipeline to be open and accessible.
-Open resources are actively being developed and used to improve accessibility to all the data, methods and results that feed into stock assessments [e.g.
-the Transparent Assessment Framework, https://taf.ices.dk/app/about; @magnusson2020].
-Likewise, tools are being developed to facilitate the creation of reproducible advisory documents [e.g.
-@ices2018; @anderson2020; @regular2020], and these documents can be interactive, as is the case for the VISA tool [@ices2018], or they can be supplemented by interactive tools.
-There is plenty of opportunity for synergies between these initiatives, as most are being developed using R, and all details and decisions can be traceable and accessible using an online repository such as GitHub.
-We therefore foresee a future where stock assessments become increasingly open and reproducible.
-
-Providing open access to code and data, however, does not necessarily make a stock assessment more accessible.
-We believe that interactive tools, such as those presented here and in Appendix A, represent an important a step forward in terms of open science as they make the peer review process more efficient.
-Although these tools do not reveal the entire data pipeline, a wide range of peers can easily access and review details that may be inaccessible even if the pipeline is open.
-For example, it is unlikely that an individual has the time to fully review and recreate a particular analysis in its entirety when most participants are already overcommitted [@banks2011].
-Further, in the case of NCAM, considerable statistical experience and expertise is required simply to run the model.
-However, the dashboard approach removes many of these obstacles.
-The html file contains all model inputs, outputs and diagnostics in an open, and easy to use, interactive document that can be distributed prior to meetings.
-As such, participants can assess the results and diagnostics at their leisure.
-Specific experience with the tools used to generate the results are not required to constructively and critically review of the results presented in the dashboard.
-
-# Conclusions
-
-Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information.
-However, we must acknowledge there are costs to adopting the use of interactive tools.
-The first and obvious one being that staff are required to have both the time and training to effectively develop these tools.
-Ideally, professional programmers and graphic designers would be hired to construct dashboards and other interactive info-graphics [@mcinerny2014], but fisheries management organizations rarely have the resources to fund such initiatives.
-Consequently, the process of communicating fisheries science tends to be the responsibility of fisheries scientists who lack formal training in these fields.
-In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology.
-While some learning time was needed, the learning curve was relatively minor because there was no need to learn a new programming language (e.g.
-JavaScript, CSS, etc.).
-We can confirm that an R user can construct a basic dashboard within a day, using Supplement 1 as a guide, and a more custom dashboard in a matter of weeks.
-Moreover, a growing user base means that most programming issues can be readily addressed through simple internet searches.
-There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools.
-Once over the initial learning curve, the time investment may not be much greater than one would invest into a standard presentation.
-We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents.
-It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.
-
-# Acknowledgements
-
-We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful.
-We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar, Daniel Goethel and seven anonymous reviewers.
-
-All code used to generate this paper, Supplement 1 (getting started dashboard) and Supplement 2 (NCAM explorer) are available at https://github.com/PaulRegular/interactive-stock-assessment.
-
-# Appendix A
-
-Partial list of applications developed to explore and communicate marine and fisheries science:
-
-- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/); a benthic-focused tool to assist with marine license application and compliance monitoring.
-- Species Dashboard (https://shiny.marine.ie/speciesdash/); a tool for exploring biological fisheries data collected in Irish waters
-- The Stock Book (https://shiny.marine.ie/stockbook/); a tool for accessing up-to-date information on the status of fish stocks in Irish waters.
-- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
-- Toy Tuna MSE application (https://puntapps.shinyapps.io/tunamse/); a toy MSE example built to allow users to enter various management procedures and assess their performance.
-- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.
-- EcoCast Explorer (https://coastwatch.pfeg.noaa.gov/ecocast/explorer.html); an application for displaying results from a dynamic ocean management tool focused on bycatch avoidance [see @hazen2018 for more detail].
-- B-VBGM!
-Bayesian Von Bertalanffy Growth Model tool (https://fishecology.shinyapps.io/B-VBGM/); a supplement Catalán et al.
-[-@catalan2018] that facilitates the exploration of the model described in the paper.
-- Beyond Temperature tool (https://heatherwelch.shinyapps.io/beyond_temperature/); a supplement to McHenry et al.
-[-@mchenry2019] that allows the exploration of habitat suitability hindcasts and projections by species and model type.
-- BRSApp (https://catalinagomez.shinyapps.io/BRSApp/); a supplement to Gomez et al. [-@gomez2016] that allows the exploration of behavioural response severity scores of marine mammals to noise.
-- VISA tool (e.g.
-https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice [developed to meet an "EU request on dissemination of ICES advice beyond pdf files"; @ices2018].
-
-# References
+See [Supplement 2](NCAM_dashboard.html) for a sample dashboard built to explore the inputs and outputs of an age-structured stock assessment model

### 2020-07-23 analysis/site/render_site.R

@@ -1,0 +1,13 @@
+
+
+## Copy several files from the dashboard dir. The getting_started.Rmd should capture all 
+## css etc. dependancies needed for the NCAM_dashboard.html to work
+files <- c("getting_started.Rmd", "knit_button.png", 
+           "length-weight_data.csv", "skeleton.Rmd", 
+           "NCAM_dashboard.html")
+file.copy(file.path("analysis/dashboards", files),
+          file.path("analysis/site", files))
+
+rmarkdown::render_site("analysis/site")
+
+
